---
title: "Hypothesis 1 - Mother-daughter associations"
author: 
  name: "Flore Michel"
  email: flore.michel@unil.ch
  affiliation:
    name: University of Lausanne, Departement of Ecology and Evolution
execute:                            # Global options for code chunks
  echo: true                        # Show code globally
  warning: false                    # Hide warnings globally
  freeze: auto                      # Only re-render if code changes
  cache: true                       # Cache results (best for heavy computations)
format:                             # Global options for rendering the document
  html:                              # Defines render format in this case html
    margin-top: 30px
    toc: true                       # Includes table of content
    toc-location: left              # Sets the location of the toc
    toc-depth: 2                    # Sets the number of levels for the toc
    number-sections: true           # Add automatic section number
    number-depth: 2                 # Sets the maximum numberin level e.g. 2.2.1
    #code-fold: true                 # Code chunk can be expanded
    fig-cap-location: bottom        # Location of figure captions [top, bottom]
    lightbox: true                  # When clicking on a figure it is expanded
    embed-resources: true           # Make the report a stand-alone for sharing
    theme:                          # Set the theme of the report 
      light: litera                   # [default, cerulean, cosmo, flatly, journal, ...]
      dark: slate                   # [cyborg, darkly, solar, superhero, vapor, ...]
#bibliography: bib/workshop.bib.     # References used in the document
#csl: bib/apa.csl                    # Citation style for rendering the references
bibliography: references.bib
---

# Overview

This script addresses **Hypothesis 1**: do mother–daughter pairs show stronger associations than other pair types during the daughter's first summer ?

Three pair types are compared throughout:

- mother-daughter (MD)

- mother paired with an unrelated juvenile female

- all remaining female dyads ("other")

The analysis has two stages:

1.  **QAP** (year-specific): tests whether MD dyads are more strongly associated expected by chance, within each year separately.
2.  **GLMM** (cross-year): estimates the overall effect of pair type (3 type) on association strength across all eleven years, and tests whether this effect is modulated by colony size.

Both colonies (GB2 and UA) are analysed in parallel using identical pipelines.

----------------------

```{r packages}
library(dplyr)
library(tidyr)
library(igraph)
library(asnipe)
library(ggplot2)
library(patchwork)
library(here)
library(sna)        
library(glmmTMB)    #beta GLMM
library(DHARMa)     #residual diagnostics
library(performance) #check_overdispersion, check_collinearity
library(broom.mixed) #tidy model output
library(ggeffects)  #model predictions
library(lme4)       #glmer — used only for Cook's distance diagnostic
library(cowplot)
```

---

# 1 Load data

## 1.1 Edge lists

```{r load-edges}
edges_GB2 <- read.csv(
  here("data", "saved_matrices", "GB2", "edge_list_GB2.csv"),
  header = TRUE, sep = ",", check.names = FALSE
)

edges_UA <- read.csv(
  here("data", "saved_matrices", "UA", "edge_list_UA.csv"),
  header = TRUE, sep = ",", check.names = FALSE
)
```

## 1.2 Association matrices

```{r load-networks}
#| echo: true
#GB2
net_G2013<-read.csv(here("data","saved_matrices","GB2","net_G2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2013 <- as.matrix(net_G2013)
net_G2014<-read.csv(here("data","saved_matrices","GB2","net_G2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2014 <- as.matrix(net_G2014)
net_G2015<-read.csv(here("data","saved_matrices","GB2","net_G2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2015 <- as.matrix(net_G2015)
net_G2016<-read.csv(here("data","saved_matrices","GB2","net_G2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2016 <- as.matrix(net_G2016)
net_G2017<-read.csv(here("data","saved_matrices","GB2","net_G2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2017 <- as.matrix(net_G2017)
net_G2018<-read.csv(here("data","saved_matrices","GB2","net_G2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2018 <- as.matrix(net_G2018)
net_G2019<-read.csv(here("data","saved_matrices","GB2","net_G2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2019 <- as.matrix(net_G2019)
net_G2020<-read.csv(here("data","saved_matrices","GB2","net_G2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2020 <- as.matrix(net_G2020)
net_G2021<-read.csv(here("data","saved_matrices","GB2","net_G2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2021 <- as.matrix(net_G2021)
net_G2022<-read.csv(here("data","saved_matrices","GB2","net_G2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2022 <- as.matrix(net_G2022)
net_G2023<-read.csv(here("data","saved_matrices","GB2","net_G2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2023 <- as.matrix(net_G2023)

#UA
net_U2013<-read.csv(here("data","saved_matrices","UA","net_U2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2013 <- as.matrix(net_U2013)
net_U2014<-read.csv(here("data","saved_matrices","UA","net_U2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2014 <- as.matrix(net_U2014)
net_U2015<-read.csv(here("data","saved_matrices","UA","net_U2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2015 <- as.matrix(net_U2015)
net_U2016<-read.csv(here("data","saved_matrices","UA","net_U2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2016 <- as.matrix(net_U2016)
net_U2017<-read.csv(here("data","saved_matrices","UA","net_U2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2017 <- as.matrix(net_U2017)
net_U2018<-read.csv(here("data","saved_matrices","UA","net_U2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2018 <- as.matrix(net_U2018)
net_U2019<-read.csv(here("data","saved_matrices","UA","net_U2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2019 <- as.matrix(net_U2019)
net_U2020<-read.csv(here("data","saved_matrices","UA","net_U2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2020 <- as.matrix(net_U2020)
net_U2021<-read.csv(here("data","saved_matrices","UA","net_U2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2021 <- as.matrix(net_U2021)
net_U2022<-read.csv(here("data","saved_matrices","UA","net_U2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2022 <- as.matrix(net_U2022)
net_U2023<-read.csv(here("data","saved_matrices","UA","net_U2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2023 <- as.matrix(net_U2023)
```

Create the igraph object from the network matrices
```{r}
#| echo: true
#GB2
g_G2013<- graph_from_adjacency_matrix(net_G2013, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2016<- graph_from_adjacency_matrix(net_G2016, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2015<- graph_from_adjacency_matrix(net_G2015, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2016<- graph_from_adjacency_matrix(net_G2016, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2017<- graph_from_adjacency_matrix(net_G2017, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2018<- graph_from_adjacency_matrix(net_G2018, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2019<- graph_from_adjacency_matrix(net_G2019, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2020<- graph_from_adjacency_matrix(net_G2020, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2021<- graph_from_adjacency_matrix(net_G2021, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2022<- graph_from_adjacency_matrix(net_G2022, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2023<- graph_from_adjacency_matrix(net_G2023, mode = "undirected", weighted = TRUE, diag = FALSE)

#UA
g_U2013<- graph_from_adjacency_matrix(net_U2013, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2014<- graph_from_adjacency_matrix(net_U2014, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2015<- graph_from_adjacency_matrix(net_U2015, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2016<- graph_from_adjacency_matrix(net_U2016, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2017<- graph_from_adjacency_matrix(net_U2017, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2018<- graph_from_adjacency_matrix(net_U2018, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2019<- graph_from_adjacency_matrix(net_U2019, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2020<- graph_from_adjacency_matrix(net_U2020, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2021<- graph_from_adjacency_matrix(net_U2021, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2022<- graph_from_adjacency_matrix(net_U2022, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2023<- graph_from_adjacency_matrix(net_U2023, mode = "undirected", weighted = TRUE, diag = FALSE)
```

## 1.3 GBI matrices
```{r}
#| echo: true
#GB2
GBI_G2013<-read.csv(here("data","saved_matrices","GB2","GBI_G2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2013 <- as.matrix(GBI_G2013)
GBI_G2014<-read.csv(here("data","saved_matrices","GB2","GBI_G2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2014 <- as.matrix(GBI_G2014)
GBI_G2015<-read.csv(here("data","saved_matrices","GB2","GBI_G2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2015 <- as.matrix(GBI_G2015)
GBI_G2016<-read.csv(here("data","saved_matrices","GB2","GBI_G2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2016 <- as.matrix(GBI_G2016)
GBI_G2017<-read.csv(here("data","saved_matrices","GB2","GBI_G2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2017 <- as.matrix(GBI_G2017)
GBI_G2018<-read.csv(here("data","saved_matrices","GB2","GBI_G2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2018 <- as.matrix(GBI_G2018)
GBI_G2019<-read.csv(here("data","saved_matrices","GB2","GBI_G2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2019 <- as.matrix(GBI_G2019)
GBI_G2020<-read.csv(here("data","saved_matrices","GB2","GBI_G2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2020 <- as.matrix(GBI_G2020)
GBI_G2021<-read.csv(here("data","saved_matrices","GB2","GBI_G2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2021 <- as.matrix(GBI_G2021)
GBI_G2022<-read.csv(here("data","saved_matrices","GB2","GBI_G2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2022 <- as.matrix(GBI_G2022)
GBI_G2023<-read.csv(here("data","saved_matrices","GB2","GBI_G2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2023 <- as.matrix(GBI_G2023)

#UA
GBI_U2013<-read.csv(here("data","saved_matrices","UA","GBI_U2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2013 <- as.matrix(GBI_U2013)
GBI_U2014<-read.csv(here("data","saved_matrices","UA","GBI_U2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2014 <- as.matrix(GBI_U2014)
GBI_U2015<-read.csv(here("data","saved_matrices","UA","GBI_U2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2015 <- as.matrix(GBI_U2015)
GBI_U2016<-read.csv(here("data","saved_matrices","UA","GBI_U2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2016 <- as.matrix(GBI_U2016)
GBI_U2017<-read.csv(here("data","saved_matrices","UA","GBI_U2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2017 <- as.matrix(GBI_U2017)
GBI_U2018<-read.csv(here("data","saved_matrices","UA","GBI_U2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2018 <- as.matrix(GBI_U2018)
GBI_U2019<-read.csv(here("data","saved_matrices","UA","GBI_U2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2019 <- as.matrix(GBI_U2019)
GBI_U2020<-read.csv(here("data","saved_matrices","UA","GBI_U2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2020 <- as.matrix(GBI_U2020)
GBI_U2021<-read.csv(here("data","saved_matrices","UA","GBI_U2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2021 <- as.matrix(GBI_U2021)
GBI_U2022<-read.csv(here("data","saved_matrices","UA","GBI_U2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2022 <- as.matrix(GBI_U2022)
GBI_U2023<-read.csv(here("data","saved_matrices","UA","GBI_U2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2023 <- as.matrix(GBI_U2023)
```

## 1.4 colony size

```{r load-colony-size}
colony_size_GB2 <- read.csv(
  here("data", "saved_matrices", "GB2", "colony_size_df.csv"),
  header = TRUE, sep = ","
) %>% mutate(year = as.numeric(year))

colony_size_UA <- read.csv(
  here("data", "saved_matrices", "UA", "colony_size_UA.csv"),
  header = TRUE, sep = ","
) %>% mutate(year = as.numeric(year))
```

## 1.5 Mother–unrelated-daughter pair lists

```{r load-mom-juv}
mom_juv_GB2 <- read.csv(
  here("data", "saved_matrices", "GB2", "mom_juv_GB2.csv"),
  header = TRUE, sep = ",", check.names = FALSE
) %>%
  mutate(pair_type_2 = recode(pair_type_2,
                              "mother-random-juvenile" = "mother-unrelated-daughter"))

mom_juv_UA <- read.csv(
  here("data", "saved_matrices", "UA", "mom_juv_UA.csv"),
  header = TRUE, sep = ",", check.names = FALSE
) %>%
  mutate(pair_type_2 = recode(pair_type_2,
                              "mother-random-juvenile" = "mother-unrelated-daughter"))
```

------------------------------------------------------------------------
# 2  QAP analysis (year-specific)
Tests whether MD pairs are more strongly associated than all other dyads within
each year. This is a binary comparison (MD vs. non-MD)
## 2.1 functions
```{r}
#----- build M predictor matrix for QAP 
# create matrix with 1 for mother-daughter dyads, 0 otherwise.
build_m_matrix <- function(association_matrix, edges, the_year) {

  ids <- rownames(association_matrix)
  #edges for this year + the desired pair_type
  edges_subset <- subset(edges, year == the_year & pair_type == "mother-daughter")
  
  #M matrix of zeros with same row/col
  M <- matrix(0, nrow = length(ids), ncol = length(ids),
              dimnames = list(ids, ids))
  
  #M matrix with 1 for mother–daughter
  for (r in seq_len(nrow(edges_subset))) {
    bat_i <- edges_subset$from[r]
    bat_j <- edges_subset$to[r]
    #only if both bats
    if (bat_i %in% ids && bat_j %in% ids) {
      M[bat_i, bat_j] <- 1
      M[bat_j, bat_i] <- 1
    }
  }
  
  return(M)
}

#-----  run QAP for all years 
# separately for each year - 10 000 permutations
run_qap <- function(net_list,edges){
  results_list <- list()
  
  #for each year
  for (the_year in names(net_list)) {
    A_yr <- net_list[[the_year]]
  
    #build predictor matrix
    M <- build_m_matrix(A_yr, edges, the_year) 
    g <- list(A_yr, M)
    
    #run QAP
    qap_res <- qaptest(g,FUN = gcor,g1  = 1,g2  = 2,reps = 10000)
    
    #extract statistics
    s <- summary(qap_res)
    test_val <- s$testval
    p_greq   <- s$pgreq
    p_leeq   <- s$pleeq
    
  #create result table 
    results_df <- data.frame(
      year        = the_year,
      correlation = test_val,
      pval_upper  = p_greq,
      pval_lower  = p_leeq,
      stringsAsFactors = FALSE
    )
    
    results_list[[the_year]] <- results_df
  }
  
  all_results_df <- do.call(rbind, results_list)
  rownames(all_results_df) <- NULL
  
  return(all_results_df)
}
#----- create list of all the networks for each years

net_list_GB2 <- list(
  "2013" = net_G2013,"2014" = net_G2014,"2015" = net_G2015,"2016" = net_G2016,"2017" = net_G2017,"2018" = net_G2018,"2019" = net_G2019,"2020" = net_G2020,"2021" = net_G2021,"2022" = net_G2022,"2023" = net_G2023
)

net_list_UA <- list(
  "2013" = net_U2013,"2014" = net_U2014,"2015" = net_U2015,"2016" = net_U2016,"2017" = net_U2017,"2018" = net_U2018,"2019" = net_U2019,"2020" = net_U2020,"2021" = net_U2021,"2022" = net_U2022,"2023" = net_U2023
)
```

## 2.2 run QAP
```{r qap-run}
QAP_GB2 <- run_qap(net_list_GB2, edges_GB2)
QAP_UA  <- run_qap(net_list_UA,  edges_UA)
```

```{r qap-results}
cat("GB2\n"); print(QAP_GB2)
cat("\n UA \n");  print(QAP_UA)
```
## 2.1  Visualisation — association strength by year

Stars indicate QAP significance: * p < 0.05, ** p < 0.01, *** p < 0.001.

```{r qap-plot}
add_sig_stars <- function(qap_df) {
  qap_df %>%
    mutate(star = case_when(
      pval_upper < 0.001 ~ "***",
      pval_upper < 0.01  ~ "**",
      pval_upper < 0.05  ~ "*",
      TRUE            ~ ""
    ))
}

make_qap_plot <- function(edges, qap_df, colony_label) {
  edges_plot <- edges %>%
    mutate(year = as.character(year),
           pair_type = factor(pair_type,
                              levels = c("other", "mother-daughter")))
  stars <- add_sig_stars(qap_df) %>%
    mutate(year = as.character(year))

  ggplot(edges_plot, aes(x = year, y = weight, fill = pair_type)) +
    geom_boxplot(width = 0.6, outlier.size = 0.5, outlier.colour = "#c4c2c2") +
    geom_text(data = stars, aes(x = year, y = 0.97, label = star),
              inherit.aes = FALSE, size = 5) +
    scale_fill_manual(values = c("mother-daughter" = "#d18caf",
                                 "other"           = "#c4c2c2")) +
    labs(title = colony_label, x = NULL,
         y = "Association strength (SRI)", fill = "pair type") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5))
}

p_qap_GB2 <- make_qap_plot(edges_GB2, QAP_GB2, "Colony GB2")
p_qap_UA  <- make_qap_plot(edges_UA,  QAP_UA,  "Colony UA")

combined_qap <- (p_qap_GB2 / p_qap_UA) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

combined_qap

ggsave(here("results", "figures", "H1_QAP_combined.png"),
       plot = combined_qap, width = 8, height = 6, dpi = 300)
```

---

# 3  GLMM — cross-year analysis

## 3.1 Functions
```{r}
# Prepare 3 pair type edge list ----- 
#join all pair labels in the full edge liss -> unlabelled dyads become "other"
build_three_type_edges <- function(edges, mom_juv, colony_size) {
  edges %>%
    left_join(colony_size, by = "year") %>%
    left_join(mom_juv %>% select(pairID, year, pair_type_2),
              by = c("pairID", "year")) %>%
    mutate(
      pair_type_2 = if_else(is.na(pair_type_2), "other", pair_type_2),
      pair_type_2 = factor(pair_type_2,
                           levels = c("other", "mother-daughter",
                                      "mother-unrelated-daughter")),
      colony_size_c = as.numeric(scale(colony_size, center = TRUE, scale = FALSE))
    )
}

#node-label permutation p-values (two-tailed)----- 
compute_perm_pvalues <- function(obs_coef, perm_matrix) {
  sapply(names(obs_coef), function(nm)
    mean(abs(perm_matrix[, nm]) >= abs(obs_coef[nm]))
  )
}

#permutation table (observed vs permuted distribution)----- 
make_perm_table <- function(obs_coef, perm_matrix, p_values) {
  obs_df <- data.frame(term = names(obs_coef),
                       observed = as.numeric(obs_coef),
                       stringsAsFactors = FALSE)
  perm_summary <- as.data.frame(perm_matrix) %>%
    pivot_longer(everything(), names_to = "term", values_to = "value") %>%
    group_by(term) %>%
    summarise(perm_mean = mean(value), perm_sd = sd(value), .groups = "drop")
  obs_df %>%
    left_join(perm_summary, by = "term") %>%
    mutate(p_value = p_values[term]) %>%
    arrange(desc(observed))
}
```

## 3.2 edge lists - add pair types

```{r}
edge_GB2 <- build_three_type_edges(edges_GB2, mom_juv_GB2, colony_size_GB2)
edge_UA  <- build_three_type_edges(edges_UA,  mom_juv_UA,  colony_size_UA)
```

## 3.3  check for influential years (cook's distance)

Use of lme4 here in order to measure cook's distance (not possible with glmmTMB)

```{r cooks distance}
# GB2 -----
model_diag_GB2 <- glmer(
  weight ~ pair_type_2 * colony_size_c + (1 | pairID) + (1 | year),
  data = edge_GB2, family = binomial()
)
infl_GB2 <- influence(model_diag_GB2, group = "year")
plot(infl_GB2, which = "cook")
cooks.distance(infl_GB2)

# UA -----
model_diag_UA <- glmer(
  weight ~ pair_type_2 * colony_size_c + (1 | pairID) + (1 | year),
  data = edge_UA, family = binomial()
)
infl_UA <- influence(model_diag_UA, group = "year")
plot(infl_UA, which = "cook")
cooks.distance(infl_UA)
```

```{r remove years}
# GB2: Year 2019 identified as influential -> excluded
# UA: Year 2016 identified as influential -> excluded
edge_GB2_clean <- edge_GB2 %>% filter(year != 2019)
edge_UA_clean  <- edge_UA  %>% filter(year != 2016)
```

## 3.4 Weight transformation before GLMM 
Beta distribution requires values strictly between 0 and 1. Tranform following: y' = (y*(n-1) + 0.5) / n  (Smithson & Verkuilen, 2006)
```{r transformation}
#GB2
n <- nrow(edge_GB2_clean)
edge_GB2_clean$weight <- (edge_GB2_clean$weight * (n - 1) + 0.5) / n

#UA
n <- nrow(edge_UA_clean)
edge_UA_clean$weight <- (edge_UA_clean$weight * (n - 1) + 0.5) / n
```


## 3.5 final models

Two models per colony: one without interaction (main effects only) and one with
a pair type × colony size interaction to test whether the effect of pair type depends on demographic context

```{r models}
# GB2 -------
model_GB2 <- glmmTMB(
  weight ~ pair_type_2 + colony_size_c + (1 | pairID) + (1 | year),
  data   = edge_GB2_clean,
  family = beta_family()
)

model_GB2_int <- glmmTMB(
  weight ~ pair_type_2 * colony_size_c + (1 | pairID) + (1 | year),
  data   = edge_GB2_clean,
  family = beta_family()
)

# UA ------
model_UA <- glmmTMB(
  weight ~ pair_type_2 + colony_size_c + (1 | year),
  data   = edge_UA_clean,
  family = beta_family()
)

model_UA_int <- glmmTMB(
  weight ~ pair_type_2 * colony_size_c + (1 | year),
  data   = edge_UA_clean,
  family = beta_family()
)
```

```{r model summaries}
summary(model_GB2);     summary(model_GB2_int)
summary(model_UA);      summary(model_UA_int)
```

## 3.6 model diagnostics

```{r diagnostics}
check_model <- function(model, label) {
  cat("\n──", label, "──\n")
  print(check_overdispersion(model))
  print(check_collinearity(model))
  sim <- simulateResiduals(model, n = 1000)
  plot(sim, main = label)
}

check_model(model_GB2,     "GB2 — main effects")
check_model(model_GB2_int, "GB2 — interaction")
check_model(model_UA,      "UA  — main effects")
check_model(model_UA_int,  "UA  — interaction")
```

---

# 4  Permutation (node-label permutations)

For each permutation, pair-type labels are shuffled within year while preserving the number of pairs per type The GLMM is refitted on each permutations, and the coefficients are stored. P values are measured as the proportion of permuted coefficients that are equal or exceeding the observed coefficient

We did n = 1000 permutations for each model

## 4.1  GB2

```{r perm GB2}
set.seed(23)
n_perm <- 1000

#coefficients 
coef_names_GB2 <- c(
  "pair_type_2mother-daughter",
  "pair_type_2mother-unrelated-daughter",
  "colony_size_c"
)
coef_names_GB2_int <- c(
  "pair_type_2mother-daughter",
  "pair_type_2mother-unrelated-daughter",
  "colony_size_c",
  "pair_type_2mother-daughter:colony_size_c",
  "pair_type_2mother-unrelated-daughter:colony_size_c"
)

obs_coef_GB2     <- fixef(model_GB2)$cond[coef_names_GB2]
obs_coef_GB2_int <- fixef(model_GB2_int)$cond[coef_names_GB2_int]

perm_GB2     <- matrix(NA, n_perm, length(coef_names_GB2),
                       dimnames = list(NULL, coef_names_GB2))
perm_GB2_int <- matrix(NA, n_perm, length(coef_names_GB2_int),
                       dimnames = list(NULL, coef_names_GB2_int))

for (i in seq_len(n_perm)) {
  d <- edge_GB2_clean %>%
    group_by(year) %>%
    mutate(pair_type_2 = sample(pair_type_2)) %>%
    ungroup()

  m <- glmmTMB(weight ~ pair_type_2 + colony_size_c + (1 | pairID) + (1 | year),
               family = beta_family(), data = d)
  perm_GB2[i, ] <- fixef(m)$cond[coef_names_GB2]

  m_int <- glmmTMB(weight ~ pair_type_2 * colony_size_c + (1 | pairID) + (1 | year),
                   family = beta_family(), data = d)
  perm_GB2_int[i, ] <- fixef(m_int)$cond[coef_names_GB2_int]
}

save(perm_GB2, perm_GB2_int,
     file = here("results", "H1_permutations_GB2.RData"))
```

```{r perm-gb2-pvalues}
p_GB2     <- compute_perm_pvalues(obs_coef_GB2,     perm_GB2)
p_GB2_int <- compute_perm_pvalues(obs_coef_GB2_int, perm_GB2_int)

table_GB2     <- make_perm_table(obs_coef_GB2,     perm_GB2,     p_GB2)
table_GB2_int <- make_perm_table(obs_coef_GB2_int, perm_GB2_int, p_GB2_int)

cat("── GB2 main effects ──\n");  print(table_GB2)
cat("\n── GB2 interaction ──\n"); print(table_GB2_int)

write.csv(table_GB2,
          here("results", "tables", "H1_perm_GB2_main.csv"),     row.names = FALSE)
write.csv(table_GB2_int,
          here("results", "tables", "H1_perm_GB2_interaction.csv"), row.names = FALSE)
```

## 4.2  UA

```{r perm-ua, cache=TRUE}
set.seed(23)

#coefficients 
coef_names_UA <- c(
  "pair_type_2mother-daughter",
  "pair_type_2mother-unrelated-daughter",
  "colony_size_c"
)
coef_names_UA_int <- c(
  "pair_type_2mother-daughter",
  "pair_type_2mother-unrelated-daughter",
  "colony_size_c",
  "pair_type_2mother-daughter:colony_size_c",
  "pair_type_2mother-unrelated-daughter:colony_size_c"
)

obs_coef_UA     <- fixef(model_UA)$cond[coef_names_UA]
obs_coef_UA_int <- fixef(model_UA_int)$cond[coef_names_UA_int]

perm_UA     <- matrix(NA, n_perm, length(coef_names_UA),
                      dimnames = list(NULL, coef_names_UA))
perm_UA_int <- matrix(NA, n_perm, length(coef_names_UA_int),
                      dimnames = list(NULL, coef_names_UA_int))

for (i in seq_len(n_perm)) {
  d <- edge_UA_clean %>%
    group_by(year) %>%
    mutate(pair_type_2 = sample(pair_type_2)) %>%
    ungroup()

  m <- glmmTMB(weight ~ pair_type_2 + colony_size_c + (1 | year),
               family = beta_family(), data = d)
  perm_UA[i, ] <- fixef(m)$cond[coef_names_UA]

  m_int <- glmmTMB(weight ~ pair_type_2 * colony_size_c + (1 | year),
                   family = beta_family(), data = d)
  perm_UA_int[i, ] <- fixef(m_int)$cond[coef_names_UA_int]
}

save(perm_UA, perm_UA_int,
     file = here("results", "H1_permutations_UA.RData"))
```

```{r perm-ua-pvalues}
p_UA     <- compute_perm_pvalues(obs_coef_UA,     perm_UA)
p_UA_int <- compute_perm_pvalues(obs_coef_UA_int, perm_UA_int)

table_UA     <- make_perm_table(obs_coef_UA,     perm_UA,     p_UA)
table_UA_int <- make_perm_table(obs_coef_UA_int, perm_UA_int, p_UA_int)

cat(" UA main effects ----\n");  print(table_UA)
cat("\nUA interaction ----\n"); print(table_UA_int)

write.csv(table_UA,
          here("results", "tables", "H1_perm_UA_main.csv"),        row.names = FALSE)
write.csv(table_UA_int,
          here("results", "tables", "H1_perm_UA_interaction.csv"), row.names = FALSE)
```

---

# 5  Visualisation — model predictions
Using ggpredict
```{r model-predictions}
pred_GB2 <- ggpredict(model_GB2_int,
                      terms = c("colony_size_c [-25:25 by=1]", "pair_type_2"),
                      bias_correction = TRUE)

pred_UA  <- ggpredict(model_UA,
                      terms = c("colony_size_c [-10:10 by=1]", "pair_type_2"),
                      bias_correction = TRUE)

pair_colours <- c(
  "mother-daughter"           = "#C2347A",
  "mother-unrelated-daughter" = "#004488",
  "other"                     = "#999999"
)

make_pred_plot <- function(pred, title, show_y = TRUE) {
  ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
                alpha = 0.2, colour = NA) +
    geom_line(linewidth = 1) +
    scale_colour_manual(values = pair_colours) +
    scale_fill_manual(values = pair_colours) +
    labs(title = title,
         x = "Centred colony size",
         y = if (show_y) "Predicted association strength (SRI)" else NULL,
         colour = "Pair type", fill = "Pair type") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5))
}

p_pred_GB2 <- make_pred_plot(pred_GB2, "Colony GB2", show_y = TRUE)
p_pred_UA  <- make_pred_plot(pred_UA,  "Colony UA",  show_y = FALSE)

combined_pred <- p_pred_GB2 + p_pred_UA +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

combined_pred

ggsave(here("results", "figures", "H1_model_predictions.png"),
       plot = combined_pred, width = 8, height = 4, dpi = 300)
```

---

# 6  other descriptive statistics

```{r stats}
# mean + SD of SRI per pair type per year, and colony size per year
summarise_edges <- function(edges, colony_label) {
  cat("\nSRI summary by year and pair type —", colony_label, "--\n")
  edges %>%
    group_by(year, pair_type_2) %>%
    summarise(mean_sri = round(mean(weight), 3),
              sd_sri   = round(sd(weight),   3),
              n_dyads  = n(), .groups = "drop") %>%
    print()

  cat("\nColony size per year —", colony_label, "--\n")
  edges %>%
    group_by(year) %>%
    summarise(colony_size = mean(colony_size, na.rm = TRUE), .groups = "drop") %>%
    print()

  cat("\nNumber of MD pairs per year —", colony_label, "--\n")
  edges %>%
    filter(pair_type_2 == "mother-daughter") %>%
    group_by(year) %>%
    summarise(n_md = n(), .groups = "drop") %>%
    print()
}

summarise_edges(edge_GB2_clean, "GB2")
summarise_edges(edge_UA_clean,  "UA")
```

```{r density}
# mean network density across years
mean_density <- function(net_list, label) {
  d <- sapply(net_list, function(net) {
    g <- igraph::graph_from_adjacency_matrix(net, mode = "undirected",
                                             weighted = TRUE, diag = FALSE)
    igraph::edge_density(g, loops = FALSE)
  })
  cat("mean network density —", label, ":", round(mean(d), 4), "\n")
}

mean_density(net_list_GB2, "GB2")
mean_density(net_list_UA,  "UA")
```

---

# 7  Network plot
Import mother-daughter file to get juveniles ID
```{r}
mother_daughter_GB2 <- read.csv(here("data","mother_daughter_GB2.csv"))
mother_daughter_UA <- read.csv(here("data","mother_daughter_UA.csv"))
```

```{r}
# create function to create network plot
network_plot <- function(edge_df, net_matrix, md_file, yr, title) {
  
  edges_yr <- edge_df %>%
    filter(year == yr) %>%
    mutate(pair_type_2 = as.character(pair_type_2))
  
  juv <- md_file %>% filter(year == yr) %>% pull(rfid_off) %>% unique()
  
  g <- graph_from_adjacency_matrix(net_matrix, mode = "undirected",
                                   weighted = TRUE, diag = FALSE)
  set.seed(42)
  lay <- layout_with_fr(g, weights = E(g)$weight)
  
  # node colour (pair type) + shape (adult/juvenile)
  md_females <- unique(c(
    edges_yr$from[edges_yr$pair_type_2 == "mother-daughter"],
    edges_yr$to[edges_yr$pair_type_2   == "mother-daughter"]
  ))
  murd_females <- unique(c(
    edges_yr$from[edges_yr$pair_type_2 == "mother-unrelated-daughter"],
    edges_yr$to[edges_yr$pair_type_2   == "mother-unrelated-daughter"]
  ))
  
  nodes <- data.frame(name = V(g)$name, x = lay[, 1], y = lay[, 2]) %>%
    mutate(
      node_type = case_when(
        name %in% md_females   ~ "mother-daughter",
        name %in% murd_females ~ "mother-unrelated-daughter",
        TRUE                   ~ "other"
      ),
      is_juv = name %in% juv
    )
  
  el  <- as_edgelist(g) %>% as.data.frame() %>% rename(from = V1, to = V2)
  fwd <- el %>% left_join(edges_yr %>% select(from, to, pair_type_2), by = c("from", "to"))
  rev <- el %>% left_join(edges_yr %>% select(from, to, pair_type_2), by = c("from" = "to", "to" = "from"))
  
  edges_full <- fwd %>%
    mutate(pair_type_2 = if_else(is.na(pair_type_2), rev$pair_type_2, pair_type_2),
           pair_type_2 = if_else(is.na(pair_type_2), "other", pair_type_2)) %>%
    left_join(nodes %>% rename(from = name, x1 = x, y1 = y), by = "from") %>%
    left_join(nodes %>% rename(to   = name, x2 = x, y2 = y), by = "to")
  
  e_other <- edges_full %>% filter(pair_type_2 == "other")
  e_murd  <- edges_full %>% filter(pair_type_2 == "mother-unrelated-daughter")
  e_md    <- edges_full %>% filter(pair_type_2 == "mother-daughter")
  
  # split nodes by type and age for layered drawing
  n_other_adult <- nodes %>% filter(node_type == "other", !is_juv)
  n_other_juv   <- nodes %>% filter(node_type == "other",  is_juv)
  n_murd_adult  <- nodes %>% filter(node_type == "mother-unrelated-daughter", !is_juv)
  n_murd_juv    <- nodes %>% filter(node_type == "mother-unrelated-daughter",  is_juv)
  n_md_adult    <- nodes %>% filter(node_type == "mother-daughter", !is_juv)
  n_md_juv      <- nodes %>% filter(node_type == "mother-daughter",  is_juv)
  
  ggplot() +
    #edges with other behind and MD on top
    geom_segment(data = e_other, aes(x = x1, y = y1, xend = x2, yend = y2),
                 colour = pair_colours["other"], alpha = 0.3, linewidth = 0.2) +
    geom_segment(data = e_murd,  aes(x = x1, y = y1, xend = x2, yend = y2),
                 colour = pair_colours["mother-unrelated-daughter"],
                 alpha = 0.7, linewidth = 0.8) +
    geom_segment(data = e_md,    aes(x = x1, y = y1, xend = x2, yend = y2),
                 colour = pair_colours["mother-daughter"],
                 alpha = 0.9, linewidth = 1.1) +
    #other nodes
    geom_point(data = n_other_adult, aes(x = x, y = y), shape = 21, size = 2.3,
               fill = pair_colours["other"], colour = "white", stroke = 0.4) +
    geom_point(data = n_other_juv,   aes(x = x, y = y), shape = 24, size = 2.5,
               fill = pair_colours["other"], colour = "white", stroke = 0.4) +
    #M-NRD nodes
    geom_point(data = n_murd_adult,  aes(x = x, y = y), shape = 21, size = 2.5,
               fill = pair_colours["mother-unrelated-daughter"], colour = "white", stroke = 0.5) +
    geom_point(data = n_murd_juv,    aes(x = x, y = y), shape = 24, size = 2.7,
               fill = pair_colours["mother-unrelated-daughter"], colour = "white", stroke = 0.5) +
    # MD nodes on top
    geom_point(data = n_md_adult,    aes(x = x, y = y), shape = 21, size = 2.5,
               fill = pair_colours["mother-daughter"], colour = "white", stroke = 0.5) +
    geom_point(data = n_md_juv,      aes(x = x, y = y), shape = 24, size = 2.7,
               fill = pair_colours["mother-daughter"], colour = "white", stroke = 0.5) +
    labs(title = title) +
    theme_void(base_size = 11) +
    theme(plot.title = element_text(hjust = 0.5, size = 11, face = "plain"))
}

# function for histogram plot—————————
histogram_plot <- function(edge_df, yr) {
  
  edges_yr <- edge_df %>% filter(year == yr)
  
  m <- mean(edges_yr$weight, na.rm = TRUE)
  
  ggplot(edges_yr, aes(x = weight)) +
    geom_histogram(bins = 25, fill = "#BBBBBB", colour = "white",
                   linewidth = 0.15) +
    geom_vline(xintercept = m, colour = "red", linewidth = 0.6) +
    annotate("text", x = m + 0.02, y = Inf,
             label = paste0("mean = ", round(m, 2)),
             colour = "red", size = 2.8, hjust = 0, vjust = 1.6) +
    labs(x = "Association strength (SRI)", y = "Count") +
    theme_minimal(base_size = 10) +
    theme(legend.position  = "none",
          panel.grid.minor = element_blank(),
          axis.title       = element_text(size = 9),
          axis.text        = element_text(size = 8))
}

# create legends
leg_edges <- get_legend(
  ggplot(edge_GB2_clean %>%
           filter(year == 2023) %>%
           mutate(pair_type_2 = factor(pair_type_2,
                                       levels = c("other",
                                                  "mother-unrelated-daughter",
                                                  "mother-daughter"))),
         aes(x = weight, fill = pair_type_2)) +
    geom_histogram() +
    scale_fill_manual(values = pair_colours,
                      name   = "pair type:",
                      labels = c("other",
                                 "mother–unrelated juvenile",
                                 "mother–daughter")) +
    guides(fill = guide_legend(override.aes = list(alpha = 1))) +
    theme_minimal() +
    theme(legend.position = "bottom",
          legend.title    = element_text(size = 10),
          legend.text     = element_text(size = 9),
          legend.key.size = unit(0.4, "cm"))
)

leg_nodes <- get_legend(
  ggplot(data.frame(x = 1:2, y = 1, type = c("adult", "juvenile")),
         aes(x = x, y = y, colour = type)) +
    geom_point(alpha = 0) +
    scale_colour_manual(
      values = c("adult" = "#333333", "juvenile" = "#333333"),
      name   = "Individual",
      labels = c("adult", "juvenile"),
      guide  = guide_legend(
        override.aes = list(shape  = c(21, 24),
                            fill   = c("#333333", "#333333"),
                            colour = "white",
                            size   = 3.5,
                            alpha  = 1,
                            stroke = 0.5)
      )
    ) +
    theme_void() +
    theme(legend.position = "bottom",
          legend.title    = element_text(size = 10),
          legend.text     = element_text(size = 9),
          legend.key.size = unit(0.5, "cm"))
)

leg_combined <- plot_grid(leg_edges, leg_nodes, ncol = 1, rel_heights = c(1, 0.8))

# create both plots for both colonies —————————————
p_net_GB2  <- network_plot(edge_GB2_clean, net_G2023, mother_daughter_GB2, 2023, "Colony GB2")
p_hist_GB2 <- histogram_plot(edge_GB2_clean, 2023)

p_net_UA   <- network_plot(edge_UA_clean,  net_U2023, mother_daughter_UA,  2023, "Colony UA")
p_hist_UA  <- histogram_plot(edge_UA_clean, 2023)

row_GB2 <- p_net_GB2 + p_hist_GB2 + plot_layout(widths = c(2.5, 1))
row_UA  <- p_net_UA  + p_hist_UA  + plot_layout(widths = c(2.5, 1))

# combine with patchwork
combined <- wrap_elements(row_GB2) /
            wrap_elements(row_UA)  /
            wrap_elements(leg_combined) +
  plot_layout(heights = c(1, 1, 0.18))

combined

#save———————————
ggsave(here("results", "figures", "Fig1_networks_histograms.png"),
       plot = combined, width = 9, height = 7, dpi = 300)
```








