---
title: "Hypothesis 2 -  Associations stability"
author: 
  name: "Flore Michel"
  email: flore.michel@unil.ch
  affiliation:
    name: University of Lausanne, Departement of Ecology and Evolution
execute:                            # Global options for code chunks
  echo: true                        # Show code globally
  warning: false                    # Hide warnings globally
  message: false
  freeze: auto                      # Only re-render if code changes
  cache: true                       # Cache results (best for heavy computations)
format:                             # Global options for rendering the document
  html:                              # Defines render format in this case html
    margin-top: 30px
    toc: true                       # Includes table of content
    toc-location: left              # Sets the location of the toc
    toc-depth: 2                    # Sets the number of levels for the toc
    number-sections: true           # Add automatic section number
    number-depth: 2                 # Sets the maximum numberin level e.g. 2.2.1
    #code-fold: true                 # Code chunk can be expanded
    fig-cap-location: bottom        # Location of figure captions [top, bottom]
    lightbox: true                  # When clicking on a figure it is expanded
    embed-resources: true           # Make the report a stand-alone for sharing
    theme:                          # Set the theme of the report 
      light: litera                   # [default, cerulean, cosmo, flatly, journal, ...]
      dark: slate                   # [cyborg, darkly, solar, superhero, vapor, ...]
#csl: bib/apa.csl                    # Citation style for rendering the references
    bibliography: references.bib
---

# overview

This script addresses **Hypothesis 3**: do bonds formed early in life persist and strengthen across years, and does this differ between pair types? It also assess how temporal pattern can be influenced by demographic context, by comparing two colonies. 

In this analysis, we measured association strength from the entire year (in contrast to H1). The daughter's first year is therefore excluded from the analysis, since juveniles are marked in late August and do not have a full summer of presence to measure association strength.

Three pair types are compared:

mother-daughter (MD), mother paired with an unrelated juvenile female and all remaining female dyads (other)

The predictor in this analysis is **relationship age**, which is defined differently per pair type:
- MD pairs: daughter age (years since birth year recorded in mother-daughter file)
- all other pairs: years since the pair was first observed together in the dataset
  
Pairs already present in 2013 are marked with "unknown_age" as a control, since their true relationship age is unknown.

---

```{r packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(here)
library(glmmTMB)
library(DHARMa)
library(performance)
library(ggeffects)
library(broom.mixed)
```

---

# 1  load data (full years)

## 1.1 yearly edge list 
```{r}
#| echo: true
#GB2 —————
edge_G2013<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2013.csv"))
edge_G2013 <- edge_G2013 %>% mutate(year = 2013)
edge_G2014<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2014.csv"));
edge_G2014 <- edge_G2014 %>% mutate(year = 2014)
edge_G2015<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2015.csv"));
edge_G2015 <- edge_G2015 %>% mutate(year = 2015)
edge_G2016<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2016.csv"));edge_G2016 <- edge_G2016 %>% select(from,to,weight, pair_type);
edge_G2016 <- edge_G2016 %>% mutate(year = 2016)
edge_G2017<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2017.csv"));
edge_G2017 <- edge_G2017 %>% mutate(year = 2017)
edge_G2018<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2018.csv"));
edge_G2018 <- edge_G2018 %>% mutate(year = 2018)
edge_G2019<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2019.csv"));
edge_G2019 <- edge_G2019 %>% mutate(year = 2019)
edge_G2020<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2020.csv"));
edge_G2020 <- edge_G2020 %>% mutate(year = 2020)
edge_G2021<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2021.csv"));
edge_G2021 <- edge_G2021 %>% mutate(year = 2021)
edge_G2022<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2022.csv")); 
edge_G2022 <- edge_G2022 %>% mutate(year = 2022)
edge_G2023<-read.csv(here("data","saved_matrices","GB2","full","full_edge_G2023.csv"));
edge_G2023 <- edge_G2023 %>% mutate(year = 2023)

#create global edge list
edge_GB2_list<-list("2013"=edge_G2013,"2014"=edge_G2014,"2015"=edge_G2015,"2016"=edge_G2016,"2017"=edge_G2017,"2018"=edge_G2018,"2019"=edge_G2019,"2020"=edge_G2020,"2021"=edge_G2021,"2022"=edge_G2022,"2023"=edge_G2023)

edges_GB2_full<-rbind(edge_G2013,edge_G2014,edge_G2015,edge_G2016,edge_G2017,edge_G2018,edge_G2019,edge_G2020,edge_G2021,edge_G2022,edge_G2023)

edges_GB2_full$weight<-as.numeric(edges_GB2_full$weight)

#UA ————
edge_U2013<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2013.csv"))
edge_U2013 <- edge_U2013 %>% mutate(year = 2013)
edge_U2014<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2014.csv"));
edge_U2014 <- edge_U2014 %>% mutate(year = 2014)
edge_U2015<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2015.csv"));
edge_U2015 <- edge_U2015 %>% mutate(year = 2015)
edge_U2016<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2016.csv"));edge_U2016 <- edge_U2016 %>% select(from,to,weight, pair_type);
edge_U2016 <- edge_U2016 %>% mutate(year = 2016)
edge_U2017<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2017.csv"));
edge_U2017 <- edge_U2017 %>% mutate(year = 2017)
edge_U2018<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2018.csv"));
edge_U2018 <- edge_U2018 %>% mutate(year = 2018)
edge_U2019<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2019.csv"));
edge_U2019 <- edge_U2019 %>% mutate(year = 2019)
edge_U2020<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2020.csv"));
edge_U2020 <- edge_U2020 %>% mutate(year = 2020)
edge_U2021<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2021.csv"));
edge_U2021 <- edge_U2021 %>% mutate(year = 2021)
edge_U2022<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2022.csv")); 
edge_U2022 <- edge_U2022 %>% mutate(year = 2022)
edge_U2023<-read.csv(here("data","saved_matrices","UA","full","full_edges_U2023.csv"));
edge_U2023 <- edge_U2023 %>% mutate(year = 2023)

#create global edge list
edge_UA_list<-list("2013"=edge_U2013,"2014"=edge_U2014,"2015"=edge_U2015,"2016"=edge_U2016,"2017"=edge_U2017,"2018"=edge_U2018,"2019"=edge_U2019,"2020"=edge_U2020,"2021"=edge_U2021,"2022"=edge_U2022,"2023"=edge_U2023)
edges_UA_full<-rbind(edge_U2013,edge_U2014,edge_U2015,edge_U2016,edge_U2017,edge_U2018,edge_U2019,edge_U2020,edge_U2021,edge_U2022,edge_U2023)
edges_UA_full$weight<-as.numeric(edges_UA_full$weight)
```

## 1.2  mother-daughter files

```{r load}
mother_daughter_GB2 <- read.csv(here("data", "mother_daughter_GB2.csv")) %>%
  mutate(pairID = as.character(pairID))

mother_daughter_UA <- read.csv(here("data", "mother_daughter_UA.csv")) %>%
  mutate(pairID = as.character(pairID))
```

## 1.3  mother-unrelated-daughter files

```{r load}
mom_juv_GB2 <- read.csv(here("data", "saved_matrices", "GB2", "mom_juv_GB2.csv")) %>%
  mutate(pair_type_2 = recode(pair_type_2,
                              "mother-random-juvenile" = "mother-unrelated-daughter"))

mom_juv_UA <- read.csv(here("data", "saved_matrices", "UA", "mom_juv_UA.csv")) %>%
  mutate(pair_type_2 = recode(pair_type_2,
                              "mother-random-juvenile" = "mother-unrelated-daughter"))
```

## 1.4  colony size

```{r load-colony-size}
colony_size_GB2 <- read.csv(
  here("data", "saved_matrices", "GB2", "colony_size_df.csv")
) %>% mutate(year = as.numeric(year))

colony_size_UA <- read.csv(
  here("data", "saved_matrices", "UA", "colony_size_UA.csv")
) %>% mutate(year = as.numeric(year))
```

---

# 2  Prepare edge lists

# 2.1 function
This add pairID, relation_age and pair type labels to the edge lists
- relation_age for MD pairs = daughter age (years since birth)
- relation_age for other pairs = years since first observed together

add "unkown_age" column for the control 

and remove daughter's first summer (daughter_age == 1) (since the year is incomplete)
```{r}
prepare_edges <- function(edges_full, md_file, mom_juv, colony_size_df) {
  
  # create sorted pairID
  df <- edges_full %>%
    mutate(pairID = if_else(from < to,
                            paste(from, to, sep = "-"),
                            paste(to,   from, sep = "-")))
  
  # label MD pairs
  df <- df %>%
    mutate(ultimate_pair_type = if_else(
      pairID %in% md_file$pairID, "mother-daughter", "other"
    ))
  
  # add daughter age for MD pairs
  df <- df %>%
    group_by(pairID) %>%
    mutate(
      first_year_md = md_file$year[match(pairID, md_file$pairID)],
      daughter_age  = if_else(
        is.na(first_year_md), NA_real_, as.numeric(year - first_year_md + 1)
      )
    ) %>%
    ungroup()
  
  # remove first summer of MD pairs
  df <- df %>%
    filter(!(ultimate_pair_type == "mother-daughter" & daughter_age == 1))
  
  # relation age: daughter_age for MD, years-since-first for others
  df <- df %>%
    group_by(pairID) %>%
    mutate(first_year_overall = min(year)) %>%
    ungroup() %>%
    mutate(relation_age = if_else(
      ultimate_pair_type == "mother-daughter",
      daughter_age,
      as.numeric(year - first_year_overall + 1)
    ))
  
  # unknown_age flag: other pairs already present in 2013
  df <- df %>%
    group_by(pairID) %>%
    mutate(unknown_age = factor(if_else(
      ultimate_pair_type == "other" & first_year_overall == 2013, TRUE, FALSE
    ))) %>%
    ungroup()
  
  # add mother-unrelated daughter labels
  df <- df %>%
    left_join(mom_juv %>% select(pairID, pair_type_2), by = "pairID") %>%
    mutate(ultimate_pair_type = case_when(
      pair_type_2 == "mother-daughter"           ~ "mother-daughter",
      pair_type_2 == "mother-unrelated-daughter" ~ "mother-unrelated-daughter",
      TRUE                                       ~ ultimate_pair_type
    ))
  
  # add colony size and center it
  df <- df %>%
    left_join(colony_size_df, by = "year") %>%
    mutate(colony_size_c = as.numeric(
      scale(colony_size, center = TRUE, scale = FALSE)
    ))
  
  # beta transformation before model
  n <- nrow(df)
  df <- df %>%
    mutate(weight_beta = (weight * (n - 1) + 0.5) / n)
  
  # pair type as factor and relevel 
  df <- df %>%
    mutate(
      ultimate_pair_type = factor(ultimate_pair_type,
                                  levels = c("other",
                                             "mother-unrelated-daughter",
                                             "mother-daughter")),
      # center relation_age to reduce collinearity in interaction model
      relation_age_c = relation_age - mean(relation_age, na.rm = TRUE)
    )
  
  df
}
```

# 2.2 edge list
```{r}
edges_GB2_final <- prepare_edges(edges_GB2_full, mother_daughter_GB2,
                                 mom_juv_GB2, colony_size_GB2)

edges_UA_final  <- prepare_edges(edges_UA_full,  mother_daughter_UA,
                                 mom_juv_UA,  colony_size_UA)
```

# 2.3 check 
How many "unkown_age" 
```{r}
# GB2
edges_GB2_final %>%
  summarise(
    total_dyads   = n_distinct(pairID),
    unknown_age_dyads = n_distinct(pairID[unknown_age == TRUE]),
    prop_unknown  = round(unknown_age_dyads / total_dyads, 2)
  )

# UA
edges_UA_final %>%
  summarise(
    total_dyads   = n_distinct(pairID),
    unknown_age_dyads = n_distinct(pairID[unknown_age == TRUE]),
    prop_unknown  = round(unknown_age_dyads / total_dyads, 2)
  )
```

Check relationship age range
```{r}
edges_GB2_final %>% summarise(max_ra = max(relation_age, na.rm = TRUE),
                               mean_ra = mean(relation_age, na.rm = TRUE))
edges_UA_final %>%  summarise(max_ra = max(relation_age, na.rm = TRUE),
                               mean_ra = mean(relation_age, na.rm = TRUE))
```


---

# 3  Fit models
Two models per colony: main effects (no interaction) and interaction between relation_age and pair type. We are mainly interested in the interaction model, to test whether the temporal trajectory of association strength differs between pair type. However, we are checking whether the interaction is meaningful here

```{r fit-models}
#GB2—————
model_age_GB2 <- glmmTMB(
  weight_beta ~ relation_age_c + ultimate_pair_type + colony_size_c + unknown_age +
    (1 | pairID) + (1 | year),
  data   = edges_GB2_final,
  family = beta_family()
)

model_age_int_GB2 <- glmmTMB(
  weight_beta ~ relation_age_c * ultimate_pair_type + colony_size_c + unknown_age +
    (1 | pairID) + (1 | year),
  data   = edges_GB2_final,
  family = beta_family()
)

# UA—————
model_age_UA <- glmmTMB(
  weight_beta ~ relation_age_c + ultimate_pair_type + colony_size_c + unknown_age +
    (1 | pairID) + (1 | year),
  data   = edges_UA_final,
  family = beta_family()
)

model_age_int_UA <- glmmTMB(
  weight_beta ~ relation_age_c * ultimate_pair_type + colony_size_c + unknown_age +
    (1 | pairID) + (1 | year),
  data   = edges_UA_final,
  family = beta_family()
)
```

Summaries
```{r}
summary(model_age_GB2);    summary(model_age_int_GB2)
summary(model_age_UA);     summary(model_age_int_UA)
```

# 4 Models diagnostics

```{r diagnostics}
check_model <- function(model, label) {
  cat("\n──", label, "──\n")
  print(check_overdispersion(model))
  print(check_collinearity(model))
  sim <- simulateResiduals(model, n = 1000)
  plot(sim, main = label)
}

check_model(model_age_GB2,     "GB2 — main effects")
check_model(model_age_int_GB2, "GB2 — interaction")
check_model(model_age_UA,      "UA  — main effects")
check_model(model_age_int_UA,  "UA  — interaction")
```

KS test and outlier test are significant across all models but is expected given the large sample size. permutation-based inference further accounts for any residual non-independence in our data

---

# 5  Permutations 
Node-label permutations within each year (1000 permutations) 

## 5.1 Functions
#### p-value function (two-tailed)
```{r}
#two-tailed permutation p-values
compute_perm_pvalues <- function(obs_coef, perm_matrix) {
  sapply(names(obs_coef), function(nm)
    mean(abs(perm_matrix[, nm]) >= abs(obs_coef[nm]))
  )
}
```

#### Permutation results table
```{r}
make_perm_table <- function(obs_coef, perm_matrix, p_values) {
  data.frame(term = names(obs_coef), observed = as.numeric(obs_coef),
             stringsAsFactors = FALSE) %>%
    left_join(
      as.data.frame(perm_matrix) %>%
        pivot_longer(everything(), names_to = "term", values_to = "value") %>%
        group_by(term) %>%
        summarise(perm_mean = mean(value), perm_sd = sd(value), .groups = "drop"),
      by = "term"
    ) %>%
    mutate(p_value = p_values[term]) %>%
    arrange(desc(observed))
}
```

## 5.2  GB2
```{r}
set.seed(23)
n_perm <- 1000

coef_names_GB2 <- c(
  "relation_age_c",
  "ultimate_pair_typemother-unrelated-daughter",
  "ultimate_pair_typemother-daughter",
  "colony_size_c"
)
coef_names_int_GB2 <- c(
  "relation_age_c",
  "ultimate_pair_typemother-unrelated-daughter",
  "ultimate_pair_typemother-daughter",
  "colony_size_c",
  "relation_age_c:ultimate_pair_typemother-unrelated-daughter",
  "relation_age_c:ultimate_pair_typemother-daughter"
)

obs_GB2     <- fixef(model_age_GB2)$cond[coef_names_GB2]
obs_int_GB2 <- fixef(model_age_int_GB2)$cond[coef_names_int_GB2]

perm_GB2     <- matrix(NA, n_perm, length(coef_names_GB2),
                       dimnames = list(NULL, coef_names_GB2))
perm_int_GB2 <- matrix(NA, n_perm, length(coef_names_int_GB2),
                       dimnames = list(NULL, coef_names_int_GB2))

for (i in seq_len(n_perm)) {
  d <- edges_GB2_final %>%
    group_by(year) %>%
    mutate(ultimate_pair_type = sample(ultimate_pair_type)) %>%
    ungroup()
  
  m <- glmmTMB(
    weight_beta ~ relation_age_c + ultimate_pair_type + colony_size_c + unknown_age +
      (1 | pairID) + (1 | year),
    family = beta_family(), data = d
  )
  perm_GB2[i, ] <- fixef(m)$cond[coef_names_GB2]
  
  m_int <- glmmTMB(
    weight_beta ~ relation_age_c * ultimate_pair_type + colony_size_c + unknown_age +
      (1 | pairID) + (1 | year),
    family = beta_family(), data = d
  )
  perm_int_GB2[i, ] <- fixef(m_int)$cond[coef_names_int_GB2]
}

save(perm_GB2, perm_int_GB2,
     file = here("results", "H3_permutations_GB2.RData"))
```

```{r perm gb2 - pval}
p_GB2     <- compute_perm_pvalues(obs_GB2,     perm_GB2)
p_int_GB2 <- compute_perm_pvalues(obs_int_GB2, perm_int_GB2)

table_GB2     <- make_perm_table(obs_GB2,     perm_GB2,     p_GB2)
table_int_GB2 <- make_perm_table(obs_int_GB2, perm_int_GB2, p_int_GB2)

cat("GB2 main effects ---\n");   print(table_GB2)
cat("\nGB2 interaction ---\n");  print(table_int_GB2)

write.csv(table_GB2,
          here("results", "tables", "H3_perm_GB2_main.csv"),        row.names = FALSE)
write.csv(table_int_GB2,
          here("results", "tables", "H3_perm_GB2_interaction.csv"), row.names = FALSE)
```

## 6.2  UA

```{r perm ua, cache=TRUE}
set.seed(23)

#coeff
coef_names_UA <- c(
  "relation_age_c",
  "ultimate_pair_typemother-unrelated-daughter",
  "ultimate_pair_typemother-daughter",
  "colony_size_c"
)
coef_names_int_UA <- c(
  "relation_age_c",
  "ultimate_pair_typemother-unrelated-daughter",
  "ultimate_pair_typemother-daughter",
  "colony_size_c",
  "relation_age_c:ultimate_pair_typemother-unrelated-daughter",
  "relation_age_c:ultimate_pair_typemother-daughter"
)

obs_UA     <- fixef(model_age_UA)$cond[coef_names_UA]
obs_int_UA <- fixef(model_age_int_UA)$cond[coef_names_int_UA]

perm_UA     <- matrix(NA, n_perm, length(coef_names_UA),
                      dimnames = list(NULL, coef_names_UA))
perm_int_UA <- matrix(NA, n_perm, length(coef_names_int_UA),
                      dimnames = list(NULL, coef_names_int_UA))

for (i in seq_len(n_perm)) {
  d <- edges_UA_final %>%
    group_by(year) %>%
    mutate(ultimate_pair_type = sample(ultimate_pair_type)) %>%
    ungroup()
  
  m <- glmmTMB(
    weight_beta ~ relation_age_c + ultimate_pair_type + colony_size_c + unknown_age +
      (1 | pairID) + (1 | year),
    family = beta_family(), data = d
  )
  perm_UA[i, ] <- fixef(m)$cond[coef_names_UA]
  
  m_int <- glmmTMB(
    weight_beta ~ relation_age_c * ultimate_pair_type + colony_size_c + unknown_age +
      (1 | pairID) + (1 | year),
    family = beta_family(), data = d
  )
  perm_int_UA[i, ] <- fixef(m_int)$cond[coef_names_int_UA]
}

save(perm_UA, perm_int_UA,
     file = here("results", "H3_permutations_UA.RData"))
```

```{r perm-ua-pvalues}
p_UA     <- compute_perm_pvalues(obs_UA,     perm_UA)
p_int_UA <- compute_perm_pvalues(obs_int_UA, perm_int_UA)

table_UA     <- make_perm_table(obs_UA,     perm_UA,     p_UA)
table_int_UA <- make_perm_table(obs_int_UA, perm_int_UA, p_int_UA)

cat("UA main effects ---\n");    print(table_UA)
cat("\n UA interaction ---\n");   print(table_int_UA)

write.csv(table_UA,
          here("results", "tables", "H3_perm_UA_main.csv"),        row.names = FALSE)
write.csv(table_int_UA,
          here("results", "tables", "H3_perm_UA_interaction.csv"), row.names = FALSE)
```

---

# 7  plot — model predictions

Relation age is back-transform to real years by adding the mean relation age back
```{r mean}
cat("mean relation_age GB2:", round(mean(edges_GB2_final$relation_age, na.rm = TRUE), 2), "\n")
cat("mean relation_age UA: ", round(mean(edges_UA_final$relation_age,  na.rm = TRUE), 2), "\n")
```

```{r predictions}
pair_colours <- c(
  "mother-daughter"           = "#C2347A",
  "mother-unrelated-daughter" = "#004488",
  "other"                     = "#BBBBBB"
)

pred_GB2 <- ggpredict(model_age_int_GB2,
                      terms = c("relation_age_c", "ultimate_pair_type"),
                      bias_correction = TRUE)

pred_UA  <- ggpredict(model_age_int_UA,
                      terms = c("relation_age_c", "ultimate_pair_type"),
                      bias_correction = TRUE)

make_pred_plot <- function(pred, title, show_y = TRUE) {
  ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
                alpha = 0.2, colour = NA) +
    geom_line(linewidth = 1) +
    scale_colour_manual(values = pair_colours,
                        name   = "pair type",
                        labels = c("other",
                                   "mother–unrelated juvenile",
                                   "mother–daughter")) +
    scale_fill_manual(values = pair_colours,
                      name   = "pair type",
                      labels = c("other",
                                 "mother–unrelated juvenile",
                                 "mother–daughter")) +
    
    scale_x_continuous(breaks = seq(-3, 8, by = 1), 
                       labels = seq(-3, 8, by = 1) + 3) +

    
    coord_cartesian(ylim = c(0.1, 0.9)) +
  
    labs(title = title,
         x = "relationship age (years)",
         y = if (show_y) "Predicted association strength (SRI)" else NULL) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, size = 12, face = "plain"))
}

p1 <- make_pred_plot(pred_GB2, "Colony GB2", show_y = TRUE)
p2 <- make_pred_plot(pred_UA,  "Colony UA",  show_y = FALSE)

combined_plot <- p1 + p2 +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

combined_plot

ggsave(here("results", "figures", "H3_age_modelprediction.png"),
       plot = combined_plot, width = 7, height = 4, dpi = 300)
```

---

