---
title: "Thesis-Analysis-GB2"
output: word_document
date: "2025-02-04"
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/flo_r/Desktop/master thesis/data/data in CSV/")

library(igraph)
library(asnipe)
library(sna)
library(assocInd)
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(tidyverse)
library(viridis)
library(patchwork)
library(hrbrthemes)
library(igraph)
library(ggraph)
library(colormap)
library(tidyr)
library(dplyr)
library(data.table)
library(tidygraph)
library(RColorBrewer)
```

# GB2 2013 — (￣\^￣)ゞ

```{r}
GB2_2013<-read.csv("CSV_Dayroosts_GB2_2013_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)
GB2_2013<-GB2_2013[c(0:30)]
GB2_2013 <- GB2_2013[, -4]
GB2_2013 <- GB2_2013[, -3]
GB2_2013 <- GB2_2013[, -2]
colnames(GB2_2013)[colnames(GB2_2013) == "number"] <- "date"
```

##GBI matrix
```{r}
data_G2013<- GB2_2013 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2013$daygroup<-as.factor(paste(data_G2013$date, data_G2013$box, sep="_"))

ind_G2013<-data.table(ID=c(data_G2013$ID), group=data_G2013$daygroup, day=c(data_G2013$date)) #df with 3 col: ID, group and day 
GBI_G2013<- get_group_by_individual(ind_G2013, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2013 <- as.matrix(GBI_G2013)
```

## network
```{r}
net_G2013<-get_network(GBI_G2013,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2013 <- graph_from_adjacency_matrix(net_G2013,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2013,vertex.label = V(g_G2013)$name,edge.width = E(g_G2013)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue") 
```
##### threshold?
```{r}
threshold_G2013 <- mean(net_G2013, na.rm = TRUE) +  sd(net_G2013, na.rm = TRUE)

gfilter_G2013 <- delete_edges(g_G2013, E(g_G2013)[weight < threshold_G2013]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2013, edge.width = E(gfilter_G2013)$weight*6,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7)
```

##vertex attributes "mother" "daughter" or "other"
```{r}
mother_daughter_pairs_G2013<-data.frame(mother=c("B587","6D0D","07F9","0D1B","24E9"),daughter=c("5672","4C7B","BF04","4FE4","41C1"))
mother_daughter_pairs_G2013$pairID<-c(1:nrow(mother_daughter_pairs_G2013))

mother_ids_G2013<- mother_daughter_pairs_G2013$mother
daughter_ids_G2013<- mother_daughter_pairs_G2013$daughter

V(g_G2013)$type <- "other"  # default value
V(g_G2013)[V(g_G2013)$name %in% mother_ids_G2013]$type   <- "mother-daughter"
V(g_G2013)[V(g_G2013)$name %in% daughter_ids_G2013]$type <- "mother-daughter"

plot(gfilter_G2013, edge.width = E(g_G2013)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2013)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2013)$name)
```
### "pair ID" and add the color
```{r}
md_lookup_G2013<- mother_daughter_pairs_G2013 %>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2013<- match(V(g_G2013)$name, md_lookup_G2013$ID)   # returns NA if the name is not found
V(g_G2013)$pairID <- md_lookup_G2013$pairID[IDs_G2013] 

pair_IDs_G2013<- factor(V(g_G2013)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2013<-length(levels(na.omit(pair_IDs_G2013)))
my_palette_G2013<-brewer.pal(n = max(3,n_pairs_G2013), name = "Set2") 

# Map each vertex to the correct color
pair_colors_G2013 <- my_palette_G2013[pair_IDs_G2013]
pair_colors_G2013[is.na(pair_colors_G2013)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2013, edge.width = E(gfilter_G2013)$weight*6,edge.curved=F,vertex.color =pair_colors_G2013, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2013 <- levels(na.omit(pair_IDs_G2013))
legend_colors_G2013 <- my_palette_G2013[seq_along(pair_levels_G2013)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2013), "random pair"),
  col    = c(legend_colors_G2013, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2013 <-  igraph::as_data_frame(g_G2013, what="edges")

vtype_G2013<- V(g_G2013)$type
names(vtype_G2013) <- V(g_G2013)$name

from_type_G2013<- vtype_G2013[ edge_G2013$from ]
to_type_G2013<- vtype_G2013[ edge_G2013$to   ]

edge_G2013$pair_type <- ifelse(
  from_type_G2013 == "mother-daughter" & to_type_G2013 == "mother-daughter",
  "mother-daughter",
  "other"
)

```

### SRI plot
```{r}
sri_G2013<- as.vector(net_G2013)
sri_G2013 <- sri_G2013[sri_G2013 > 0]  # Remove zero values
meansri_G2013<- mean(sri_G2013)


hist(sri_G2013, breaks=30, main="Distribution of association index (SRI) for GB2 in 2013", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2013, col="blue", lwd=2)
```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
####observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2013$from ]
  to_pairid<- pairid[ edge_G2013$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2013$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2013<- mean_pair_association(g_G2013)
```
si marche pas c'est qu'il faut import edges dans la loop
####permutation
```{r}
num_permutations<- 5000
perm_E_G2013<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2013
  permuted_edge_list$pair_type <- sample(edge_G2013$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2013[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2013<-mean(perm_E_G2013 >= obs_value_G2013)

cat("Mean of observed association =",obs_value_G2013,"\n")
cat("Mean of permuted association =",mean(perm_E_G2013),"\n")
cat("P-value is",pvalue_e_G2013, "\t")
if (pvalue_e_G2013< 0.05) cat("significative") else
  cat("not significative")

```
```{r}
hist(perm_E_G2013, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.3,0.48))
abline(v=obs_value_G2013, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mothjer-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

#### permutation
```{r}
original_pairid_G2013<- V(g_G2013)$pairID

num_perm<- 5000
perm_V_G2013<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2013, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2013
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2013[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2013 <- mean(perm_V_G2013 >= obs_value_G2013)
cat("Mean of observed association =",obs_value_G2013,"\n")
cat("Mean of permuted association =",mean(perm_V_G2013),"\n")
cat("P-value is",pvalue_v_G2013, "\t")
if (pvalue_v_G2013 < 0.05) cat("significative") else
  cat("not significative")
```
```{r}
hist(perm_V_G2013, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.2,0.6))
abline(v=obs_value_G2013, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)

mean of association for mother-daughter pairs is already measured

####permute
```{r}
network_perm_G2013<-network_permutation(association_data=GBI_G2013,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2013,identities=colnames(GBI_G2013),days= 1:nrow(GBI_G2013))

```

```{r}
individuals_G2013<-colnames(GBI_G2013)
dimnames(network_perm_G2013)[[2]] <- individuals_G2013
dimnames(network_perm_G2013)[[3]] <- individuals_G2013
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2013 <- apply(network_perm_G2013, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2013)
})

pvalue_N_G2013<- mean(MD_perm_G2013 >= obs_value_G2013)

cat("Mean of observed association =",obs_value_G2013,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2013),"\n")
cat("P-value is",pvalue_N_G2013,"\t")
if (pvalue_N_G2013 < 0.05) cat("significative") else
  cat("not significative")

```

```{r}
hist(MD_perm_G2013, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.3,0.5))
abline(v=mean(obs_value_G2013), col="red", lwd=3)
```
# GB2_2014 — (ﾉ\^_^)ﾉ—————————————

```{r}
GB2_2014<-read.csv("CSV_Dayroosts_GB2_2014_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)
GB2_2014<-GB2_2014[c(0:40)] #filter columns
GB2_2014<-GB2_2014[1:51,]
GB2_2014 <- GB2_2014[, -4]
GB2_2014 <- GB2_2014[, -3]
GB2_2014 <- GB2_2014[, -2]
colnames(GB2_2014)[colnames(GB2_2014) == "number"] <- "date"
GB2_2014 <- GB2_2014[, !(colnames(GB2_2014) %in% c("E4D5", "B661"))]
```

```{r}
data_G2014<- GB2_2014 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2014$daygroup<-as.factor(paste(data_G2014$date, data_G2014$box, sep="_"))

ind_G2014<-data.table(ID=c(data_G2014$ID), group=data_G2014$daygroup, day=c(data_G2014$date)) #df with 3 col: ID, group and day 
GBI_G2014<- get_group_by_individual(ind_G2014, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2014 <- as.matrix(GBI_G2014)
```

## network
```{r}
net_G2014<-get_network(GBI_G2014,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2014 <- graph_from_adjacency_matrix(net_G2014,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2014,vertex.label = V(g_G2014)$name,edge.width = E(g_G2014)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue") 
```
#####plot + threshold
```{r}
threshold_G2014 <- mean(net_G2014, na.rm = TRUE) +  sd(net_G2014, na.rm = TRUE)

gfilter_G2014<- delete_edges(g_G2014, E(g_G2014)[weight < threshold_G2014]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2014, edge.width = E(gfilter_G2014)$weight*6,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7)
```

##vertex attributes
```{r}
mother_daughter_pairs_G2014<-data.frame(mother=c("B587","4ECA","AA57","BF04","5672","377E"),daughter=c("879F", "29B2", "076C", "EE00", "22B0","E063"))
mother_daughter_pairs_G2014$pairID<-c(1:nrow(mother_daughter_pairs_G2014))

mother_ids_G2014<- mother_daughter_pairs_G2014$mother
daughter_ids_G2014<- mother_daughter_pairs_G2014$daughter

V(g_G2014)$type <- "other"  # default value
V(g_G2014)[V(g_G2014)$name %in% mother_ids_G2014]$type   <- "mother-daughter"
V(g_G2014)[V(g_G2014)$name %in% daughter_ids_G2014]$type <- "mother-daughter"

plot(gfilter_G2014, edge.width = E(g_G2014)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2014)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2014)$name)
```
```{r}
md_lookup_G2014<- mother_daughter_pairs_G2014%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2014<- match(V(g_G2014)$name, md_lookup_G2014$ID)   # returns NA if the name is not found
V(g_G2014)$pairID <- md_lookup_G2014$pairID[IDs_G2014] 

pair_IDs_G2014<- factor(V(g_G2014)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2014<-length(levels(na.omit(pair_IDs_G2014)))
my_palette_G2014<-brewer.pal(n = max(3,n_pairs_G2014), name = "Set2") 

# Map each vertex to the correct color
pair_colors_G2014<- my_palette_G2014[pair_IDs_G2014]
pair_colors_G2014[is.na(pair_colors_G2014)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2014, edge.width = E(gfilter_G2014)$weight*6,edge.curved=F,vertex.color =pair_colors_G2014, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2014 <- levels(na.omit(pair_IDs_G2014))
legend_colors_G2014 <- my_palette_G2014[seq_along(pair_levels_G2014)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2014), "random pair"),
  col    = c(legend_colors_G2014, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2014 <-  igraph::as_data_frame(g_G2014, what="edges")

vtype_G2014<- V(g_G2014)$type
names(vtype_G2014) <- V(g_G2014)$name

from_type_G2014<- vtype_G2014[ edge_G2014$from ]
to_type_G2014<- vtype_G2014[ edge_G2014$to   ]

edge_G2014$pair_type <- ifelse(
  from_type_G2014 == "mother-daughter" & to_type_G2014 == "mother-daughter",
  "mother-daughter",
  "other"
)

```

### SRI plot
```{r}
sri_G2014<- as.vector(net_G2014)
sri_G2014 <- sri_G2014[sri_G2014 > 0]  # Remove zero values
meansri_G2014<- mean(sri_G2014)


hist(sri_G2014, breaks=30, main="Distribution of association index (SRI) for GB2 in 2014", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2014, col="blue", lwd=2)
```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
####observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2014$from ]
  to_pairid<- pairid[ edge_G2014$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2014$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2014<- mean_pair_association(g_G2014)
```
si marche pas c'est qu'il faut import edges dans la loop
####permutation
```{r}
num_permutations<- 5000
perm_E_G2014<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2014
  permuted_edge_list$pair_type <- sample(edge_G2014$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2014[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2014<-mean(perm_E_G2014 >= obs_value_G2014)

cat("Mean of observed association =",obs_value_G2014,"\n")
cat("Mean of permuted association =",mean(perm_E_G2014),"\n")
cat("P-value is",pvalue_e_G2014, "\t")
if (pvalue_e_G2014< 0.05) cat("significative") else
  cat("not significative")

```
```{r}
hist(perm_E_G2014, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.3,0.48))
abline(v=obs_value_G2014, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mothjer-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

#### permutation
```{r}
original_pairid_G2014<- V(g_G2014)$pairID

num_perm<- 5000
perm_V_G2014<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2014, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2014
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2014[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2014<- mean(perm_V_G2014 >= obs_value_G2014)
cat("Mean of observed association =",obs_value_G2014,"\n")
cat("Mean of permuted association =",mean(perm_V_G2014),"\n")
cat("P-value is",pvalue_v_G2014, "\t")
if (pvalue_v_G2014 < 0.05) cat("significative") else
  cat("not significative")
```

```{r}
hist(perm_V_G2014, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.2,0.6))
abline(v=obs_value_G2014, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured

####permute
```{r}
network_perm_G2014<-network_permutation(association_data=GBI_G2014,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2014,identities=colnames(GBI_G2014),days= 1:nrow(GBI_G2014))

```

```{r}
individuals_G2014<-colnames(GBI_G2014)
dimnames(network_perm_G2014)[[2]] <- individuals_G2014
dimnames(network_perm_G2014)[[3]] <- individuals_G2014
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2014<-apply(network_perm_G2014, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2014)
})

pvalue_N_G2014<- mean(MD_perm_G2014>= obs_value_G2014)

cat("Mean of observed association =",obs_value_G2014,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2014),"\n")
cat("P-value is",pvalue_N_G2014,"\t")
if (pvalue_N_G2014 < 0.05) cat("significative") else
  cat("not significative")

```

```{r}
hist(MD_perm_G2014, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.3,0.5))
abline(v=mean(obs_value_G2014), col="red", lwd=3)
```
# GB2 2015 (◡‿◡✿)-----------------------

```{r}
GB2_2015<-read.csv("Dayroosts_GB2_2015_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)

GB2_2015<-GB2_2015[c(0:38)] #filter columns
GB2_2015<-GB2_2015[1:45,]

colnames(GB2_2015)[colnames(GB2_2015) == "number"] <- "date"

#remove one individual
GB2_2015<- GB2_2015[, !(colnames(GB2_2015) %in% "3D98")]
```

```{r}
data_G2015<- GB2_2015 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2015$daygroup<-as.factor(paste(data_G2015$date, data_G2015$box, sep="_"))

ind_G2015<-data.table(ID=c(data_G2015$ID), group=data_G2015$daygroup, day=c(data_G2015$date)) #df with 3 col: ID, group and day 
GBI_G2015<- get_group_by_individual(ind_G2015, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2015<- as.matrix(GBI_G2015)
```

## network
```{r}
net_G2015<-get_network(GBI_G2015,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2015<- graph_from_adjacency_matrix(net_G2015,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2015,vertex.label = V(g_G2015)$name,edge.width = E(g_G2015)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
#####plot + threshold
```{r}
threshold_G2015<- mean(net_G2015, na.rm = TRUE) +  sd(net_G2015, na.rm = TRUE)

gfilter_G2015<- delete_edges(g_G2015, E(g_G2015)[weight < threshold_G2015]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2015, edge.width = E(gfilter_G2015)$weight*6,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

##vertex attributes
```{r}
mother_daughter_pairs_G2015<-data.frame(mother=c("B75C","3F49","1731"),daughter=c("663E", "EC1A", "256A"))
mother_daughter_pairs_G2015$pairID<-c(1:nrow(mother_daughter_pairs_G2015))

mother_ids_G2015<- mother_daughter_pairs_G2015$mother
daughter_ids_G2015<- mother_daughter_pairs_G2015$daughter

V(g_G2015)$type <- "other"  # default value
V(g_G2015)[V(g_G2015)$name %in% mother_ids_G2015]$type   <- "mother-daughter"
V(g_G2015)[V(g_G2015)$name %in% daughter_ids_G2015]$type <- "mother-daughter"

plot(gfilter_G2015, edge.width = E(gfilter_G2015)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2015)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2015)$name,edge.color="plum3")
```
```{r}
md_lookup_G2015<- mother_daughter_pairs_G2015%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2015<- match(V(g_G2015)$name, md_lookup_G2015$ID)   # returns NA if the name is not found
V(g_G2015)$pairID <- md_lookup_G2015$pairID[IDs_G2015] 

pair_IDs_G2015<- factor(V(g_G2015)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2015<-length(levels(na.omit(pair_IDs_G2015)))
my_palette_G2015<-brewer.pal(n = max(3,n_pairs_G2015), name = "Set2") 

# Map each vertex to the correct color
pair_colors_G2015<- my_palette_G2015[pair_IDs_G2015]
pair_colors_G2015[is.na(pair_colors_G2015)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2015, edge.width = E(gfilter_G2015)$weight*6,edge.curved=F,vertex.color =pair_colors_G2015, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2015<- levels(na.omit(pair_IDs_G2015))
legend_colors_G2015 <- my_palette_G2015[seq_along(pair_levels_G2015)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2015), "random pair"),
  col    = c(legend_colors_G2015, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2015<-  igraph::as_data_frame(g_G2015, what="edges")

vtype_G2015<- V(g_G2015)$type
names(vtype_G2015) <- V(g_G2015)$name

from_type_G2015<- vtype_G2015[ edge_G2015$from ]
to_type_G2015<- vtype_G2015[ edge_G2015$to   ]

edge_G2015$pair_type <- ifelse(
  from_type_G2015== "mother-daughter" & to_type_G2015 == "mother-daughter",
  "mother-daughter",
  "other"
)

```

### SRI plot
```{r}
sri_G2015<- as.vector(net_G2015)
sri_G2015 <- sri_G2015[sri_G2015 > 0]  # Remove zero values
meansri_G2015<- mean(sri_G2015)


hist(sri_G2015, breaks=30, main="Distribution of association index (SRI) for GB2 in 2015", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2015, col="blue", lwd=2)
```
Two different hist for the two types of pairs
```{r}
sri_md_G2015<- edge_G2015$weight[edge_G2015$pair_type == "mother-daughter" ]


hist_all_G2015<-hist(net_G2015, breaks = 30, plot = FALSE)
hist_md_G2015<- hist(sri_md_G2015, breaks = 30, plot = FALSE)

plot(hist_all_G2015, col = "lightblue", border = "black",
     main = "Distribution of Association Index (SRI) for GB2 in 2015",
     xlab = "Association index (SRI)", ylab = "Frequency", xlim = c(0,1))
plot(hist_md_G2015, col = adjustcolor("plum4", alpha.f = 0.5), border = "plum4", add = TRUE)

abline(v = mean(sri_G2015), col = "blue", lwd = 2)
abline(v = mean(sri_md_G2015), col = "plum4", lwd = 2, lty = 2)

```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2015$from ]
  to_pairid<- pairid[ edge_G2015$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2015$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2015<- mean_pair_association(g_G2015)
```

permutation
```{r}
num_permutations<- 5000
perm_E_G2015<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2015
  permuted_edge_list$pair_type <- sample(edge_G2015$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2015[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2015<-mean(perm_E_G2015>= obs_value_G2015)

cat("Mean of observed association =",obs_value_G2015,"\n")
cat("Mean of permuted association =",mean(perm_E_G2015),"\n")
cat("P-value is",pvalue_e_G2015, "\t")
if (pvalue_e_G2015< 0.05) cat("significative") else
  cat("not significative")

```
```{r}
hist(perm_E_G2015, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.05,0.35))
abline(v=obs_value_G2015, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mothjer-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value


```{r}
original_pairid_G2015<- V(g_G2015)$pairID

num_perm<- 5000
perm_V_G2015<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2015, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2015
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2015[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2015<- mean(perm_V_G2015 >= obs_value_G2015)
cat("Mean of observed association =",obs_value_G2015,"\n")
cat("Mean of permuted association =",mean(perm_V_G2015),"\n")
cat("P-value is",pvalue_v_G2015, "\t")
if (pvalue_v_G2015< 0.05) cat("significative") else
  cat("not significative")
```

```{r}
hist(perm_V_G2015, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.6))
abline(v=obs_value_G2015, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured
```{r}
network_perm_G2015<-network_permutation(association_data=GBI_G2015,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2015,identities=colnames(GBI_G2015),days= 1:nrow(GBI_G2015))

```

```{r}
individuals_G2015<-colnames(GBI_G2015)
dimnames(network_perm_G2015)[[2]] <- individuals_G2015
dimnames(network_perm_G2015)[[3]] <- individuals_G2015
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2015<-apply(network_perm_G2015, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2015)
})

pvalue_N_G2015<- mean(MD_perm_G2015>= obs_value_G2015)

cat("Mean of observed association =",obs_value_G2015,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2015),"\n")
cat("P-value is",pvalue_N_G2015,"\t")
if (pvalue_N_G2015 < 0.05) cat("significative") else
  cat("not significative")

```

```{r}
hist(MD_perm_G2015, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.1,0.32))
abline(v=mean(obs_value_G2015), col="red", lwd=3)
```


# GB2 -2016 ———————————————–

Only 1 pair of mother-daughter... 3 juveniles were born but not recorded?

```{r}
GB2_2016<-read.csv("CSV_Dayroosts_GB2_2016_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)

GB2_2016<-GB2_2016[c(0:25)] #filter columns

```

```{r}
data_G2016<- GB2_2016%>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2016$daygroup<-as.factor(paste(data_G2016$date, data_G2016$box, sep="_"))

ind_G2016<-data.table(ID=c(data_G2016$ID), group=data_G2016$daygroup, day=c(data_G2016$date)) #df with 3 col: ID, group and day 
GBI_G2016<- get_group_by_individual(ind_G2016, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2016<- as.matrix(GBI_G2016)
```

## network
```{r}
net_G2016<-get_network(GBI_G2016,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2016<- graph_from_adjacency_matrix(net_G2016,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2016,vertex.label = V(g_G2016)$name,edge.width = E(g_G2016)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
#####plot + threshold
```{r}
threshold_G2016<- mean(net_G2016, na.rm = TRUE) +  sd(net_G2016, na.rm = TRUE)

gfilter_G2016<- delete_edges(g_G2016, E(g_G2016)[weight < threshold_G2016]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2016, edge.width = E(gfilter_G2016)$weight*6,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

##vertex attributes
```{r}
mother_daughter_pairs_G2016<-data.frame(mother=c("29B2"),daughter=c("0758"))
mother_daughter_pairs_G2016$pairID<-c(1:nrow(mother_daughter_pairs_G2016))

mother_ids_G2016<- mother_daughter_pairs_G2016$mother
daughter_ids_G2016<- mother_daughter_pairs_G2016$daughter

V(g_G2016)$type <- "other"  # default value
V(g_G2016)[V(g_G2016)$name %in% mother_ids_G2016]$type   <- "mother-daughter"
V(g_G2016)[V(g_G2016)$name %in% daughter_ids_G2016]$type <- "mother-daughter"

plot(gfilter_G2016, edge.width = E(gfilter_G2016)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2016)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2016)$name,edge.color="plum3")
```
```{r}
md_lookup_G2016<- mother_daughter_pairs_G2016%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2016<- match(V(g_G2016)$name, md_lookup_G2016$ID)   # returns NA if the name is not found
V(g_G2016)$pairID <- md_lookup_G2016$pairID[IDs_G2016] 

pair_IDs_G2016<- factor(V(g_G2016)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2016<-length(levels(na.omit(pair_IDs_G2016)))
my_palette_G2016<-brewer.pal(n = max(3,n_pairs_G2016), name = "Set2") 

# Map each vertex to the correct color
pair_colors_G2016<- my_palette_G2016[pair_IDs_G2016]
pair_colors_G2016[is.na(pair_colors_G2016)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2016, edge.width = E(gfilter_G2016)$weight*6,edge.curved=F,vertex.color =pair_colors_G2016, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2016<- levels(na.omit(pair_IDs_G2016))
legend_colors_G2016 <- my_palette_G2016[seq_along(pair_levels_G2016)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2016), "random pair"),
  col    = c(legend_colors_G2016, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2016<-  igraph::as_data_frame(g_G2016, what="edges")

vtype_G2016<- V(g_G2016)$type
names(vtype_G2016) <- V(g_G2016)$name

from_type_G2016<- vtype_G2016[ edge_G2016$from ]
to_type_G2016<- vtype_G2016[ edge_G2016$to   ]

edge_G2016$pair_type <- ifelse(
  from_type_G2016== "mother-daughter" & to_type_G2016 == "mother-daughter",
  "mother-daughter",
  "other"
)

```

### SRI plot
```{r}
sri_G2016<- as.vector(net_G2016)
sri_G2016<- sri_G2016[sri_G2016 > 0]  # Remove zero values
meansri_G2016<- mean(sri_G2016)


hist(sri_G2016, breaks=30, main="Distribution of association index (SRI) for GB2 in 2016", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2016, col="blue", lwd=2)
```
Two different hist for the two types of pairs
```{r}


```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2016$from ]
  to_pairid<- pairid[ edge_G2016$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2016$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2016<- mean_pair_association(g_G2016)
```

permutation
```{r}
num_permutations<- 5000
perm_E_G2016<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2016
  permuted_edge_list$pair_type <- sample(edge_G2016$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2016[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2016<-mean(perm_E_G2016>= obs_value_G2016)

cat("Mean of observed association =",obs_value_G2016,"\n")
cat("Mean of permuted association =",mean(perm_E_G2016),"\n")
cat("P-value is",pvalue_e_G2016, "\t")
if (pvalue_e_G2016< 0.05) cat("significative") else
  cat("not significative")

```
```{r}
hist(perm_E_G2016, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.9))
abline(v=obs_value_G2016, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mothjer-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

permutation
```{r}
original_pairid_G2016<- V(g_G2016)$pairID

num_perm<- 5000
perm_V_G2016<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2016, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2016
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2016[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2016<- mean(perm_V_G2016>= obs_value_G2016)
cat("Mean of observed association =",obs_value_G2016,"\n")
cat("Mean of permuted association =",mean(perm_V_G2016),"\n")
cat("P-value is",pvalue_v_G2016, "\t")
if (pvalue_v_G2016< 0.05) cat("significative") else
  cat("not significative")
```

```{r}
hist(perm_V_G2016, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.9))
abline(v=obs_value_G2016, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured

permute
```{r}
network_perm_G2016<-network_permutation(association_data=GBI_G2016,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2016,identities=colnames(GBI_G2016),days= 1:nrow(GBI_G2016))

```

```{r}
individuals_G2016<-colnames(GBI_G2016)
dimnames(network_perm_G2016)[[2]] <- individuals_G2016
dimnames(network_perm_G2016)[[3]] <- individuals_G2016
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2016<-apply(network_perm_G2016, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2016)
})

pvalue_N_G2016<- mean(MD_perm_G2016>= obs_value_G2016)

cat("Mean of observed association =",obs_value_G2016,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2016),"\n")
cat("P-value is",pvalue_N_G2016,"\t")
if (pvalue_N_G2016 < 0.05) cat("significative") else
  cat("not significative")

```

```{r}
hist(MD_perm_G2016, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.1,0.9))
abline(v=mean(obs_value_G2016), col="red", lwd=3)
```
# GB2 2017 ---------------------

```{r}
GB2_2017<-read.csv("CSV_Dayroosts_GB2_2017_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)

GB2_2017<-GB2_2017[c(0:36)] #filter columns
GB2_2017<-GB2_2017[1:42,]

```

```{r}
data_G2017<- GB2_2017%>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2017$daygroup<-as.factor(paste(data_G2017$date, data_G2017$box, sep="_"))

ind_G2017<-data.table(ID=c(data_G2017$ID), group=data_G2017$daygroup, day=c(data_G2017$date)) #df with 3 col: ID, group and day 
GBI_G2017<- get_group_by_individual(ind_G2017, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2017<- as.matrix(GBI_G2017)
```

## network
```{r}
net_G2017<-get_network(GBI_G2017,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2017<- graph_from_adjacency_matrix(net_G2017,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2017,vertex.label = V(g_G2017)$name,edge.width = E(g_G2017)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
#####plot + threshold
```{r}
threshold_G2017<- mean(net_G2017, na.rm = TRUE) +  sd(net_G2017, na.rm = TRUE)

gfilter_G2017<- delete_edges(g_G2017, E(g_G2017)[weight < mean(net_G2017, na.rm = TRUE)]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2017, edge.width = E(gfilter_G2017)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

##vertex attributes
```{r}
mother_daughter_pairs_G2017<-data.frame(mother=c("41C1","879F","29B2","663E","256A","0758","377E","24E9","4FE4","F377"),daughter=c("9FB7","2F8D","D629","A539","DE72","C5F2","DCCF","44BD","F44A","ACF4"))
mother_daughter_pairs_G2017$pairID<-c(1:nrow(mother_daughter_pairs_G2017))

mother_ids_G2017<- mother_daughter_pairs_G2017$mother
daughter_ids_G2017<- mother_daughter_pairs_G2017$daughter

V(g_G2017)$type <- "other"  # default value
V(g_G2017)[V(g_G2017)$name %in% mother_ids_G2017]$type   <- "mother-daughter"
V(g_G2017)[V(g_G2017)$name %in% daughter_ids_G2017]$type <- "mother-daughter"

plot(gfilter_G2017, edge.width = E(gfilter_G2017)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2017)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2017)$name,edge.color="plum3")
```
```{r}
md_lookup_G2017<- mother_daughter_pairs_G2017%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2017<- match(V(g_G2017)$name, md_lookup_G2017$ID)   # returns NA if the name is not found
V(g_G2017)$pairID <- md_lookup_G2017$pairID[IDs_G2017] 

pair_IDs_G2017<- factor(V(g_G2017)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2017<-length(levels(na.omit(pair_IDs_G2017)))
my_palette_G2017<-brewer.pal(n = max(3,n_pairs_G2017), name = "Paired") 

# Map each vertex to the correct color
pair_colors_G2017<- my_palette_G2017[pair_IDs_G2017]
pair_colors_G2017[is.na(pair_colors_G2017)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2017, edge.width = E(gfilter_G2017)$weight*6,edge.curved=F,vertex.color =pair_colors_G2017, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2017<- levels(na.omit(pair_IDs_G2017))
legend_colors_G2017 <- my_palette_G2017[seq_along(pair_levels_G2017)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2017), "random pair"),
  col    = c(legend_colors_G2017, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2017<-  igraph::as_data_frame(g_G2017, what="edges")

vtype_G2017<- V(g_G2017)$type
names(vtype_G2017) <- V(g_G2017)$name

from_type_G2017<-vtype_G2017[ edge_G2017$from ]
to_type_G2017<-vtype_G2017[ edge_G2017$to   ]

edge_G2017$pair_type <- ifelse(
  from_type_G2017== "mother-daughter" & to_type_G2017 == "mother-daughter",
  "mother-daughter",
  "other"
)

```

### SRI plot
```{r}
sri_G2017<- as.vector(net_G2017)
sri_G2017 <- sri_G2017[sri_G2017 > 0]  # Remove zero values
meansri_G2017<- mean(sri_G2017)


hist(sri_G2017, breaks=30, main="Distribution of association index (SRI) for GB2 in 2017", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2017, col="blue", lwd=2)
```
Two different hist for the two types of pairs
```{r}
sri_md_G2017<- edge_G2017$weight[edge_G2017$pair_type == "mother-daughter" ]


hist_all_G2017<-hist(net_G2017, breaks = 30, plot = FALSE)
hist_md_G2017<- hist(sri_md_G2017, breaks = 30, plot = FALSE)

plot(hist_all_G2017, col = "lightblue", border = "black",
     main = "Distribution of Association Index (SRI) for GB2 in 2017",
     xlab = "Association index (SRI)", ylab = "Frequency", xlim = c(0,1))
plot(hist_md_G2017, col = adjustcolor("plum4", alpha.f = 0.5), border = "plum4", add = TRUE)

abline(v = mean(sri_G2017), col = "blue", lwd = 2)
abline(v = mean(sri_md_G2017), col = "plum4", lwd = 2, lty = 2)

```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2017$from ]
  to_pairid<- pairid[ edge_G2017$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2017$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2017<- mean_pair_association(g_G2017)
```

permutation
```{r}
num_permutations<- 5000
perm_E_G2017<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2017
  permuted_edge_list$pair_type <- sample(edge_G2017$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2017[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2017<-mean(perm_E_G2017>= obs_value_G2017)

cat("Mean of observed association =",obs_value_G2017,"\n")
cat("Mean of permuted association =",mean(perm_E_G2017),"\n")
cat("P-value is",pvalue_e_G2017, "\t")
if (pvalue_e_G2017< 0.05) cat("significative") else
  cat("not significative")

```
```{r}
hist(perm_E_G2017, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.35,0.48))
abline(v=obs_value_G2017, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mothjer-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

permutation
```{r}
original_pairid_G2017<- V(g_G2017)$pairID

num_perm<- 5000
perm_V_G2017<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2017, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2017
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2017[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2017<- mean(perm_V_G2017 >= obs_value_G2017)
cat("Mean of observed association =",obs_value_G2017,"\n")
cat("Mean of permuted association =",mean(perm_V_G2017),"\n")
cat("P-value is",pvalue_v_G2017, "\t")
if (pvalue_v_G2017< 0.05) cat("significative") else
  cat("not significative")
```

```{r}
hist(perm_V_G2017, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.2,0.6))
abline(v=obs_value_G2017, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured

permute
```{r}
network_perm_G2017<-network_permutation(association_data=GBI_G2017,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2017,identities=colnames(GBI_G2017),days= 1:nrow(GBI_G2017))

```

```{r}
individuals_G2017<-colnames(GBI_G2017)
dimnames(network_perm_G2017)[[2]] <- individuals_G2017
dimnames(network_perm_G2017)[[3]] <- individuals_G2017
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2017<-apply(network_perm_G2017, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2017)
})

pvalue_N_G2017<- mean(MD_perm_G2017>= obs_value_G2017)

cat("Mean of observed association =",obs_value_G2017,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2017),"\n")
cat("P-value is",pvalue_N_G2017,"\t")
if (pvalue_N_G2017 < 0.05) cat("significative") else
  cat("not significative")

```

```{r}
hist(MD_perm_G2017, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.35,0.5))
abline(v=mean(obs_value_G2017), col="red", lwd=3)
```

# GB2 2018 ---------------------
not a lot of data

```{r}
GB2_2018<-read.csv("CSV_Dayroosts_GB2_2018_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)

GB2_2018<-GB2_2018[c(1:39)] #filter columns
GB2_2018<-GB2_2018[1:17,]

GB2_2018<- GB2_2018[, !(colnames(GB2_2018) %in% "F44A")]
```

```{r}
data_G2018<- GB2_2018%>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2018$daygroup<-as.factor(paste(data_G2018$date, data_G2018$box, sep="_"))

ind_G2018<-data.table(ID=c(data_G2018$ID), group=data_G2018$daygroup, day=c(data_G2018$date)) #df with 3 col: ID, group and day 
GBI_G2018<- get_group_by_individual(ind_G2018, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2018<- as.matrix(GBI_G2018)
```

## network
```{r}
net_G2018<-get_network(GBI_G2018,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2018<- graph_from_adjacency_matrix(net_G2018,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2018,vertex.label = V(g_G2018)$name,edge.width = E(g_G2018)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
#####plot + threshold
```{r}
threshold_G2018<- mean(net_G2018, na.rm = TRUE) +  sd(net_G2018, na.rm = TRUE)

gfilter_G2018<- delete_edges(g_G2018, E(g_G2018)[weight < threshold_G2018]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2018, edge.width = E(gfilter_G2018)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

##vertex attributes
```{r}
mother_daughter_pairs_G2018<-data.frame(mother=c("39ED","AC35","377E","7C6F"),daughter=c("80F1","AA3C","5939","CB2B"))
mother_daughter_pairs_G2018$pairID<-c(1:nrow(mother_daughter_pairs_G2018))

mother_ids_G2018<- mother_daughter_pairs_G2018$mother
daughter_ids_G2018<- mother_daughter_pairs_G2018$daughter

V(g_G2018)$type <- "other"  # default value
V(g_G2018)[V(g_G2018)$name %in% mother_ids_G2018]$type   <- "mother-daughter"
V(g_G2018)[V(g_G2018)$name %in% daughter_ids_G2018]$type <- "mother-daughter"

plot(gfilter_G2018, edge.width = E(gfilter_G2018)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2018)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2018)$name,edge.color="plum3")
```
```{r}
md_lookup_G2018<- mother_daughter_pairs_G2018%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2018<- match(V(g_G2018)$name, md_lookup_G2018$ID)   # returns NA if the name is not found
V(g_G2018)$pairID <- md_lookup_G2018$pairID[IDs_G2018] 

pair_IDs_G2018<- factor(V(g_G2018)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2018<-length(levels(na.omit(pair_IDs_G2018)))
my_palette_G2018<-brewer.pal(n = max(3,n_pairs_G2018), name = "Paired") 

# Map each vertex to the correct color
pair_colors_G2018<- my_palette_G2018[pair_IDs_G2018]
pair_colors_G2018[is.na(pair_colors_G2018)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2018, edge.width = E(gfilter_G2018)$weight*6,edge.curved=F,vertex.color =pair_colors_G2018, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2018<- levels(na.omit(pair_IDs_G2018))
legend_colors_G2018 <- my_palette_G2018[seq_along(pair_levels_G2018)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2018), "random pair"),
  col    = c(legend_colors_G2018, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2018<-  igraph::as_data_frame(g_G2018, what="edges")

vtype_G2018<- V(g_G2018)$type
names(vtype_G2018) <- V(g_G2018)$name

from_type_G2018<-vtype_G2018[ edge_G2018$from ]
to_type_G2018<-vtype_G2018[ edge_G2018$to   ]

edge_G2018$pair_type <- ifelse(
  from_type_G2018== "mother-daughter" & to_type_G2018 == "mother-daughter",
  "mother-daughter",
  "other"
)

```

### SRI plot
```{r}
sri_G2018<- as.vector(net_G2018)
sri_G2018 <- sri_G2018[sri_G2018 > 0]  # Remove zero values
meansri_G2018<- mean(sri_G2018)


hist(sri_G2018, breaks=30, main="Distribution of association index (SRI) for GB2 in 2018", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2018, col="blue", lwd=2)
```
Two different hist for the two types of pairs
```{r}

```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2018$from ]
  to_pairid<- pairid[ edge_G2018$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2018$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2018<- mean_pair_association(g_G2018)
```

permutation
```{r}
num_permutations<- 5000
perm_E_G2018<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2018
  permuted_edge_list$pair_type <- sample(edge_G2018$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2018[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2018<-mean(perm_E_G2018>= obs_value_G2018)

cat("Mean of observed association =",obs_value_G2018,"\n")
cat("Mean of permuted association =",mean(perm_E_G2018),"\n")
cat("P-value is",pvalue_e_G2018, "\t")
if (pvalue_e_G2018< 0.05) cat("significative") else
  cat("not significative")

```
```{r}
hist(perm_E_G2018, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.2,0.48))
abline(v=obs_value_G2018, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mother-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

permutation
```{r}
original_pairid_G2018<- V(g_G2018)$pairID

num_perm<- 5000
perm_V_G2018<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2018, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2018
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2018[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2018<- mean(perm_V_G2018 >= obs_value_G2018)
cat("Mean of observed association =",obs_value_G2018,"\n")
cat("Mean of permuted association =",mean(perm_V_G2018),"\n")
cat("P-value is",pvalue_v_G2018, "\t")
if (pvalue_v_G2018< 0.05) cat("significative") else
  cat("not significative")
```

```{r}
hist(perm_V_G2018, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.6))
abline(v=obs_value_G2018, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured

permute
```{r}
network_perm_G2018<-network_permutation(association_data=GBI_G2018,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2018,identities=colnames(GBI_G2018),days= 1:nrow(GBI_G2018))

```

```{r}
individuals_G2018<-colnames(GBI_G2018)
dimnames(network_perm_G2018)[[2]] <- individuals_G2018
dimnames(network_perm_G2018)[[3]] <- individuals_G2018
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2018<-apply(network_perm_G2018, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2018)
})

pvalue_N_G2018<- mean(MD_perm_G2018>= obs_value_G2018)

cat("Mean of observed association =",obs_value_G2018,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2018),"\n")
cat("P-value is",pvalue_N_G2018,"\t")
if (pvalue_N_G2018 < 0.05) cat("significative") else
  cat("not significative")

```

```{r}
hist(MD_perm_G2018, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.5))
abline(v=mean(obs_value_G2018), col="red", lwd=3)
```

# GB2 2019 ---------------------

```{r}
GB2_2019<-read.csv("CSV_Dayroosts_GB2_2019_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)

GB2_2019<-GB2_2019[c(1:41)] #filter columns


GB2_2018<- GB2_2018[, !(colnames(GB2_2018) %in% "F44A")]
```

```{r}
colnames(GB2_2019)<-as.character(colnames(GB2_2019))
data_G2019<- GB2_2019 %>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2019$daygroup<-as.factor(paste(data_G2019$date, data_G2019$box, sep="_"))

ind_G2019<-data.table(ID=c(data_G2019$ID), group=data_G2019$daygroup, day=c(data_G2019$date)) #df with 3 col: ID, group and day 
GBI_G2019<- get_group_by_individual(ind_G2019, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2019<- as.matrix(GBI_G2019)
```

## network
```{r}
net_G2019<-get_network(GBI_G2019,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2019<- graph_from_adjacency_matrix(net_G2019,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2019,vertex.label = V(g_G2019)$name,edge.width = E(g_G2019)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
#####plot + threshold
```{r}
threshold_G2019<- mean(net_G2019, na.rm = TRUE) +  sd(net_G2019, na.rm = TRUE)

gfilter_G2019<- delete_edges(g_G2019, E(g_G2019)[weight <threshold_G2019]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2019, edge.width = E(gfilter_G2019)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

##vertex attributes
```{r}
mother_daughter_pairs_G2019<-data.frame(mother=c("41C1","F44A","377E"),daughter=c("A24D","2D31","B8A2"))
mother_daughter_pairs_G2019$pairID<-c(1:nrow(mother_daughter_pairs_G2019))

mother_ids_G2019<- mother_daughter_pairs_G2019$mother
daughter_ids_G2019<- mother_daughter_pairs_G2019$daughter

V(g_G2019)$type <- "other"  # default value
V(g_G2019)[V(g_G2019)$name %in% mother_ids_G2019]$type   <- "mother-daughter"
V(g_G2019)[V(g_G2019)$name %in% daughter_ids_G2019]$type <- "mother-daughter"

plot(gfilter_G2019, edge.width = E(gfilter_G2019)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2019)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2019)$name,edge.color="plum3")
```
```{r}
md_lookup_G2019<- mother_daughter_pairs_G2019%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2019<- match(V(g_G2019)$name, md_lookup_G2019$ID)   # returns NA if the name is not found
V(g_G2019)$pairID <- md_lookup_G2019$pairID[IDs_G2019] 

pair_IDs_G2019<- factor(V(g_G2019)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2019<-length(levels(na.omit(pair_IDs_G2019)))
my_palette_G2019<-brewer.pal(n = max(3,n_pairs_G2019), name = "Paired") 

# Map each vertex to the correct color
pair_colors_G2019<- my_palette_G2019[pair_IDs_G2019]
pair_colors_G2019[is.na(pair_colors_G2019)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2019, edge.width = E(gfilter_G2019)$weight*6,edge.curved=F,vertex.color =pair_colors_G2019, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2019<- levels(na.omit(pair_IDs_G2019))
legend_colors_G2019 <- my_palette_G2019[seq_along(pair_levels_G2019)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2019), "random pair"),
  col    = c(legend_colors_G2019, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2019<-  igraph::as_data_frame(g_G2019, what="edges")

vtype_G2019<- V(g_G2019)$type
names(vtype_G2019) <- V(g_G2019)$name

from_type_G2019<-vtype_G2019[ edge_G2019$from ]
to_type_G2019<-vtype_G2019[ edge_G2019$to   ]

edge_G2019$pair_type <- ifelse(
  from_type_G2019== "mother-daughter" & to_type_G2019 == "mother-daughter",
  "mother-daughter",
  "other"
)

```

### SRI plot
```{r}
sri_G2019<- as.vector(net_G2019)
sri_G2019 <- sri_G2019[sri_G2019 > 0]  # Remove zero values
meansri_G2019<- mean(sri_G2019)


hist(sri_G2019, breaks=30, main="Distribution of association index (SRI) for GB2 in 2019", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2019, col="blue", lwd=2)
```
Two different hist for the two types of pairs
```{r}

```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2019$from ]
  to_pairid<- pairid[ edge_G2019$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2019$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2019<- mean_pair_association(g_G2019)
```

permutation
```{r}
num_permutations<- 5000
perm_E_G2019<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2019
  permuted_edge_list$pair_type <- sample(edge_G2019$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2019[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2019<-mean(perm_E_G2019>= obs_value_G2019)

cat("Mean of observed association =",obs_value_G2019,"\n")
cat("Mean of permuted association =",mean(perm_E_G2019),"\n")
cat("P-value is",pvalue_e_G2019, "\t")
if (pvalue_e_G2019< 0.05) cat("significative") else
  cat("not significative")

```
```{r}
hist(perm_E_G2019, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.2,0.7))
abline(v=obs_value_G2019, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mother-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

permutation
```{r}
original_pairid_G2019<- V(g_G2019)$pairID

num_perm<- 5000
perm_V_G2019<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2019, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2019
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2019[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2019<- mean(perm_V_G2019>= obs_value_G2019)
cat("Mean of observed association =",obs_value_G2019,"\n")
cat("Mean of permuted association =",mean(perm_V_G2019),"\n")
cat("P-value is",pvalue_v_G2019, "\t")
if (pvalue_v_G2019< 0.05) cat("significative") else
  cat("not significative")
```

```{r}
hist(perm_V_G2019, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.8))
abline(v=obs_value_G2019, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured

permute
```{r}
network_perm_G2019<-network_permutation(association_data=GBI_G2019,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2019,identities=colnames(GBI_G2019),days= 1:nrow(GBI_G2019))

```

```{r}
individuals_G2019<-colnames(GBI_G2019)
dimnames(network_perm_G2019)[[2]] <- individuals_G2019
dimnames(network_perm_G2019)[[3]] <- individuals_G2019
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2019<-apply(network_perm_G2019, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2019)
})

pvalue_N_G2019<- mean(MD_perm_G2019>= obs_value_G2019)

cat("Mean of observed association =",obs_value_G2019,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2019),"\n")
cat("P-value is",pvalue_N_G2019,"\t")
if (pvalue_N_G2019 < 0.05) cat("significative") else
  cat("not significative")

```

```{r}
hist(MD_perm_G2019, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.3,0.7))
abline(v=mean(obs_value_G2019), col="red", lwd=3)
```
# GB2 2020 ---------------------

```{r}
GB2_2020<-read.csv("CSV_Dayroosts_GB2_2020_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)

GB2_2020<-GB2_2020[c(1:51)] #filter columns
GB2_2020<-GB2_2020[1:27,]

GB2_2020 <- GB2_2020 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

```

```{r}
colnames(GB2_2020)<-as.character(colnames(GB2_2020))
data_G2020<- GB2_2020 %>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2020$daygroup<-as.factor(paste(data_G2020$date, data_G2020$box, sep="_"))

ind_G2020<-data.table(ID=c(data_G2020$ID), group=data_G2020$daygroup, day=c(data_G2020$date)) #df with 3 col: ID, group and day 
GBI_G2020<- get_group_by_individual(ind_G2020, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2020<- as.matrix(GBI_G2020)
```

## network
```{r}
net_G2020<-get_network(GBI_G2020,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2020<- graph_from_adjacency_matrix(net_G2020,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2020,vertex.label = V(g_G2020)$name,edge.width = E(g_G2020)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
#####plot + threshold
```{r}
threshold_G2020<- mean(net_G2020, na.rm = TRUE) +  sd(net_G2020, na.rm = TRUE)

gfilter_G2020<- delete_edges(g_G2020, E(g_G2020)[weight <threshold_G2020]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2020, edge.width = E(gfilter_G2020)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

##vertex attributes
```{r}
mother_daughter_pairs_G2020<-data.frame(mother=c("4ECA","41C1","BF04","879F","EC1A","663E","DCCF","80F1","CB2B","1731","4FE4"),daughter=c("C809","B5D9","D9EF","AC37","82C7",
"0387","3CC1","9A5D","8787","D4DC","418B"))

mother_daughter_pairs_G2020$pairID<-c(1:nrow(mother_daughter_pairs_G2020))

mother_ids_G2020<- mother_daughter_pairs_G2020$mother
daughter_ids_G2020<- mother_daughter_pairs_G2020$daughter

V(g_G2020)$type <- "other"  # default value
V(g_G2020)[V(g_G2020)$name %in% mother_ids_G2020]$type   <- "mother-daughter"
V(g_G2020)[V(g_G2020)$name %in% daughter_ids_G2020]$type <- "mother-daughter"

plot(gfilter_G2020, edge.width = E(gfilter_G2020)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2020)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2020)$name,edge.color="plum3")
```
```{r}
md_lookup_G2020<- mother_daughter_pairs_G2020%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2020<- match(V(g_G2020)$name, md_lookup_G2020$ID)   # returns NA if the name is not found
V(g_G2020)$pairID <- md_lookup_G2020$pairID[IDs_G2020] 

pair_IDs_G2020<- factor(V(g_G2020)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2020<-length(levels(na.omit(pair_IDs_G2020)))
my_palette_G2020<-brewer.pal(n = max(3,n_pairs_G2020), name = "Paired") 

# Map each vertex to the correct color
pair_colors_G2020<- my_palette_G2020[pair_IDs_G2020]
pair_colors_G2020[is.na(pair_colors_G2020)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2020, edge.width = E(gfilter_G2020)$weight*6,edge.curved=F,vertex.color =pair_colors_G2020, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2020<- levels(na.omit(pair_IDs_G2020))
legend_colors_G2020 <- my_palette_G2020[seq_along(pair_levels_G2020)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2020), "random pair"),
  col    = c(legend_colors_G2020, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2020<-  igraph::as_data_frame(g_G2020, what="edges")

vtype_G2020<- V(g_G2020)$type
names(vtype_G2020) <- V(g_G2020)$name

from_type_G2020<-vtype_G2020[ edge_G2020$from ]
to_type_G2020<-vtype_G2020[ edge_G2020$to   ]

edge_G2020$pair_type <- ifelse(
  from_type_G2020== "mother-daughter" & to_type_G2020 == "mother-daughter",
  "mother-daughter",
  "other"
)

```

### SRI plot
```{r}
sri_G2020<- as.vector(net_G2020)
sri_G2020 <- sri_G2020[sri_G2020 > 0]  # Remove zero values
meansri_G2020<- mean(sri_G2020)


hist(sri_G2020, breaks=30, main="Distribution of association index (SRI) for GB2 in 2020", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2020, col="blue", lwd=2)
```
Two different hist for the two types of pairs
```{r}

```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2020$from ]
  to_pairid<- pairid[ edge_G2020$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2020$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2020<- mean_pair_association(g_G2020)
```

permutation
```{r}
num_permutations<- 5000
perm_E_G2020<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2020
  permuted_edge_list$pair_type <- sample(edge_G2020$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2020[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2020<-mean(perm_E_G2020>= obs_value_G2020)

cat("Mean of observed association =",obs_value_G2020,"\n")
cat("Mean of permuted association =",mean(perm_E_G2020),"\n")
cat("P-value is",pvalue_e_G2020, "\t")
if (pvalue_e_G2020< 0.05) cat("significative") else
  cat("not significative")

```
```{r}
hist(perm_E_G2020, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.2,0.4))
abline(v=obs_value_G2020, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mother-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

permutation
```{r}
original_pairid_G2020<- V(g_G2020)$pairID

num_perm<- 5000
perm_V_G2020<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2020, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2020
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2020[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2020<- mean(perm_V_G2020>= obs_value_G2020)
cat("Mean of observed association =",obs_value_G2020,"\n")
cat("Mean of permuted association =",mean(perm_V_G2020),"\n")
cat("P-value is",pvalue_v_G2020, "\t")
if (pvalue_v_G2020< 0.05) cat("significative") else
  cat("not significative")
```

```{r}
hist(perm_V_G2020, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.6))
abline(v=obs_value_G2020, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured

permute
```{r}
network_perm_G2020<-network_permutation(association_data=GBI_G2020,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2020,identities=colnames(GBI_G2020),days= 1:nrow(GBI_G2020))

```

```{r}
individuals_G2020<-colnames(GBI_G2020)
dimnames(network_perm_G2020)[[2]] <- individuals_G2020
dimnames(network_perm_G2020)[[3]] <- individuals_G2020
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2020<-apply(network_perm_G2020, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2020)
})

pvalue_N_G2020<- mean(MD_perm_G2020>= obs_value_G2020)

cat("Mean of observed association =",obs_value_G2020,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2020),"\n")
cat("P-value is",pvalue_N_G2020,"\t")
if (pvalue_N_G2020 < 0.05) cat("-> significative") else
  cat("-> not significative")

```

```{r}
hist(MD_perm_G2020, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.1,0.4))
abline(v=mean(obs_value_G2020), col="red", lwd=3)
```
# GB2 2021 ---------------------

```{r}
GB2_2021<-read.csv("CSV_Dayroosts_GB2_2021_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)

GB2_2021<-GB2_2021[c(1:56)] #filter columns
GB2_2021<-GB2_2021[1:37,]

GB2_2021 <- GB2_2021 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

colnames(GB2_2021)[colnames(GB2_2021) == "Date"] <- "date"
```

```{r}
colnames(GB2_2021)<-as.character(colnames(GB2_2021))
data_G2021<- GB2_2021%>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2021$daygroup<-as.factor(paste(data_G2021$date, data_G2021$box, sep="_"))

ind_G2021<-data.table(ID=c(data_G2021$ID), group=data_G2021$daygroup, day=c(data_G2021$date)) #df with 3 col: ID, group and day 
GBI_G2021<- get_group_by_individual(ind_G2021, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2021<- as.matrix(GBI_G2021)
```

## network
```{r}
net_G2021<-get_network(GBI_G2021,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2021<- graph_from_adjacency_matrix(net_G2021,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2021,vertex.label = V(g_G2021)$name,edge.width = E(g_G2021)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
#####plot + threshold
```{r}
threshold_G2021<- mean(net_G2021, na.rm = TRUE) +  sd(net_G2021, na.rm = TRUE)

gfilter_G2021<- delete_edges(g_G2021, E(g_G2021)[weight <threshold_G2021]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2021, edge.width = E(gfilter_G2021)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

##vertex attributes
```{r}
mother_daughter_pairs_G2021<-data.frame(mother=c("AC35","41C1","076C","F44A","44BD","D629","C5F2","AA3C","15BF"),daughter=c("A816","34C9","494F","9A80","0734","8106","0145","A42D","8BE4"))

mother_daughter_pairs_G2021$pairID<-c(1:nrow(mother_daughter_pairs_G2021))

mother_ids_G2021<- mother_daughter_pairs_G2021$mother
daughter_ids_G2021<- mother_daughter_pairs_G2021$daughter

V(g_G2021)$type <- "other"  # default value
V(g_G2021)[V(g_G2021)$name %in% mother_ids_G2021]$type   <- "mother-daughter"
V(g_G2021)[V(g_G2021)$name %in% daughter_ids_G2021]$type <- "mother-daughter"

plot(gfilter_G2021, edge.width = E(gfilter_G2021)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2021)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2021)$name,edge.color="plum3")
```
```{r}
md_lookup_G2021<- mother_daughter_pairs_G2021%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2021<- match(V(g_G2021)$name, md_lookup_G2021$ID)   # returns NA if the name is not found
V(g_G2021)$pairID <- md_lookup_G2021$pairID[IDs_G2021] 

pair_IDs_G2021<- factor(V(g_G2021)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2021<-length(levels(na.omit(pair_IDs_G2021)))
my_palette_G2021<-brewer.pal(n = max(3,n_pairs_G2021), name = "Paired") 

# Map each vertex to the correct color
pair_colors_G2021<- my_palette_G2021[pair_IDs_G2021]
pair_colors_G2021[is.na(pair_colors_G2021)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2021, edge.width = E(gfilter_G2021)$weight*6,edge.curved=F,vertex.color =pair_colors_G2021, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2021<- levels(na.omit(pair_IDs_G2021))
legend_colors_G2021 <- my_palette_G2021[seq_along(pair_levels_G2021)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2021), "random pair"),
  col    = c(legend_colors_G2021, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2021<-  igraph::as_data_frame(g_G2021, what="edges")

vtype_G2021<- V(g_G2021)$type
names(vtype_G2021) <- V(g_G2021)$name

from_type_G2021<-vtype_G2021[ edge_G2021$from ]
to_type_G2021<-vtype_G2021[ edge_G2021$to   ]

edge_G2021$pair_type <- ifelse(
  from_type_G2021== "mother-daughter" & to_type_G2021 == "mother-daughter",
  "mother-daughter",
  "other"
)
```

### SRI plot
```{r}
sri_G2021<- as.vector(net_G2021)
sri_G2021 <- sri_G2021[sri_G2021 > 0]  # Remove zero values
meansri_G2021<- mean(sri_G2021)


hist(sri_G2021, breaks=30, main="Distribution of association index (SRI) for GB2 in 2021", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2021, col="blue", lwd=2)
```
Two different hist for the two types of pairs
```{r}

```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2021$from ]
  to_pairid<- pairid[ edge_G2021$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2021$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2021<- mean_pair_association(g_G2021)
```

permutation
```{r}
num_permutations<- 5000
perm_E_G2021<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2021
  permuted_edge_list$pair_type <- sample(edge_G2021$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2021[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2021<-mean(perm_E_G2021>= obs_value_G2021)

cat("Mean of observed association =",obs_value_G2021,"\n")
cat("Mean of permuted association =",mean(perm_E_G2021),"\n")
cat("P-value is",pvalue_e_G2021, "\t")
if (pvalue_e_G2021< 0.05) cat("-> significative") else
  cat("-> not significative")

```
```{r}
hist(perm_E_G2021, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.2,0.5))
abline(v=obs_value_G2021, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mother-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

permutation
```{r}
original_pairid_G2021<- V(g_G2021)$pairID

num_perm<- 5000
perm_V_G2021<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2021, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2021
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2021[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2021<- mean(perm_V_G2021>= obs_value_G2021)
cat("Mean of observed association =",obs_value_G2021,"\n")
cat("Mean of permuted association =",mean(perm_V_G2021),"\n")
cat("P-value is",pvalue_v_G2021, "\t")
if (pvalue_v_G2021< 0.05) cat("-> significative") else
  cat("-> not significative")
```

```{r}
hist(perm_V_G2021, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.6))
abline(v=obs_value_G2021, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured

permute
```{r}
network_perm_G2021<-network_permutation(association_data=GBI_G2021,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2021,identities=colnames(GBI_G2021),days= 1:nrow(GBI_G2021))

```

```{r}
individuals_G2021<-colnames(GBI_G2021)
dimnames(network_perm_G2021)[[2]] <- individuals_G2021
dimnames(network_perm_G2021)[[3]] <- individuals_G2021
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2021<-apply(network_perm_G2021, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2021)
})

pvalue_N_G2021<- mean(MD_perm_G2021>= obs_value_G2021)

cat("Mean of observed association =",obs_value_G2021,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2021),"\n")
cat("P-value is",pvalue_N_G2021,"\t")
if (pvalue_N_G2021 < 0.05) cat("-> significative") else
  cat("-> not significative")

```

```{r}
hist(MD_perm_G2021, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.2,0.55))
abline(v=mean(obs_value_G2021), col="red", lwd=3)
```
# GB2 2022 ---------------------

```{r}
GB2_2022<-read.csv("CSV_Dayroosts_GB2_2022_juv.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)

GB2_2022<-GB2_2022[c(1:60)]#filter columns

GB2_2022 <-GB2_2022 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(GB2_2022)[colnames(GB2_2022) == "Date"] <- "date"
```

```{r}
colnames(GB2_2022)<-as.character(colnames(GB2_2022))
data_G2022<- GB2_2022%>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2022$daygroup<-as.factor(paste(data_G2022$date, data_G2022$box, sep="_"))

ind_G2022<-data.table(ID=c(data_G2022$ID), group=data_G2022$daygroup, day=c(data_G2022$date)) #df with 3 col: ID, group and day 
GBI_G2022<- get_group_by_individual(ind_G2022, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2022<- as.matrix(GBI_G2022)
```

## network
```{r}
net_G2022<-get_network(GBI_G2022,data_format = "GBI",association_index = "SRI")
```
```{r}
g_G2022<- graph_from_adjacency_matrix(net_G2022,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2022,vertex.label = V(g_G2022)$name,edge.width = E(g_G2022)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
#####plot + threshold
```{r}
threshold_G2022<- mean(net_G2022, na.rm = TRUE) +  sd(net_G2022, na.rm = TRUE)

gfilter_G2022<- delete_edges(g_G2022, E(g_G2022)[weight <threshold_G2022]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2022, edge.width = E(gfilter_G2022)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

##vertex attributes
```{r}
mother_daughter_pairs_G2022<-data.frame(mother=c("41C1","BF04","879F","E063","22B0","663E","F44A","A539","44BD","D629","AA3C","5939","2D31","D4DC","418B","8787","A816","F377","D14D"),daughter=c("DFAD","0636","12D7","0F3D","CC06","8C52","0896","922C","FB4B","F4D9","054D","0A62","8B80","F948","B484","D6D5","B0C8","EB80","BA7C"))

mother_daughter_pairs_G2022$pairID<-c(1:nrow(mother_daughter_pairs_G2022))

mother_ids_G2022<- mother_daughter_pairs_G2022$mother
daughter_ids_G2022<- mother_daughter_pairs_G2022$daughter

V(g_G2022)$type <- "other"  # default value
V(g_G2022)[V(g_G2022)$name %in% mother_ids_G2022]$type   <- "mother-daughter"
V(g_G2022)[V(g_G2022)$name %in% daughter_ids_G2022]$type <- "mother-daughter"

plot(gfilter_G2022, edge.width = E(gfilter_G2022)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2022)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2022)$name,edge.color="plum3")
```
```{r}
md_lookup_G2022<- mother_daughter_pairs_G2022%>%pivot_longer(cols = c("mother","daughter"), values_to = "ID") %>%select(ID, pairID)
#pivot the matrix so that each ID has the pair ID separately

IDs_G2022<- match(V(g_G2022)$name, md_lookup_G2022$ID)   # returns NA if the name is not found
V(g_G2022)$pairID <- md_lookup_G2022$pairID[IDs_G2022] 

pair_IDs_G2022<- factor(V(g_G2022)$pairID)

# Create a palette for however many pairIDs you have (excluding NA)
n_pairs_G2022<-length(levels(na.omit(pair_IDs_G2022)))
my_palette_G2022 <-  rainbow(19) 

# Map each vertex to the correct color
pair_colors_G2022<- my_palette_G2022[pair_IDs_G2022]
pair_colors_G2022[is.na(pair_colors_G2022)] <- "gray80"
```

##big plot
```{r}
plot(gfilter_G2022, edge.width = E(gfilter_G2022)$weight*6,edge.curved=F,vertex.color =pair_colors_G2022, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2022<- levels(na.omit(pair_IDs_G2022))
legend_colors_G2022 <- my_palette_G2022[seq_along(pair_levels_G2022)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2022), "random pair"),
  col    = c(legend_colors_G2022, "gray80"),
  pch    = 19,
  pt.cex = 2,
  cex = 0.6,
  bty    = "n"
)
```
###edge list
```{r}
edge_G2022<-  igraph::as_data_frame(g_G2022, what="edges")

vtype_G2022<- V(g_G2022)$type
names(vtype_G2022) <- V(g_G2022)$name

from_type_G2022<-vtype_G2022[ edge_G2022$from ]
to_type_G2022<-vtype_G2022[ edge_G2022$to   ]

edge_G2022$pair_type <- ifelse(
  from_type_G2022== "mother-daughter" & to_type_G2022 == "mother-daughter",
  "mother-daughter",
  "other"
)
```

### SRI plot
```{r}
sri_G2022<- as.vector(net_G2022)
sri_G2022 <- sri_G2022[sri_G2022 > 0]  # Remove zero values
meansri_G2022<- mean(sri_G2022)


hist(sri_G2022, breaks=30, main="Distribution of association index (SRI) for GB2 in 2022", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2022, col="blue", lwd=2)
```
Two different hist for the two types of pairs
```{r}

```
## EDGE PERMUTATION
Metrics of interest for quantifying: 
1- The mean edge weight among mother–daughter vertices
2- the difference in average association between mother–daughter pairs versus all other pairs

1-
observed value
```{r}
mean_pair_association <- function(g) {
  pairid <- V(g)$pairID #new variable with the pair ID
  names(pairid) <- V(g)$name #get the pair ID individual's name 
  #find the pairID of both individual
  from_pairid<- pairid[ edge_G2022$from ]
  to_pairid<- pairid[ edge_G2022$to   ]
  
  #has to have the same ID and not NA
  MDpairs <- (from_pairid == to_pairid) & !is.na(from_pairid)
  
  #isolate the weight of association and measure mean
  md_weights <- edge_G2022$weight[MDpairs]
  mean(md_weights, na.rm=TRUE)
}
obs_value_G2022<- mean_pair_association(g_G2022)
```

permutation
```{r}
num_permutations<- 5000
perm_E_G2022<-numeric(num_permutations)

for (i in seq_len(num_permutations)) {
  # Randomly permute the pair_type labels
  permuted_edge_list<-edge_G2022
  permuted_edge_list$pair_type <- sample(edge_G2022$pair_type)
  
  # Calculate the statistic for permuted mother-daughter pairs
  permuted_md_edges <- permuted_edge_list[permuted_edge_list$pair_type == "mother-daughter", ]
  perm_E_G2022[i] <- mean(permuted_md_edges$weight, na.rm = TRUE)
}
```

```{r}
pvalue_e_G2022<-mean(perm_E_G2022>= obs_value_G2022)

cat("Mean of observed association =",obs_value_G2022,"\n")
cat("Mean of permuted association =",mean(perm_E_G2022),"\n")
cat("P-value is",pvalue_e_G2022, "\t")
if (pvalue_e_G2022< 0.05) cat("-> significative") else
  cat("-> not significative")

```
```{r}
hist(perm_E_G2022, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(edge label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0.1,0.3))
abline(v=obs_value_G2022, col="red", lwd=3)
```
Good, however: maybe a vertex-label permutation would be better than a edge-label because mother-daughter is a fundamental property of individuals.

In edge-label, we shuffle which edges are labeled "MD", and in a vertex, we would shuffle the type of the vertex and then redefine the MD edges

vertex-label permutations best capture the idea that we’re testing whether these specific mother–daughter individuals have unusually strong connections compared to any random subset of individuals
But we would measure all the mother with all the mothers association?

## VERTEX PERMUTATION

I already calculated the observed value

permutation
```{r}
original_pairid_G2022<- V(g_G2022)$pairID

num_perm<- 5000
perm_V_G2022<- numeric(num_perm)

for (i in seq_len(num_perm)) {
  #shuffle the entire vector of pairIDs
  new_pid<- sample(original_pairid_G2022, replace = FALSE)
  
  #copy the original graph and assign the new id
  g_perm <- g_G2022
  V(g_perm)$pairID <- new_pid
  
  #mean association
  perm_V_G2022[i] <- mean_pair_association(g_perm)
}

```

```{r}
pvalue_v_G2022<- mean(perm_V_G2022>= obs_value_G2022)
cat("Mean of observed association =",obs_value_G2022,"\n")
cat("Mean of permuted association =",mean(perm_V_G2022),"\n")
cat("P-value is",pvalue_v_G2022, "\t")
if (pvalue_v_G2022< 0.05) cat("-> significative") else
  cat("-> not significative")
```

```{r}
hist(perm_V_G2022, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(vertex label permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.5))
abline(v=obs_value_G2022, col="red", lwd=3)
```

##DATA RANDOMIZATION 
(randomization of the entire network, then calculation of the differences between md pairs and non-md pairs to see if it is different than what is observed)
mean of association for mother-daughter pairs is already measured

permute
```{r}
network_perm_G2022<-network_permutation(association_data=GBI_G2022,data_format = "GBI",permutations=5000,association_index = "SRI",association_matrix = net_G2022,identities=colnames(GBI_G2022),days= 1:nrow(GBI_G2022))

```

```{r}
individuals_G2022<-colnames(GBI_G2022)
dimnames(network_perm_G2022)[[2]] <- individuals_G2022
dimnames(network_perm_G2022)[[3]] <- individuals_G2022
#network_perm[1,,]

md_mean_from_adj <- function(adj_mat, md_pairs) {
  # 'adj_mat': a 26x26 adjacency matrix with row/col names = individual IDs
  # 'md_pairs': a data frame with columns "mother", "daughter" 
  #    e.g. mother = "B587", daughter = "5672", etc.

  n_pairs <- nrow(md_pairs)
  vals <- numeric(n_pairs)

  for (i in seq_len(n_pairs)) {
    mother_id <- md_pairs$mother[i]
    daughter_id <- md_pairs$daughter[i]

    # match them to the adjacency
    row_idx <- match(mother_id, rownames(adj_mat))
    col_idx <- match(daughter_id, colnames(adj_mat))

    # If they exist in the names, get the weight
    # If either is NA, the pair is not found, so set to NA
    if (!is.na(row_idx) && !is.na(col_idx)) {
      vals[i] <- adj_mat[row_idx, col_idx]
    } else {
      vals[i] <- NA
    }
  }

  # Return the mean edge weight across all real MD pairs
  mean(vals, na.rm = TRUE)
}

```

```{r}
#measure the association only for the md in the entire permutations
MD_perm_G2022<-apply(network_perm_G2022, 1, function(adj_mat) {
  md_mean_from_adj(adj_mat, mother_daughter_pairs_G2022)
})

pvalue_N_G2022<- mean(MD_perm_G2022>= obs_value_G2022)

cat("Mean of observed association =",obs_value_G2022,"\n")
cat("Mean of permuted association =",mean(MD_perm_G2022),"\n")
cat("P-value is",pvalue_N_G2022,"\t")
if (pvalue_N_G2022 < 0.05) cat("-> significative") else
  cat("-> not significative")

```

```{r}
hist(MD_perm_G2022, breaks=30, col="lightblue",
     main="Distribution of the permuted mother-daughter association \n(whole network permutation)",
     xlab="Mean association among mother-daughter vertices (permuted)",xlim = c(0,0.2))
abline(v=mean(obs_value_G2022), col="red", lwd=3)
```

