---
title: "Data Preparation for Network Analysis "
author: 
  name: "Flore Michel"
  email: flore.michel@unil.ch
  affiliation:
    name: University of Lausanne, Departement of Ecology and Evolution
execute:                            # Global options for code chunks
  echo: true                        # Show code globally
  warning: false                    # Hide warnings globally
  freeze: auto                      # Only re-render if code changes
  cache: true                       # Cache results (best for heavy computations)
format:                             # Global options for rendering the document
  html:                              # Defines render format in this case html
    margin-top: 30px
    toc: true                       # Includes table of content
    toc-location: left              # Sets the location of the toc
    toc-depth: 2                    # Sets the number of levels for the toc
    number-sections: true           # Add automatic section number
    number-depth: 2                 # Sets the maximum numberin level e.g. 2.2.1
    #code-fold: true                 # Code chunk can be expanded
    fig-cap-location: bottom        # Location of figure captions [top, bottom]
    lightbox: true                  # When clicking on a figure it is expanded
    embed-resources: true           # Make the report a stand-alone for sharing
    theme:                          # Set the theme of the report 
      light: litera                   # [default, cerulean, cosmo, flatly, journal, ...]
      dark: slate                   # [cyborg, darkly, solar, superhero, vapor, ...]
#bibliography: bib/workshop.bib.     # References used in the document
#csl: bib/apa.csl                    # Citation style for rendering the references
bibliography: references.bib
---

# Introduction

This first script imports, clean and prepares raw association data collected from *Myotis bechsteinii* in two colonies (GB2 and UA), near Wurzburg in Germany between 2013 and 2023.

### Data description

-   **Raw roost datasets**: one data set for each year, each row representing a day of observation, column names represent the individuals in the colony. Number in the table represent the roost number where the individual spent the day.

-   **Mother-daughter file**: One for each colony, each row is a mother-daughter pair. Column name "rfid_mom" (mother ID), "rfid_off" (daughter ID), year and a unique pairID composed of the rfid_mom and rfid_off, order alphabetically to ensure universal ID.

### Load packages

```{r}
#| label: packages
#| message: false
#| warning: false
#code-fold: true
library(dplyr)
library(here)
library(igraph)
library(asnipe)
library(sna)
library(ggplot2)
library(tidyverse)
library(igraph)
library(data.table)
```

# Load datasets

### Mother-daughter lists

```{r}
mother_daughter_GB2 <- read.csv(here("data","mother_daughter_GB2.csv"))
mother_daughter_UA <- read.csv(here("data","mother_daughter_UA.csv"))
```

# Load yearly roost data

## GB2

```{r}
#| label: load roost df
#| output: false
#| code-fold: true
#load data set and clean unnecessary columns
# 2013
GB2_2013<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2013_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2013<-GB2_2013[c(0:30)]
GB2_2013 <- GB2_2013[, -4]
GB2_2013 <- GB2_2013[, -3]
GB2_2013 <- GB2_2013[, -2]
colnames(GB2_2013)[colnames(GB2_2013) == "number"] <- "date"
#_______________________________________________________________
#2014
GB2_2014<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2014_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2014<-GB2_2014[c(0:40)] #filter columns
GB2_2014<-GB2_2014[1:51,]
GB2_2014 <- GB2_2014[, -4]
GB2_2014 <- GB2_2014[, -3]
GB2_2014 <- GB2_2014[, -2]
colnames(GB2_2014)[colnames(GB2_2014) == "number"] <- "date"
GB2_2014 <- GB2_2014[, !(colnames(GB2_2014) %in% c("E4D5", "B661"))] #fix mismatching in individual ID
#_______________________________________________________________
#2015
GB2_2015<-read.csv(here("data","raw_roost_data","Dayroosts_GB2_2015_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2015<-GB2_2015[c(0:38)] #filter columns
GB2_2015<-GB2_2015[1:45,]

colnames(GB2_2015)[colnames(GB2_2015) == "number"] <- "date"

#remove one individual (not enough data)
GB2_2015<- GB2_2015[, !(colnames(GB2_2015) %in% "3D98")]
#_______________________________________________________________
#2016
GB2_2016<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2016_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2016<-GB2_2016[c(0:25)] #filter columns
colnames(GB2_2016)<-as.character(colnames(GB2_2016))

#_______________________________________________________________
#2017
GB2_2017<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2017_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2017<-GB2_2017[c(0:36)] #filter columns
GB2_2017<-GB2_2017[1:42,]

#_______________________________________________________________
#2018
GB2_2018<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2018_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2018<-GB2_2018[c(1:39)] #filter columns
GB2_2018<-GB2_2018[1:17,]

GB2_2018<- GB2_2018[, !(colnames(GB2_2018) %in% "F44A")]

#filter individuals that are observed less than 2 times
#count number of non-NA presences per individual
presence_counts_G2018 <- colSums(!is.na(GB2_2018[,-1]))  # exclude date column

# Keep only individuals with >= 2 or 3 presences
keep_IDs_G2018 <- names(presence_counts_G2018[presence_counts_G2018 >= 2])
print(ncol(GB2_2018[,-1]));print(length(keep_IDs_G2018))

GB2_2018 <- GB2_2018[, c("date", keep_IDs_G2018)]

#_______________________________________________________________
#2019
GB2_2019<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2019_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2019<-GB2_2019[c(1:41)] #filter columns

GB2_2019 <- GB2_2019 %>%
  mutate(across(-date, as.character))

#_______________________________________________________________
#2020
GB2_2020<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2020_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2020<-GB2_2020[c(1:51)] #filter columns
GB2_2020<-GB2_2020[1:27,]

GB2_2020 <- GB2_2020 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

GB2_2020 <- GB2_2020 %>%
  mutate(across(-date, as.character))
#_______________________________________________________________
#2021
GB2_2021<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2021_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)

GB2_2021<-GB2_2021[c(1:56)] #filter columns
GB2_2021<-GB2_2021[1:37,]

GB2_2021 <- GB2_2021 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

colnames(GB2_2021)[colnames(GB2_2021) == "Date"] <- "date"

GB2_2021<- GB2_2021 %>%
  mutate(across(-date, as.character))

#_______________________________________________________________
#2022
GB2_2022<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2022_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)

GB2_2022<-GB2_2022[c(1:60)]#filter columns

GB2_2022 <-GB2_2022 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(GB2_2022)[colnames(GB2_2022) == "Date"] <- "date"

#filter individuals that are observed less than 2 times
#count number of non-NA presences per individual
presence_counts_G2022 <- colSums(!is.na(GB2_2022[,-1]))  # exclude date column

# Keep only individuals with >= 2 or 3 presences
keep_IDs_G2022 <- names(presence_counts_G2022[presence_counts_G2022 >= 2])
print(ncol(GB2_2022[,-1]));print(length(keep_IDs_G2022))

GB2_2022 <- GB2_2022[, c("date", keep_IDs_G2022)]

GB2_2022<- GB2_2022 %>%
  mutate(across(-date, as.character))

#_______________________________________________________________
#2023
GB2_2023<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2023_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2023<-GB2_2023[c(1:70)] #filter columns

GB2_2023<-GB2_2023 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

#GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(GB2_2023)[colnames(GB2_2023) == "Date"] <- "date"

GB2_2023<- GB2_2023 %>%
  mutate(across(-date, as.character))
```

# Create GBI and association matrices, and edge lists from raw association datasets

The **Group-by-Individual matrix** (GBI), is the format required for most function in the <igraph> and <asnipe> packages. We can create the GBI matrix from raw roost data by merging day and box number together, creating unique "group" of individuals that roosted all together in the same day in the same roost. This transforms the dataset into a presence/absence matrix. The column names remain the individual ID.  

We can also create the **association matrices** that is a 

```{r}
#Create a function that will create the GBI, the association matrices (net), the network objects and the edge list for all years 
create_matrices <- function(data, year, colony, mother_daughter_df) {
  
  # Create GBI matrix
  data_long <- data %>%
    pivot_longer(
      cols = -c(date),
      names_to = "ID",
      values_to = "box"
    ) %>%
    drop_na(box)
  
  data_long$daygroup <- as.factor(paste(data_long$date, data_long$box, sep = "_"))
  
  ind <- data.table(
    ID = data_long$ID,
    group = data_long$daygroup,
    day = data_long$date
  )
  
  #then use the function in the 'asnipe' package
  GBI <- get_group_by_individual(ind, data_format = "individuals")
  GBI <- as.matrix(GBI)
  write.csv(GBI, here("data", "saved_matrices", colony, paste0("GBI_G", year, ".csv")), row.names = TRUE)
  
  # Create network
  net <- get_network(GBI, data_format = "GBI", association_index = "SRI")
  write.csv(net, here("data", "saved_matrices", colony, paste0("net_G", year, ".csv")), row.names = TRUE)
  
  # Create edge list
  g <- graph_from_adjacency_matrix(net, mode = "undirected", weighted = TRUE, diag = FALSE)
  edge <- igraph::as_data_frame(g, what = "edges")
  
  edge <- edge %>%
    mutate(
      pairID = ifelse(
        from < to,
        paste(from, to, sep = "-"),
        paste(to, from, sep = "-")
      )
    )
  
  # Add pair type, either "mother-daughter" if the pairID is present in the mother_daughter df, otherwise "other"
  edge <- edge %>%
    mutate(
      pair_type = ifelse(
        pairID %in% mother_daughter_df$pairID[mother_daughter_df$year == year], 
        "mother-daughter",
        "other"
      )
    )
  write.csv(edge, here("data", "saved_matrices", colony, paste0("edge_G", year, ".csv")), row.names = FALSE)
  
  return(edge)
}
```

Run the function
```{r}
#| label: run function
#| message: false
#| warning: false
#| output: false
#| 
years <- 2013:2023

for (y in 2013:2023) {
  create_matrices(
    data = get(paste0("GB2_", y)),
    year = y,
    colony = "GB2",
    mother_daughter_df = mother_daughter_GB2
  )
}
```

#### Create the edge list of all the years together
```{r}
edges_all <- lapply(2013:2023, function(y) {
  #load the created edge lists for each year
  df <- read.csv(here("data", "saved_matrices", "GB2", paste0("edge_G", y, ".csv")))
  # add the year in a new column
  df$year <- y
  return(df)
})

edges_GB2 <- bind_rows(edges_all)
#and save it 
write.csv(edges_GB2, here("data","saved_matrices","GB2","edge_list_GB2.csv"), row.names = FALSE)
```

Then, we can do the same for the other colony

## UA 
#### load raw association data
```{r}
#| label: load roost df
#| output: false
#| code-fold: true
#load data set and clean unnecessary columns
# 2013
UA_2013<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2013_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2013<-UA_2013[c(1:24)]
colnames(UA_2013)[colnames(UA_2013) == "number"] <- "date"
#_______________________________________________________________
#2014
UA_2014<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2014_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2014<-UA_2014[c(1:30)] #filter columns
UA_2014<-UA_2014[1:55,]

colnames(UA_2014)[colnames(UA_2014) == "number"] <- "date"
UA_2014<- UA_2014[, !(colnames(UA_2014) %in% "Year")]
UA_2014<- UA_2014[, !(colnames(UA_2014) %in% "Month")]
UA_2014<- UA_2014[, !(colnames(UA_2014) %in% "Day")]
#_______________________________________________________________
#2015
UA_2015<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2015_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2015<-UA_2015[c(1:23)] #filter columns
UA_2015<-UA_2015[2:51,]

UA_2015<-UA_2015 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

#GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(UA_2015)[colnames(UA_2015) == "number"] <- "date"
#_______________________________________________________________
#2016
UA_2016<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2016_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2016<-UA_2016[c(1:27)] #filter columns
UA_2016<-UA_2016[2:39,]

#_______________________________________________________________
#2017
UA_2017<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2017_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2017<-UA_2017[c(1:30)] #filter columns

#_______________________________________________________________
#2018
UA_2018<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2018_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2018<-UA_2018[c(1:27)] #filter columns
names(UA_2018)[names(UA_2018) == "D6CC"] <- "FA5B"
names(UA_2018)[names(UA_2018) == "0B43"] <- "101F"

#_______________________________________________________________
#2019
UA_2019<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2019_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2019<-UA_2019[c(1:23)] #filter columns
UA_2019<-UA_2019[1:41,]

UA_2019<-UA_2019 %>%
 mutate(across(everything(), ~ replace(., . == "", NA)))
UA_2019 <- UA_2019 %>%
  mutate(across(-date, as.character))

#_______________________________________________________________
#2020
UA_2020<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2020_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2020<-UA_2020[c(1:24)] #filter columns
UA_2020<-UA_2020[1:28,]

UA_2020 <- UA_2020 %>%
  mutate(across(-date, as.character))
#_______________________________________________________________
#2021
UA_2021<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2021_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)

UA_2021<-UA_2021[c(1:25)] #filter columns

UA_2021<-UA_2021 %>%
 mutate(across(everything(), ~ replace(., . == "", NA)))


colnames(UA_2021)[colnames(UA_2021) == "Date"] <- "date"
names(UA_2021)[names(UA_2021) == "9B39"] <- "F50E"
names(UA_2021)[names(UA_2021) == "37F2"] <- "100F"

UA_2021<- UA_2021 %>%
  mutate(across(-date, as.character))

#_______________________________________________________________
#2022
UA_2022<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2022_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2022<-UA_2022[c(1:22)] #filter columns
UA_2022<-UA_2022[1:29,]

colnames(UA_2022)[colnames(UA_2022) == "Date"] <- "date"

UA_2022<- UA_2022 %>%
  mutate(across(-date, as.character))

#_______________________________________________________________
#2023
UA_2023<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2023_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2023<-UA_2023[c(1:29)]
colnames(UA_2023)[colnames(UA_2023) == "Date"] <- "date"
```

# Create GBI and association matrices, and edge lists from raw association datasets
```{r}
#Create a function that will create the GBI, the association matrices (net), the network objects and the edge list for all years 
create_matrices_UA <- function(data, year, colony, mother_daughter_df) {
  
  # Create GBI matrix
  data_long <- data %>%
    pivot_longer(
      cols = -c(date),
      names_to = "ID",
      values_to = "box"
    ) %>%
    drop_na(box)
  
  data_long$daygroup <- as.factor(paste(data_long$date, data_long$box, sep = "_"))
  
  ind <- data.table(
    ID = data_long$ID,
    group = data_long$daygroup,
    day = data_long$date
  )
  
  #then use the function in the 'asnipe' package
  GBI <- get_group_by_individual(ind, data_format = "individuals")
  GBI <- as.matrix(GBI)
  write.csv(GBI, here("data", "saved_matrices", colony, paste0("GBI_U", year, ".csv")), row.names = TRUE)
  
  # Create network
  net <- get_network(GBI, data_format = "GBI", association_index = "SRI")
  write.csv(net, here("data", "saved_matrices", colony, paste0("net_U", year, ".csv")), row.names = TRUE)
  
  # Create edge list
  g <- graph_from_adjacency_matrix(net, mode = "undirected", weighted = TRUE, diag = FALSE)
  edge <- igraph::as_data_frame(g, what = "edges")
  
  edge <- edge %>%
    mutate(
      pairID = ifelse(
        from < to,
        paste(from, to, sep = "-"),
        paste(to, from, sep = "-")
      )
    )
  
  # Add pair type, either "mother-daughter" if the pairID is present in the mother_daughter df, otherwise "other"
  edge <- edge %>%
    mutate(
      pair_type = ifelse(
        pairID %in% mother_daughter_df$pairID[mother_daughter_df$year == year], 
        "mother-daughter",
        "other"
      )
    )
  write.csv(edge, here("data", "saved_matrices", colony, paste0("edge_U", year, ".csv")), row.names = FALSE)
  
  return(edge)
}
```

Run the function
```{r}
#| label: run function
#| message: false
#| warning: false
#| output: false
#| 
years <- 2013:2023

for (y in 2013:2023) {
  create_matrices_UA(
    data = get(paste0("UA_", y)),
    year = y,
    colony = "UA",
    mother_daughter_df = mother_daughter_UA
  )
}
```

#### Create the edge list of all the years together
```{r}
edges_all <- lapply(2013:2023, function(y) {
  #load the created edge lists for each year
  df <- read.csv(here("data", "saved_matrices", "UA", paste0("edge_U", y, ".csv")))
  # add the year in a new column
  df$year <- y
  return(df)
})

edges_UA <- bind_rows(edges_all)
#and save it 
write.csv(edges_UA, here("data","saved_matrices","UA","edge_list_UA.csv"), row.names = FALSE)
```

# Mother - non-related daughter datasets
Creation of a lists that state associations between mothers and non-related daughters, by creating a function that goes into the mother-daughter df, and assign them in a role for a specific year, either "mother" or "daughter"

```{r}
make_mom_juv_for_year <- function(y, edges, md_table) {
  md_y <- md_table %>% filter(year == y)

  # bats that matter in this year -> mothers and daughters only
  bats_y <- unique(c(md_y$rfid_mom, md_y$rfid_off))

  edges_y <- edges %>%
    filter(year == y) %>%
    filter(from %in% bats_y | to %in% bats_y)

  # lookup table with roles
  lookup_y <- md_y %>%
    pivot_longer(
      cols = c(rfid_mom, rfid_off),
      names_to = "role",
      values_to = "batID"
    ) %>%
    distinct(batID, role)

  edges_y <- edges_y %>%
    left_join(lookup_y, by = c("from" = "batID")) %>%
    rename(from_role = role) %>%
    left_join(lookup_y, by = c("to" = "batID")) %>%
    rename(to_role = role)

  # ordered pairID, to ensure correct matching
  make_pairID <- function(x, y) {
    ifelse(x < y, paste(x, y, sep = "-"), paste(y, x, sep = "-"))
  }

  edges_y <- edges_y %>%
    mutate(pairID = make_pairID(from, to))

  # identify true MD pairs for this year
  true_MD_y <- md_y %>%
    mutate(pairID = make_pairID(rfid_mom, rfid_off)) %>%
    select(pairID)

  edges_y <- edges_y %>%
    mutate(
    pair_type_2 = case_when(
      # true mother–daughter
      pairID %in% true_MD_y$pairID ~ "mother-daughter",

      # mother with a non-own juvenile
      ((from_role == "rfid_mom" & to_role == "rfid_off") |
        (from_role == "rfid_off" & to_role == "rfid_mom")
      ) ~ "mother-random-juvenile",

      # everything else
      TRUE ~ "other"
    )
  )

}
```

Run function for both colonies
```{r}
years <- sort(unique(edges_GB2$year))

#for GB2
mom_juv_GB2 <- map_dfr(
  years,
  make_mom_juv_for_year,
  edges = edges_GB2,
  md_table = mother_daughter_GB2
)


mom_juv_UA <- map_dfr(
  years,
  make_mom_juv_for_year,
  edges = edges_UA,
  md_table = mother_daughter_UA
)
#save datasets
write.csv(mom_juv_GB2, here("data","saved_matrices","GB2","mom_juv_GB2.csv"), row.names = FALSE)
write.csv(mom_juv_UA, here("data","saved_matrices","UA","mom_juv_UA.csv"), row.names = FALSE)
```
