---
title: "Master_hypothesis3"
author: "Flore Michel"
date: "2025-02-19"
output: html_document
---

```{r setup, include=FALSE}
library(here)
library(igraph)
library(asnipe)
library(sna)
library(assocInd)
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(tidyverse)
library(viridis)
library(patchwork)
library(igraph)
library(ggraph)
library(colormap)
library(tidyr)
library(dplyr)
library(data.table)
library(glmmTMB)
library(DHARMa)
library(influence.ME)
library(MuMIn)
library(performance)
library(car)
library(ggrepel)
library(ggplot2)
library(patchwork)      
library(ggraph)
library(igraph)
library(dplyr)
library(ggsignif)
#library(lemon)     
#library(scales)
```

Load
```{r }
edges_GB2<-read.csv(here("data","saved_matrices","GB2","full","full_edges_GB2.csv"),header=TRUE, row.names=NULL,sep=",",check.names = F)
mother_daughter_GB2<-read.csv(here("data","mother_daughter_GB2.csv"),header=TRUE, row.names=NULL,sep=",",check.names = F)
```

```{r load the networls, include= FALSE}
net_G2013<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2013<- as.matrix(net_G2013) 
net_G2014<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2014 <- as.matrix(net_G2014)
net_G2015<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2015 <- as.matrix(net_G2015)
net_G2016<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2016 <- as.matrix(net_G2016)
net_G2017<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2017 <- as.matrix(net_G2017)
net_G2018<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2018 <- as.matrix(net_G2018)
net_G2019<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2019 <- as.matrix(net_G2019)
net_G2020<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2020 <- as.matrix(net_G2020)
net_G2021<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2021 <- as.matrix(net_G2021)
net_G2022<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2022 <- as.matrix(net_G2022)
net_G2023<-read.csv(here("data","saved_matrices","GB2","full","full_net_G2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2023 <- as.matrix(net_G2023)
```

```{r}
g_G2013<- graph_from_adjacency_matrix(net_G2013, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2014<- graph_from_adjacency_matrix(net_G2014, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2015<- graph_from_adjacency_matrix(net_G2015, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2016<- graph_from_adjacency_matrix(net_G2016, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2017<- graph_from_adjacency_matrix(net_G2017, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2018<- graph_from_adjacency_matrix(net_G2018, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2019<- graph_from_adjacency_matrix(net_G2019, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2020<- graph_from_adjacency_matrix(net_G2020, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2021<- graph_from_adjacency_matrix(net_G2021, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2022<- graph_from_adjacency_matrix(net_G2022, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2023<- graph_from_adjacency_matrix(net_G2023, mode = "undirected", weighted = TRUE, diag = FALSE)
```

# Hypothesis 

*How do these mother-daughter associations influences the network structure?*
Do mothers that have a lot of daughters over the years are more central/important in the network than others?

#---------- colony GB2

#offspring table
```{r}
edges_MD_GB2 <- edges_GB2 %>%
  filter(pairID %in% mother_daughter_GB2$pairID)
```

```{r}
offspring_table_GB2 <- edges_MD_GB2 %>%
  select(from, to, year, pairID) %>%
   pivot_longer(cols = c("from", "to"), values_to = "ID") %>%
   select(ID, year, pairID) %>%
   distinct()  # no duplicate rows

#then we can add a new column "is mother", to be sure that the individual is the mother in the pair (look in mother-daughter df)
offspring_table_GB2 <- offspring_table_GB2 %>%
  left_join(
    mother_daughter_GB2 %>%
      select(pairID, rfid_mom) , by = "pairID" )%>%
      mutate(is_mother = ifelse(ID == rfid_mom, "mother", "daughter")
  )
```

Then, we can measure the number of offsprings for each mother
```{r}
offspring_counts_GB2 <- offspring_table_GB2 %>%
  filter(is_mother == "mother") %>% 
  group_by(ID, year) %>%
  summarise(total_offspring = n_distinct(pairID, na.rm = TRUE), .groups = "drop")

offspring_table_GB2 <- left_join(offspring_table_GB2, offspring_counts_GB2, by = c("ID", "year"))

offspring_table_GB2$total_offspring[is.na(offspring_table_GB2$total_offspring)] <- 0

offspring_table_GB2 <- offspring_table_GB2 %>%
  mutate(across(c(ID, pairID), as.character)) 
#write.csv(offspring_table, "variables_GB2/offspring_table.csv", row.names = FALSE,quote = TRUE)
```

ADD mother AGE as a control !!!
```{r}
mother_df <- read.csv(here("data", "raw_roost_data","mother daughter pairs_2013-2023.csv"),
                      header = TRUE,
                      sep    = ";",
                      check.names = FALSE) %>% 
  select(Year, Colony, TAG_mom, birth_mom) %>% 
  #birth as integrer 
  mutate(Year = as.integer(Year),
         birth_mom = as.integer(birth_mom)
  )

#rename some individual in TAG_mom
mother_df$TAG_mom[mother_df$TAG_mom == "B597=B587"] <- "B587"
mother_df$TAG_mom[mother_df$TAG_mom == "6C0D=6D0D"] <- "6D0D"
mother_df$TAG_mom[mother_df$TAG_mom == "07F9=51C3"] <- "07F9"
mother_df$TAG_mom[mother_df$TAG_mom == "EE00=7C6F"] <- "7C6F"
mother_df$TAG_mom[mother_df$TAG_mom == "5E-BF3C=A4A4"] <- "5EBF3C"
mother_df$TAG_mom[mother_df$TAG_mom == "4FCA=7CCA"] <- "4FCA"
mother_df$TAG_mom[mother_df$TAG_mom == "F50E=9B39=AE33"] <- "F50E"
mother_df$TAG_mom[mother_df$TAG_mom == "37F2=100F"] <- "100F"
mother_df$TAG_mom[mother_df$TAG_mom == "DE72=15BF"] <-"15BF"

```

Then, create a df with ALL females age (add the daughters age aswell)
```{r}
female_birth_year <- bind_rows(
  mother_df %>%
    filter(Colony == "GB2") %>%
    transmute(
      ID = as.character(TAG_mom),
      birth_year = as.integer(birth_mom)
    ),
  mother_daughter_GB2 %>%
    transmute(
      ID = as.character(rfid_off),
      birth_year = as.integer(year)
    )
) %>%
  distinct(ID, birth_year)
```

join to get mother age at each year (Year - birth year),
```{r}
offspring_table_GB2 <- offspring_table_GB2 %>%
  mutate(ID = as.character(ID)) %>%
  left_join(
    female_birth_year %>%
      distinct(ID, birth_year) %>%
      mutate(
        ID        = as.character(ID),
        birth_year = as.integer(birth_year)
      ) %>%
      select(ID, birth_year),
    by = "ID"
  )

#create mother_age
offspring_table_GB2 <- offspring_table_GB2 %>%
  mutate(female_age = year - birth_year)

#put "unkown age" when we don't have the birth year of the mother (only where is_mother is "mother)
offspring_table_GB2 <- offspring_table_GB2 %>%
  mutate(unkown_age = ifelse(is_mother == "mother" & is.na(female_age), TRUE, FALSE)
  )
```
Now we can join the info on the offsprings on the edges list (total)
--> first need to creat ethe node list and not the edge list 
```{r}
nodes_GB2 <- edges_GB2 %>%
  select(year, from, to) %>%
  pivot_longer(cols = c(from, to), names_to = "role",
    values_to = "ID"
  ) %>%
  select(year, ID) %>%
  distinct()

nodes_GB2 <- nodes_GB2 %>%
  left_join(
    offspring_table_GB2 %>%
      select(
        ID, year,
        is_mother,
        total_offspring,
        unkown_age
      ),
    by = c("ID", "year")
  ) %>%
  mutate(
    is_mother = if_else(is.na(is_mother), "non_mother", is_mother),
    total_offspring = replace_na(total_offspring, 0),
    unkown_age = replace_na(unkown_age, 0L)
  )

nodes_GB2 <- nodes_GB2 %>%
  left_join(female_birth_year, by = "ID") %>%
  mutate(
    female_age = year - birth_year
  )

nodes_GB2 <- nodes_GB2 %>%
  distinct(
    year, ID,
    .keep_all = TRUE
  )

nodes_GB2 <- nodes_GB2 %>%
  mutate(unknown_age = ifelse(is.na(birth_year), TRUE, FALSE)) %>% 
  select(-unkown_age)
```


# Measure all three metrics and put in the same table
#metrics
```{r}
network_list_GB2 <- list(
  "2013" = net_G2013, "2014" = net_G2014, "2015" = net_G2015,
  "2016" = net_G2016, "2017" = net_G2017, "2018" = net_G2018,
  "2019" = net_G2019, "2020" = net_G2020, "2021" = net_G2021,
  "2022" = net_G2022, "2023" = net_G2023
)
# each year, build the graph once and compute all three metrics
centrality_list_GB2 <- lapply(names(network_list_GB2), function(year_char) {

  year_int <- as.integer(year_char)
  
  # Build the undirected, weighted igraph for that year
  mat <- network_list_GB2[[year_char]]
  g    <- graph_from_adjacency_matrix(
    mat,
    mode     = "undirected",
    weighted = TRUE,
    diag     = FALSE
  )
  
  # closeness 
  E(g)$dist <- 1 / E(g)$weight
  
  clos_vals <- igraph::closeness(
  g,
  vids = V(g),
  weights = E(g)$dist,
  normalized = TRUE
)
  
  #eigenvector centrality
  eig_vals <- igraph::eigen_centrality(
    g,
    directed = FALSE,
    weights  = E(g)$weight
  )$vector
  
  #Strength (weighted degre)
  str_vals <- igraph::strength(
    g,
    v       = V(g),
    mode    = "all",
    weights = E(g)$weight
  )
  
  # create df
  data.frame(
    ID            = names(clos_vals),
    year          = year_int,
    closeness     = as.numeric(clos_vals),
    eigen_cent    = as.numeric(eig_vals),
    strength      = as.numeric(str_vals),
    stringsAsFactors = FALSE
  )
})

centrality_metrics_GB2 <- bind_rows(centrality_list_GB2)

#merge
centrality_GB2 <- nodes_GB2 %>%
  mutate(
    year = as.integer(year),
    ID   = as.character(ID)
  ) %>%

  left_join(
    centrality_metrics_GB2 %>%
      mutate(
        year = as.integer(year),
        ID   = as.character(ID)
      ),
    by = c("ID", "year")
  )

#add colon ysize
colony_size_GB2 <- read.csv(here("data","saved_matrices","GB2","colony_size_df.csv"),
  header      = TRUE,
  sep         = ",",
  check.names = FALSE
) %>%
  mutate(year = as.integer(year))

centrality_GB2 <- centrality_GB2 %>%
  left_join(colony_size_GB2, by = "year")

centrality_GB2$total_offspring <- as.numeric(centrality_GB2$total_offspring)
centrality_GB2$female_age <- as.numeric(centrality_GB2$female_age)

#recalculate female age
centrality_GB2 <- centrality_GB2 %>% 
  mutate(female_age = year - birth_year)

#add "offspring cat", categorical variable with 3 categories: 0, 1, 2+ (2 or more)
centrality_GB2 <- centrality_GB2 %>%
  mutate(offspring_cat = case_when(
    total_offspring == 0 ~ "0",
    total_offspring == 1 ~ "1",
    total_offspring == 2 ~ "2",
    total_offspring >= 3 ~ "3+"
  ))

centrality_GB2 <- centrality_GB2 %>%
  group_by(year) %>%
  mutate(
    strength_z   = as.numeric(scale(strength)),
    eigen_z      = as.numeric(scale(eigen_cent)),
    closeness_z  = as.numeric(scale(closeness))
  ) %>%
  ungroup()

```
Check correlation between the three metrics
```{r}
cor(centrality_GB2[, c("strength","eigen_cent","closeness")],
    use="complete.obs")

#and scaled metrics too 
cor(centrality_GB2[, c("strength_z","eigen_z","closeness_z")],
    use="complete.obs")
```

#Analysis
### first check the distribution

```{r}
ggplot(centrality_GB2, aes(strength)) +
  geom_histogram(bins = 30) +
  facet_wrap(~year)

ggplot(centrality_GB2, aes(eigen_cent)) +
  geom_histogram(bins = 30) +
  facet_wrap(~year)

ggplot(centrality_GB2, aes(closeness)) +
  geom_histogram(bins = 30) +
  facet_wrap(~year)
```


```{r}
#check only 1 year
centrality_GB2_2013<- centrality_GB2 %>%
  filter(year == 2013)

#model to investigate eigenvector centrality with number of offsrping, while controling for mother_age 
m <- lm(eigen_cent ~ total_offspring + female_age + unknown_age, data = centrality_GB2_2013)
summary(m)

plot(m)

m2 <- lm(eigen_cent ~ offspring_cat + female_age, data = centrality_GB2_2013)
summary(m2)

plot(m2)
```


```{r}
coef(m)[c("total_offspring","female_age")]
#check final model
simulationOutput_eigen_GB2 <- simulateResiduals(fittedModel = m, n = 1000)
plot(simulationOutput_eigen_GB2)
testUniformity(simulationOutput_eigen_GB2)   # KS test for uniformity
testDispersion(simulationOutput_eigen_GB2)     # Test for over/under dispersion
testZeroInflation(simulationOutput_eigen_GB2) 


par(mfrow = c(1,2))
plot(m, which = 1, main = "Residuals")
plot(m, which = 2, main = "QQ")

df_cor <- centrality_GB2 %>%
  filter(total_offspring > 0) %>%
  filter(!is.na(female_age), !is.na(total_offspring))

cor(df_cor$female_age, df_cor$total_offspring)
```

#### distribution
```{r}
hist(centrality_GB2$eigen_cent); range(centrality_GB2$eigen_cent) #between 0 and 1 --> need to change the model
hist(centrality_GB2$closeness); range(centrality_GB2$closeness)
hist(centrality_GB2$strength); range(centrality_GB2$strength)

#of scaled metrics 
hist(centrality_GB2$eigen_z); range(centrality_GB2$eigen_z) #between 0 and 1 --> need to change the model
hist(centrality_GB2$closeness_z); range(centrality_GB2$closeness_z)
hist(centrality_GB2$strength_z); range(centrality_GB2$strength_z)
```

#Permutations 
```{r perm function for all metrics}
run_permutation_by_year <- function(df,
                                    years  = 2013:2023,
                                    n_perm = 1000) {

  metrics <- c("eigen_cent", "closeness", "strength")

  out_list <- vector("list", length(years) * length(metrics))
  idx <- 1

  for (yr in years) {

    df_model <- df %>%
      dplyr::filter(year == yr)

    for (m in metrics) {

      # observed model
      form <- as.formula(
        paste(m, "~ total_offspring + female_age")
      )

      mod_obs <- lm(form, data = df_model)

      beta_obs_off <- as.numeric(coef(mod_obs)["total_offspring"])
      beta_obs_age <- as.numeric(coef(mod_obs)["female_age"])

      # permutation storage
      beta_perm_off <- numeric(n_perm)
      beta_perm_age <- numeric(n_perm)

      for (i in seq_len(n_perm)) {

        perm_df <- df_model %>%
          dplyr::mutate(
            total_offspring = sample(total_offspring)
          )

        mod_p <- lm(form, data = perm_df)

        beta_perm_off[i] <- as.numeric(coef(mod_p)["total_offspring"])
        beta_perm_age[i] <- as.numeric(coef(mod_p)["female_age"])
      }

      # permutation p-value (offspring only)
      p_perm_off <- mean(abs(beta_perm_off) >= abs(beta_obs_off))
      p_perm_age <- mean(abs(beta_perm_age) >= abs(beta_obs_age))

      out_list[[idx]] <- data.frame(
        year = yr,
        metric = m,
        beta_offspring = beta_obs_off,
        p_perm_off = p_perm_off,
        beta_age = beta_obs_age,
        p_perm_age = p_perm_age
      )

      idx <- idx + 1
    }
  }

  dplyr::bind_rows(out_list)
}
```

```{r}
centrality_perm_GB2 <-run_permutation_by_year(
  df = centrality_GB2,
  years = 2013:2023,
  n_perm = 1000
)

centrality_perm_GB2 <- centrality_perm_GB2 %>%
  mutate(
    metric = dplyr::case_when(
      metric == "eigen_cent" ~ "Eigenvector",
      metric == "closeness"  ~ "Closeness",
      metric == "strength"   ~ "Strength",
      TRUE ~ metric
    )) %>% 
      mutate(beta_offspring = round(beta_offspring,3)) %>% 
  mutate(beta_age = round(beta_age, 3)) 

centrality_table_wide <- centrality_perm_GB2 %>%
  tidyr::pivot_wider(
    id_cols = year,
    names_from = metric,
    values_from = c(beta_offspring, p_perm_off, beta_age, p_perm_age),
    names_glue = "{metric}_{.value}"
  ) %>%
  dplyr::arrange(year)

centrality_table_wide <- centrality_table_wide %>% 
  select(Eigenvector_beta_offspring,Eigenvector_p_perm_off,Eigenvector_beta_age,Eigenvector_p_perm_age,
          Closeness_beta_offspring, Closeness_p_perm_off, Closeness_beta_age, Closeness_p_perm_age,
         Strength_beta_offspring,Strength_p_perm_off,Strength_beta_age,Strength_p_perm_age)


centrality_table_wide

write.csv(centrality_table_wide,here("results","tables","centrality_perm_GB2.csv"))
```


### Visualization
```{r}
df14 <- centrality_GB2 %>%                   
  filter(year == 2014) %>% 
  mutate(
    d_cat = factor(pmin(total_offspring, 5),
                   levels = 0:5,
                   labels = c("0","1","2","3","4","5+"))
  )

V(g_G2014)$d_cat <- df14$d_cat[match(V(g_G2014)$name, df14$ID)]


pal <- pal <- c("#e8f1ff", "#c2d8ff", "#9cc0ff", "#729aff", "#3f6cff", "#103fbc") 

centrality_G2014<-igraph::eigen_centrality(g_G2014)
```

```{r}
# compute eigenvector centrality
eig_cent <- igraph::eigen_centrality(g_G2023, weights = E(g_G2023)$weight)$vector

# store it as a node attribute (recommended)
V(g_G2023)$eigen_cent <- eig_cent

# daughters category (cap at 5+)
V(g_G2023)$d_cat <- factor(
  pmin(V(g_G2023)$total_offspring, 5),
  levels = 0:5,
  labels = c("0", "1", "2", "3", "4", "5+")
)

# simple centrality-based network plot
ggraph(g_G2023,
       layout = "centrality",
       centrality = eig_cent) +
  geom_edge_link(
    colour = "grey80",
    width = 0.4,
    alpha = 0.6
  ) +
  geom_node_point(
    aes(fill = d_cat),
    shape = 21,
    size = 4,
    colour = "grey30"
  ) +
  scale_fill_manual(
    values = pal,
    name = "Number of daughters"
  ) +
  geom_node_text(
    aes(label = name),
    repel = TRUE,
    size = 2.5
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold")
  ) +
  ggtitle("Network 2023")

plot_network_centrality <- function(g,
                                    title = NULL,
                                    palette = c("#e8f1ff", "#c2d8ff", "#9cc0ff",
                                                "#729aff", "#3f6cff", "#103fbc")) {

  # cap daughters at 5+
  V(g)$d_cat <- factor(
    pmin(V(g)$total_offspring, 5),
    levels = 0:5,
    labels = c("0", "1", "2", "3", "4", "5+")
  )

  # layout that pulls central nodes to the middle
  lay <- ggraph::create_layout(
    g,
    layout = "stress",
    weights = E(g)$weight
  )

  ggraph(lay) +
    geom_edge_link(
      colour = "grey80",
      width = 0.3,
      alpha = 0.6
    ) +
    geom_node_point(
      aes(
        fill = d_cat,
        size = eigen_cent        # centrality → size
      ),
      shape = 21,
      colour = "grey30",
      stroke = 0.3
    ) +
    scale_fill_manual(
      values = palette,
      name = "Number of daughters"
    ) +
    scale_size_continuous(
      range = c(2.5, 7),
      guide = "none"
    ) +
    coord_equal() +
    theme_void() +
    theme(
      legend.position = "right",
      plot.title = element_text(
        size = 14,
        face = "bold",
        hjust = 0.5
      )
    ) +
    labs(title = title)
}
```


#CATEGORIC analysis --> not enough mothers with 3+ daughters
```{r}
centrality_GB2 <- centrality_GB2 %>%
  dplyr::mutate(
    offspring_cat = dplyr::case_when(
      total_offspring == 0 ~ "0",
      total_offspring == 1 ~ "1",
      total_offspring == 2 ~ "2",
      total_offspring >= 3 ~ "3"
    ),
    offspring_cat = factor(
      offspring_cat,
      levels = c("0", "1", "2","3")
    )
  )
```

#Permutations 
```{r perm function - categorical}
run_permutation_cat <- function(df,
                                years  = 2013:2023,
                                n_perm = 1000) {

  metrics <- c("eigen_cent", "closeness", "strength")
  out_list <- list()
  idx <- 1

  for (yr in years) {

    df_year <- df %>%
      dplyr::filter(year == yr) %>%
      dplyr::filter(
        !is.na(offspring_cat),
        !is.na(female_age)
      )

    # skip empty or non-identifiable years
    if (nrow(df_year) < 5 ||
        dplyr::n_distinct(df_year$offspring_cat) < 2) {
      next
    }

    for (m in metrics) {

      form <- as.formula(
        paste(m, "~ offspring_cat + female_age + unknown_age")
      )

      # observed model
      mod_obs <- lm(form, data = df_year)

      off_levels <- grep("^offspring_cat", names(coef(mod_obs)), value = TRUE)

      beta_obs_off <- coef(mod_obs)[off_levels]
      beta_obs_age <- as.numeric(coef(mod_obs)["female_age"])

      # permutation storage
      beta_perm_off <- matrix(
        NA,
        nrow = n_perm,
        ncol = length(off_levels),
        dimnames = list(NULL, off_levels)
      )

      beta_perm_age <- numeric(n_perm)

      for (i in seq_len(n_perm)) {

        perm_df <- df_year %>%
          dplyr::mutate(
            offspring_cat = sample(offspring_cat)
          )

        mod_p <- lm(form, data = perm_df)

        beta_perm_off[i, ] <- coef(mod_p)[off_levels]
        beta_perm_age[i]   <- as.numeric(coef(mod_p)["female_age"])
      }

      # permutation p-values (one per contrast)
      p_perm_off <- sapply(seq_along(off_levels), function(j) {
        mean(abs(beta_perm_off[, j]) >= abs(beta_obs_off[j]))
      })

      # optional diagnostic p-value for age
      p_perm_age <- mean(abs(beta_perm_age) >= abs(beta_obs_age))

      # store results: ONE ROW PER CONTRAST
      for (j in seq_along(off_levels)) {

        out_list[[idx]] <- data.frame(
          year        = yr,
          Metric      = m,
          Contrast    = off_levels[j],
          beta_off    = beta_obs_off[j],
          p_perm_off  = p_perm_off[j],
          beta_age    = beta_obs_age,
          p_perm_age  = p_perm_age,
          stringsAsFactors = FALSE
        )

        idx <- idx + 1
      }
    }
  }

  dplyr::bind_rows(out_list)
}
```


```{r perm function for all metrics}
centrality_cat_perm_GB2 <-run_permutation_cat(
  df = centrality_GB2,
  years = 2013:2023,
  n_perm = 1000
)
centrality_cat_perm_GB2

```

```{r}
centrality_cat_perm_GB2 <- centrality_cat_perm_GB2 %>%
  mutate(
    Metric = dplyr::case_when(
      Metric == "eigen_cent" ~ "Eigenvector",
      Metric == "closeness"  ~ "Closeness",
      Metric == "strength"   ~ "Strength",
      TRUE ~ Metric
    )) %>% 
      mutate(beta_off = round(beta_off,3)) %>% 
  mutate(beta_age = round(beta_age, 3)) 

centrality_table_wide_cat_GB2 <- centrality_cat_perm_GB2 %>%
  tidyr::pivot_wider(
    id_cols = c(year, Contrast),
    names_from = Metric,
    values_from = c(beta_off, p_perm_off),
    names_glue = "{Metric}_{.value}"
  ) %>%
  dplyr::arrange(year, Contrast)

centrality_table_wide_cat_GB2 <- centrality_table_wide_cat_GB2 %>%
  select(
    year,
    Contrast,

    Eigenvector_beta_off,
    Eigenvector_p_perm_off,

    Closeness_beta_off,
    Closeness_p_perm_off,

    Strength_beta_off,
    Strength_p_perm_off
  )


centrality_table_wide_cat_GB2
write.csv(centrality_table_wide_cat_GB2,here("results","tables","centrality_perm_GB2_categorical.csv"))
```

#////////////
#ALL years analysis 
######numerical
```{r}
run_permutation_pooled_numeric_GB2 <- function(df,
                                               responses = c("strength_z",
                                                             "eigen_z",
                                                             "closeness_z"),
                                               n_perm = 1000) {

  results_list <- list()
  null_list <- list()

  for (resp in responses) {

    form <- as.formula(
      paste(resp,
            "~ total_offspring + female_age + colony_size + (1 | ID)")
    )

    # observed model
    mod_obs <- lmer(form, data = df, REML = FALSE)

    beta_obs <- fixef(mod_obs)["total_offspring"]

    beta_perm <- numeric(n_perm)

    for (i in seq_len(n_perm)) {

      perm_df <- df %>%
        group_by(year) %>%
        mutate(total_offspring = sample(total_offspring)) %>%
        ungroup()

      mod_p <- lmer(form, data = perm_df, REML = FALSE)

      beta_perm[i] <- fixef(mod_p)["total_offspring"]
    }

    p_perm <- mean(abs(beta_perm) >= abs(beta_obs))
    
        # ---- store short summary ----
    results_list[[resp]] <- data.frame(
      metric = resp,
      beta_observed = beta_obs,
      beta_perm = mean(abs(beta_perm)),
      p_perm = p_perm
    )

    # ---- store full null distribution ----
    null_list[[resp]] <- beta_perm
  }

  return(list(
    summary = bind_rows(results_list),
    null_dist = null_list
  ))
}
```

run 
```{r}
centrality_perm_pooled_numeric_GB2 <- run_permutation_pooled_numeric_GB2(
  df = centrality_GB2,
  responses = c("strength_z", "eigen_z", "closeness_z"),
  n_perm = 1000
)
```

Results
```{r}
centrality_perm_pooled_numeric_GB2$null_dist

centrality_result_pooled_numeric_GB2 <- centrality_perm_pooled_numeric_GB2$summary %>%
  mutate(metric = case_when(
    metric == "strength_z"  ~ "Strength",
    metric == "eigen_z"     ~ "Eigenvector",
    metric == "closeness_z" ~ "Closeness"
  ))

centrality_result_pooled_numeric_GB2
p.adjust(centrality_result_pooled_numeric_GB2$p_perm, method = "fdr")
```

### hist plot
histogram of the distribution of the permuted coefficients, with a vertical line for the observed coefficient, for each metric
```{r}
null_df <- bind_rows(
  lapply(names(centrality_perm_pooled_numeric_GB2$null_dist), function(m) {
    data.frame(
      metric = m,
      beta_perm = centrality_perm_pooled_numeric_GB2$null_dist[[m]]
    )
  })
)

null_df <- null_df %>%
  mutate(metric = case_when(
    metric == "strength_z"  ~ "Strength",
    metric == "eigen_z"     ~ "Eigenvector",
    metric == "closeness_z" ~ "Closeness"
  ))

obs_df <- centrality_perm_pooled_numeric_GB2$summary %>%
  mutate(metric = case_when(
    metric == "strength_z"  ~ "Strength",
    metric == "eigen_z"     ~ "Eigenvector",
    metric == "closeness_z" ~ "Closeness"
  ))
```
plot
```{r}
hist <- ggplot(null_df, aes(x = beta_perm)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "grey80",
                 color = "black",
                 linewidth = 0.3) +
  geom_vline(data = obs_df,
             aes(xintercept = beta_observed),
             color = "red",
             linewidth = 1.2) +
  geom_text(data = obs_df,
            aes(x = -0.1,
                y = 8.5,
                label = paste0("p = ", round(p_perm, 3))),
            color = "red",
            size = 4) +
  facet_wrap(~ metric, scales = "free") +
  labs(x = "Permuted effect",
       y = "Density") +
  theme_classic(base_size = 13) +
  theme(
    strip.background = element_blank(),      # removes beige box
    strip.text = element_text(face = "bold", size = 13),
    panel.spacing = unit(1.2, "lines"),
    plot.title = element_text(face = "bold", size = 14)
  )
hist

ggsave(here("results","figures","Q3_perm_distribution.png"), plot = hist,
  width = 8,
  height = 3
)
```

### perm function for categorical variable - al
```{r}
run_permutation_pooled_multi <- function(df,
                                         responses = c("strength_z",
                                                       "eigen_z",
                                                       "closeness_z"),
                                         n_perm = 1000) {

  results_list <- list()

  for (resp in responses) {

    form <- as.formula(
      paste(resp,
            "~ offspring_cat + female_age + colony_size + (1 | ID)")
    )

    # observed model
    mod_obs <- lmer(form, data = df, REML = FALSE)

    beta_obs <- fixef(mod_obs)[grep("offspring_cat",
                                    names(fixef(mod_obs)))]

    beta_perm <- matrix(NA, n_perm, length(beta_obs))
    colnames(beta_perm) <- names(beta_obs)

    for (i in seq_len(n_perm)) {

      perm_df <- df %>%
        group_by(year) %>%
        mutate(offspring_cat = sample(offspring_cat)) %>%
        ungroup()

      mod_p <- lmer(form, data = perm_df, REML = FALSE)

      beta_perm[i, ] <- fixef(mod_p)[names(beta_obs)]
    }

    p_perm <- colMeans(abs(beta_perm) >= abs(beta_obs))

    res_df <- data.frame(
      metric = resp,
      coefficient = names(beta_obs),
      beta_observed = beta_obs,
      p_perm = p_perm
    )

    results_list[[resp]] <- res_df
  }

  bind_rows(results_list)
}
```
 run it 
```{r}
centrality_perm_pooled_GB2 <- run_permutation_pooled_multi(
  df = centrality_GB2,
  responses = c("strength_z", "eigen_z", "closeness_z"),
  n_perm = 100
)
```

```{r}
centrality_perm_pooled_GB2 <- centrality_perm_pooled_GB2 %>%
  mutate(metric = case_when(
    metric == "strength_z"  ~ "Strength",
    metric == "eigen_z"     ~ "Eigenvector",
    metric == "closeness_z" ~ "Closeness"
  ))

centrality_perm_pooled_GB2
```


#----------- UA
Load
```{r }
edges_UA<-read.csv(here("data","saved_matrices","UA","full","full_edges_UA.csv"),header=TRUE, row.names=NULL,sep=",",check.names = F)
mother_daughter_UA<-read.csv(here("data","mother_daughter_UA.csv"),header=TRUE, row.names=NULL,sep=",",check.names = F)

edges_UA <- edges_UA %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
```

```{r load the networls, include= FALSE}
GBI_U2013 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2013<-get_network(GBI_U2013,data_format = "GBI",association_index = "SRI")
GBI_U2014 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2014<-get_network(GBI_U2014,data_format = "GBI",association_index = "SRI")
GBI_U2015 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2015<-get_network(GBI_U2015,data_format = "GBI",association_index = "SRI")
GBI_U2016 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2016<-get_network(GBI_U2016,data_format = "GBI",association_index = "SRI")
GBI_U2017 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2017<-get_network(GBI_U2017,data_format = "GBI",association_index = "SRI")
GBI_U2018 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2018<-get_network(GBI_U2018,data_format = "GBI",association_index = "SRI")
GBI_U2019 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2019<-get_network(GBI_U2019,data_format = "GBI",association_index = "SRI")
GBI_U2020 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2020<-get_network(GBI_U2020,data_format = "GBI",association_index = "SRI")
GBI_U2021 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2021<-get_network(GBI_U2021,data_format = "GBI",association_index = "SRI")
GBI_U2022 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2022<-get_network(GBI_U2022,data_format = "GBI",association_index = "SRI")
GBI_U2023 <- read.csv(here("data","saved_matrices","UA","full","fullGBI_U2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);
net_U2023<-get_network(GBI_U2023,data_format = "GBI",association_index = "SRI")
```

```{r}
g_U2015<- graph_from_adjacency_matrix(net_U2015, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2014<- graph_from_adjacency_matrix(net_U2014, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2015<- graph_from_adjacency_matrix(net_U2015, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2016<- graph_from_adjacency_matrix(net_U2016, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2017<- graph_from_adjacency_matrix(net_U2017, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2018<- graph_from_adjacency_matrix(net_U2018, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2019<- graph_from_adjacency_matrix(net_U2019, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2020<- graph_from_adjacency_matrix(net_U2020, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2021<- graph_from_adjacency_matrix(net_U2021, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2022<- graph_from_adjacency_matrix(net_U2022, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2023<- graph_from_adjacency_matrix(net_U2023, mode = "undirected", weighted = TRUE, diag = FALSE)
```

#offspring table
```{r}
edges_MD_UA <- edges_UA %>%
  filter(pairID %in% mother_daughter_UA$pairID)
```

```{r}
offspring_table_UA <- edges_MD_UA %>%
  select(from, to, year, pairID) %>%
   pivot_longer(cols = c("from", "to"), values_to = "ID") %>%
   select(ID, year, pairID) %>%
   distinct()  # no duplicate rows

#then we can add a new column "is mother", to be sure that the individual is the mother in the pair (look in mother-daughter df)
offspring_table_UA <- offspring_table_UA %>%
  left_join(
    mother_daughter_UA %>%
      select(pairID, rfid_mom) , by = "pairID" )%>%
      mutate(is_mother = ifelse(ID == rfid_mom, "mother", "daughter")
  )
```

Then, we can measure the number of offsprings for each mother
```{r}
offspring_counts_UA <- offspring_table_UA %>%
  filter(is_mother == "mother") %>% 
  group_by(ID, year) %>%
  summarise(total_offspring = n_distinct(pairID, na.rm = TRUE), .groups = "drop")

offspring_table_UA <- left_join(offspring_table_UA, offspring_counts_UA, by = c("ID", "year"))

offspring_table_UA$total_offspring[is.na(offspring_table_UA$total_offspring)] <- 0

offspring_table_UA <- offspring_table_UA %>%
  mutate(across(c(ID, pairID), as.character)) 
#write.csv(offspring_table, "variables_UA/offspring_table.csv", row.names = FALSE,quote = TRUE)
```

ADD mother AGE as a control !!!
Then, create a df with ALL females age (add the daughters age aswell)
```{r}
female_birth_year_UA <- bind_rows(
  mother_df %>%
    filter(Colony == "UA") %>%
    transmute(
      ID = as.character(TAG_mom),
      birth_year = as.integer(birth_mom)
    ),
  mother_daughter_UA %>%
    transmute(
      ID = as.character(rfid_off),
      birth_year = as.integer(year)
    )
) %>%
  distinct(ID, birth_year)
```

join to get mother age at each year (Year - birth year),
```{r}
offspring_table_UA <- offspring_table_UA %>%
  mutate(ID = as.character(ID)) %>%
  left_join(
    mother_df %>%
      filter(Colony == "UA") %>%
      distinct(TAG_mom, birth_mom) %>%
      mutate(
        ID        = as.character(TAG_mom),
        birth_mom = as.integer(birth_mom)
      ) %>%
      select(ID, birth_mom),
    by = "ID"
  )

offspring_table_UA <- offspring_table_UA %>%
  mutate(unkown_age = ifelse(is_mother == "mother" & is.na(birth_mom), TRUE, FALSE)  )
```

Now we can join the info on the offsprings on the edges list (total)
--> first need to creat ethe node list and not the edge list 
```{r}
nodes_UA <- edges_UA %>%
  select(year, from, to) %>%
  pivot_longer(cols = c(from, to), names_to = "role",
    values_to = "ID"
  ) %>%
  select(year, ID) %>%
  distinct()

nodes_UA <- nodes_UA %>%
  left_join(
    offspring_table_UA %>%
      select(
        ID, year,
        is_mother,
        total_offspring,
        unkown_age
      ),
    by = c("ID", "year")
  ) %>%
  mutate(
    is_mother = if_else(is.na(is_mother), "non_mother", is_mother),
    total_offspring = replace_na(total_offspring, 0),
    unkown_age = replace_na(unkown_age, 0L)
  )

nodes_UA <- nodes_UA %>%
  left_join(female_birth_year_UA, by = "ID") %>%
  mutate(
    female_age = year - birth_year
  )

nodes_UA <- nodes_UA %>%
  distinct(
    year, ID,
    .keep_all = TRUE
  )

nodes_UA <- nodes_UA %>%
  mutate(unknown_age = ifelse(is.na(birth_year), TRUE, FALSE)) %>% 
  select(-unkown_age)
```


# Measure all three metrics and put in the same table
#metrics
```{r}
network_list_UA <- list(
  "2013" = net_U2013, "2014" = net_U2014, "2015" = net_U2015,
  "2016" = net_U2016, "2017" = net_U2017, "2018" = net_U2018,
  "2019" = net_U2019, "2020" = net_U2020, "2021" = net_U2021,
  "2022" = net_U2022, "2023" = net_U2023
)
# each year, build the graph once and compute all three metrics
centrality_list_UA <- lapply(names(network_list_UA), function(year_char) {

  year_int <- as.integer(year_char)
  
  # Build the undirected, weighted igraph for that year
  mat <- network_list_UA[[year_char]]
  g    <- graph_from_adjacency_matrix(
    mat,
    mode     = "undirected",
    weighted = TRUE,
    diag     = FALSE
  )
  
  # closeness 
  E(g)$dist <- 1 / E(g)$weight
  
  clos_vals <- igraph::closeness(
  g,
  vids = V(g),
  weights = E(g)$dist,
  normalized = TRUE
)
  
  #eigenvector centrality
  eig_vals <- igraph::eigen_centrality(
    g,
    directed = FALSE,
    weights  = E(g)$weight
  )$vector
  
  #Strength (weighted degre)
  str_vals <- igraph::strength(
    g,
    v       = V(g),
    mode    = "all",
    weights = E(g)$weight
  )
  
  # create df
  data.frame(
    ID            = names(clos_vals),
    year          = year_int,
    closeness     = as.numeric(clos_vals),
    eigen_cent    = as.numeric(eig_vals),
    strength      = as.numeric(str_vals),
    stringsAsFactors = FALSE
  )
})

centrality_metrics_UA <- bind_rows(centrality_list_UA)

#merge
centrality_UA <- nodes_UA %>%
  mutate(
    year = as.integer(year),
    ID   = as.character(ID)
  ) %>%

  left_join(
    centrality_metrics_UA %>%
      mutate(
        year = as.integer(year),
        ID   = as.character(ID)
      ),
    by = c("ID", "year")
  )

#add colon ysize
colony_size_UA <- read.csv(here("data","saved_matrices","UA","colony_size_UA.csv"),
  header      = TRUE,
  sep         = ",",
  check.names = FALSE
) %>%
  mutate(year = as.integer(year))

centrality_UA <- centrality_UA %>%
  left_join(colony_size_UA, by = "year")

centrality_UA <- centrality_UA %>%
  group_by(year) %>%
  mutate(
    strength_z   = as.numeric(scale(strength)),
    eigen_z      = as.numeric(scale(eigen_cent)),
    closeness_z  = as.numeric(scale(closeness))
  ) %>%
  ungroup()

```

# *Analysis* 
### first to check

```{r}
#check only 1 year
centrality_UA_2015<- centrality_UA %>%
  filter(year == 2015)

#model to investigate eigenvector centrality with number of offsrping, while controling for mother_age 
m <- lm(eigen_cent ~ total_offspring + female_age, data = centrality_UA_2015)
summary(m)

coef(m)[c("total_offspring","female_age")]
#check final model
simulationOutput_eigen_UA <- simulateResiduals(fittedModel = m, n = 1000)
plot(simulationOutput_eigen_UA)
testUniformity(simulationOutput_eigen_UA)   # KS test for uniformity
testDispersion(simulationOutput_eigen_UA)     # Test for over/under dispersion
testZeroInflation(simulationOutput_eigen_UA) 


par(mfrow = c(1,2))
plot(m, which = 1, main = "Residuals")
plot(m, which = 2, main = "QQ")
```
Check all years combined
```{r}
m_allyears<- lm(eigen_z ~ total_offspring + female_age + colony_size, data = centrality_UA)
summary(m_allyears)

simulationOutput_eigen_allyears <- simulateResiduals(fittedModel = m_allyears, n = 1000)
plot(simulationOutput_eigen_allyears)
```


#permutation
```{r}
centrality_perm_UA <-run_permutation_by_year(
  df = centrality_UA,
  years = 2013:2023,
  n_perm = 1000
)
centrality_perm_UA
```

```{r}
centrality_perm_UA <- centrality_perm_UA %>%
  mutate(
    metric = dplyr::case_when(
      metric == "eigen_cent" ~ "Eigenvector",
      metric == "closeness"  ~ "Closeness",
      metric == "strength"   ~ "Strength",
      TRUE ~ metric
    )) %>% 
      mutate(beta_offspring = round(beta_offspring,3)) %>% 
  mutate(beta_age = round(beta_age, 3)) 

centrality_table_wide <- centrality_perm_UA %>%
  tidyr::pivot_wider(
    id_cols = year,
    names_from = metric,
    values_from = c(beta_offspring, p_perm_off, beta_age, p_perm_age),
    names_glue = "{metric}_{.value}"
  ) %>%
  dplyr::arrange(year)

centrality_table_wide <- centrality_table_wide %>% 
  select(year, Eigenvector_beta_offspring,Eigenvector_p_perm_off,Eigenvector_beta_age,Eigenvector_p_perm_age,
          Closeness_beta_offspring, Closeness_p_perm_off, Closeness_beta_age, Closeness_p_perm_age,
         Strength_beta_offspring,Strength_p_perm_off,Strength_beta_age,Strength_p_perm_age)


centrality_table_wide

write.csv(centrality_table_wide,here("results","tables","centrality_perm_UA.csv"))
```
#CATEGORIC analysis --> not enough mothers with 3+ daughters
```{r}
centrality_UA <- centrality_UA %>%
  dplyr::mutate(
    offspring_cat = dplyr::case_when(
      total_offspring == 0 ~ "0",
      total_offspring == 1 ~ "1",
      total_offspring == 2 ~ "2",
      total_offspring >= 3 ~ "3"
    ),
    offspring_cat = factor(
      offspring_cat,
      levels = c("0", "1", "2","3")
    )
  )
```

```{r perm function for all metrics}
centrality_cat_perm_UA <-run_permutation_cat(
  df = centrality_UA,
  years = 2013:2023,
  n_perm = 1000
)
centrality_cat_perm_UA
```
```{r}
centrality_cat_perm_UA <- centrality_cat_perm_UA %>%
  mutate(
    Metric = dplyr::case_when(
      Metric == "eigen_cent" ~ "Eigenvector",
      Metric == "closeness"  ~ "Closeness",
      Metric == "strength"   ~ "Strength",
      TRUE ~ Metric
    )) %>% 
      mutate(beta_off = round(beta_off,3)) %>% 
  mutate(beta_age = round(beta_age, 3)) 

centrality_table_wide_cat_UA <- centrality_cat_perm_UA %>%
  tidyr::pivot_wider(
    id_cols = c(year, Contrast),
    names_from = Metric,
    values_from = c(beta_off, p_perm_off),
    names_glue = "{Metric}_{.value}"
  ) %>%
  dplyr::arrange(year, Contrast)

centrality_table_wide_cat_UA <- centrality_table_wide_cat_UA %>%
  select(
    year,
    Contrast,

    Eigenvector_beta_off,
    Eigenvector_p_perm_off,

    Closeness_beta_off,
    Closeness_p_perm_off,

    Strength_beta_off,
    Strength_p_perm_off
  )


centrality_table_wide_cat_UA
write.csv(centrality_table_wide_cat_UA,here("results","tables","centrality_perm_UA_categorical.csv"))
```
#////////////
#ALL years analysis 
######numerical
```{r}
run_permutation_pooled_numeric_UA <- function(df,
                                              responses = c("strength_z",
                                                            "eigen_z",
                                                            "closeness_z"),
                                              n_perm = 1000) {

  summary_list <- list()
  null_list <- list()

  for (resp in responses) {

    form <- as.formula(
      paste(resp,
            "~ total_offspring + female_age + colony_size")
    )

    # ---- observed model ----
    mod_obs <- lm(form, data = df)

    beta_obs <- coef(mod_obs)["total_offspring"]

    beta_perm <- numeric(n_perm)

    # ---- permutation loop ----
    for (i in seq_len(n_perm)) {

      perm_df <- df %>%
        group_by(year) %>%
        mutate(total_offspring = sample(total_offspring)) %>%
        ungroup()

      mod_p <- lm(form, data = perm_df)

      beta_perm[i] <- coef(mod_p)["total_offspring"]
    }

    p_perm <- mean(abs(beta_perm) >= abs(beta_obs))

    summary_list[[resp]] <- data.frame(
      metric = resp,
      beta_observed = beta_obs,
      p_perm = p_perm
    )

    null_list[[resp]] <- beta_perm
  }

  return(list(
    summary = bind_rows(summary_list),
    null_dist = null_list
  ))
}

```

run 
```{r}
centrality_perm_pooled_numeric_UA <- run_permutation_pooled_numeric_UA(
  df = centrality_UA,
  responses = c("strength_z", "eigen_z", "closeness_z"),
  n_perm = 1000
)
```

Results
```{r}
centrality_perm_pooled_numeric_UA$null_dist

centrality_result_pooled_numeric_UA <- centrality_perm_pooled_numeric_UA$summary %>%
  mutate(metric = case_when(
    metric == "strength_z"  ~ "Strength",
    metric == "eigen_z"     ~ "Eigenvector",
    metric == "closeness_z" ~ "Closeness"
  ))

p.adjust(centrality_result_pooled_numeric_UA$p_perm, method = "fdr")
centrality_result_pooled_numeric_UA
```

### hist plot
histogram of the distribution of the permuted coefficients, with a vertical line for the observed coefficient, for each metric
```{r}
null_df_UA <- bind_rows(
  lapply(names(centrality_perm_pooled_numeric_UA$null_dist), function(m) {
    data.frame(
      metric = m,
      beta_perm = centrality_perm_pooled_numeric_UA$null_dist[[m]]
    )
  })
)

null_df_UA <- null_df_UA %>%
  mutate(metric = case_when(
    metric == "strength_z"  ~ "Strength",
    metric == "eigen_z"     ~ "Eigenvector",
    metric == "closeness_z" ~ "Closeness"
  ))

obs_df_UA <- centrality_perm_pooled_numeric_UA$summary %>%
  mutate(metric = case_when(
    metric == "strength_z"  ~ "Strength",
    metric == "eigen_z"     ~ "Eigenvector",
    metric == "closeness_z" ~ "Closeness"
  ))
```
plot
```{r}
hist <- ggplot(null_df_UA, aes(x = beta_perm)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "grey80",
                 color = "black",
                 linewidth = 0.3) +
  geom_vline(data = obs_df_UA,
             aes(xintercept = beta_observed),
             color = "red",
             linewidth = 1.2) +
  geom_text(data = obs_df_UA,
            aes(x = -0.1,
                y = 5,
                label = paste0("p = ", round(p_perm, 3))),
            color = "red",
            size = 4) +
  facet_wrap(~ metric, scales = "free") +
  labs(x = "Permuted effect",
       y = "Density") +
  theme_classic(base_size = 13) +
  theme(
    strip.background = element_blank(),      # removes beige box
    strip.text = element_text(face = "bold", size = 13),
    panel.spacing = unit(1.2, "lines"),
    plot.title = element_text(face = "bold", size = 14)
  )
hist

ggsave(here("results","figures","Q3_perm_distribution_UA.png"), plot = hist,
  width = 8,
  height = 3
)
```

```{r}
year_slopes <- centrality_GB2 %>%
  group_by(year) %>%
  do(tidy(lm(strength_z ~ total_offspring + female_age, data = .))) %>%
  filter(term == "total_offspring")

table(year_slopes$estimate > 0)
```

# /////////////
#PLOT result

First the panels plot
```{r}
df_beta_GB2 <- centrality_perm_GB2 %>%
  filter(metric == "Strength") %>%
  mutate(
    signif = p_perm_off < 0.05,
    colony = "GB2"
  )

df_beta_UA <- centrality_perm_UA %>%
  filter(metric == "Strength") %>%
  mutate(
    signif = p_perm_off < 0.05,
    colony = "UA"
  )

df_beta_all <- bind_rows(df_beta_GB2, df_beta_UA)

df_beta_all <- df_beta_all %>%
  mutate(year = factor(year))
```
plot the two together
```{r}
p_beta <- ggplot(df_beta_all,
                 aes(x = year,
                     y = beta_offspring)) +

  geom_hline(yintercept = 0,
             colour = "grey75",
             linewidth = 0.4) +

  geom_point(aes(shape = signif),
             size = 2.5,
             colour = "black") +
  scale_shape_manual(
  values = c("TRUE" = 16, "FALSE" = 1),
  labels = c("FALSE" = "n.s.",
             "TRUE"  = "p < 0.05"),
  name = NULL
) +

  facet_wrap(~ colony, ncol = 1) +

  labs(
    x = "Year",
    y = expression("observed coefficient")
  ) +

  theme_classic(base_size = 13) +

  theme(
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45,
                             hjust = 1,
                             vjust = 1),
    strip.background = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )
p_beta

```
And then the graphs
GB2

```{r}
# select graph
g_G2021
vertex_attr_names(g_G2021) #only name 

#take node attributes
nodes_2021 <- centrality_GB2 %>%
  dplyr::filter(year == 2021) %>%
  dplyr::select(
    ID,
    strength,
    offspring_cat,
    female_age
  )

#add node attibutes on the igraph object 
V(g_G2021)$strength <- nodes_2021$strength[
  match(V(g_G2021)$name, nodes_2021$ID)
]

V(g_G2021)$offspring_cat <- nodes_2021$offspring_cat[
  match(V(g_G2021)$name, nodes_2021$ID)
]

V(g_G2021)$female_age <- nodes_2021$female_age[
  match(V(g_G2021)$name, nodes_2021$ID)
]

# colour palette (reuse yours if you want)
pal <- c(
  "0" = "#e8f1ff",
  "1" = "#9cc0ff",
  "2" = "#3f6cff",
  "3" = "#103fbc"
)

p_net_GB2 <- ggraph(
  g_G2021,
  layout = "centrality",
  centrality = V(g_G2021)$strength
) +
  geom_edge_link(
    colour = "grey80",
    width = 0.4,
    alpha = 0.6
  ) +
   geom_node_point(
  aes(fill = offspring_cat),
  shape = 21,
  size = 4,
  stroke = 0.15
  ) +
  scale_fill_manual(
    values = pal,
    name = "Number of daughters"
  ) +
  scale_size_continuous(
    range = c(3, 7),
    guide = "none"
  ) +
  theme_void() +
  ggtitle("GB2 - 2021")+ 
theme(
  plot.title = element_text(face = "bold", size = 12),
  legend.position = "bottom"
)

p_net_GB2
```

UA
```{r}
# select graph
g_U2021

#take node attributes
nodes_2021_UA <- centrality_UA%>%
  dplyr::filter(year == 2021) %>%
  dplyr::select(
    ID,
    strength,
    offspring_cat,
    female_age
  )

#add node attibutes on the igraph object 
V(g_U2021)$strength <- nodes_2021_UA$strength[
  match(V(g_U2021)$name, nodes_2021_UA$ID)
]

V(g_U2021)$offspring_cat <- nodes_2021_UA$offspring_cat[
  match(V(g_U2021)$name, nodes_2021_UA$ID)
]

V(g_U2021)$female_age <- nodes_2021_UA$female_age[
  match(V(g_U2021)$name, nodes_2021_UA$ID)
]

# colour palette (reuse yours if you want)
pal <- c(
  "0" = "#e8f1ff",
  "1" = "#9cc0ff",
  "2" = "#3f6cff",
  "3" = "#103fbc"
)

p_net_UA <- ggraph(
  g_U2021,
  layout = "centrality",
  centrality = V(g_U2021)$strength
) +
  geom_edge_link(
    colour = "grey80",
    width = 0.4,
    alpha = 0.6
  ) +
  geom_node_point(
  aes(fill = offspring_cat),
  shape = 21,
  size = 4,
  stroke = 0.15
  ) +
  scale_fill_manual(
    values = pal,
    name = "Number of daughters"
  ) +
  scale_size_continuous(
    range = c(3, 7),
    guide = "none"
  ) +
  theme_void() +
  ggtitle("UA - 2021")+ 
theme(
  plot.title = element_text(face = "bold", size = 12),
  legend.position = "bottom"
)

p_net_UA
```

Stack networks
```{r}
#remove legend
p_net_GB2 <- p_net_GB2 +
  theme(legend.position = "none")

#get more place for UA
networks_stacked <- p_net_GB2 /
                    p_net_UA +
  plot_layout(heights = c(1, 1.4))
```


```{r}
#combine with the year panel, networks takes more place
main_panel <- networks_stacked | p_beta 
main_panel

ggsave(here("results","figures","Q3_yearly.png"), plot = main_panel,
  width = 10,
  height = 5
)
```



# Predicted plot (results)
```{r}
library(ggeffects)
#STRENGTH
mod_GB2_strength  <- lmer(strength_z ~ total_offspring + female_age + colony_size + (1 | ID),
                data = centrality_GB2)

mod_UA_strength  <- lm(strength_z ~ total_offspring + female_age + colony_size, 
                data = centrality_UA)

pred_GB2_strength <- ggpredict(mod_GB2_strength, terms = "total_offspring")
pred_UA_strength  <- ggpredict(mod_UA_strength,  terms = "total_offspring")

#EIGENVECTOR 
mod_GB2_eigenvector  <- lmer(eigen_z ~ total_offspring + female_age + colony_size + (1 | ID),
                data = centrality_GB2)

mod_UA_eigenvector  <- lm(eigen_z ~ total_offspring + female_age + colony_size, 
                data = centrality_UA)

pred_GB2_eigenvector <- ggpredict(mod_GB2_eigenvector, terms = "total_offspring")
pred_UA_eigenvector  <- ggpredict(mod_UA_eigenvector,  terms = "total_offspring")

#CLOSENESS
mod_GB2_closeness  <- lmer(closeness_z ~ total_offspring + female_age + colony_size + (1 | ID),
                data = centrality_GB2)

mod_UA_closeness  <- lm(closeness_z ~ total_offspring + female_age + colony_size, 
                data = centrality_UA)

pred_GB2_closeness <- ggpredict(mod_GB2_closeness, terms = "total_offspring")
pred_UA_closeness  <- ggpredict(mod_UA_closeness,  terms = "total_offspring")
```
Add metrics
```{r}
pred_GB2_strength$metric <- "Strength"
pred_UA_strength$metric <- "Strength"

pred_GB2_eigenvector$metric <- "Eigenvector"
pred_UA_eigenvector$metric <- "Eigenvector"

pred_GB2_closeness$metric <- "Closeness"
pred_UA_closeness$metric <- "Closeness"

pred_GB2 <- rbind(pred_GB2_strength, pred_GB2_eigenvector, pred_GB2_closeness)
pred_UA  <- rbind(pred_UA_strength, pred_UA_eigenvector, pred_UA_closeness)

#and add colony size
pred_GB2$colony <- "GB2"
pred_UA$colony  <- "UA"

pred_all <- rbind(pred_GB2, pred_UA)
```

#plot all
```{r}
plot_predicted <- ggplot(pred_all, aes(x = x, y = predicted, color = metric, fill = metric)) +

  geom_line(size = 1.2) +
  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.15,
              colour = NA) +
  facet_wrap(~ colony) +

  labs(x = "Number of daughters present",
       y = "Standardized centrality",
       color = "Metric",
       fill  = "Metric") +
  theme_classic(base_size = 14)
```

+ add the data points
```{r}
#combine the two datasets (centrality)
centrality_GB2$colony <- "GB2"
centrality_UA$colony  <- "UA"
data_all <- rbind(centrality_GB2, centrality_UA)

plot_predicted <- ggplot(pred_all, aes(x = x, y = predicted, color = metric, fill = metric)) +

  geom_line(size = 1.2) +
  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.15,
              colour = NA) +
  facet_wrap(~ colony) +

  labs(x = "Number of daughters present",
       y = "Standardized centrality",
       color = "Metric",
       fill  = "Metric") +
  theme_classic(base_size = 14) +
geom_point(data = data_all,
           aes(x = total_offspring, y = strength_z),
           alpha = 0.25,
           inherit.aes = FALSE)

plot_predicted
```
#plot only strength
```{r}
plot_GB2 <- ggplot(pred_GB2_strength, aes(x = x, y = predicted)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.2) +
  labs(x = "Number of daughters present",
       y = "Predicted centrality",
       title = "GB2 colony") +
  theme_classic(base_size = 14)

plot_GB2 <- plot_GB2 + geom_point(data = centrality_GB2,
             aes(x = total_offspring, y = strength_z),
             alpha = 0.2)
#UA
plot_UA <- ggplot(pred_UA_strength, aes(x = x, y = predicted)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.2) +
  labs(x = "Number of daughters present",
       y = "Predicted centrality",
       title = "UA colony") +
  theme_classic(base_size = 14)

plot_UA <- plot_UA + geom_point(data = centrality_UA,
             aes(x = total_offspring, y = strength_z),
             alpha = 0.2)

final_predicted <- (plot_GB2 | plot_UA)

ggsave(here("results","figures","Q3_LM_predicted.png"), plot = final_predicted,
  width = 8,
  height = 4
)
```
