---
title: "Data Prep"
author: "Flore Michel"
date: "2026-01-23"
output: html_document
---
```{r}
library(dplyr)
library(here)
library(igraph)
library(asnipe)
library(sna)
library(assocInd)
library(dplyr)
#library(lme4)
library(ggplot2)
library(tidyverse)
library(viridis)
library(patchwork)
library(igraph)
library(ggraph)
library(data.table)
```

Load both mother-daughter datasets for GB2 and UA
```{r}
mother_daughter_GB2 <- read.csv(here("data","mother_daughter_GB2.csv"))
mother_daughter_UA <- read.csv(here("data","mother_daughter_UA.csv"))
```

#Load years and create GBI and edges lists 
# GB2
##### 2013
```{r}
GB2_2013<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2013_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2013<-GB2_2013[c(0:30)]
GB2_2013 <- GB2_2013[, -4]
GB2_2013 <- GB2_2013[, -3]
GB2_2013 <- GB2_2013[, -2]
colnames(GB2_2013)[colnames(GB2_2013) == "number"] <- "date"
```

Create the GBI matrix 
```{r}
data_G2013<- GB2_2013 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2013$daygroup<-as.factor(paste(data_G2013$date, data_G2013$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2013<-data.table(ID=data_G2013$ID, group=data_G2013$daygroup, day=c(data_G2013$date))
         
GBI_G2013<- get_group_by_individual(ind_G2013, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2013 <- as.matrix(GBI_G2013)

write.csv(GBI_G2013, here("data","saved_matrices","GB2","GBI_G2013.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2013 <- get_network(GBI_G2013, data_format = "GBI", association_index = "SRI")
write.csv(net_G2013, here("data","saved_matrices","GB2","net_G2013.csv"), row.names = TRUE)
```
Create the edge list
```{r}
#first create the igraph object 
g_G2013 <- graph_from_adjacency_matrix(net_G2013,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2013 <- igraph::as_data_frame(g_G2013, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2013 <- edge_G2013 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2013 <- edge_G2013 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2013],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2013),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2013$pairID
  }
)
write.csv(edge_G2013, here("data","saved_matrices","GB2","edge_G2013.csv"), row.names = FALSE)
```
##### 2014
```{r}
GB2_2014<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2014_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2014<-GB2_2014[c(0:40)] #filter columns
GB2_2014<-GB2_2014[1:51,]
GB2_2014 <- GB2_2014[, -4]
GB2_2014 <- GB2_2014[, -3]
GB2_2014 <- GB2_2014[, -2]
colnames(GB2_2014)[colnames(GB2_2014) == "number"] <- "date"
GB2_2014 <- GB2_2014[, !(colnames(GB2_2014) %in% c("E4D5", "B661"))]
```

Create the GBI matrix 
```{r}
data_G2014<- GB2_2014 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2014$daygroup<-as.factor(paste(data_G2014$date, data_G2014$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2014<-data.table(ID=data_G2014$ID, group=data_G2014$daygroup, day=c(data_G2014$date))
         
GBI_G2014<- get_group_by_individual(ind_G2014, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2014 <- as.matrix(GBI_G2014)

write.csv(GBI_G2014, here("data","saved_matrices","GB2","GBI_G2014.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2014 <- get_network(GBI_G2014, data_format = "GBI", association_index = "SRI")
write.csv(net_G2014, here("data","saved_matrices","GB2","net_G2014.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2014 <- graph_from_adjacency_matrix(net_G2014,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2014 <- igraph::as_data_frame(g_G2014, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2014 <- edge_G2014 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2014 <- edge_G2014 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2014],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2014),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2014$pairID
  }
)
write.csv(edge_G2014, here("data","saved_matrices","GB2","edge_G2014.csv"), row.names = FALSE)
```
##### 2015
```{r}
GB2_2015<-read.csv(here("data","raw_roost_data","Dayroosts_GB2_2015_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2015<-GB2_2015[c(0:38)] #filter columns
GB2_2015<-GB2_2015[1:45,]

colnames(GB2_2015)[colnames(GB2_2015) == "number"] <- "date"

#remove one individual
GB2_2015<- GB2_2015[, !(colnames(GB2_2015) %in% "3D98")]
```

Create the GBI matrix 
```{r}
data_G2015<- GB2_2015 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2015$daygroup<-as.factor(paste(data_G2015$date, data_G2015$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2015<-data.table(ID=data_G2015$ID, group=data_G2015$daygroup, day=c(data_G2015$date))
         
GBI_G2015<- get_group_by_individual(ind_G2015, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2015 <- as.matrix(GBI_G2015)

write.csv(GBI_G2015, here("data","saved_matrices","GB2","GBI_G2015.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2015 <- get_network(GBI_G2015, data_format = "GBI", association_index = "SRI")
write.csv(net_G2015, here("data","saved_matrices","GB2","net_G2015.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2015 <- graph_from_adjacency_matrix(net_G2015,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2015 <- igraph::as_data_frame(g_G2015, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2015 <- edge_G2015 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2015, here("data","saved_matrices","GB2","edge_G2015.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2015 <- edge_G2015 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2015],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2015),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2015$pairID
  }
)
write.csv(edge_G2015, here("data","saved_matrices","GB2","edge_G2015.csv"), row.names = FALSE)
```
##### 2016
```{r}
GB2_2016<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2016_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2016<-GB2_2016[c(0:25)] #filter columns
colnames(GB2_2016)<-as.character(colnames(GB2_2016))
```

Create the GBI matrix 
```{r}
data_G2016<- GB2_2016 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2016$daygroup<-as.factor(paste(data_G2016$date, data_G2016$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2016<-data.table(ID=data_G2016$ID, group=data_G2016$daygroup, day=c(data_G2016$date))
         
GBI_G2016<- get_group_by_individual(ind_G2016, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2016 <- as.matrix(GBI_G2016)

write.csv(GBI_G2016, here("data","saved_matrices","GB2","GBI_G2016.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2016 <- get_network(GBI_G2016, data_format = "GBI", association_index = "SRI")
write.csv(net_G2016, here("data","saved_matrices","GB2","net_G2016.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2016 <- graph_from_adjacency_matrix(net_G2016,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2016 <- igraph::as_data_frame(g_G2016, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2016 <- edge_G2016 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2016, here("data","saved_matrices","GB2","edge_G2016.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2016 <- edge_G2016 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2016],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2016),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2016$pairID
  }
)
write.csv(edge_G2016, here("data","saved_matrices","GB2","edge_G2016.csv"), row.names = FALSE)
```
##### 2017
```{r}
GB2_2017<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2017_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2017<-GB2_2017[c(0:36)] #filter columns
GB2_2017<-GB2_2017[1:42,]
```

Create the GBI matrix 
```{r}
data_G2017<- GB2_2017 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2017$daygroup<-as.factor(paste(data_G2017$date, data_G2017$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2017<-data.table(ID=data_G2017$ID, group=data_G2017$daygroup, day=c(data_G2017$date))
         
GBI_G2017<- get_group_by_individual(ind_G2017, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2017 <- as.matrix(GBI_G2017)

write.csv(GBI_G2017, here("data","saved_matrices","GB2","GBI_G2017.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2017 <- get_network(GBI_G2017, data_format = "GBI", association_index = "SRI")
write.csv(net_G2017, here("data","saved_matrices","GB2","net_G2017.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2017 <- graph_from_adjacency_matrix(net_G2017,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2017 <- igraph::as_data_frame(g_G2017, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2017 <- edge_G2017 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2017, here("data","saved_matrices","GB2","edge_G2017.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2017 <- edge_G2017 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2017],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2017),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2017$pairID
  }
)
write.csv(edge_G2017, here("data","saved_matrices","GB2","edge_G2017.csv"), row.names = FALSE)
```
##### 2018
```{r}
GB2_2018<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2018_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2018<-GB2_2018[c(1:39)] #filter columns
GB2_2018<-GB2_2018[1:17,]

GB2_2018<- GB2_2018[, !(colnames(GB2_2018) %in% "F44A")]

#filter individuals that are observed less than 2 times
#count number of non-NA presences per individual
presence_counts_G2018 <- colSums(!is.na(GB2_2018[,-1]))  # exclude date column

# Keep only individuals with >= 2 or 3 presences
keep_IDs_G2018 <- names(presence_counts_G2018[presence_counts_G2018 >= 2])
print(ncol(GB2_2018[,-1]));print(length(keep_IDs_G2018))

GB2_2018_filtered <- GB2_2018[, c("date", keep_IDs_G2018)]
```

Create the GBI matrix 
```{r}
data_G2018<- GB2_2018_filtered %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2018$daygroup<-as.factor(paste(data_G2018$date, data_G2018$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2018<-data.table(ID=data_G2018$ID, group=data_G2018$daygroup, day=c(data_G2018$date))
         
GBI_G2018<- get_group_by_individual(ind_G2018, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2018 <- as.matrix(GBI_G2018)

write.csv(GBI_G2018, here("data","saved_matrices","GB2","GBI_G2018.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2018 <- get_network(GBI_G2018, data_format = "GBI", association_index = "SRI")
write.csv(net_G2018, here("data","saved_matrices","GB2","net_G2018.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2018 <- graph_from_adjacency_matrix(net_G2018,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2018 <- igraph::as_data_frame(g_G2018, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2018 <- edge_G2018 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2018, here("data","saved_matrices","GB2","edge_G2018.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2018 <- edge_G2018 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2018],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2018),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2018$pairID
  }
)
write.csv(edge_G2018, here("data","saved_matrices","GB2","edge_G2018.csv"), row.names = FALSE)
```
##### 2019
```{r}
GB2_2019<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2019_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2019<-GB2_2019[c(1:41)] #filter columns

GB2_2019 <- GB2_2019 %>%
  mutate(across(-date, as.character))

```

Create the GBI matrix 
```{r}
data_G2019<- GB2_2019 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2019$daygroup<-as.factor(paste(data_G2019$date, data_G2019$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2019<-data.table(ID=data_G2019$ID, group=data_G2019$daygroup, day=c(data_G2019$date))
         
GBI_G2019<- get_group_by_individual(ind_G2019, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2019 <- as.matrix(GBI_G2019)

write.csv(GBI_G2019, here("data","saved_matrices","GB2","GBI_G2019.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2019 <- get_network(GBI_G2019, data_format = "GBI", association_index = "SRI")
write.csv(net_G2019, here("data","saved_matrices","GB2","net_G2019.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2019 <- graph_from_adjacency_matrix(net_G2019,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2019 <- igraph::as_data_frame(g_G2019, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2019 <- edge_G2019 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2019, here("data","saved_matrices","GB2","edge_G2019.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2019 <- edge_G2019 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2019],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2019),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2019$pairID
  }
)
write.csv(edge_G2019, here("data","saved_matrices","GB2","edge_G2019.csv"), row.names = FALSE)
```
##### 2020
```{r}
GB2_2020<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2020_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2020<-GB2_2020[c(1:51)] #filter columns
GB2_2020<-GB2_2020[1:27,]

GB2_2020 <- GB2_2020 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

GB2_2020 <- GB2_2020 %>%
  mutate(across(-date, as.character))
```

Create the GBI matrix 
```{r}
data_G2020<- GB2_2020 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2020$daygroup<-as.factor(paste(data_G2020$date, data_G2020$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2020<-data.table(ID=data_G2020$ID, group=data_G2020$daygroup, day=c(data_G2020$date))
         
GBI_G2020<- get_group_by_individual(ind_G2020, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2020 <- as.matrix(GBI_G2020)

write.csv(GBI_G2020, here("data","saved_matrices","GB2","GBI_G2020.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2020 <- get_network(GBI_G2020, data_format = "GBI", association_index = "SRI")
write.csv(net_G2020, here("data","saved_matrices","GB2","net_G2020.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2020 <- graph_from_adjacency_matrix(net_G2020,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2020 <- igraph::as_data_frame(g_G2020, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2020 <- edge_G2020 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2020, here("data","saved_matrices","GB2","edge_G2020.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2020 <- edge_G2020 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2020],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2020),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2020$pairID
  }
)
write.csv(edge_G2020, here("data","saved_matrices","GB2","edge_G2020.csv"), row.names = FALSE)
```
##### 2021
```{r}
GB2_2021<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2021_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)

GB2_2021<-GB2_2021[c(1:56)] #filter columns
GB2_2021<-GB2_2021[1:37,]

GB2_2021 <- GB2_2021 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

colnames(GB2_2021)[colnames(GB2_2021) == "Date"] <- "date"

GB2_2021<- GB2_2021 %>%
  mutate(across(-date, as.character))
```

Create the GBI matrix 
```{r}
data_G2021<- GB2_2021 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2021$daygroup<-as.factor(paste(data_G2021$date, data_G2021$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2021<-data.table(ID=data_G2021$ID, group=data_G2021$daygroup, day=c(data_G2021$date))
         
GBI_G2021<- get_group_by_individual(ind_G2021, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2021 <- as.matrix(GBI_G2021)

write.csv(GBI_G2021, here("data","saved_matrices","GB2","GBI_G2021.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2021 <- get_network(GBI_G2021, data_format = "GBI", association_index = "SRI")
write.csv(net_G2021, here("data","saved_matrices","GB2","net_G2021.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2021 <- graph_from_adjacency_matrix(net_G2021,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2021 <- igraph::as_data_frame(g_G2021, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2021 <- edge_G2021 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2021, here("data","saved_matrices","GB2","edge_G2021.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2021 <- edge_G2021 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2021],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2021),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2021$pairID
  }
)
write.csv(edge_G2021, here("data","saved_matrices","GB2","edge_G2021.csv"), row.names = FALSE)
```
##### 2022
```{r}
GB2_2022<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2022_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)

GB2_2022<-GB2_2022[c(1:60)]#filter columns

GB2_2022 <-GB2_2022 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(GB2_2022)[colnames(GB2_2022) == "Date"] <- "date"

#filter individuals that are observed less than 2 times
#count number of non-NA presences per individual
presence_counts_G2022 <- colSums(!is.na(GB2_2022[,-1]))  # exclude date column

# Keep only individuals with >= 2 or 3 presences
keep_IDs_G2022 <- names(presence_counts_G2022[presence_counts_G2022 >= 2])
print(ncol(GB2_2022[,-1]));print(length(keep_IDs_G2022))

GB2_2022_filtered <- GB2_2022[, c("date", keep_IDs_G2022)]

GB2_2022_filtered<- GB2_2022_filtered %>%
  mutate(across(-date, as.character))
```

Create the GBI matrix 
```{r}
data_G2022<- GB2_2022_filtered %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2022$daygroup<-as.factor(paste(data_G2022$date, data_G2022$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2022<-data.table(ID=data_G2022$ID, group=data_G2022$daygroup, day=c(data_G2022$date))
         
GBI_G2022<- get_group_by_individual(ind_G2022, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2022 <- as.matrix(GBI_G2022)

write.csv(GBI_G2022, here("data","saved_matrices","GB2","GBI_G2022.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2022 <- get_network(GBI_G2022, data_format = "GBI", association_index = "SRI")
write.csv(net_G2022, here("data","saved_matrices","GB2","net_G2022.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2022 <- graph_from_adjacency_matrix(net_G2022,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2022 <- igraph::as_data_frame(g_G2022, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2022 <- edge_G2022 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2022, here("data","saved_matrices","GB2","edge_G2022.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2022 <- edge_G2022 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2022],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2022),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2022$pairID
  }
)
write.csv(edge_G2022, here("data","saved_matrices","GB2","edge_G2022.csv"), row.names = FALSE)
```
##### 2023
```{r}
GB2_2023<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_GB2_2023_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
GB2_2023<-GB2_2023[c(1:70)] #filter columns

GB2_2023<-GB2_2023 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

#GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(GB2_2023)[colnames(GB2_2023) == "Date"] <- "date"

GB2_2023<- GB2_2023 %>%
  mutate(across(-date, as.character))
```

Create the GBI matrix 
```{r}
data_G2023<- GB2_2023 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2023$daygroup<-as.factor(paste(data_G2023$date, data_G2023$box, sep="_"))

#df with 3 col: ID, group and day 
ind_G2023<-data.table(ID=data_G2023$ID, group=data_G2023$daygroup, day=c(data_G2023$date))
         
GBI_G2023<- get_group_by_individual(ind_G2023, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2023 <- as.matrix(GBI_G2023)

write.csv(GBI_G2023, here("data","saved_matrices","GB2","GBI_G2023.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_G2023 <- get_network(GBI_G2023, data_format = "GBI", association_index = "SRI")
write.csv(net_G2023, here("data","saved_matrices","GB2","net_G2023.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_G2023 <- graph_from_adjacency_matrix(net_G2023,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_G2023 <- igraph::as_data_frame(g_G2023, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_G2023 <- edge_G2023 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_G2023, here("data","saved_matrices","GB2","edge_G2023.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_G2023 <- edge_G2023 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_GB2$pairID[mother_daughter_GB2$year == 2023],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_GB2$year == 2023),
  function(i) {
    md_pair <- sort(c(mother_daughter_GB2$rfid_mom[i], mother_daughter_GB2$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_G2023$pairID
  }
)
write.csv(edge_G2023, here("data","saved_matrices","GB2","edge_G2023.csv"), row.names = FALSE)
```

#Combined edge list 
We can rbind all the edge lists together to have one file that captures all the associations between the females over the 11 years
```{r}
edge_G2013$year<-2013;edge_G2014$year<-2014;edge_G2015$year<-2015;edge_G2016$year<-2016;edge_G2017$year<-2017;edge_G2018$year<-2018;edge_G2019$year<-2019;edge_G2020$year<-2020;edge_G2021$year<-2021;edge_G2022$year<-2022;edge_G2023$year<-2023

edges_GB2<- rbind(edge_G2013, edge_G2014, edge_G2015, edge_G2016, edge_G2017, edge_G2018,
                        edge_G2019, edge_G2020, edge_G2021, edge_G2022, edge_G2023)
#save it 
write.csv(edges_GB2, here("data","saved_matrices","GB2","edge_list_GB2.csv"), row.names = FALSE)
```

#UA
##### 2013
```{r}
UA_2013<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2013_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2013<-UA_2013[c(1:24)]
colnames(UA_2013)[colnames(UA_2013) == "number"] <- "date"
```

Create the GBI matrix 
```{r}
data_U2013<- UA_2013 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2013$daygroup<-as.factor(paste(data_U2013$date, data_U2013$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2013<-data.table(ID=data_U2013$ID, group=data_U2013$daygroup, day=c(data_U2013$date))
         
GBI_U2013<- get_group_by_individual(ind_U2013, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2013 <- as.matrix(GBI_U2013)

write.csv(GBI_U2013, here("data","saved_matrices","UA","GBI_U2013.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2013 <- get_network(GBI_U2013, data_format = "GBI", association_index = "SRI")
write.csv(net_U2013, here("data","saved_matrices","UA","net_U2013.csv"), row.names = TRUE)
```
Create the edge list
```{r}
#first create the igraph object 
g_U2013 <- graph_from_adjacency_matrix(net_U2013,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2013 <- igraph::as_data_frame(g_U2013, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2013 <- edge_U2013 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2013 <- edge_U2013 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2013],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2013),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2013$pairID
  }
)
write.csv(edge_U2013, here("data","saved_matrices","UA","edge_U2013.csv"), row.names = FALSE)
```
##### 2014
```{r}
UA_2014<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2014_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2014<-UA_2014[c(1:30)] #filter columns
UA_2014<-UA_2014[1:55,]

colnames(UA_2014)[colnames(UA_2014) == "number"] <- "date"
UA_2014<- UA_2014[, !(colnames(UA_2014) %in% "Year")]
UA_2014<- UA_2014[, !(colnames(UA_2014) %in% "Month")]
UA_2014<- UA_2014[, !(colnames(UA_2014) %in% "Day")]
```

Create the GBI matrix 
```{r}
data_U2014<- UA_2014 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2014$daygroup<-as.factor(paste(data_U2014$date, data_U2014$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2014<-data.table(ID=data_U2014$ID, group=data_U2014$daygroup, day=c(data_U2014$date))
         
GBI_U2014<- get_group_by_individual(ind_U2014, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2014 <- as.matrix(GBI_U2014)

write.csv(GBI_U2014, here("data","saved_matrices","UA","GBI_U2014.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2014 <- get_network(GBI_U2014, data_format = "GBI", association_index = "SRI")
write.csv(net_U2014, here("data","saved_matrices","UA","net_U2014.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2014 <- graph_from_adjacency_matrix(net_U2014,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2014 <- igraph::as_data_frame(g_U2014, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2014 <- edge_U2014 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2014 <- edge_U2014 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2014],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2014),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2014$pairID
  }
)
write.csv(edge_U2014, here("data","saved_matrices","UA","edge_U2014.csv"), row.names = FALSE)
```
##### 2015
```{r}
UA_2015<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2015_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2015<-UA_2015[c(1:23)] #filter columns
UA_2015<-UA_2015[2:51,]

UA_2015<-UA_2015 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

#GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(UA_2015)[colnames(UA_2015) == "number"] <- "date"
```

Create the GBI matrix 
```{r}
data_U2015<- UA_2015 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2015$daygroup<-as.factor(paste(data_U2015$date, data_U2015$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2015<-data.table(ID=data_U2015$ID, group=data_U2015$daygroup, day=c(data_U2015$date))
         
GBI_U2015<- get_group_by_individual(ind_U2015, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2015 <- as.matrix(GBI_U2015)

write.csv(GBI_U2015, here("data","saved_matrices","UA","GBI_U2015.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2015 <- get_network(GBI_U2015, data_format = "GBI", association_index = "SRI")
write.csv(net_U2015, here("data","saved_matrices","UA","net_U2015.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2015 <- graph_from_adjacency_matrix(net_U2015,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2015 <- igraph::as_data_frame(g_U2015, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2015 <- edge_U2015 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2015, here("data","saved_matrices","UA","edge_U2015.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2015 <- edge_U2015 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2015],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2015),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2015$pairID
  }
)
write.csv(edge_U2015, here("data","saved_matrices","UA","edge_U2015.csv"), row.names = FALSE)
```
##### 2016
```{r}
UA_2016<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2016_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2016<-UA_2016[c(1:27)] #filter columns
UA_2016<-UA_2016[2:39,]
```

Create the GBI matrix 
```{r}
data_U2016<- UA_2016 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2016$daygroup<-as.factor(paste(data_U2016$date, data_U2016$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2016<-data.table(ID=data_U2016$ID, group=data_U2016$daygroup, day=c(data_U2016$date))
         
GBI_U2016<- get_group_by_individual(ind_U2016, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2016 <- as.matrix(GBI_U2016)

write.csv(GBI_U2016, here("data","saved_matrices","UA","GBI_U2016.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2016 <- get_network(GBI_U2016, data_format = "GBI", association_index = "SRI")
write.csv(net_U2016, here("data","saved_matrices","UA","net_U2016.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2016 <- graph_from_adjacency_matrix(net_U2016,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2016 <- igraph::as_data_frame(g_U2016, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2016 <- edge_U2016 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2016, here("data","saved_matrices","UA","edge_U2016.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2016 <- edge_U2016 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2016],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2016),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2016$pairID
  }
)
write.csv(edge_U2016, here("data","saved_matrices","UA","edge_U2016.csv"), row.names = FALSE)
```
##### 2017
```{r}
UA_2017<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2017_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2017<-UA_2017[c(1:30)] #filter columns
```

Create the GBI matrix 
```{r}
data_U2017<- UA_2017 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2017$daygroup<-as.factor(paste(data_U2017$date, data_U2017$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2017<-data.table(ID=data_U2017$ID, group=data_U2017$daygroup, day=c(data_U2017$date))
         
GBI_U2017<- get_group_by_individual(ind_U2017, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2017 <- as.matrix(GBI_U2017)

write.csv(GBI_U2017, here("data","saved_matrices","UA","GBI_U2017.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2017 <- get_network(GBI_U2017, data_format = "GBI", association_index = "SRI")
write.csv(net_U2017, here("data","saved_matrices","UA","net_U2017.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2017 <- graph_from_adjacency_matrix(net_U2017,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2017 <- igraph::as_data_frame(g_U2017, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2017 <- edge_U2017 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2017, here("data","saved_matrices","UA","edge_U2017.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2017 <- edge_U2017 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2017],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2017),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2017$pairID
  }
)
write.csv(edge_U2017, here("data","saved_matrices","UA","edge_U2017.csv"), row.names = FALSE)
```
##### 2018
```{r}
UA_2018<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2018_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2018<-UA_2018[c(1:27)] #filter columns
names(UA_2018)[names(UA_2018) == "D6CC"] <- "FA5B"
names(UA_2018)[names(UA_2018) == "0B43"] <- "101F"
```

Create the GBI matrix 
```{r}
data_U2018<- UA_2018 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2018$daygroup<-as.factor(paste(data_U2018$date, data_U2018$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2018<-data.table(ID=data_U2018$ID, group=data_U2018$daygroup, day=c(data_U2018$date))
         
GBI_U2018<- get_group_by_individual(ind_U2018, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2018 <- as.matrix(GBI_U2018)

write.csv(GBI_U2018, here("data","saved_matrices","UA","GBI_U2018.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2018 <- get_network(GBI_U2018, data_format = "GBI", association_index = "SRI")
write.csv(net_U2018, here("data","saved_matrices","UA","net_U2018.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2018 <- graph_from_adjacency_matrix(net_U2018,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2018 <- igraph::as_data_frame(g_U2018, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2018 <- edge_U2018 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2018, here("data","saved_matrices","UA","edge_U2018.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2018 <- edge_U2018 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2018],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2018),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2018$pairID
  }
)
write.csv(edge_U2018, here("data","saved_matrices","UA","edge_U2018.csv"), row.names = FALSE)
```
##### 2019
```{r}
UA_2019<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2019_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2019<-UA_2019[c(1:23)] #filter columns
UA_2019<-UA_2019[1:41,]

UA_2019<-UA_2019 %>%
 mutate(across(everything(), ~ replace(., . == "", NA)))
UA_2019 <- UA_2019 %>%
  mutate(across(-date, as.character))

```

Create the GBI matrix 
```{r}
data_U2019<- UA_2019 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2019$daygroup<-as.factor(paste(data_U2019$date, data_U2019$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2019<-data.table(ID=data_U2019$ID, group=data_U2019$daygroup, day=c(data_U2019$date))
         
GBI_U2019<- get_group_by_individual(ind_U2019, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2019 <- as.matrix(GBI_U2019)

write.csv(GBI_U2019, here("data","saved_matrices","UA","GBI_U2019.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2019 <- get_network(GBI_U2019, data_format = "GBI", association_index = "SRI")
write.csv(net_U2019, here("data","saved_matrices","UA","net_U2019.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2019 <- graph_from_adjacency_matrix(net_U2019,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2019 <- igraph::as_data_frame(g_U2019, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2019 <- edge_U2019 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2019, here("data","saved_matrices","UA","edge_U2019.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2019 <- edge_U2019 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2019],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2019),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2019$pairID
  }
)
write.csv(edge_U2019, here("data","saved_matrices","UA","edge_U2019.csv"), row.names = FALSE)
```
##### 2020
```{r}
UA_2020<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2020_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2020<-UA_2020[c(1:24)] #filter columns
UA_2020<-UA_2020[1:28,]

UA_2020 <- UA_2020 %>%
  mutate(across(-date, as.character))
```

Create the GBI matrix 
```{r}
data_U2020<- UA_2020 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2020$daygroup<-as.factor(paste(data_U2020$date, data_U2020$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2020<-data.table(ID=data_U2020$ID, group=data_U2020$daygroup, day=c(data_U2020$date))
         
GBI_U2020<- get_group_by_individual(ind_U2020, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2020 <- as.matrix(GBI_U2020)

write.csv(GBI_U2020, here("data","saved_matrices","UA","GBI_U2020.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2020 <- get_network(GBI_U2020, data_format = "GBI", association_index = "SRI")
write.csv(net_U2020, here("data","saved_matrices","UA","net_U2020.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2020 <- graph_from_adjacency_matrix(net_U2020,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2020 <- igraph::as_data_frame(g_U2020, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2020 <- edge_U2020 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2020, here("data","saved_matrices","UA","edge_U2020.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2020 <- edge_U2020 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2020],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2020),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2020$pairID
  }
)
write.csv(edge_U2020, here("data","saved_matrices","UA","edge_U2020.csv"), row.names = FALSE)
```
##### 2021
```{r}
UA_2021<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2021_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)

UA_2021<-UA_2021[c(1:25)] #filter columns

UA_2021<-UA_2021 %>%
 mutate(across(everything(), ~ replace(., . == "", NA)))

#GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(UA_2021)[colnames(UA_2021) == "Date"] <- "date"
names(UA_2021)[names(UA_2021) == "9B39"] <- "F50E"
names(UA_2021)[names(UA_2021) == "37F2"] <- "100F"

UA_2021<- UA_2021 %>%
  mutate(across(-date, as.character))
```

Create the GBI matrix 
```{r}
data_U2021<- UA_2021 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2021$daygroup<-as.factor(paste(data_U2021$date, data_U2021$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2021<-data.table(ID=data_U2021$ID, group=data_U2021$daygroup, day=c(data_U2021$date))
         
GBI_U2021<- get_group_by_individual(ind_U2021, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2021 <- as.matrix(GBI_U2021)

write.csv(GBI_U2021, here("data","saved_matrices","UA","GBI_U2021.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2021 <- get_network(GBI_U2021, data_format = "GBI", association_index = "SRI")
write.csv(net_U2021, here("data","saved_matrices","UA","net_U2021.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2021 <- graph_from_adjacency_matrix(net_U2021,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2021 <- igraph::as_data_frame(g_U2021, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2021 <- edge_U2021 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2021, here("data","saved_matrices","UA","edge_U2021.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2021 <- edge_U2021 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2021],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2021),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2021$pairID
  }
)
write.csv(edge_U2021, here("data","saved_matrices","UA","edge_U2021.csv"), row.names = FALSE)
```
##### 2022
```{r}
UA_2022<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2022_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2022<-UA_2022[c(1:22)] #filter columns
UA_2022<-UA_2022[1:29,]

#GB2_2023<-GB2_2023 %>%
 # mutate(across(everything(), ~ replace(., . == "", NA)))

#GB2_2022<-GB2_2022[, !(colnames(GB2_2022) %in% "15BF")]

colnames(UA_2022)[colnames(UA_2022) == "Date"] <- "date"

UA_2022<- UA_2022 %>%
  mutate(across(-date, as.character))
```

Create the GBI matrix 
```{r}
data_U2022<- UA_2022 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2022$daygroup<-as.factor(paste(data_U2022$date, data_U2022$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2022<-data.table(ID=data_U2022$ID, group=data_U2022$daygroup, day=c(data_U2022$date))
         
GBI_U2022<- get_group_by_individual(ind_U2022, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2022 <- as.matrix(GBI_U2022)

write.csv(GBI_U2022, here("data","saved_matrices","UA","GBI_U2022.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2022 <- get_network(GBI_U2022, data_format = "GBI", association_index = "SRI")
write.csv(net_U2022, here("data","saved_matrices","UA","net_U2022.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2022 <- graph_from_adjacency_matrix(net_U2022,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2022 <- igraph::as_data_frame(g_U2022, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2022 <- edge_U2022 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2022, here("data","saved_matrices","UA","edge_U2022.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2022 <- edge_U2022 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2022],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2022),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2022$pairID
  }
)
write.csv(edge_U2022, here("data","saved_matrices","UA","edge_U2022.csv"), row.names = FALSE)
```
##### 2023
```{r}
UA_2023<-read.csv(here("data","raw_roost_data","CSV_Dayroosts_UA_2023_juv.csv"),header = TRUE,sep = ";",row.names = NULL,check.names = FALSE)
UA_2023<-UA_2023[c(1:29)]


colnames(UA_2023)[colnames(UA_2023) == "Date"] <- "date"
mother_daughter_UA$rfid_off[mother_daughter_UA$rfid_off == "AE33"] <- "F50E"


```

Create the GBI matrix 
```{r}
data_U2023<- UA_2023 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2023$daygroup<-as.factor(paste(data_U2023$date, data_U2023$box, sep="_"))

#df with 3 col: ID, group and day 
ind_U2023<-data.table(ID=data_U2023$ID, group=data_U2023$daygroup, day=c(data_U2023$date))
         
GBI_U2023<- get_group_by_individual(ind_U2023, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2023 <- as.matrix(GBI_U2023)

write.csv(GBI_U2023, here("data","saved_matrices","UA","GBI_U2023.csv"), row.names = TRUE)
```

Create network using the SRI association index
```{r}
net_U2023 <- get_network(GBI_U2023, data_format = "GBI", association_index = "SRI")
write.csv(net_U2023, here("data","saved_matrices","UA","net_U2023.csv"), row.names = TRUE)
```

Create the edge list
```{r}
#first create the igraph object 
g_U2023 <- graph_from_adjacency_matrix(net_U2023,mode = "undirected",weighted = TRUE,diag = FALSE) 

#then edge list 
edge_U2023 <- igraph::as_data_frame(g_U2023, what = "edges")

#create a column with the pair_ID
# very important : have to be ordered to match everytime !
edge_U2023 <- edge_U2023 %>%
  mutate(
    pairID = ifelse(
      from < to,
      paste(from, to, sep = "-"),
      paste(to, from, sep = "-")
    )
  )
write.csv(edge_U2023, here("data","saved_matrices","UA","edge_U2023.csv"), row.names = FALSE)
```

Add a "pair_type" column in the edge list, by discriminating each pair to "mother-daughter" or "other"
```{r}
edge_U2023 <- edge_U2023 %>%
  mutate(
    pair_type = ifelse(
      pairID %in% mother_daughter_UA$pairID[mother_daughter_UA$year == 2023],
      "mother-daughter",
      "other"
    )
  )
#rapid check if all pairs are present 
sapply(
  which(mother_daughter_UA$year == 2023),
  function(i) {
    md_pair <- sort(c(mother_daughter_UA$rfid_mom[i], mother_daughter_UA$rfid_off[i]))
    md_pairID <- paste(md_pair, collapse = "-")
    md_pairID %in% edge_U2023$pairID
  }
)
write.csv(edge_U2023, here("data","saved_matrices","UA","edge_U2023.csv"), row.names = FALSE)
```

#Combined edge list 
We can rbind all the edge lists together to have one file that captures all the associations between the females over the 11 years
```{r}
edge_U2013$year<-2013;edge_U2014$year<-2014;edge_U2015$year<-2015;edge_U2016$year<-2016;edge_U2017$year<-2017;edge_U2018$year<-2018;edge_U2019$year<-2019;edge_U2020$year<-2020;edge_U2021$year<-2021;edge_U2022$year<-2022;edge_U2023$year<-2023

edges_UA<- rbind(edge_U2013, edge_U2014, edge_U2015, edge_U2016, edge_U2017, edge_U2018, edge_U2019, edge_U2020, edge_U2021, edge_U2022, edge_U2023)
#save it 
write.csv(edges_UA, here("data","saved_matrices","UA","edge_list_UA.csv"), row.names = FALSE)
```

#create the mom-random juv dataframe
```{r}
make_mom_juv_for_year <- function(y, edges, md_table) {

  md_y <- md_table %>% filter(year == y)

  # bats that matter in this year
  bats_y <- unique(c(md_y$rfid_mom, md_y$rfid_off))

  edges_y <- edges %>%
    filter(year == y) %>%
    filter(from %in% bats_y | to %in% bats_y)

  # lookup table with roles
  lookup_y <- md_y %>%
    pivot_longer(
      cols = c(rfid_mom, rfid_off),
      names_to = "role",
      values_to = "batID"
    ) %>%
    distinct(batID, role)

  edges_y <- edges_y %>%
    left_join(lookup_y, by = c("from" = "batID")) %>%
    rename(from_role = role) %>%
    left_join(lookup_y, by = c("to" = "batID")) %>%
    rename(to_role = role)

  # ordered pairID
  make_pairID <- function(x, y) {
    ifelse(x < y, paste(x, y, sep = "-"), paste(y, x, sep = "-"))
  }

  edges_y <- edges_y %>%
    mutate(pairID = make_pairID(from, to))

  # identify true MD pairs for this year
  true_MD_y <- md_y %>%
    mutate(pairID = make_pairID(rfid_mom, rfid_off)) %>%
    select(pairID)

  edges_y <- edges_y %>%
    mutate(
    pair_type_2 = case_when(
      # true mother–daughter
      pairID %in% true_MD_y$pairID ~ "mother-daughter",

      # mother with a non-own juvenile
      (
        (from_role == "rfid_mom" & to_role == "rfid_off") |
        (from_role == "rfid_off" & to_role == "rfid_mom")
      ) ~ "mother-random-juvenile",

      # everything else
      TRUE ~ "other"
    )
  )

}
```

```{r}
years_UA <- sort(unique(edges_UA$year))

mom_juv_UA <- map_dfr(
  years_UA,
  make_mom_juv_for_year,
  edges = edges_UA,
  md_table = mother_daughter_UA
)

write.csv(mom_juv_UA, here("data","saved_matrices","UA","mom_juv_UA.csv"), row.names = FALSE)
```



