---
title: "Manuscript_hypothesis1"
author: "Flore Michel"
date: "2025-10-09"
output: html_document
---
```{r setup, include=FALSE}
library(here)
library(igraph)
library(asnipe)
library(sna)
library(assocInd)
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(tidyverse)
library(viridis)
library(patchwork)
#library(hrbrthemes)
library(igraph)
library(ggraph)
library(colormap)
library(tidyr)
library(dplyr)
library(data.table)
library(glmmTMB)
library(sjPlot)
library(DHARMa)
library(influence.ME)
library(MuMIn)
library(performance)
library(car)
library(ggrepel)
library(broom.mixed)
library(gt)
```

###load
```{r load edges}
edges_GB2<-read.csv(here("data", "saved_matrices", "GB2", "edge_list_GB2.csv"),header=TRUE, row.names=NULL,sep=",",check.names = F)
```

```{r load the networks, include= FALSE}
net_G2013<-read.csv(here("data","saved_matrices","GB2","net_G2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2013 <- as.matrix(net_G2013)
net_G2016<-read.csv(here("data","saved_matrices","GB2","net_G2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2016 <- as.matrix(net_G2016)
net_G2015<-read.csv(here("data","saved_matrices","GB2","net_G2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2015 <- as.matrix(net_G2015)
net_G2016<-read.csv(here("data","saved_matrices","GB2","net_G2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2016 <- as.matrix(net_G2016)
net_G2017<-read.csv(here("data","saved_matrices","GB2","net_G2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2017 <- as.matrix(net_G2017)
net_G2018<-read.csv(here("data","saved_matrices","GB2","net_G2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2018 <- as.matrix(net_G2018)
net_G2019<-read.csv(here("data","saved_matrices","GB2","net_G2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2019 <- as.matrix(net_G2019)
net_G2020<-read.csv(here("data","saved_matrices","GB2","net_G2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2020 <- as.matrix(net_G2020)
net_G2021<-read.csv(here("data","saved_matrices","GB2","net_G2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2021 <- as.matrix(net_G2021)
net_G2022<-read.csv(here("data","saved_matrices","GB2","net_G2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2022 <- as.matrix(net_G2022)
net_G2023<-read.csv(here("data","saved_matrices","GB2","net_G2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_G2023 <- as.matrix(net_G2023)
```

Create the igraph object from the network matrices
```{r}
g_G2013<- graph_from_adjacency_matrix(net_G2013, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2016<- graph_from_adjacency_matrix(net_G2016, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2015<- graph_from_adjacency_matrix(net_G2015, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2016<- graph_from_adjacency_matrix(net_G2016, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2017<- graph_from_adjacency_matrix(net_G2017, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2018<- graph_from_adjacency_matrix(net_G2018, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2019<- graph_from_adjacency_matrix(net_G2019, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2020<- graph_from_adjacency_matrix(net_G2020, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2021<- graph_from_adjacency_matrix(net_G2021, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2022<- graph_from_adjacency_matrix(net_G2022, mode = "undirected", weighted = TRUE, diag = FALSE)
g_G2023<- graph_from_adjacency_matrix(net_G2023, mode = "undirected", weighted = TRUE, diag = FALSE)
```

Load the GBI matrices
```{r}
GBI_G2013<-read.csv(here("data","saved_matrices","GB2","GBI_G2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2013 <- as.matrix(GBI_G2013)
GBI_G2014<-read.csv(here("data","saved_matrices","GB2","GBI_G2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2014 <- as.matrix(GBI_G2014)
GBI_G2015<-read.csv(here("data","saved_matrices","GB2","GBI_G2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2015 <- as.matrix(GBI_G2015)
GBI_G2016<-read.csv(here("data","saved_matrices","GB2","GBI_G2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2016 <- as.matrix(GBI_G2016)
GBI_G2017<-read.csv(here("data","saved_matrices","GB2","GBI_G2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2017 <- as.matrix(GBI_G2017)
GBI_G2018<-read.csv(here("data","saved_matrices","GB2","GBI_G2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2018 <- as.matrix(GBI_G2018)
GBI_G2019<-read.csv(here("data","saved_matrices","GB2","GBI_G2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2019 <- as.matrix(GBI_G2019)
GBI_G2020<-read.csv(here("data","saved_matrices","GB2","GBI_G2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2020 <- as.matrix(GBI_G2020)
GBI_G2021<-read.csv(here("data","saved_matrices","GB2","GBI_G2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2021 <- as.matrix(GBI_G2021)
GBI_G2022<-read.csv(here("data","saved_matrices","GB2","GBI_G2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2022 <- as.matrix(GBI_G2022)
GBI_G2023<-read.csv(here("data","saved_matrices","GB2","GBI_G2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_G2023 <- as.matrix(GBI_G2023)
```

```{r}
mother_daughter_GB2 <- read.csv(here("data","mother_daughter_GB2.csv"))
mother_daughter_UA <- read.csv(here("data","mother_daughter_UA.csv"))
```

```{r}
mom_juv_GB2<-read.csv(here("data","saved_matrices","GB2","mom_juv_GB2.csv"),header=TRUE,sep=",",check.names = F) 
```

### Do mother-daughters pairs have higher association than other pairs

### GB2 ----------------

## QAP 
*(quadratic assignment procedure) analysis*
This analysis need two matrices: one with the associations, and another with the predictor (here "mother-daughter pairs", binary (md = 1, other = 0)). The QAP analysis will test if the mother-daughter pairs are more associated than other pairs, by permuting the rows and columns of th predictor matrix. This is a yearly analysis, that will inform us on the strength of monther-daughter associations each year, but also the consistency of this pattern across years.

```{r functin to create M}
#create function to create the M matrix, binary with 1 for mother-daughter and 0 for other pairs
build_m_matrix <- function(association_matrix, edges, the_year) {

  ids <- rownames(association_matrix)
  #edges for this year + the desired pair_type
  edges_subset <- subset(edges, year == the_year & pair_type == "mother-daughter")
  
  #M matrix of zeros with same row/col
  M <- matrix(0, nrow = length(ids), ncol = length(ids),
              dimnames = list(ids, ids))
  
  #M matrix with 1 for mother–daughter
  for (r in seq_len(nrow(edges_subset))) {
    bat_i <- edges_subset$from[r]
    bat_j <- edges_subset$to[r]
    #only if both bats
    if (bat_i %in% ids && bat_j %in% ids) {
      M[bat_i, bat_j] <- 1
      M[bat_j, bat_i] <- 1
    }
  }
  
  return(M)
}
```

```{r}
net_list_GB2 <- list(
  "2013" = net_G2013,"2014" = net_G2014,"2015" = net_G2015,"2016" = net_G2016,"2017" = net_G2017,"2018" = net_G2018,"2019" = net_G2019,"2020" = net_G2020,"2021" = net_G2021,"2022" = net_G2022,"2023" = net_G2023
)
```

```{r function for QAP}
qap_all_years <- function(net_list,edges){
  results_list <- list()
  
  for (the_year in names(net_list)) {
    A_yr <- net_list[[the_year]]
  
    M <- build_m_matrix(A_yr, edges, the_year) 
    g <- list(A_yr, M)
    
    #run QAP
    qap_res <- qaptest(g,FUN = gcor,g1  = 1,g2  = 2,reps = 10000)
    
    s <- summary(qap_res)
    test_val <- s$testval
    p_greq   <- s$pgreq
    p_leeq   <- s$pleeq
    

    results_df <- data.frame(
      year        = the_year,
      correlation = test_val,
      pval_upper  = p_greq,
      pval_lower  = p_leeq,
      stringsAsFactors = FALSE
    )
    
    results_list[[the_year]] <- results_df
  }
  
  all_results_df <- do.call(rbind, results_list)
  rownames(all_results_df) <- NULL
  
  return(all_results_df)
}
```

```{r}
QAP_GB2 <- qap_all_years(
  net_list = net_list_GB2,
  edges    = edges_GB2)

QAP_GB2
```
#GLMM 
*Analysis across the 11 years* 
To compare the strength of associations across all years for entire data set, incorporating all the years.

Prepare the edge list for the model
```{r}
edges_GB2$pair_type<-as.factor(edges_GB2$pair_type)
edges_GB2$pair_type <- relevel(edges_GB2$pair_type, ref = "other")

# re scale to fit the beta distribution
edges_GB2$weight <- (edges_GB2$weight * (nrow(edges_GB2) - 1) + 0.5) / nrow(edges_GB2)
```

Add the colony size 
```{r}
colony_size_GB2<-read.csv(here("data","saved_matrices","GB2","colony_size_df.csv"),header=TRUE,sep=",",check.names = F)
colony_size_GB2 <- colony_size_GB2 %>%
  mutate(year = as.numeric(year))

edges_GB2 <- edges_GB2 %>%
  left_join(colony_size_GB2, by = "year")
```

check assumptions before the model
```{r}
assoc_mother_daughter_GB2<-edges_GB2$weight[edges_GB2$pair_type=="mother-daughter"]
assoc_other_GB2<-edges_GB2$weight[edges_GB2$pair_type!="mother-daughter"]

hist(assoc_mother_daughter_GB2, main="Distribution of associations for mother-daughter pairs", xlab="association strength")
hist(assoc_other_GB2,main="Distribution of associations for other pairs", xlab="association strength")
hist(edges_GB2$weight,main="Distribution of associations of all pairs", xlab="association strength")

ggplot(edges_GB2, aes(x = pair_type, y = weight)) +
  geom_boxplot()  +
  theme_minimal() +
  labs(title = "Associations strength of mother-daughter pairs and other pairs", 
       x = "pair type", y = "strength")

#check distribution of the weight
qqnorm(edges_GB2$weight, main="QQ-Plot of association strength")
qqline(edges_GB2$weight, col = "red")
```

##model
```{r}
model_interaction_GB2 <- glmmTMB(weight ~ pair_type * colony_size + (1 | pairID) +  (1|year) ,  data = edges_GB2, family= beta_family())
summary(model_interaction_GB2)

model_GB2 <- glmmTMB(weight ~ pair_type + colony_size + (1 | pairID) + (1|year) ,  data = edges_GB2, family= beta_family())
summary(model_GB2)
```

run diagnostics
```{r}
simulationOutput <- simulateResiduals(fittedModel = model_GB2, n = 1000)
plot(simulationOutput)

simulationOutput_interaction <- simulateResiduals(fittedModel = model_interaction_GB2, n = 1000)
plot(simulationOutput_interaction)
```

```{r}
plotResiduals(simulationOutput, form = edges_GB2$colony_size)
plotResiduals(simulationOutput, form = edges_GB2$pair_type)
plotResiduals(simulationOutput, form = edges_GB2$year)
```
```{r}
plotResiduals(simulationOutput_interaction, form = edges_GB2$colony_size)
plotResiduals(simulationOutput_interaction, form = edges_GB2$pair_type)
plotResiduals(simulationOutput_interaction, form = edges_GB2$year)
```

####run diagnostics 
```{r}
r.squaredGLMM(model)
r.squaredGLMM(model_interaction)

check_overdispersion(model)
check_overdispersion(model_interaction)

#multicollinearity
check_collinearity(model)
check_collinearity(model_interaction)

anova(model, model_interaction, test="Chisq")
```


```{r}
#check for influential data points --> cooks distance
##-> need to use lme4 to do this (only to check)
model_lmer <- glmer(weight ~ pair_type * colony_size + (1 | pairID) + (1 | year),
                    data = edges_GB2,
                    family = binomial())

# influential years
infl_year <- influence(model_lmer, group = "year")
plot(infl_year, which = "cook")

# cooks distance but numerical
cooks.distance(infl_year)

```
```{r}
edges_GB2_new <- edges_GB2 %>%
  filter(year != 2019)

#center the numeric variable
edges_GB2_new$colony_size_c <- scale(edges_GB2_new$colony_size, center = TRUE, scale = FALSE)
```

```{r}
#Refit the models without 2019, and with centered colony size
model2_interaction_GB2 <- glmmTMB(weight ~ pair_type * colony_size_c + (1 | pairID) + (1|year) ,  data = edges_GB2_new, family= beta_family())

summary(model2_interaction_GB2)

model2_GB2 <- glmmTMB(weight ~ pair_type + colony_size_c + (1 | pairID) + (1|year) ,  data = edges_GB2_new, family= beta_family())
summary(model2_GB2)

```
The model with interaction show better fit (AIC and BIC)

We can check again the diagnostics for the final model
```{r}
check_overdispersion(model2_interaction_GB2) #no overdispersion
check_collinearity(model2_interaction_GB2) #low correlation

simulationOutput_interaction_2 <- simulateResiduals(fittedModel = model2_interaction_GB2, n = 1000)
plot(simulationOutput_interaction_2)
```

#mom -juv interactions
```{r}
mom_juv_GB2$pair_type_2 <- mom_juv_GB2$pair_type

#check
table(mom_juv_GB2$pair_type_2)
# 89    m-d      and          848 mother-random-juvenile

#rename every "mother-random-juvenile" to "mother-unrelated-daughter"
mom_juv_GB2$pair_type_2[mom_juv_GB2$pair_type_2 == "mother-random-juvenile"] <- "mother-unrelated-daughter"
table(mom_juv_GB2$pair_type_2)

#add the "mother-random-juvenile" to the edge list
edge_GB2_mom_juv <- edges_GB2_new %>%
  left_join(
    mom_juv_GB2 %>% select(pairID, year, pair_type_2),
    by = c("pairID", "year")
  )
#replace NA with "other" using %>% 
edge_GB2_mom_juv <- edge_GB2_mom_juv %>%
  mutate(pair_type_2 = ifelse(is.na(pair_type_2), "other", pair_type_2))

#change to factor
edge_GB2_mom_juv$pair_type_2<-as.factor(edge_GB2_mom_juv$pair_type_2)
edge_GB2_mom_juv$pair_type_2 <- relevel(edge_GB2_mom_juv$pair_type_2, ref = "other")
```

#final model with mom-juv
```{r}
model_mom_juv_GB2 <- glmmTMB(weight ~ pair_type_2 + colony_size_c + (1|year) + (1|pairID),  data = edge_GB2_mom_juv, family= beta_family())
summary(model_mom_juv_GB2)

model_interaction_mom_juv_GB2 <- glmmTMB(weight ~ pair_type_2 * colony_size_c + (1|year) + (1|pairID),  data = edge_GB2_mom_juv, family= beta_family())
summary(model_interaction_mom_juv_GB2)
```
#check 
```{r}
check_overdispersion(model_mom_juv_GB2) #no over dispersion
check_collinearity(model_mom_juv_GB2) #low correlation

check_overdispersion(model_interaction_mom_juv_GB2) #no over dispersion
check_collinearity(model_interaction_mom_juv_GB2) #low correlation

simulationOutput<- simulateResiduals(fittedModel = model_mom_juv_GB2, n = 1000)
plot(simulationOutput)
simulationOutput_int<- simulateResiduals(fittedModel = model_interaction_mom_juv_GB2, n = 1000)
plot(simulationOutput_int)
```


##final model
Keep both models, one without interaction and one with 
```{r}
#withtout interaction
model_final_GB2 <- glmmTMB(weight ~ pair_type_2 + colony_size_c + (1|year) + (1|pairID),  data = edge_GB2_mom_juv, family= beta_family())
summary(model_final_GB2)

#with interaction
model_final_interaction_GB2 <- glmmTMB(weight ~ pair_type_2 * colony_size_c + (1|year) + (1|pairID),  data = edge_GB2_mom_juv, family= beta_family())
summary(model_final_interaction_GB2)
```

#Permutation - models
Would we observe an effect of “mother–daughter” of this magnitude if mother–daughter labels were randomly assigned under the same social structure?

We can do a *label permutation*, --> Randomly reassign the “mother–daughter” labels among dyads, within year, while keeping the number of MD pairs per year fixed 

First, we can save the observed coefficients, for both models 
```{r}
#store coefficients
obs_coef <- fixef(model_final_GB2)$cond[c(
  "pair_type_2mother-daughter",
  "pair_type_2mother-unrelated-daughter",
  "colony_size_c"
)]

obs_coef_interaction <- fixef(model_final_interaction_GB2)$cond[c(
  "pair_type_2mother-daughter",
  "pair_type_2mother-unrelated-daughter",
  "colony_size_c",
  "pair_type_2mother-daughter:colony_size_c",
  "pair_type_2mother-unrelated-daughter:colony_size_c"
)]

```

*Label Permutations*
#### without interaction
```{r}
# Prepare storage
n_perm <- 5
perm_coefs <- matrix(NA, nrow = n_perm, ncol = length(obs_coef))
colnames(perm_coefs) <- names(obs_coef)

set.seed(123)

for (i in 1:n_perm) {
  # Permute pair_type2 within year
  perm_data <- edge_GB2_mom_juv %>%
    group_by(year) %>%
    mutate(pair_type2_perm = sample(pair_type_2)) %>%
    ungroup()
  
  # Refit model
  perm_model <- glmmTMB(
    weight ~ pair_type2_perm + colony_size_c + (1 | pairID) + (1 | year),
    family = beta_family(),
    data = perm_data
  )
  
  # Save the coefficients of interest
  perm_coefs[i, ] <- fixef(perm_model)$cond[c(
  "pair_type2_permmother-daughter",
  "pair_type2_permmother-unrelated-daughter",
  "colony_size_c"
  )]
}

# Convert to data frame
perm_coefs_df <- as.data.frame(perm_coefs)

save(perm_coefs_df, file = here("results", "permutation_results_GB2_mom_juv.RData"))
```
load permutations
```{r}
load(file = here("results", "permutation_results_GB2_mom_juv.RData"))
```


P-values 
```{r}
p_values <- sapply(names(obs_coef), function(name) {
  mean(perm_coefs_df[[name]] >= obs_coef[name])
})

#create result table 
# Convert obs_coef to a data frame
obs_df <- data.frame(
  term = names(obs_coef),
  observed = as.numeric(obs_coef),
  stringsAsFactors = FALSE
)

# Compute permutation summary (mean, SD) for each coefficient
perm_summary <- perm_coefs_df %>%
  pivot_longer(cols = everything(), names_to = "term", values_to = "value") %>%
  group_by(term) %>%
  summarise(
    perm_mean = mean(value),
    perm_sd   = sd(value),
    .groups = "drop"
  )

# Combine with observed and p-values
final_table <- obs_df %>%
  left_join(perm_summary, by = "term") %>%
  mutate(p_value = p_values[term])

# Optional: order by observed effect
final_table <- final_table %>% arrange(desc(observed))

final_table
write.csv(final_table,here("results","tables","Q1_perm_strength_GB2_table.csv"))
```

### visualize 
the permutation results for each coefficient: pair type, colony size,
```{r}
perm_coefs_df <- perm_coefs_df %>%
  rename(
    `mother-daughter` = `pair_type_2mother-daughter`,
    `other` = `pair_type_2other`)
    
   
perm_strength_GB2<-ggplot(perm_coefs_df, aes(x = `mother-daughter`)) +
  geom_histogram(bins = 30, fill = "darkgrey", color = "black") +
  geom_vline(xintercept = obs_coef["pair_type_2mother-daughter"], color = "red", size=1.1) +
  labs(title = "Pair type - mother-daughter",
       x = "Coefficient value",
       y = "Frequency") +
  theme_minimal()

ggplot(perm_coefs_df, aes(x = `other`)) +
  geom_histogram(bins = 30, fill = "darkgrey", color = "black") +
  geom_vline(xintercept = obs_coef["pair_type_2other"], color = "red", size=1.1) +
  labs(title = "Distribution of permuted coefficient - other",
       x = "Coefficient value",
       y = "Frequency") +
  theme_minimal()
save

ggplot(perm_coefs_df, aes(x = `colony_size_c`)) +
  geom_histogram(bins = 30, fill = "darkgrey", color = "black") +
  geom_vline(xintercept = obs_coef["colony_size_c"], color = "red", size=1.1) +
  labs(title = "Distribution of permuted coefficient - colony size",
       x = "Coefficient value",
       y = "Frequency") +
  theme_minimal()

ggsave(here("results","figures","perm_strength_GB2.png"), plot = perm_strength_GB2,width = 5, height = 5, dpi = 300)
```
#### with interaction
```{r}
n_perm <- 500
set.seed(123)

coef_names <- names(obs_coef_interaction)

perm_coefs_interaction <- matrix(
  NA,
  nrow = n_perm,
  ncol = length(coef_names)
)

colnames(perm_coefs_interaction) <- coef_names

for (i in 1:n_perm) {
  
  perm_data_interaction <- edge_GB2_mom_juv %>%
    group_by(year) %>%
    mutate(pair_type_2 = sample(pair_type_2)) %>%   # overwrite original
    ungroup()
  
  perm_model_interaction <- glmmTMB(
    weight ~ pair_type_2 * colony_size_c +
      (1 | pairID) + (1 | year),
    family = beta_family(),
    data = perm_data_interaction
  )
  
  perm_fixef <- fixef(perm_model_interaction)$cond
  
  perm_coefs_interaction[i, ] <- perm_fixef[coef_names]
}

perm_coefs_df_interaction <- as.data.frame(perm_coefs_interaction)

save(perm_coefs_df_interaction, file = here("results", "permutation_interaction_results_GB2_mom_juv.RData"))
```



```{r}
#load
load(file = here("results", "permutation_interaction_results_GB2_mom_juv.RData"))
perm_coefs_df_interaction
```


P-values 
```{r}
p_values <- sapply(names(obs_coef_interaction), function(name) {
  mean(perm_coefs_df_interaction[[name]] >= obs_coef_interaction[name])
})

#create result table 
# Convert obs_coef to a data frame
obs_df_interaction <- data.frame(
  term = names(obs_coef_interaction),
  observed = as.numeric(obs_coef_interaction),
  stringsAsFactors = FALSE
)

# Compute permutation summary (mean, SD) for each coefficient
perm_summary_interaction <- perm_coefs_df_interaction %>%
  pivot_longer(cols = everything(), names_to = "term", values_to = "value") %>%
  group_by(term) %>%
  summarise(
    perm_mean = mean(value),
    perm_sd   = sd(value),
    .groups = "drop"
  )

# Combine with observed and p-values
final_table_interaction <- obs_df_interaction %>%
  left_join(perm_summary_interaction, by = "term") %>%
  mutate(p_value = p_values[term])

# Optional: order by observed effect
final_table_interaction <- final_table_interaction %>% arrange(desc(observed))

final_table_interaction
```
### visualize 
the permutation results for each coefficient: pair type, colony size,
```{r}
perm_coefs_df_interaction <- perm_coefs_df_interaction %>%
  rename(
    `mother-daughter:colony_size` = `pair_type_2mother-daughter:colony_size_c`,
    `other:colony_size` = `pair_type_2other:colony_size_c`)
    
   
perm_interactionMDcolony<-ggplot(perm_coefs_df_interaction, aes(x = `mother-daughter:colony_size`)) +
  geom_histogram(bins = 30, fill = "darkgrey", color = "black") +
  geom_vline(xintercept = obs_coef["pair_type_2mother-daughter"], color = "red", size=1.1) +
  labs(title = "mother-daughter:colony size",
       x = "Coefficient value",
       y = "Frequency") +
  theme_minimal()

ggplot(perm_coefs_df_interaction, aes(x = `other:colony_size`)) +
  geom_histogram(bins = 30, fill = "darkgrey", color = "black") +
  geom_vline(xintercept = obs_coef["pair_type_2other"], color = "red", size=1.1) +
  labs(title = "Distribution of permuted coefficient - other:colony size",
       x = "Coefficient value",
       y = "Frequency") +
  theme_minimal()

ggsave(here("results","figures","perm_interactionMDcolony_GB2.png"), plot = perm_interactionMDcolony,width = 5, height = 5, dpi = 300)
```
### result table (both models)
```{r}
#without interaction
obs_model <- tidy(
  model_final_GB2,
  effects = "fixed",
  conf.int = FALSE
) %>%
  select(
    term,
    estimate,
    std.error,
    statistic
  ) %>%
  rename(
    obs_estimate = estimate,
    SE = std.error,
    z = statistic
  )

final_table <- final_table %>%
  select(
    term,
    perm_mean,
    p_value) %>%
  rename(
    perm_estimate = perm_mean,
    p = p_value
  )

table_main <- obs_model %>%
  left_join(final_table, by = "term") %>%
  mutate(
    model = "strength ~ pair_type + colony_size"
  )
```
```{r}
#with interaction
obs_model_interaction <- tidy(
  model_final_interaction_GB2,
  effects = "fixed",
  conf.int = FALSE
) %>%
  select(
    term,
    estimate,
    std.error,
    statistic
  ) %>%
  rename(
    obs_estimate = estimate,
    SE = std.error,
    z = statistic
  )

final_table_interaction <- final_table_interaction %>%
  select(
    term,
    perm_mean,
    p_value
  ) %>%
  rename(
    perm_estimate = perm_mean,
    p = p_value
  )

table_main_interaction <- obs_model_interaction %>%
  left_join(final_table_interaction, by = "term") %>%
  mutate(
    model = "strength ~ pair_type * colony_size"
  )
```

Final table 
```{r}
final_results_table <- bind_rows(
  table_main,
  table_main_interaction
) %>%
  relocate(model, .before = term)

final_results_table
#save it as a table
write.csv(final_results_table,
  here("results","tables","Q1_GLMM_table_GB2.csv"),
  row.names = FALSE
)
```

## UA ----------------
###load
```{r load edges}
edges_UA<-read.csv(here("data", "saved_matrices", "UA", "edge_list_UA.csv"),header=TRUE, row.names=NULL,sep=",",check.names = F)
```

```{r load the networks, include= FALSE}
net_U2013<-read.csv(here("data","saved_matrices","UA","net_U2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2013 <- as.matrix(net_U2013)
net_U2014<-read.csv(here("data","saved_matrices","UA","net_U2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2014 <- as.matrix(net_U2014)
net_U2015<-read.csv(here("data","saved_matrices","UA","net_U2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2015 <- as.matrix(net_U2015)
net_U2016<-read.csv(here("data","saved_matrices","UA","net_U2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2016 <- as.matrix(net_U2016)
net_U2017<-read.csv(here("data","saved_matrices","UA","net_U2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2017 <- as.matrix(net_U2017)
net_U2018<-read.csv(here("data","saved_matrices","UA","net_U2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2018 <- as.matrix(net_U2018)
net_U2019<-read.csv(here("data","saved_matrices","UA","net_U2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2019 <- as.matrix(net_U2019)
net_U2020<-read.csv(here("data","saved_matrices","UA","net_U2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2020 <- as.matrix(net_U2020)
net_U2021<-read.csv(here("data","saved_matrices","UA","net_U2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2021 <- as.matrix(net_U2021)
net_U2022<-read.csv(here("data","saved_matrices","UA","net_U2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2022 <- as.matrix(net_U2022)
net_U2023<-read.csv(here("data","saved_matrices","UA","net_U2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);net_U2023 <- as.matrix(net_U2023)

```

Create the igraph object from the network matrices
```{r}
g_U2013<- graph_from_adjacency_matrix(net_U2013, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2014<- graph_from_adjacency_matrix(net_U2014, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2015<- graph_from_adjacency_matrix(net_U2015, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2016<- graph_from_adjacency_matrix(net_U2016, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2017<- graph_from_adjacency_matrix(net_U2017, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2018<- graph_from_adjacency_matrix(net_U2018, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2019<- graph_from_adjacency_matrix(net_U2019, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2020<- graph_from_adjacency_matrix(net_U2020, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2021<- graph_from_adjacency_matrix(net_U2021, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2022<- graph_from_adjacency_matrix(net_U2022, mode = "undirected", weighted = TRUE, diag = FALSE)
g_U2023<- graph_from_adjacency_matrix(net_U2023, mode = "undirected", weighted = TRUE, diag = FALSE)

```

Load the GBI matrices
```{r}
GBI_U2013<-read.csv(here("data","saved_matrices","UA","GBI_U2013.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2013 <- as.matrix(GBI_U2013)
GBI_U2014<-read.csv(here("data","saved_matrices","UA","GBI_U2014.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2014 <- as.matrix(GBI_U2014)
GBI_U2015<-read.csv(here("data","saved_matrices","UA","GBI_U2015.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2015 <- as.matrix(GBI_U2015)
GBI_U2016<-read.csv(here("data","saved_matrices","UA","GBI_U2016.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2016 <- as.matrix(GBI_U2016)
GBI_U2017<-read.csv(here("data","saved_matrices","UA","GBI_U2017.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2017 <- as.matrix(GBI_U2017)
GBI_U2018<-read.csv(here("data","saved_matrices","UA","GBI_U2018.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2018 <- as.matrix(GBI_U2018)
GBI_U2019<-read.csv(here("data","saved_matrices","UA","GBI_U2019.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2019 <- as.matrix(GBI_U2019)
GBI_U2020<-read.csv(here("data","saved_matrices","UA","GBI_U2020.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2020 <- as.matrix(GBI_U2020)
GBI_U2021<-read.csv(here("data","saved_matrices","UA","GBI_U2021.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2021 <- as.matrix(GBI_U2021)
GBI_U2022<-read.csv(here("data","saved_matrices","UA","GBI_U2022.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2022 <- as.matrix(GBI_U2022)
GBI_U2023<-read.csv(here("data","saved_matrices","UA","GBI_U2023.csv"),header=TRUE,sep=",",check.names = F,row.names = 1);GBI_U2023 <- as.matrix(GBI_U2023)
```

## QAP 
```{r}
net_list_UA <- list(
  "2013" = net_U2013,"2014" = net_U2014,"2015" = net_U2015,"2016" = net_U2016,"2017" = net_U2017,"2018" = net_U2018,"2019" = net_U2019,"2020" = net_U2020,"2021" = net_U2021,"2022" = net_U2022,"2023" = net_U2023
)
```

```{r}
QAP_UA <- qap_all_years(
  net_list = net_list_UA,
  edges    = edges_UA)

QAP_UA
```
#GLMM 
*Analysis across the 11 years* 
To compare the strength of associations across all years for entire data set, incorporating all the years.

Prepare the edge list for the model
```{r}
edges_UA$pair_type<-as.factor(edges_UA$pair_type)
edges_UA$pair_type <- relevel(edges_UA$pair_type, ref = "other")

# re scale to fit the beta distribution
edges_UA$weight <- (edges_UA$weight * (nrow(edges_UA) - 1) + 0.5) / nrow(edges_UA)
```

Add the colony size 
```{r}
colony_size_UA<-read.csv(here("data","saved_matrices","UA","colony_size_UA.csv"),header=TRUE,sep=",",check.names = F)
colony_size_UA<- colony_size_UA %>%
  mutate(year = as.numeric(year))

edges_UA <- edges_UA %>%
  left_join(colony_size_UA, by = "year")
```

check assumptions before the model
```{r}
assoc_mother_daughter_UA<-edges_UA$weight[edges_UA$pair_type=="mother-daughter"]
assoc_other_UA<-edges_UA$weight[edges_UA$pair_type!="mother-daughter"]

hist(assoc_mother_daughter_UA, main="Distribution of associations for mother-daughter pairs", xlab="association strength")
hist(assoc_other_UA,main="Distribution of associations for other pairs", xlab="association strength")
hist(edges_UA$weight,main="Distribution of associations of all pairs", xlab="association strength")

ggplot(edges_UA, aes(x = pair_type, y = weight)) +
  geom_boxplot()  +
  theme_minimal() +
  labs(title = "Associations strength of mother-daughter pairs and other pairs", 
       x = "pair type", y = "strength")

#check distribution of the weight
qqnorm(edges_UA$weight, main="QQ-Plot of association strength")
qqline(edges_UA$weight, col = "red")
```
```{r}
#center the colony variable
edges_UA$colony_size_c <- scale(edges_UA$colony_size, center = TRUE, scale = FALSE)
```

#mom -juv interactions
```{r}
#load
mom_juv_UA<-read.csv(here("data","saved_matrices","UA","mom_juv_UA.csv"),header=TRUE,sep=",",check.names = F)

table(mom_juv_UA$pair_type_2)
# 45                    155   (random juv)                1551 (other)

#rename every "mother-random-juvenile" to "mother-unrelated-daughter"
mom_juv_UA$pair_type_2[mom_juv_UA$pair_type_2 == "mother-random-juvenile"] <- "mother-unrelated-daughter"
table(mom_juv_UA$pair_type_2)

#add the "mother-random-juvenile" to the edge list
edge_UA_mom_juv <- edges_UA %>%
  left_join(
    mom_juv_UA %>% select(pairID, year, pair_type_2),
    by = c("pairID", "year")
  )
#replace NA with "other" using %>% 
edge_UA_mom_juv <- edge_UA_mom_juv %>%
  mutate(pair_type_2 = ifelse(is.na(pair_type_2), "other", pair_type_2))

#change to factor
edge_UA_mom_juv$pair_type_2<-as.factor(edge_UA_mom_juv$pair_type_2)
edge_UA_mom_juv$pair_type_2 <- relevel(edge_UA_mom_juv$pair_type_2, ref = "other")
```

Models
```{r}
model_mom_juv_UA <- glmmTMB(weight ~ pair_type_2 + colony_size_c + (1|year) + (1|pairID),  data = edge_UA_mom_juv, family= beta_family())
summary(model_mom_juv_UA)

model_interaction_mom_juv_UA <- glmmTMB(weight ~ pair_type_2 * colony_size_c + (1|year) + (1|pairID),  data = edge_UA_mom_juv, family= beta_family())
summary(model_interaction_mom_juv_UA)
```

#check 
```{r}
check_overdispersion(model_mom_juv_UA) #no over dispersion
check_collinearity(model_mom_juv_UA) #low correlation

check_overdispersion(model_interaction_mom_juv_UA) #no over dispersion
check_collinearity(model_interaction_mom_juv_UA) #low correlation

simulationOutput<- simulateResiduals(fittedModel = model_mom_juv_UA, n = 1000)
plot(simulationOutput)
simulationOutput_int<- simulateResiduals(fittedModel = model_interaction_mom_juv_UA, n = 1000)
plot(simulationOutput_int)
```

```{r}
# influential years
model_lmer <- glmer(weight ~ pair_type + colony_size + (1 | pairID) + (1 | year),
                    data = edge_UA_mom_juv,
                    family = binomial())

infl_year <- influence(model_lmer, group = "year")
plot(infl_year, which = "cook")
```
Remove 2016
```{r}
edge_UA_mom_juv_new <- edge_UA_mom_juv %>%
  filter(year != 2016)
```

#final model
```{r}
model_final_UA <- glmmTMB(weight ~ pair_type_2 + colony_size_c + (1|year),  data = edge_UA_mom_juv_new, family= beta_family())
summary(model_final_UA)

model_final_interaction_UA <- glmmTMB(weight ~ pair_type_2 * colony_size_c + (1|year),  data = edge_UA_mom_juv_new, family= beta_family())
summary(model_final_interaction_UA)
```
#Permutation - models
Would we observe an effect of “mother–daughter” of this magnitude if mother–daughter labels were randomly assigned under the same social structure?

We can do a *label permutation*, --> Randomly reassign the “mother–daughter” labels among dyads, within year, while keeping the number of MD pairs per year fixed 

First, we can save the observed coefficients, for both models 
```{r}
#store coefficients
obs_coef_UA <- fixef(model_final_UA)$cond[c(
  "pair_type_2mother-daughter",
  "pair_type_2mother-unrelated-daughter",
  "colony_size_c"
)]

obs_coef_interaction_UA <- fixef(model_final_interaction_UA)$cond[c(
  "pair_type_2mother-daughter",
  "pair_type_2other:colony_size_c"
)]

```

*Label Permutations*
#### without interaction
```{r}
# Prepare storage
n_perm <- 50
perm_coefs_UA <- matrix(NA, nrow = n_perm, ncol = length(obs_coef_UA))

colnames(perm_coefs_UA) <- names(obs_coef_UA)

set.seed(123)

for (i in 1:n_perm) {
  # Permute pair_type2 within year
  perm_data <- edge_UA_mom_juv_new %>%
    group_by(year) %>%
    mutate(pair_type_2 = sample(pair_type_2)) %>%
    ungroup()
  
  # Refit model
  perm_model <- glmmTMB(
    weight ~ pair_type_2 + colony_size_c + (1 | pairID) + (1 | year),
    family = beta_family(),
    data = perm_data
  )
  
  # Save the coefficients of interest
  perm_coefs_UA[i, ] <- fixef(perm_model)$cond[c(
    "pair_type_2mother-daughter",
    "pair_type_2mother-unrelated-daughter",
    "colony_size_c"
  )]

}

# Convert to data frame
perm_coefs_df_UA <- as.data.frame(perm_coefs_UA)

save(perm_coefs_df_UA, file = here("results", "permutation_results_UA_mom_juv.RData"))
```

P-values 
```{r}
p_values_UA <- sapply(names(obs_coef_UA), function(name) {
  mean(perm_coefs_df_UA[[name]] >= obs_coef_UA[name])
})

#create result table 
# Convert obs_coef to a data frame
obs_df_UA <- data.frame(
  term = names(obs_coef_UA),
  observed = as.numeric(obs_coef_UA),
  stringsAsFactors = FALSE
)

# Compute permutation summary (mean, SD) for each coefficient
perm_summary_UA <- perm_coefs_df_UA %>%
  pivot_longer(cols = everything(), names_to = "term", values_to = "value") %>%
  group_by(term) %>%
  summarise(
    perm_mean = mean(value),
    perm_sd   = sd(value),
    .groups = "drop"
  )
# Combine with observed and p-values
final_table_UA <- obs_df_UA %>%
  left_join(perm_summary_UA, by = "term") %>%
  mutate(p_value = p_values_UA[term])

# Optional: order by observed effect
final_table_UA <- final_table_UA %>% arrange(desc(observed))

final_table_UA
```

### visualize 
the permutation results for each coefficient: pair type, colony size,
```{r}
perm_coefs_df_UA <- perm_coefs_df_UA %>%
  rename(
    `mother-daughter` = `pair_type_2mother-daughter`,
    `other` = `pair_type_2other`)
    
   
ggplot(perm_coefs_df_UA, aes(x = `mother-daughter`)) +
  geom_histogram(bins = 30, fill = "darkgrey", color = "black") +
  geom_vline(xintercept = obs_coef["pair_type_2mother-daughter"], color = "red", size=1.1) +
  labs(title = "Distribution of permuted coefficient - mother-daughter",
       x = "Coefficient value",
       y = "Frequency") +
  theme_minimal()

ggplot(perm_coefs_df_UA, aes(x = `other`)) +
  geom_histogram(bins = 30, fill = "darkgrey", color = "black") +
  geom_vline(xintercept = obs_coef["pair_type_2other"], color = "red", size=1.1) +
  labs(title = "Distribution of permuted coefficient - other",
       x = "Coefficient value",
       y = "Frequency") +
  theme_minimal()

ggplot(perm_coefs_df_UA, aes(x = `colony_size_c`)) +
  geom_histogram(bins = 30, fill = "darkgrey", color = "black") +
  geom_vline(xintercept = obs_coef["colony_size_c"], color = "red", size=1.1) +
  labs(title = "Distribution of permuted coefficient - colony size",
       x = "Coefficient value",
       y = "Frequency") +
  theme_minimal()

```

#### with interaction
```{r}
# Prepare storage
n_perm <- 1000
perm_coefs_interaction_UA <- matrix(NA, nrow = n_perm, ncol = length(obs_coef_interaction_UA))
colnames(perm_coefs_interaction_UA) <- names(obs_coef_interaction_UA)

set.seed(123)

for (i in 1:n_perm) {
  # Permute pair_type2 within year
  perm_data_interaction <- edge_UA_mom_juv_new %>%
    group_by(year) %>%
    mutate(pair_type2_perm = sample(pair_type_2)) %>%
    ungroup()
  
  # Refit model
  perm_model_interaction <- glmmTMB(
    weight ~ pair_type2_perm * colony_size_c + (1 | pairID) + (1 | year),
    family = beta_family(),
    data = perm_data_interaction
  )
  
  # Save the coefficients of interest
  perm_coefs_interaction_UA[i, ] <- fixef(perm_model_interaction)$cond[c(
    "pair_type2_permmother-daughter:colony_size_c",
    "pair_type2_permother:colony_size_c"
  )]
}
# Convert to data frame
perm_coefs_df_interaction_UA <- as.data.frame(perm_coefs_interaction_UA)

save(perm_coefs_df_interaction, file = here("results", "permutation_interaction_results_UA_mom_juv.RData"))
```
P-values 
```{r}
p_values_UA <- sapply(names(obs_coef_interaction_UA), function(name) {
  mean(perm_coefs_df_interaction_UA[[name]] >= obs_coef_interaction_UA[name])
})

#create result table 
# Convert obs_coef to a data frame
obs_df_interaction_UA <- data.frame(
  term = names(obs_coef_interaction_UA),
  observed = as.numeric(obs_coef_interaction_UA),
  stringsAsFactors = FALSE
)

# Compute permutation summary (mean, SD) for each coefficient
perm_summary_interaction_UA <- perm_coefs_df_interaction_UA %>%
  pivot_longer(cols = everything(), names_to = "term", values_to = "value") %>%
  group_by(term) %>%
  summarise(
    perm_mean = mean(value),
    perm_sd   = sd(value),
    .groups = "drop"
  )

# Combine with observed and p-values
final_table_interaction_UA <- obs_df_interaction_UA %>%
  left_join(perm_summary_interaction_UA, by = "term") %>%
  mutate(p_value = p_values_UA[term])

# Optional: order by observed effect
final_table_interaction_UA <- final_table_interaction_UA %>% arrange(desc(observed))

final_table_interaction_UA
```

#-------global RESULT 
# Plot - model prediction

PLOT
```{r}
library(ggeffects)
pred_GB2 <- ggpredict(model_final_interaction_GB2, terms = c("colony_size_c [-25:25 by=1]", "pair_type_2"), bias_correction = TRUE)
pred_UA <- ggpredict(model_final_UA, terms = c("colony_size_c [-10:10 by=1]", "pair_type_2"), bias_correction = TRUE)

#simple plot
plot(pred_GB2) + ggtitle("Colony GB2")
plot(pred_UA) + ggtitle("Colony UA")
```
```{r}
p1 <- ggplot(pred_GB2, aes(x = x, y = predicted, color = group, fill = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  labs(
    title = "Colony GB2",
    x = "Centered colony size",
    y = "Predicted association strength"
  ) +
  scale_color_manual(values = c("mother-daughter" = "#2A9D8F", "mother-unrelated-daughter"="goldenrod1","other" = "#8c2b94")) +
  scale_fill_manual(values  = c("mother-daughter" = "#2A9D8F", "mother-unrelated-daughter"="goldenrod1","other" = "#8c2b94")) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 13, face = "plain", hjust = 0.5)
  )
  

p2 <- ggplot(pred_UA, aes(x = x, y = predicted, color = group, fill = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  labs(
    title = "Colony UA",
    x = "Centered colony size",
    y = NULL
  ) +
  scale_color_manual(values = c("mother-daughter" = "#2A9D8F", "mother-unrelated-daughter"="goldenrod1","other" = "#8c2b94")) +
  scale_fill_manual(values  = c("mother-daughter" = "#2A9D8F", "mother-unrelated-daughter"="goldenrod1","other" = "#8c2b94")) +
    theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 13, face = "plain", hjust = 0.5)
  )


combined_plot <- p1 + p2 +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "Model-predicted association strength by colony size",
    theme = theme(
      plot.title = element_text(size = 15, face = "bold", hjust = 0.5)
    )
  ) &
  theme(legend.position = "bottom")

combined_plot
ggsave(here("results","figures","Q1_modelprediction.png"), plot = combined_plot,width = 6, height = 4, dpi = 300)
```

## QAP visualization 
### for GB2
```{r}
#remove year 2016
edges_plot_GB2 <- edges_GB2 %>%
  filter(year != "2016")

edges_plot_GB2$pair_type<-as.factor(edges_plot_GB2$pair_type)
edges_plot_GB2$pair_type <- relevel(edges_plot_GB2$pair_type, ref = "other")

# re scale to fit the beta distribution
edges_plot_GB2$weight <- (edges_plot_GB2$weight * (nrow(edges_plot_GB2) - 1) + 0.5) / nrow(edges_plot_GB2)

edges_plot_GB2$year <- as.character(edges_plot_GB2$year)

sig_labels <- QAP_GB2 %>%
  mutate(significant = pval_upper < 0.05,
         #star according to the p-value
         star = ifelse(pval_upper < 0.001, "***",
                       ifelse(pval_upper < 0.01, "**",
                              ifelse(pval_upper < 0.05, "*", "")))) %>%
  select(year, star)

# Merge with edges_GB2 to get positions for the stars
# We'll place the star slightly above the max observed weight for each year
star_positions <- edges_plot_GB2 %>%
  group_by(year) %>%
  summarise(y_pos = max(weight, na.rm = TRUE) + 0.05, .groups = "drop") %>%
  left_join(sig_labels, by = "year")

# Plot: boxplots of association strength by year and pair type
QAP_plot_GB2 <- ggplot(edges_plot_GB2, aes(x = factor(year), y = weight, fill = pair_type)) +
  geom_boxplot(width = 0.6, outlier.size = 0.5, outlier.colour = "#b69cb8") +
  # Add stars for significance above boxes
  geom_text(data = star_positions, aes(x = factor(year), y = 0.97, label = star),
            inherit.aes = FALSE, size = 5) +
  scale_fill_manual(values = c("mother-daughter" = "#98b8b4", "other" = "#b69cb8")) +
  labs(
    x = "Year",
    y = NULL,
    fill = "pair type",
    title = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    axis.title.x = element_blank(),
    plot.margin = margin(3, 5, 3, 5),
    panel.spacing = unit(0.2, "lines"),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5)
  )

QAP_plot_GB2

ggsave(here("results","figures","Q1_boxplot_QAP_GB2.png"), width = 6, height = 4, dpi = 300)
```

### for UA
```{r}
edges_UA$pair_type<-as.factor(edges_UA$pair_type)
edges_UA$pair_type <- relevel(edges_UA$pair_type, ref = "other")

edges_plot_UA<-edges_UA

edges_plot_UA$year <- as.character(edges_UA$year)

sig_labels_UA <- QAP_UA %>%
  mutate(significant = pval_upper < 0.05,
         #star according to the p-value
         star = ifelse(pval_upper < 0.001, "***",
                       ifelse(pval_upper < 0.01, "**",
                              ifelse(pval_upper < 0.05, "*", "")))) %>%
  select(year, star)

# Merge with edges_GB2 to get positions for the stars
# We'll place the star slightly above the max observed weight for each year
star_positions_UA <- edges_plot_UA %>%
  group_by(year) %>%
  summarise(y_pos = max(weight, na.rm = TRUE) + 0.05, .groups = "drop") %>%
  left_join(sig_labels_UA, by = "year")

# Plot: boxplots of association strength by year and pair type
QAP_plot_UA <- ggplot(edges_plot_UA, aes(x = factor(year), y = weight, fill = pair_type)) +
  geom_boxplot(width = 0.6, outlier.size = 0.5, outlier.colour = "#b69cb8") +
  # Add stars for significance above boxes
  geom_text(data = star_positions_UA, aes(x = factor(year), y = 0.97, label = star),
            inherit.aes = FALSE, size = 5) +
  scale_fill_manual(values = c("mother-daughter" = "#98b8b4", "other" = "#b69cb8")) +
  labs(
    x = "Year",
    y = NULL,
    fill = "pair type",
    title = ""
  ) +
  theme_minimal(base_size = 12) +
 theme(
    legend.position = "bottom",
    axis.title.x = element_blank(),
    plot.margin = margin(3, 5, 3, 5),
    panel.spacing = unit(0.2, "lines"),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5)
  )
ggsave(here("results","figures","Q1_boxplot_QAP_UA.png"), width = 6, height = 4, dpi = 300)
```

Combined plot QAP, one on top and one at the bottom, and add only one y axis for both plots
```{r}
#create the y label
y_label_plot <- ggplot() +
  annotate("text", x = 1, y = 1, label = "Association strength (SRI)", angle = 90, size = 4) +
  theme_void()

combined_QAP_plot <-
  y_label_plot + (QAP_plot_GB2 / QAP_plot_UA) +
  plot_layout(widths = c(0.05, 1), guides = "collect") &
  plot_annotation(
    title = "Association strength by year and pair type (QAP results)",
    theme = theme(
      plot.title = element_text(size = 15, face = "bold", hjust = 0.3, vjust = -3
    )
  ) &
 theme(legend.position = "bottom"))

combined_QAP_plot

ggsave(here("results","figures","Q1_boxplot_QAP_combined.png"), plot = combined_QAP_plot,width = 7, height = 5, dpi = 300)
```

# Mean network density for all years
```{r}
mean_density_GB2 <- sapply(net_list_GB2, function(net) {
  graph <- graph_from_adjacency_matrix(net, mode = "undirected", weighted = TRUE, diag = FALSE)
  edge_density(graph, loops = FALSE)
})
mean(mean_density_GB2)

mean_density_UA <- sapply(net_list_UA, function(net) {
  graph <- graph_from_adjacency_matrix(net, mode = "undirected", weighted = TRUE, diag = FALSE)
  edge_density(graph, loops = FALSE)
})
mean(mean_density_UA)
```

#Number of md pairs for each year
```{r}
md_pairs_count_GB2 <- edges_GB2 %>%
  filter(pair_type == "mother-daughter") %>%
  group_by(year) %>%
  summarise(md_count = n(), .groups = "drop")

md_pairs_count_GB2

md_pairs_count_UA <- edges_UA %>%
  filter(pair_type == "mother-daughter") %>%
  group_by(year) %>%
  summarise(md_count = n(), .groups = "drop")

md_pairs_count_UA
```
#colony size for each year
```{r}
colony_size_summary_GB2 <- edges_GB2 %>%
  group_by(year) %>%
  summarise(mean_colony_size = mean(colony_size, na.rm = TRUE), .groups = "drop")
colony_size_summary_GB2
colony_size_summary_UA <- edges_UA %>%
  group_by(year) %>%
  summarise(mean_colony_size = mean(colony_size, na.rm = TRUE), .groups = "drop")
colony_size_summary_UA
```
#compute for each year the mean SRI for mother-daughter (+ SD) and for other pairs
```{r}
sri_summary_GB2 <- edges_GB2 %>%
  group_by(year, pair_type) %>%
  summarise(mean_sri = mean(weight, na.rm = TRUE), sd_sri = sd(weight, na.rm = TRUE), .groups = "drop")
sri_summary_GB2
sri_summary_UA <- edges_UA %>%
  group_by(year, pair_type) %>%
  summarise(mean_sri = mean(weight, na.rm = TRUE), sd_sri = sd(weight, na.rm = TRUE), .groups = "drop")
sri_summary_UA
```

