---
title: "Master Thesis - General analysis"
author: "Flore Michel"
date: "2025-02-17"
output: html_document
---
```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir = "C:/Users/flo_r/Desktop/master thesis/data/data in CSV/")
knitr::opts_knit$set(root.dir = "/Users/floremichel/Desktop/THESIS/analysis/")

library(igraph)
library(asnipe)
library(sna)
library(assocInd)
library(dplyr)
#library(lme4)
library(lmerTest)
library(ggplot2)
library(tidyverse)
library(viridis)
library(patchwork)
library(hrbrthemes)
library(igraph)
library(ggraph)
library(colormap)
library(tidyr)
library(dplyr)
library(data.table)
library(tidygraph)
library(RColorBrewer)
library(threejs)
library(patchwork)
```

```{r import, include=F}
net_G2013<-read.csv("variables_GB2/2013/net_G2013.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_G2013 <- as.matrix(net_G2013)
net_G2014<-read.csv("variables_GB2/2014/net_G2014.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2014 <- as.matrix(net_G2014)
net_G2015<-read.csv("variables_GB2/2015/net_G2015.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2015 <- as.matrix(net_G2015)
net_G2016<-read.csv("variables_GB2/2016/net_G2016.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2016 <- as.matrix(net_G2016)
net_G2017<-read.csv("variables_GB2/2017/net_G2017.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2017 <- as.matrix(net_G2017)
net_G2018<-read.csv("variables_GB2/2018/net_G2018.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2018 <- as.matrix(net_G2018)
net_G2019<-read.csv("variables_GB2/2019/net_G2019.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2019 <- as.matrix(net_G2019)
net_G2020<-read.csv("variables_GB2/2020/net_G2020.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2020 <- as.matrix(net_G2020)
net_G2021<-read.csv("variables_GB2/2021/net_G2021.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2021 <- as.matrix(net_G2021)
net_G2022<-read.csv("variables_GB2/2022/net_G2022.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2022 <- as.matrix(net_G2022)
net_G2023<-read.csv("variables_GB2/2023/net_G2023.csv",row.names=1,header=TRUE, sep=",",check.names = F); net_G2023 <- as.matrix(net_G2023)
```

## First visualization of the network
```{r}
par(mfrow=c(1,2))
par(mar=c(1,1,1,1))
g_G2013<-graph_from_adjacency_matrix(net_G2013,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2013,vertex.label = V(g_G2013)$name,edge.width = E(g_G2013)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2013") 
#graphjs(g_G2013) plot interactif
g_G2014<-graph_from_adjacency_matrix(net_G2014,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2014,vertex.label = V(g_G2014)$name,edge.width = E(g_G2013)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2014") 
g_G2015<-graph_from_adjacency_matrix(net_G2015,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2015,vertex.label = V(g_G2015)$name,edge.width = E(g_G2015)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2015")
g_G2016<-graph_from_adjacency_matrix(net_G2016,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2016,vertex.label = V(g_G2016)$name,edge.width = E(g_G2016)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2016")
g_G2017<-graph_from_adjacency_matrix(net_G2017,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2017,vertex.label = V(g_G2017)$name,edge.width = E(g_G2017)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2017")
g_G2018<-graph_from_adjacency_matrix(net_G2018,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2018,vertex.label = V(g_G2018)$name,edge.width = E(g_G2018)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2018")
g_G2019<-graph_from_adjacency_matrix(net_G2019,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2019,vertex.label = V(g_G2019)$name,edge.width = E(g_G2019)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2019")
g_G2020<-graph_from_adjacency_matrix(net_G2020,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2020,vertex.label = V(g_G2020)$name,edge.width = E(g_G2020)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2020")
g_G2021<-graph_from_adjacency_matrix(net_G2021,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2021,vertex.label = V(g_G2021)$name,edge.width = E(g_G2021)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2021")
g_G2022<-graph_from_adjacency_matrix(net_G2022,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2022,vertex.label = V(g_G2022)$name,edge.width = E(g_G2022)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2022")
g_G2023<-graph_from_adjacency_matrix(net_G2023,mode = "undirected",weighted = TRUE,diag = FALSE)
plot(g_G2023,vertex.label = V(g_G2023)$name,edge.width = E(g_G2023)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",main="Network GB2 2023")
```
Better visualization with ggraph package: 
```{r}
par(mfrow=c(1,2))
par(mar=c(1,1,1,1))

ggraph(g_G2013, layout = "fr") +
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2013")
ggraph(g_G2014, layout = "fr") +
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2014")
ggraph(g_G2015, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2015")
ggraph(g_G2016, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2016")
ggraph(g_G2017, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2017")
ggraph(g_G2018, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2018")
ggraph(g_G2019, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2019")
ggraph(g_G2020, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2020")
ggraph(g_G2021, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2021")
ggraph(g_G2022, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2022")
ggraph(g_G2023, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Network - 2023")

```
We can see that for each year, the networks are very dense because every individual is associated at least once with another individual. 
To enhance visualization, we can filter the network to keep only the strongest associations.To do so, we can add a threshold of the mean association in the network. 

### Thresholding the network
```{r fig.width=10, fig.height=6}

par(mfrow=c(1,2))
threshold<-function(net){
  mean_weight<-mean(net[net!=0])
  net[net<mean_weight]<-0
  return(net)
}
filternet_G2013<-threshold(net_G2013)
g_filter_G2013<-graph_from_adjacency_matrix(filternet_G2013,mode = "undirected",weighted = TRUE,diag = FALSE) 
ggraph(filternet_G2013, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2013")
filternet_G2014<-threshold(net_G2014)
g_filter_G2014<-graph_from_adjacency_matrix(filternet_G2014,mode = "undirected",weighted = TRUE,diag = FALSE) 
ggraph(filternet_G2014, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2014")
filternet_G2015<-threshold(net_G2015)
g_filter_G2015<-graph_from_adjacency_matrix(filternet_G2015,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(filternet_G2015, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2015")
filternet_G2016<-threshold(net_G2016)
g_filter_G2016<-graph_from_adjacency_matrix(filternet_G2016,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(filternet_G2016, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2016")
filternet_G2017<-threshold(net_G2017)
g_filter_G2017<-graph_from_adjacency_matrix(filternet_G2017,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(filternet_G2017, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2017")
filternet_G2018<-threshold(net_G2018)
g_filter_G2018<-graph_from_adjacency_matrix(filternet_G2018,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(filternet_G2018, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2018")
filternet_G2019<-threshold(net_G2019)
g_filter_G2019<-graph_from_adjacency_matrix(filternet_G2019,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(filternet_G2019, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2019")
filternet_G2020<-threshold(net_G2020)
g_filter_G2020<-graph_from_adjacency_matrix(filternet_G2020,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(filternet_G2020, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2020")
filternet_G2021<-threshold(net_G2021)
g_filter_G2021<-graph_from_adjacency_matrix(filternet_G2021,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(filternet_G2021, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2021")
filternet_G2022<-threshold(net_G2022)
g_filter_G2022<-graph_from_adjacency_matrix(filternet_G2022,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(filternet_G2022, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2022")

filternet_G2023<-threshold(net_G2023)
g_filter_G2023<-graph_from_adjacency_matrix(filternet_G2023,mode = "undirected",weighted = TRUE,diag = FALSE)
ggraph(g_G2013, layout = "fr") + 
  geom_edge_link(aes(width = weight), alpha = 0.3, color="darkgrey") +
  scale_edge_width(range = c(0.1, 2))+ #smaller weight 
  geom_node_point(color = "cornflowerblue", size = 5) +  
  geom_node_text(aes(label = name), repel = TRUE, size = 4,fontface="bold") + 
  theme_void() +
  ggtitle("Filtered network - 2023")

```
## Vertex Attributes
We can add a new vertex attributes "mother-daughter" to discriminate our pairs of interest with the others.

```{r import, include = FALSE}
mother_daughter_pairs_G2013<-read.csv("variables_GB2/2013/mother_daughter_pairs_G2013.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character");mother_daughter_pairs_G2014<-read.csv("variables_GB2/2014/mother_daughter_pairs_G2014.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2015<-read.csv("variables_GB2/2015/mother_daughter_pairs_G2015.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2016<-read.csv("variables_GB2/2016/mother_daughter_pairs_G2016.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2017<-read.csv("variables_GB2/2017/mother_daughter_pairs_G2017.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2018<-read.csv("variables_GB2/2018/mother_daughter_pairs_G2018.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2019<-read.csv("variables_GB2/2019/mother_daughter_pairs_G2019.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2020<-read.csv("variables_GB2/2020/mother_daughter_pairs_G2020.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2021<-read.csv("variables_GB2/2021/mother_daughter_pairs_G2021.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2022<-read.csv("variables_GB2/2022/mother_daughter_pairs_G2022.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character"); mother_daughter_pairs_G2023<-read.csv("variables_GB2/2023/mother_daughter_pairs_G2023.csv",header=TRUE, row.names=NULL,sep=",",check.names = F,colClasses = "character")

```

```{r}
edges_2023 <- edges_GB2 %>% filter(year == 2023)

# Plot histogram
p <-ggplot(edges_2023, aes(x = weight)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black") +
  labs(title = "Distribution of association strengh in 2023",
       x = "SRI values",
       y = "Frequency") +
  theme_minimal()

ggsave("SRI_2023_histogram.png", plot = p, width = 4, height = 3, dpi = 300)
```


```{r add vertex attribute}
highlight_mother_daughter <- function(graph, mother_daughter_pairs) {
  V(graph)$type <- "other"
  
  mother_ids <- mother_daughter_pairs$rfid_mom
  daughter_ids <- mother_daughter_pairs$rfid_off
  
  V(graph)[V(graph)$name %in% mother_ids]$type <- "mother-daughter"
  V(graph)[V(graph)$name %in% daughter_ids]$type <- "mother-daughter"
  return(graph)
}

g_G2013 <- highlight_mother_daughter(g_G2013, mother_daughter_pairs_G2013);g_G2014 <-highlight_mother_daughter(g_G2014, mother_daughter_pairs_G2014); g_G2015 <-highlight_mother_daughter(g_G2015, mother_daughter_pairs_G2015); g_G2016 <-highlight_mother_daughter(g_G2016, mother_daughter_pairs_G2016); g_G2017 <-highlight_mother_daughter(g_G2017, mother_daughter_pairs_G2017); g_G2018 <-highlight_mother_daughter(g_G2018, mother_daughter_pairs_G2018); g_G2019 <-highlight_mother_daughter(g_G2019, mother_daughter_pairs_G2019); g_G2020 <-highlight_mother_daughter(g_G2020, mother_daughter_pairs_G2020); g_G2021 <-highlight_mother_daughter(g_G2021, mother_daughter_pairs_G2021); g_G2022 <-highlight_mother_daughter(g_G2022, mother_daughter_pairs_G2022); g_G2023 <-highlight_mother_daughter(g_G2023, mother_daughter_pairs_G2023)
```
### Plot the mother-daughter pairs in the network: 
```{r plot md}
par(mfrow=c(2,2))

g_G2013<- as_tbl_graph(filternet_G2013) %>%
  mutate(type = V(g_G2013)$type) 

E(g_G2013)$highlight <- ifelse(
  ends(g_G2013, E(g_G2013))[,1] %in% mother_daughter_pairs_G2013$mother & 
  ends(g_G2013, E(g_G2013))[,2] %in% mother_daughter_pairs_G2013$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2013, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2013")

g_G2014<- as_tbl_graph(filternet_G2014) %>%
  mutate(type = V(g_G2014)$type) 

E(g_G2014)$highlight <- ifelse(
  ends(g_G2014, E(g_G2014))[,1] %in% mother_daughter_pairs_G2014$mother & 
  ends(g_G2014, E(g_G2014))[,2] %in% mother_daughter_pairs_G2014$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2014, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2014")

g_G2015 <- as_tbl_graph(filternet_G2015) %>%
  mutate(type = V(g_G2015)$type) 

E(g_G2015)$highlight <- ifelse(
  ends(g_G2015, E(g_G2015))[,1] %in% mother_daughter_pairs_G2015$mother & 
  ends(g_G2015, E(g_G2015))[,2] %in% mother_daughter_pairs_G2015$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2015, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2015")

g_G2016 <- as_tbl_graph(filternet_G2016) %>%
  mutate(type = V(g_G2016)$type) 

E(g_G2016)$highlight <- ifelse(
  ends(g_G2016, E(g_G2016))[,1] %in% mother_daughter_pairs_G2016$mother & 
  ends(g_G2016, E(g_G2016))[,2] %in% mother_daughter_pairs_G2016$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2016, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2016")

g_G2017 <- as_tbl_graph(filternet_G2017) %>% 
  mutate(type = V(g_G2017)$type)

E(g_G2017)$highlight <- ifelse(
  ends(g_G2017, E(g_G2017))[,1] %in% mother_daughter_pairs_G2017$mother & 
  ends(g_G2017, E(g_G2017))[,2] %in% mother_daughter_pairs_G2017$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2017, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2017")

g_G2018 <- as_tbl_graph(filternet_G2018) %>%
  mutate(type = V(g_G2018)$type)

E(g_G2018)$highlight <- ifelse(
  ends(g_G2018, E(g_G2018))[,1] %in% mother_daughter_pairs_G2018$mother & 
  ends(g_G2018, E(g_G2018))[,2] %in% mother_daughter_pairs_G2018$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2018, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2018")

g_G2019 <- as_tbl_graph(filternet_G2019) %>%
  mutate(type = V(g_G2019)$type)

E(g_G2019)$highlight <- ifelse(
  ends(g_G2019, E(g_G2019))[,1] %in% mother_daughter_pairs_G2019$mother & 
  ends(g_G2019, E(g_G2019))[,2] %in% mother_daughter_pairs_G2019$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2019, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2019")

g_G2020 <- as_tbl_graph(filternet_G2020) %>%
  mutate(type = V(g_G2020)$type)

E(g_G2020)$highlight <- ifelse(
  ends(g_G2020, E(g_G2020))[,1] %in% mother_daughter_pairs_G2020$mother & 
  ends(g_G2020, E(g_G2020))[,2] %in% mother_daughter_pairs_G2020$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2020, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2020")

g_G2021 <- as_tbl_graph(filternet_G2021) %>%
  mutate(type = V(g_G2021)$type)

E(g_G2021)$highlight <- ifelse(
  ends(g_G2021, E(g_G2021))[,1] %in% mother_daughter_pairs_G2021$mother & 
  ends(g_G2021, E(g_G2021))[,2] %in% mother_daughter_pairs_G2021$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2021, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2021")

g_G2022 <- as_tbl_graph(filternet_G2022) %>%
  mutate(type = V(g_G2022)$type)

E(g_G2022)$highlight <- ifelse(
  ends(g_G2022, E(g_G2022))[,1] %in% mother_daughter_pairs_G2022$mother & 
  ends(g_G2022, E(g_G2022))[,2] %in% mother_daughter_pairs_G2022$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2022, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2022")

g_G2023_tbl <- as_tbl_graph(g_G2013)

E(g_G2023_tbl)$highlight <- ifelse(
  ends(g_G2023_tbl, E(g_G2023_tbl))[,1] %in% mother_daughter_pairs_G2023$mother & 
  ends(g_G2023_tbl, E(g_G2023_tbl))[,2] %in% mother_daughter_pairs_G2023$daughter, 
  "mother-daughter", "other"
)

ggraph(g_G2023_tbl, layout = "kk") +  
  geom_edge_link(aes(width = weight, color = highlight), alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  scale_edge_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_point(aes(color = type), size = 4) +  
  scale_color_manual(values = c("mother-daughter" = "darkcyan", "other" = "gray")) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2023")
```

## Creation of a pair label 
This will allow us to have a unique identifyer for every year.

```{r assign pair ID}
assign_pair_IDs <- function(graph, mother_daughter_pairs) {
  V(graph)$pairID <- NA

  for (i in 1:nrow(mother_daughter_pairs)) {
    mother <- mother_daughter_pairs$rfid_mom[i]
    daughter <- mother_daughter_pairs$rfid_off[i]

    if (mother %in% V(graph)$name & daughter %in% V(graph)$name) {
      pair_name <- paste(mother, daughter, sep = "-")
      
      V(graph)$pairID[V(graph)$name == mother] <- pair_name
      V(graph)$pairID[V(graph)$name == daughter] <- pair_name
    }
  }
  
  return(graph)
}

g_G2013 <- assign_pair_IDs(g_G2013,mother_daughter_pairs_G2013); g_G2014 <- assign_pair_IDs(g_G2014,mother_daughter_pairs_G2014); g_G2015 <- assign_pair_IDs(g_G2015,mother_daughter_pairs_G2015); g_G2016 <- assign_pair_IDs(g_G2016,mother_daughter_pairs_G2016); g_G2017 <- assign_pair_IDs(g_G2017,mother_daughter_pairs_G2017); g_G2018 <- assign_pair_IDs(g_G2018,mother_daughter_pairs_G2018); g_G2019 <- assign_pair_IDs(g_G2019,mother_daughter_pairs_G2019); g_G2020 <- assign_pair_IDs(g_G2020,mother_daughter_pairs_G2020); g_G2021 <- assign_pair_IDs(g_G2021,mother_daughter_pairs_G2021); g_G2022 <- assign_pair_IDs(g_G2022,mother_daughter_pairs_G2022); g_G2023 <- assign_pair_IDs(g_G2023,mother_daughter_pairs_G2023)
```

## Plot the network with mother-daughter pairs highlighted
```{r}
pair_IDs_G2013<-factor(V(g_G2013)$pairID); n_pairs_G2013 <- length(levels(na.omit(pair_IDs_G2013))); my_palette_G2013 <- brewer.pal(n = max(3, n_pairs_G2013), name = "Set2")

ggraph(g_G2013, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2013) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2013")

pair_IDs_G2014<-factor(V(g_G2014)$pairID); n_pairs_G2014 <- length(levels(na.omit(pair_IDs_G2014))); my_palette_G2014 <- brewer.pal(n = max(3, n_pairs_G2014), name = "Set2")
ggraph(g_G2014, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2014) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2014")

pair_IDs_G2015<-factor(V(g_G2015)$pairID); n_pairs_G2015 <- length(levels(na.omit(pair_IDs_G2015))); n_pairs_G2015 <- length(unique(na.omit(V(g_G2015)$pairID)));my_palette_G2015 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2015);pair_IDs_G2015 <- factor(V(g_G2015)$pairID)
pair_colors_G2015 <- setNames(my_palette_G2015, levels(pair_IDs_G2015))
ggraph(g_G2015, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2015) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2015")


pair_IDs_G2016<-factor(V(g_G2016)$pairID); n_pairs_G2016 <- length(levels(na.omit(pair_IDs_G2016))); n_pairs_G2016 <- length(unique(na.omit(V(g_G2016)$pairID)));my_palette_G2016 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2016);pair_IDs_G2016 <- factor(V(g_G2016)$pairID)
pair_colors_G2016 <- setNames(my_palette_G2016, levels(pair_IDs_G2016))
ggraph(g_G2016, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2016) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2016")

pair_IDs_G2017<-factor(V(g_G2017)$pairID); n_pairs_G2017 <- length(levels(na.omit(pair_IDs_G2017))); n_pairs_G2017 <- length(unique(na.omit(V(g_G2017)$pairID)));my_palette_G2017 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2017);pair_IDs_G2017 <- factor(V(g_G2017)$pairID)
pair_colors_G2017 <- setNames(my_palette_G2017, levels(pair_IDs_G2017))
ggraph(g_G2017, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2017) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2017")

pair_IDs_G2018<-factor(V(g_G2018)$pairID); n_pairs_G2018 <- length(levels(na.omit(pair_IDs_G2018))); n_pairs_G2018 <- length(unique(na.omit(V(g_G2018)$pairID)));my_palette_G2018 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2018);pair_IDs_G2018 <- factor(V(g_G2018)$pairID)
pair_colors_G2018 <- setNames(my_palette_G2018, levels(pair_IDs_G2018))

pair_IDs_G2018<-factor(V(g_G2018)$pairID); n_pairs_G2018 <- length(levels(na.omit(pair_IDs_G2018))); my_palette_G2018 <- brewer.pal(n = max(3, n_pairs_G2018), name = "Set2")
ggraph(g_G2018, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2018) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2018")

pair_IDs_G2019<-factor(V(g_G2019)$pairID); n_pairs_G2019 <- length(levels(na.omit(pair_IDs_G2019))); n_pairs_G2019<- length(unique(na.omit(V(g_G2019)$pairID)));my_palette_G2019 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2019);pair_IDs_G2019 <- factor(V(g_G2019)$pairID)
pair_colors_G2019 <- setNames(my_palette_G2019, levels(pair_IDs_G2019))
ggraph(g_G2019, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2019) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2019")

pair_IDs_G2020<-factor(V(g_G2020)$pairID); n_pairs_G2020 <- length(levels(na.omit(pair_IDs_G2020))); n_pairs_G2020<- length(unique(na.omit(V(g_G2020)$pairID)));my_palette_G2020 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2020);pair_IDs_G2020 <- factor(V(g_G2020)$pairID)
pair_colors_G2020 <- setNames(my_palette_G2020, levels(pair_IDs_G2020))
ggraph(g_G2020, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2020) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2020")

pair_IDs_G2021<-factor(V(g_G2021)$pairID); n_pairs_G2021 <- length(levels(na.omit(pair_IDs_G2021))); n_pairs_G2021<- length(unique(na.omit(V(g_G2021)$pairID)));my_palette_G2021 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2021);pair_IDs_G2021 <- factor(V(g_G2021)$pairID)
pair_colors_G2021 <- setNames(my_palette_G2021, levels(pair_IDs_G2021))
ggraph(g_G2021, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2021) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2021")

pair_IDs_G2022<-factor(V(g_G2022)$pairID); n_pairs_G2022 <- length(levels(na.omit(pair_IDs_G2022))); n_pairs_G2022<- length(unique(na.omit(V(g_G2022)$pairID)));my_palette_G2022 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2022);pair_IDs_G2022 <- factor(V(g_G2022)$pairID)
pair_colors_G2022 <- setNames(my_palette_G2022, levels(pair_IDs_G2022))
ggraph(g_G2022, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2022) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2022")

pair_IDs_G2023<-factor(V(g_G2023)$pairID); n_pairs_G2023 <- length(levels(na.omit(pair_IDs_G2023))); n_pairs_G2023<- length(unique(na.omit(V(g_G2023)$pairID)));my_palette_G2023 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_G2023);pair_IDs_G2023 <- factor(V(g_G2023)$pairID)
pair_colors_G2023 <- setNames(my_palette_G2023, levels(pair_IDs_G2023))
p<-ggraph(g_G2023, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_G2023) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2023")
ggsave("Mother-daughter_network_2023.png", plot = p, width = 4, height = 3, dpi = 300)

pair_IDs_G2023<-factor(V(g_G2023_tbl)$pairID);
ggraph(g_G2023_tbl, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pair_IDs_G2023), size = 4) +  
  scale_color_manual(values = my_palette_G2023) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2023")

```

```{r}
V(g_G2023)$pairID <- g_G2023_tbl$pairID[match(V(g_G2023)$name, g_G2023$name)]
p_network <- ggraph(g_G2023, layout = "nicely") +
  geom_edge_link(aes(width = weight), color = "grey70", alpha = 0.3) +
  scale_edge_width(range = c(0.1, 1.5)) +
  geom_node_point(aes(color = pairID), size = 2.5) +
  scale_color_manual(values = my_palette_G2023) +
  geom_node_text(
    data = subset(g_G2023_tbl, !is.na(pairID)),
    aes(label = name),
    repel = TRUE,
    fontface = "bold", size = 2.5
  ) +
  theme_void() +
  ggtitle("Mother-Daughter Network - 2023") +
  theme(plot.title = element_text(hjust = 0.1))

# --- SRI histogram
p_hist <- ggplot(edges_G2023, aes(x = weight)) +
  geom_histogram(bins = 30, fill = "black", color = "white") +
  labs(x = "SRI", y = "Frequency", title = "SRI distribution") +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 8)
  )

# --- Combine plots: histogram to the right with smaller width
final_plot <- p_network + p_hist + plot_layout(widths = c(3, 1))

# --- Export high-quality plot (e.g., PNG)
ggsave("network_2023_clean.png", plot = final_plot, width = 11, height = 6, dpi = 300)
```
#plot 2023 GB2
```{r}
p1<-ggraph(g_G2023, layout = "nicely") +
  geom_edge_link(aes(width = weight), color = "grey70", alpha = 0.3, show.legend = FALSE) +
  scale_edge_width(range = c(0.5, 4)) +
  geom_node_point(aes(color = pairID), size = 4) +
  scale_color_manual(values = my_palette_G2023) +
  theme_void() +
  ggtitle("GB2 - 2023") +
  theme(
    plot.title = element_text(hjust = 0.1),
    legend.position = "right",
    legend.title = element_blank()
  )

p2<-ggplot(edges_2023, aes(x = weight)) +
  geom_histogram(bins = 30, fill = "black", color = "white") +
   geom_vline(aes(xintercept = mean(weight)), 
             color = "red", size = 1) +
  labs(x = "SRI", y = "Frequency", title = "SRI distribution") +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 8)
  )
final_plot <- p1 + p2 + 
  plot_layout(ncol = 2, widths = c(3, 1))  # 3:1 ratio between network and histogram

final_plot
ggsave("network_2023.png", plot = final_plot,width = 7, height = 5, dpi = 300)
```




#import UA
```{r}
net_U2013<-read.csv("variables UA/net_U2013.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2013 <- as.matrix(net_U2013)
net_U2014<-read.csv("variables UA/net_U2014.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2014 <- as.matrix(net_U2014)
net_U2015<-read.csv("variables UA/net_U2015.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2015 <- as.matrix(net_U2015)
net_U2016<-read.csv("variables UA/net_U2016.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2016 <- as.matrix(net_U2016)
net_U2017<-read.csv("variables UA/net_U2017.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2017 <- as.matrix(net_U2017)
net_U2018<-read.csv("variables UA/net_U2018.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2018 <- as.matrix(net_U2018)
net_U2019<-read.csv("variables UA/net_U2019.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2019 <- as.matrix(net_U2019)
net_U2020<-read.csv("variables UA/net_U2020.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2020 <- as.matrix(net_U2020)
net_U2021<-read.csv("variables UA/net_U2021.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2021 <- as.matrix(net_U2021)
net_U2022<-read.csv("variables UA/net_U2022.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2022 <- as.matrix(net_U2022)
net_U2023<-read.csv("variables UA/net_U2023.csv",row.names=1,header=TRUE,sep=",",check.names = F);net_U2023 <- as.matrix(net_U2023)

g_U2023<-graph_from_adjacency_matrix(net_U2023,mode = "undirected",weighted = TRUE,diag = FALSE)
```

```{r}
mother_daughter_UA <- read.csv("variables UA/mother_daughter_UA.csv", header = TRUE, sep = ",")
```
```{r add vertex attribute}
highlight_mother_daughter_UA <- function(graph, mother_daughter_df, year) {

  year_pairs <- mother_daughter_df %>% 
    filter(year == !!year)

  V(graph)$type <- "other"
  

  mother_ids   <- year_pairs$rfid_mom
  daughter_ids <- year_pairs$rfid_off
  
  V(graph)[V(graph)$name %in% mother_ids]$type   <- "mother-daughter"
  V(graph)[V(graph)$name %in% daughter_ids]$type <- "mother-daughter"
  
  return(graph)
}
```

```{r}
g_U2013 <- highlight_mother_daughter_UA(g_U2013, mother_daughter_UA, 2013)
g_U2014 <- highlight_mother_daughter_UA(g_U2014, mother_daughter_UA, 2014)
g_U2015 <- highlight_mother_daughter_UA(g_U2015, mother_daughter_UA, 2015)
g_U2016 <- highlight_mother_daughter_UA(g_U2016, mother_daughter_UA, 2016)
g_U2017 <- highlight_mother_daughter_UA(g_U2017, mother_daughter_UA, 2017)
g_U2018 <- highlight_mother_daughter_UA(g_U2018, mother_daughter_UA, 2018)
g_U2019 <- highlight_mother_daughter_UA(g_U2019, mother_daughter_UA, 2019)
g_U2020 <- highlight_mother_daughter_UA(g_U2020, mother_daughter_UA, 2020)
g_U2021 <- highlight_mother_daughter_UA(g_U2021, mother_daughter_UA, 2021)
g_U2022 <- highlight_mother_daughter_UA(g_U2022, mother_daughter_UA, 2022)
g_U2023 <- highlight_mother_daughter_UA(g_U2023, mother_daughter_UA, 2023)
```

```{r}
assign_pair_IDs_UA <- function(graph, mother_daughter_df, year) {

  year_pairs <- mother_daughter_df %>% 
    filter(year == !!year)
  
  V(graph)$pairID <- NA
  
  for (i in 1:nrow(year_pairs)) {
    mother   <- year_pairs$rfid_mom[i]
    daughter <- year_pairs$rfid_off[i]  
    
    if (mother %in% V(graph)$name & daughter %in% V(graph)$name) {
      pair_name <- paste(mother, daughter, sep = "-")

      V(graph)$pairID[V(graph)$name == mother]   <- pair_name
      V(graph)$pairID[V(graph)$name == daughter] <- pair_name
    }
  }
  
  return(graph)
}

g_U2013 <- assign_pair_IDs_UA(g_U2013,mother_daughter_UA,2013); g_U2014 <- assign_pair_IDs_UA(g_U2014,mother_daughter_UA,2014); g_U2015 <- assign_pair_IDs_UA(g_U2015,mother_daughter_UA,2015); g_U2016 <- assign_pair_IDs_UA(g_U2016,mother_daughter_UA,2016); g_U2017 <- assign_pair_IDs_UA(g_U2017,mother_daughter_UA,2017); g_U2018 <- assign_pair_IDs_UA(g_U2018,mother_daughter_UA,2018); g_U2019 <- assign_pair_IDs_UA(g_U2019,mother_daughter_UA,2019); g_U2020 <- assign_pair_IDs_UA(g_U2020,mother_daughter_UA,2020); g_U2021 <- assign_pair_IDs_UA(g_U2021,mother_daughter_UA,2021); g_U2022 <- assign_pair_IDs_UA(g_U2022,mother_daughter_UA,2022); g_U2023 <- assign_pair_IDs_UA(g_U2023,mother_daughter_UA,2023)

```

#plot
```{r}
pair_IDs_U2013<-factor(V(g_U2013)$pairID); n_pairs_U2013 <- length(levels(na.omit(pair_IDs_U2013))); my_palette_U2013 <- brewer.pal(n = max(3, n_pairs_U2013), name = "Set2")

ggraph(g_U2013, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2013) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2013")

pair_IDs_U2014<-factor(V(g_U2014)$pairID); n_pairs_U2014 <- length(levels(na.omit(pair_IDs_U2014))); my_palette_U2014 <- brewer.pal(n = max(3, n_pairs_U2014), name = "Set2"); ggraph(g_U2014, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2014) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2014")
pair_IDs_U2015<-factor(V(g_U2015)$pairID); n_pairs_U2015 <- length(levels(na.omit(pair_IDs_U2015))); n_pairs_U2015 <- length(unique(na.omit(V(g_U2015)$pairID)));my_palette_U2015 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2015);pair_IDs_U2015 <- factor(V(g_U2015)$pairID)
ggraph(g_U2015, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2015) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2015")

pair_IDs_U2016<-factor(V(g_U2016)$pairID); n_pairs_U2016 <- length(levels(na.omit(pair_IDs_U2016))); n_pairs_U2016 <- length(unique(na.omit(V(g_U2016)$pairID)));my_palette_U2016 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2016);pair_IDs_U2016 <- factor(V(g_U2016)$pairID)
ggraph(g_U2016, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2016) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2016")

pair_IDs_U2017<-factor(V(g_U2017)$pairID); n_pairs_U2017 <- length(levels(na.omit(pair_IDs_U2017))); n_pairs_U2017 <- length(unique(na.omit(V(g_U2017)$pairID)));my_palette_U2017 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2017);pair_IDs_U2017 <- factor(V(g_U2017)$pairID)
ggraph(g_U2017, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2017) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2017")

pair_IDs_U2018<-factor(V(g_U2018)$pairID); n_pairs_U2018 <- length(levels(na.omit(pair_IDs_U2018))); n_pairs_U2018 <- length(unique(na.omit(V(g_U2018)$pairID)));my_palette_U2018 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2018);pair_IDs_U2018 <- factor(V(g_U2018)$pairID)
ggraph(g_U2018, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2018) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2018")
pair_IDs_U2019<-factor(V(g_U2019)$pairID); n_pairs_U2019 <- length(levels(na.omit(pair_IDs_U2019))); n_pairs_U2019<- length(unique(na.omit(V(g_U2019)$pairID)));my_palette_U2019 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2019);pair_IDs_U2019 <- factor(V(g_U2019)$pairID)
ggraph(g_U2019, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2019) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2019")

pair_IDs_U2020<-factor(V(g_U2020)$pairID); n_pairs_U2020 <- length(levels(na.omit(pair_IDs_U2020))); n_pairs_U2020<- length(unique(na.omit(V(g_U2020)$pairID)));my_palette_U2020 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2020);pair_IDs_U2020 <- factor(V(g_U2020)$pairID)
ggraph(g_U2020, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2020) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2020")

pair_IDs_U2021<-factor(V(g_U2021)$pairID); n_pairs_U2021 <- length(levels(na.omit(pair_IDs_U2021))); n_pairs_U2021<- length(unique(na.omit(V(g_U2021)$pairID)));my_palette_U2021 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2021);pair_IDs_U2021 <- factor(V(g_U2021)$pairID)
ggraph(g_U2021, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2021) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2021")

pair_IDs_U2022<-factor(V(g_U2022)$pairID); n_pairs_U2022 <- length(levels(na.omit(pair_IDs_U2022))); n_pairs_U2022<- length(unique(na.omit(V(g_U2022)$pairID)));my_palette_U2022 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2022);pair_IDs_U2022 <- factor(V(g_U2022)$pairID)
ggraph(g_U2022, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2022) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2022")

pair_IDs_U2023<-factor(V(g_U2023)$pairID); n_pairs_U2023 <- length(levels(na.omit(pair_IDs_U2023))); n_pairs_U2023<- length(unique(na.omit(V(g_U2023)$pairID)));my_palette_U2023 <- colorRampPalette(brewer.pal(9, "Set3"))(n_pairs_U2023);pair_IDs_U2023 <- factor(V(g_U2023)$pairID)
ggraph(g_U2023, layout = "nicely") +  
  geom_edge_link(aes(width = weight), color = "gray", alpha = 0.5) +  
  scale_edge_width(range = c(0.1, 2)) +  
  geom_node_point(aes(color = pairID), size = 4) +  
  scale_color_manual(values = my_palette_U2023) +  
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", size = 3) +  
  theme_void() +  
  ggtitle("Mother-Daughter Network - 2023")

```
#unique plot
```{r}
my_palette_U2022 <- c("#66C2A5",  "#FC8D62",  "#8DA0CB",  "#E78AC3",  "#A6D854")
pair_IDs_U2022 <- factor(V(g_U2022)$pairID)
pair_IDs_U2022_non_na <- na.omit(pair_IDs_U2022)

names(my_palette_U2022) <- levels(pair_IDs_U2022_non_na)

#apply colors 
V(g_U2022)$color <- my_palette_U2022[pair_IDs_U2022]

p1<-ggraph(g_U2022, layout = "nicely") +
  geom_edge_link(aes(width = weight), color = "grey70", alpha = 0.3, show.legend = FALSE) +
  scale_edge_width(range = c(0.5, 4)) +
  geom_node_point(aes(color = pairID), size = 4) +
  scale_color_manual(values = my_palette_U2022) +
  theme_void() +
  ggtitle("UA - 2022") +
  theme(
    plot.title = element_text(hjust = 0.1),
    legend.position = "right",
    legend.title = element_blank()
  )
edges_UA_2022<-edges_UA[edges_UA$year == 2022, ]
p2<-ggplot(edges_UA_2022, aes(x = weight)) +
  geom_histogram(bins = 30, fill = "black", color = "white") +
  geom_vline(aes(xintercept = mean(weight)), 
             color = "red", size = 1) +
  labs(x = "SRI", y = "Frequency", title = "SRI distribution") +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 8)
  )
final_plot_UA <- p1 + p2 + 
  plot_layout(ncol = 2, widths = c(2, 1))  

final_plot_UA
ggsave("network_faux_2022.png", plot = final_plot_UA,width = 7, height = 5, dpi = 300)
```
# network density
```{r}
edge_density(g_G2013)
edge_density(g_G2014)
edge_density(g_G2015)
edge_density(g_G2016)
edge_density(g_G2017)
edge_density(g_G2018)
edge_density(g_G2019)
edge_density(g_G2020)
edge_density(g_G2021)
edge_density(g_G2022)
edge_density(g_G2023)
```
```{r}
edge_density(g_U2013)
edge_density(g_U2014)
edge_density(g_U2015)
edge_density(g_U2016)
edge_density(g_U2017)
edge_density(g_U2018)
edge_density(g_U2019)
edge_density(g_U2020)
edge_density(g_U2021)
edge_density(g_U2022)
edge_density(g_U2023)
```
final plot
```{r}
g_G2023 <- graph_from_adjacency_matrix(net_G2023, mode = "undirected", weighted = TRUE, diag = FALSE)

highlight_mother_daughter <- function(graph, mother_daughter_pairs) {
  V(graph)$type <- "other"
  mother_ids <- mother_daughter_pairs$rfid_mom
  daughter_ids <- mother_daughter_pairs$rfid_off
  V(graph)[name %in% mother_ids | name %in% daughter_ids]$type <- "mother-daughter"
  return(graph)
}

g_G2023 <- highlight_mother_daughter(g_G2023, mother_daughter_pairs_G2023)

V(g_G2023)$pairID <- NA
for (i in seq_len(nrow(mother_daughter_pairs_G2023))) {
  pair <- mother_daughter_pairs_G2023[i, ]
  V(g_G2023)$pairID[V(g_G2023)$name == pair$rfid_mom] <- pair$pairID
  V(g_G2023)$pairID[V(g_G2023)$name == pair$rfid_off] <- pair$pairID
}

unique_pairs <- na.omit(unique(V(g_G2023)$pairID))
n_colors <- length(unique_pairs)
library(RColorBrewer)
my_palette_G2023 <- setNames(brewer.pal(n = max(3, min(n_colors, 8)), "Dark2")[1:n_colors], unique_pairs)

p1 <- ggraph(g_G2023, layout = "nicely") +
  geom_edge_link(aes(width = weight), color = "grey70", alpha = 0.3, show.legend = FALSE) +
  scale_edge_width(range = c(0.5, 4)) +
  geom_node_point(aes(color = pairID), size = 4) +
  scale_color_manual(values = my_palette_G2023, na.value = "lightgrey") +
  theme_void() +
  ggtitle("GB2 - 2023") +
  theme(
    plot.title = element_text(hjust = 0.1),
    legend.position = "right",
    legend.title = element_blank()
  )

p2 <- ggplot(edges_2023, aes(x = weight)) +
  geom_histogram(bins = 30, fill = "black", color = "white") +
  geom_vline(aes(xintercept = mean(weight, na.rm = TRUE)), color = "red", size = 1) +
  labs(x = "SRI", y = "Frequency", title = "SRI distribution") +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 8)
  )

final_plot <- p1 + p2 + plot_layout(ncol = 2, widths = c(3, 1))
final_plot
```
```{r}
d_by_year <- edges_UA %>%
  group_by(year, pair_type) %>%
  summarise(
    n = n(),
    mean_weight = mean(weight, na.rm = TRUE),
    sd_weight = sd(weight, na.rm = TRUE)
  ) %>%
  ungroup()
```
```{r}
md_expected_2016 <- mother_daughter_GB2 %>% filter(year == 2016)
md_actual_2016 <- summary_yearly %>% filter(year == 2016, pair_type == "mother-daughter")

unique_ids_2016 <- unique(colnames(GBI_G2016))

md_expected_2016$mother_present <- md_expected_2016$rfid_mom %in% unique_ids_2016
md_expected_2016$daughter_present <- md_expected_2016$rfid_off %in% unique_ids_2016

md_expected_2016$both_present <- md_expected_2016$mother_present & md_expected_2016$daughter_present

d_complete <- md_expected_2016 %>% filter(both_present)
md_missing <- md_expected_2016 %>% filter(!both_present)
```









