---
title: "Master Thesis - Data entire year"
author: "Flore Michel"
date: "2025-03-06"
output: html_document
---

```{r setup, include=FALSE}
#for UA
knitr::opts_knit$set(root.dir = "C:/Users/flo_r/Desktop/master thesis/data/data in CSV/UA full year/")

#knitr::opts_knit$set(root.dir = "C:/Users/flo_r/Desktop/master thesis/data/data in CSV/")
#knitr::opts_knit$set(root.dir = "/Users/floremichel/Desktop/THESIS/data/new data/")

library(igraph)
library(asnipe)
library(sna)
library(assocInd)
library(dplyr)
#library(lme4)
library(lmerTest)
library(ggplot2)
library(tidyverse)
library(viridis)
library(patchwork)
library(hrbrthemes)
library(igraph)
library(ggraph)
library(colormap)
library(tidyr)
library(dplyr)
library(data.table)
library(tidygraph)
library(RColorBrewer)
```

mother-daughter UA
```{r}
mother_daughter<-read.csv("mother-daughter-pairs.csv",header=TRUE,row.names = NULL,sep=";",check.names = F)
mother_daughter_UA<-mother_daughter[mother_daughter$Colony == "UA",]

mother_daughter_UA <- mother_daughter_UA %>%
  select(TAG_mom, lookup_formel_TAG_off, birth_off)

mother_daughter_UA <- mother_daughter_UA %>%
  rename(
    rfid_mom = TAG_mom,
    rfid_off = lookup_formel_TAG_off,
    year = birth_off
  )

mother_daughter_UA$rfid_mom[mother_daughter_UA$rfid_mom == "5E-BF3C=A4A4"] <- "5E-BF3C"
mother_daughter_UA$rfid_mom[mother_daughter_UA$rfid_mom == "5E-BF3C"] <- "5EBF3C"
mother_daughter_UA$rfid_mom[mother_daughter_UA$rfid_mom == "4FCA=7CCA"] <- "4FCA"
mother_daughter_UA$rfid_off[mother_daughter_UA$rfid_off == "FA5B=D6CC"] <- "FA5B"
mother_daughter_UA$rfid_mom[mother_daughter_UA$rfid_mom == "F50E=9B39=AE33"] <- "F50E"
mother_daughter_UA$rfid_mom[mother_daughter_UA$rfid_mom == "37F2=100F"] <- "100F"
mother_daughter_UA$rfid_off[mother_daughter_UA$rfid_off == "37F2"] <- "100F"
mother_daughter_UA$rfid_off[mother_daughter_UA$rfid_off == "37F2"] <- "100F"


mother_daughter_UA <- mother_daughter_UA %>%
  mutate(pairID = paste(rfid_mom, rfid_off, sep = "_"))

write.csv(mother_daughter_UA, "/Users/flo_r/Desktop/master thesis/analysis/variables UA/mother_daughter_UA.csv", row.names = FALSE)
```


```{r load all pairs}
mother_daugther_G2013<-extract_pair_data(mother_daugther_pairs, 2013)
mother_daugther_G2013$rfid_mom[mother_daugther_G2013$rfid_mom == "B597=B587"] <- "B587"
mother_daugther_G2013$rfid_mom[mother_daugther_G2013$rfid_mom == "6C0D=6D0D"] <- "6D0D"
mother_daugther_G2013$rfid_mom[mother_daugther_G2013$rfid_mom == "07F9=51C3"] <- "07F9"
mother_daugther_G2013 <- mother_daugther_G2013[mother_daugther_G2013$rfid_off != "f4301", ]
mother_daugther_G2013$pairID<-c(1:nrow(mother_daugther_G2013))
write.csv(mother_daugther_G2013, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2013/mother_daugther_pairs_G2013.csv", row.names = FALSE)

mother_daugther_G2014<-extract_pair_data(mother_daugther_pairs, 2014)
mother_daugther_G2014$rfid_mom[mother_daugther_G2014$rfid_mom == "B597=B587"] <- "B587"
mother_daugther_G2014$rfid_off[mother_daugther_G2014$rfid_off == "EE00=7C6F"] <- "7C6F"
mother_daugther_G2014$pairID<-c(1:nrow(mother_daugther_G2014))
write.csv(mother_daugther_G2014, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2014/mother_daugther_pairs_G2014.csv", row.names = FALSE)

mother_daugther_G2015<-extract_pair_data(mother_daugther_pairs, 2015)
mother_daugther_G2015$pairID<-c(1:nrow(mother_daugther_G2015))
write.csv(mother_daugther_G2015, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2015/mother_daugther_pairs_G2015.csv", row.names = FALSE)

mother_daugther_G2016<-extract_pair_data(mother_daugther_pairs, 2016)
mother_daugther_G2016$pairID<-c(1:nrow(mother_daugther_G2016))
write.csv(mother_daugther_G2016, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2016/mother_daugther_pairs_G2016.csv", row.names = FALSE)

mother_daugther_G2017<-extract_pair_data(mother_daugther_pairs, 2017)
mother_daugther_G2017$rfid_off[mother_daugther_G2017$rfid_off == "DE72=15BF"] <- "15BF"
mother_daugther_G2017$pairID<-c(1:nrow(mother_daugther_G2017))
write.csv(mother_daugther_G2017, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2017/mother_daugther_pairs_G2017.csv", row.names = FALSE)

mother_daugther_G2018<-extract_pair_data(mother_daugther_pairs, 2018)
mother_daugther_G2018$rfid_mom[mother_daugther_G2018$rfid_mom == "EE00=7C6F"] <- "7C6F"
mother_daugther_G2018$pairID<-c(1:nrow(mother_daugther_G2018))
write.csv(mother_daugther_G2018, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2018/mother_daugther_pairs_G2018.csv", row.names = FALSE)

mother_daugther_G2019<-extract_pair_data(mother_daugther_pairs, 2019)
mother_daugther_G2019$pairID<-c(1:nrow(mother_daugther_G2019))
write.csv(mother_daugther_G2019, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2019/mother_daugther_pairs_G2019.csv", row.names = FALSE)

mother_daugther_G2020<-extract_pair_data(mother_daugther_pairs, 2020)
mother_daugther_G2020$rfid_off[mother_daugther_G2020$rfid_off == "D9EF=F429"] <- "F429"
mother_daugther_G2020$rfid_mom[mother_daugther_G2020$rfid_mom == "DE72=15BF"] <- "15BF"
mother_daugther_G2020$pairID<-c(1:nrow(mother_daugther_G2020))
write.csv(mother_daugther_G2020, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2020/mother_daugther_pairs_G2020.csv", row.names = FALSE)

mother_daugther_G2021<-extract_pair_data(mother_daugther_pairs, 2021)
mother_daugther_G2021$rfid_mom[mother_daugther_G2021$rfid_mom == "DE72=15BF"] <- "15BF"
mother_daugther_G2021$pairID<-c(1:nrow(mother_daugther_G2021))
write.csv(mother_daugther_G2021, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2021/mother_daugther_pairs_G2021.csv", row.names = FALSE)

mother_daugther_G2022<-extract_pair_data(mother_daugther_pairs, 2022)
mother_daugther_G2022$rfid_off[mother_daugther_G2022$rfid_off == "1270=EE51"] <- "EE51"
mother_daugther_G2022<- mother_daugther_G2022[mother_daugther_G2022$rfid_off != "f5069", ]
mother_daugther_G2022$pairID<-c(1:nrow(mother_daugther_G2022))
write.csv(mother_daugther_G2022, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2022/mother_daugther_pairs_G2022.csv", row.names = FALSE)

mother_daugther_G2023<-extract_pair_data(mother_daugther_pairs, 2023)
mother_daugther_G2023$pairID<-c(1:nrow(mother_daugther_G2023))
write.csv(mother_daugther_G2023, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2023/mother_daugther_pairs_G2023.csv", row.names = FALSE)
```

# GB2 2013

```{r, include= FALSE}
GB2_2013<-read.csv("gs_GB2_2013.csv",header=TRUE,sep=",",check.names = F,row.names=NULL)
GB2_2013$date <- as.Date(GB2_2013$date, format="%d.%m.%Y")
GB2_2013$date<- format(GB2_2013$date, "%d.%m.%Y")
```

### GBI matrix

```{r GBI matrix}
data_G2013<- GB2_2013 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2013$date<- as.Date(data_G2013$date, format = "%d.%m.%Y")
data_G2013$daygroup<-as.factor(paste(data_G2013$date, data_G2013$box, sep="_"))

ind_G2013<-data.table(ID=c(data_G2013$ID), group=data_G2013$daygroup, day=c(data_G2013$date)) #df with 3 col: ID, group and day 
GBI_G2013<- get_group_by_individual(ind_G2013, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2013 <- as.matrix(GBI_G2013)

write.csv(GBI_G2013, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2013/full_GBI_G2013.csv", row.names = TRUE)

```

### network

```{r network creation}
net_G2013<-get_network(GBI_G2013,data_format = "GBI",association_index = "SRI")
write.csv(net_G2013, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2013/full_net_G2013.csv", row.names = TRUE)
```

```{r}
g_G2013 <- graph_from_adjacency_matrix(net_G2013,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2013,vertex.label = V(g_G2013)$name,edge.width = E(g_G2013)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue") 
```

As the network is too dense, I can filter the edges to keep only the
stronger ones.
### threshold

```{r threshold}
threshold_G2013 <- mean(net_G2013, na.rm = TRUE) +  sd(net_G2013, na.rm = TRUE)

gfilter_G2013 <- delete_edges(g_G2013, E(g_G2013)[weight < mean(net_G2013, na.rm = TRUE)]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2013, edge.width = E(gfilter_G2013)$weight*6,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7)
```

Now we can add a new vertex attributes, where we can identify the
mother-daughter pairs. This would allow us to color the vertices
according to their type.

###vertex attributes

```{r vertex attributes "mother" "daughter"}
mother_daughter_GB2<-read.csv("small_mother_daughter_GB2.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
mother_daughter_pairs_G2013<-mother_daughter_GB2[mother_daughter_GB2$year == 2013,]

mother_ids_G2013<- mother_daugther_G2013$mother
daughter_ids_G2013<- mother_daugther_G2013$daughter

V(g_G2013)$type <- "other"  # default value
V(g_G2013)[V(g_G2013)$name %in% mother_ids_G2013]$type   <- "mother-daughter"
V(g_G2013)[V(g_G2013)$name %in% daughter_ids_G2013]$type <- "mother-daughter"

plot(gfilter_G2013, edge.width = E(g_G2013)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2013)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2013)$name)
```

Label each mother-daughter pair in the network with a unique identifier.
This will allow us to color the vertices according to their pair ID.

```{r pair ID}
md_lookup_G2013<- mother_daugther_G2013 %>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2013<- match(V(g_G2013)$name, md_lookup_G2013$ID)
V(g_G2013)$pairID <- md_lookup_G2013$pairID[IDs_G2013] 

pair_IDs_G2013<- factor(V(g_G2013)$pairID)

n_pairs_G2013<-length(levels(na.omit(pair_IDs_G2013)))
my_palette_G2013<-brewer.pal(n = max(3,n_pairs_G2013), name = "Set2") 

pair_colors_G2013 <- my_palette_G2013[pair_IDs_G2013]
pair_colors_G2013[is.na(pair_colors_G2013)] <- "gray80"
```

Plot of the network with the vertices colored according to their pair ID

```{r plot MD}
plot(gfilter_G2013, edge.width = E(gfilter_G2013)$weight*6,edge.curved=F,vertex.color =pair_colors_G2013, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2013 <- levels(na.omit(pair_IDs_G2013))
legend_colors_G2013 <- my_palette_G2013[seq_along(pair_levels_G2013)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2013), "random pair"),
  col    = c(legend_colors_G2013, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

Creation of 
###the edge list

```{r edge list}
edge_G2013 <- igraph::as_data_frame(g_G2013, what = "edges")

pid <- V(g_G2013)$pairID
names(pid) <- V(g_G2013)$name

from_pid <- pid[ edge_G2013$from ]
to_pid   <- pid[ edge_G2013$to ]

edge_G2013$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_G2013, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2013/full_edge_G2013.csv", row.names = FALSE)
```

SRI plot

```{r SRI plot}
sri_G2013<- as.vector(net_G2013)
sri_G2013 <- sri_G2013[sri_G2013 > 0]  # Remove zero values
meansri_G2013<- mean(sri_G2013)


hist(sri_G2013, breaks=30, main="Distribution of association index (SRI) for GB2 in 2013", 
     xlab="Association index (SRI)", ylab="Frequency", col="lightblue", border="black",xlim = (0:1))

#red line for the mean
abline(v=meansri_G2013, col="blue", lwd=2)
```

# GB2_2014 
```{r import}
GB2_2014<-read.csv("gs_GB2_2014.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
```

Change some individual ID
```{r}
names(GB2_2014)[names(GB2_2014) == 'EE00'] <- '7C6F'
```


```{r GBI}
data_G2014<- GB2_2014 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2014$date<- as.Date(data_G2014$date, format = "%d.%m.%Y")
data_G2014$daygroup<-as.factor(paste(data_G2014$date, data_G2014$box, sep="_"))

ind_G2014<-data.table(ID=c(data_G2014$ID), group=data_G2014$daygroup, day=c(data_G2014$date)) #df with 3 col: ID, group and day 
GBI_G2014<- get_group_by_individual(ind_G2014, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2014 <- as.matrix(GBI_G2014)
write.csv(GBI_G2014, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2014/full_GBI_G2014.csv", row.names = TRUE)
```

###network

```{r network creation}
net_G2014<-get_network(GBI_G2014,data_format = "GBI",association_index = "SRI")
write.csv(net_G2014, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2014/full_net_G2014.csv", row.names = TRUE)
```

```{r graph}
g_G2014 <- graph_from_adjacency_matrix(net_G2014,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2014,vertex.label = V(g_G2014)$name,edge.width = E(g_G2014)$weight * 5, layout = layout_with_fr, vertex.color = "lightblue") 
```

plot + threshold

```{r threshold}
threshold_G2014 <- mean(net_G2014, na.rm = TRUE) +  sd(net_G2014, na.rm = TRUE)

gfilter_G2014<- delete_edges(g_G2014, E(g_G2014)[weight < threshold_G2014]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2014, edge.width = E(gfilter_G2014)$weight*6,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7)
```

###vertex attributes

```{r vertex attributes}
mother_ids_G2014<- mother_daugther_G2014$rfid_mom
daughter_ids_G2014<- mother_daugther_G2014$rfid_off

V(g_G2014)$type <- "other"  # default value
V(g_G2014)[V(g_G2014)$name %in% mother_ids_G2014]$type   <- "mother-daughter"
V(g_G2014)[V(g_G2014)$name %in% daughter_ids_G2014]$type <- "mother-daughter"

plot(gfilter_G2014, edge.width = E(g_G2014)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2014)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2014)$name)
```

```{r colors}
md_lookup_G2014<- mother_daugther_G2014%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2014<- match(V(g_G2014)$name, md_lookup_G2014$ID)  
V(g_G2014)$pairID <- md_lookup_G2014$pairID[IDs_G2014] 

pair_IDs_G2014<- factor(V(g_G2014)$pairID)

n_pairs_G2014<-length(levels(na.omit(pair_IDs_G2014)))
my_palette_G2014<-brewer.pal(n = max(3,n_pairs_G2014), name = "Set2") 

pair_colors_G2014<- my_palette_G2014[pair_IDs_G2014]
pair_colors_G2014[is.na(pair_colors_G2014)] <- "gray80"
```

big plot

```{r big plot}
plot(gfilter_G2014, edge.width = E(gfilter_G2014)$weight*6,edge.curved=F,vertex.color =pair_colors_G2014, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2014 <- levels(na.omit(pair_IDs_G2014))
legend_colors_G2014 <- my_palette_G2014[seq_along(pair_levels_G2014)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2014), "random pair"),
  col    = c(legend_colors_G2014, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list
```{r edge list}
edge_G2014<- igraph::as_data_frame(g_G2014, what = "edges")

pid <- V(g_G2014)$pairID
names(pid) <- V(g_G2014)$name

from_pid <- pid[ edge_G2014$from ]
to_pid   <- pid[ edge_G2014$to ]

edge_G2014$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)

write.csv(edge_G2014, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2014/full_edge_G2014.csv", row.names = FALSE)
```

# GB2 2015 

```{r import}
GB2_2015<-read.csv("gs_GB2_2015.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
```

Change some individual ID
```{r}
names(GB2_2015)[names(GB2_2015) == 'EE00'] <- '7C6F'
```

###GBI

```{r GBI matrix}
data_G2015<- GB2_2015 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2015$date<- as.Date(data_G2015$date, format = "%d.%m.%Y")
data_G2015$daygroup<-as.factor(paste(data_G2015$date, data_G2015$box, sep="_"))

ind_G2015<-data.table(ID=c(data_G2015$ID), group=data_G2015$daygroup, day=c(data_G2015$date)) #df with 3 col: ID, group and day 
GBI_G2015<- get_group_by_individual(ind_G2015, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2015<- as.matrix(GBI_G2015)


write.csv(GBI_G2015, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2015/full_GBI_G2015.csv", row.names = TRUE)
```

### network

```{r network}
net_G2015<-get_network(GBI_G2015,data_format = "GBI",association_index = "SRI")
write.csv(net_G2015, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2015/full_net_G2015.csv", row.names = TRUE)
```

```{r graph}
g_G2015<- graph_from_adjacency_matrix(net_G2015,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2015,vertex.label = V(g_G2015)$name,edge.width = E(g_G2015)$weight * 10, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```

plot + threshold

```{r threshold}
threshold_G2015<- mean(net_G2015, na.rm = TRUE) +  sd(net_G2015, na.rm = TRUE)

gfilter_G2015<- delete_edges(g_G2015, E(g_G2015)[weight < threshold_G2015]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2015, edge.width = E(gfilter_G2015)$weight*6,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

###vertex attributes

```{r vertex attributes}
mother_ids_G2015<- mother_daugther_G2015$rfid_mom
daughter_ids_G2015<- mother_daugther_G2015$rfid_off

V(g_G2015)$type <- "other"  # default value
V(g_G2015)[V(g_G2015)$name %in% mother_ids_G2015]$type   <- "mother-daughter"
V(g_G2015)[V(g_G2015)$name %in% daughter_ids_G2015]$type <- "mother-daughter"

plot(gfilter_G2015, edge.width = E(gfilter_G2015)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2015)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2015)$name,edge.color="plum3")
```

```{r MD IDs}
md_lookup_G2015<- mother_daugther_G2015%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2015<- match(V(g_G2015)$name, md_lookup_G2015$ID)  
V(g_G2015)$pairID <- md_lookup_G2015$pairID[IDs_G2015] 

pair_IDs_G2015<- factor(V(g_G2015)$pairID)

n_pairs_G2015<-length(levels(na.omit(pair_IDs_G2015)))
my_palette_G2015<-brewer.pal(n = max(3,n_pairs_G2015), name = "Set2") 

pair_colors_G2015<- my_palette_G2015[pair_IDs_G2015]
pair_colors_G2015[is.na(pair_colors_G2015)] <- "gray80"
```

big plot
```{r big plot}
plot(gfilter_G2015, edge.width = E(gfilter_G2015)$weight*6,edge.curved=F,vertex.color =pair_colors_G2015, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2015<- levels(na.omit(pair_IDs_G2015))
legend_colors_G2015 <- my_palette_G2015[seq_along(pair_levels_G2015)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2015), "random pair"),
  col    = c(legend_colors_G2015, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list
```{r edge list}
edge_G2015 <- igraph::as_data_frame(g_G2015, what = "edges")

pid <- V(g_G2015)$pairID
names(pid) <- V(g_G2015)$name

from_pid <- pid[ edge_G2015$from ]
to_pid   <- pid[ edge_G2015$to ]

edge_G2015$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)

write.csv(edge_G2015,"/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2015/full_edge_G2015.csv", row.names = FALSE)
```

# GB2 -2016

```{r import }
GB2_2016<-read.csv("gs_GB2_2016.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
colnames(GB2_2016)[colnames(GB2_2016) == "758"] <- "0758"
GB2_2016$date[33] <- "16.08.2016"
```

###GBI 

```{r}
data_G2016<- GB2_2016%>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2016$date<- as.Date(data_G2016$date, format = "%d.%m.%Y")
data_G2016$daygroup<-as.factor(paste(data_G2016$date, data_G2016$box, sep="_"))

ind_G2016<-data.table(ID=c(data_G2016$ID), group=data_G2016$daygroup, day=c(data_G2016$date)) #df with 3 col: ID, group and day 
GBI_G2016<- get_group_by_individual(ind_G2016, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2016<- as.matrix(GBI_G2016)
write.csv(GBI_G2016, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2016/full_GBI_G2016.csv", row.names = TRUE)
```

###network

```{r network}
net_G2016<-get_network(GBI_G2016,data_format = "GBI",association_index = "SRI")
write.csv(net_G2016, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2016/full_net_G2016.csv", row.names = TRUE)
```

```{r}
g_G2016<- graph_from_adjacency_matrix(net_G2016,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2016,vertex.label = V(g_G2016)$name,edge.width = E(g_G2016)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```

```{r}
threshold_G2016<- mean(net_G2016, na.rm = TRUE) +  sd(net_G2016, na.rm = TRUE)

gfilter_G2016<- delete_edges(g_G2016, E(g_G2016)[weight < threshold_G2016]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2016, edge.width = E(gfilter_G2016)$weight*6,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

###vertex attributes
```{r}
mother_ids_G2016<- mother_daugther_G2016$rfid_mom
daughter_ids_G2016<- mother_daugther_G2016$rfid_off

V(g_G2016)$type <- "other"  # default value
V(g_G2016)[V(g_G2016)$name %in% mother_ids_G2016]$type   <- "mother-daughter"
V(g_G2016)[V(g_G2016)$name %in% daughter_ids_G2016]$type <- "mother-daughter"

plot(gfilter_G2016, edge.width = E(gfilter_G2016)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2016)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2016)$name,edge.color="plum3")
```

```{r}
md_lookup_G2016<- mother_daugther_G2016%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2016<- match(V(g_G2016)$name, md_lookup_G2016$ID)  
V(g_G2016)$pairID <- md_lookup_G2016$pairID[IDs_G2016] 

pair_IDs_G2016<- factor(V(g_G2016)$pairID)

n_pairs_G2016<-length(levels(na.omit(pair_IDs_G2016)))
my_palette_G2016<-brewer.pal(n = max(3,n_pairs_G2016), name = "Set2") 

pair_colors_G2016<- my_palette_G2016[pair_IDs_G2016]
pair_colors_G2016[is.na(pair_colors_G2016)] <- "gray80"
```

big plot

```{r}
plot(gfilter_G2016, edge.width = E(gfilter_G2016)$weight*6,edge.curved=F,vertex.color =pair_colors_G2016, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2016<- levels(na.omit(pair_IDs_G2016))
legend_colors_G2016 <- my_palette_G2016[seq_along(pair_levels_G2016)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2016), "random pair"),
  col    = c(legend_colors_G2016, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_G2016 <- igraph::as_data_frame(g_G2016, what = "edges")

pid <- V(g_G2016)$pairID
names(pid) <- V(g_G2016)$name

from_pid <- pid[ edge_G2016$from ]
to_pid   <- pid[ edge_G2016$to ]

edge_G2016$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)

write.csv(edge_G2016, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2016/full_edge_G2016.csv", row.names = FALSE)
```

# GB2 2017

```{r import}
GB2_2017<-read.csv("gs_GB2_2017.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
```

Change some individual ID
```{r}
names(GB2_2015)[names(GB2_2015) == 'DE72'] <- '15BF'
```

###GBI
```{r}
data_G2017<- GB2_2017%>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2017$date<- as.Date(data_G2017$date, format = "%d.%m.%Y")
data_G2017$daygroup<-as.factor(paste(data_G2017$date, data_G2017$box, sep="_"))

ind_G2017<-data.table(ID=c(data_G2017$ID), group=data_G2017$daygroup, day=c(data_G2017$date)) #df with 3 col: ID, group and day 
GBI_G2017<- get_group_by_individual(ind_G2017, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2017<- as.matrix(GBI_G2017)

write.csv(GBI_G2017, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2017/full_GBI_G2017.csv", row.names = TRUE)
```
###network

```{r}
net_G2017<-get_network(GBI_G2017,data_format = "GBI",association_index = "SRI")
write.csv(net_G2017, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2017/full_net_G2017.csv", row.names = TRUE)
```

```{r}
g_G2017<- graph_from_adjacency_matrix(net_G2017,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2017,vertex.label = V(g_G2017)$name,edge.width = E(g_G2017)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```

plot + threshold

```{r}
threshold_G2017<- mean(net_G2017, na.rm = TRUE) +  sd(net_G2017, na.rm = TRUE)

gfilter_G2017<- delete_edges(g_G2017, E(g_G2017)[weight < mean(net_G2017, na.rm = TRUE)]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2017, edge.width = E(gfilter_G2017)$weight*5,edge.curved=F,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

###vertex attributes

```{r}
mother_ids_G2017<- mother_daugther_G2017$rfid_mom
daughter_ids_G2017<- mother_daugther_G2017$rfid_off

V(g_G2017)$type <- "other"  # default value
V(g_G2017)[V(g_G2017)$name %in% mother_ids_G2017]$type   <- "mother-daughter"
V(g_G2017)[V(g_G2017)$name %in% daughter_ids_G2017]$type <- "mother-daughter"

plot(gfilter_G2017, edge.width = E(gfilter_G2017)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2017)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2017)$name,edge.color="plum3")
```

```{r}
md_lookup_G2017<- mother_daugther_G2017%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2017<- match(V(g_G2017)$name, md_lookup_G2017$ID)
V(g_G2017)$pairID <- md_lookup_G2017$pairID[IDs_G2017] 

pair_IDs_G2017<- factor(V(g_G2017)$pairID)

n_pairs_G2017<-length(levels(na.omit(pair_IDs_G2017)))
my_palette_G2017<-brewer.pal(n = max(3,n_pairs_G2017), name = "Paired") 

pair_colors_G2017<- my_palette_G2017[pair_IDs_G2017]
pair_colors_G2017[is.na(pair_colors_G2017)] <- "gray80"
```

big plot

```{r}
plot(gfilter_G2017, edge.width = E(gfilter_G2017)$weight*6,edge.curved=F,vertex.color =pair_colors_G2017, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2017<- levels(na.omit(pair_IDs_G2017))
legend_colors_G2017 <- my_palette_G2017[seq_along(pair_levels_G2017)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2017), "random pair"),
  col    = c(legend_colors_G2017, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_G2017 <- igraph::as_data_frame(g_G2017, what = "edges")

pid <- V(g_G2017)$pairID
names(pid) <- V(g_G2017)$name

from_pid <- pid[ edge_G2017$from ]
to_pid   <- pid[ edge_G2017$to ]

edge_G2017$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_G2017, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2017/full_edge_G2017.csv", row.names = FALSE)
```

# GB2 2018

not a lot of data

```{r import}
GB2_2018<-read.csv("gs_GB2_2018.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
```

### GBI

```{r}
data_G2018<- GB2_2018%>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2018$date<- as.Date(data_G2018$date, format = "%d.%m.%Y")
data_G2018$daygroup<-as.factor(paste(data_G2018$date, data_G2018$box, sep="_"))

ind_G2018<-data.table(ID=c(data_G2018$ID), group=data_G2018$daygroup, day=c(data_G2018$date)) #df with 3 col: ID, group and day 
GBI_G2018<- get_group_by_individual(ind_G2018, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2018<- as.matrix(GBI_G2018)
write.csv(GBI_G2018, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2018/full_GBI_G2018.csv", row.names = TRUE)

```

###network

```{r}
net_G2018<-get_network(GBI_G2018,data_format = "GBI",association_index = "SRI")
write.csv(net_G2018, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2018/full_net_G2018.csv", row.names = TRUE)

```

```{r}
g_G2018<- graph_from_adjacency_matrix(net_G2018,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2018,vertex.label = V(g_G2018)$name,edge.width = E(g_G2018)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```

plot + threshold

```{r}
threshold_G2018<- mean(net_G2018, na.rm = TRUE) +  sd(net_G2018, na.rm = TRUE)

gfilter_G2018<- delete_edges(g_G2018, E(g_G2018)[weight < threshold_G2018]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2018, edge.width = E(gfilter_G2018)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

###vertex attributes

```{r}
mother_ids_G2018<- mother_daugther_G2018$rfid_mom
daughter_ids_G2018<- mother_daugther_G2018$rfid_off

V(g_G2018)$type <- "other"  # default value
V(g_G2018)[V(g_G2018)$name %in% mother_ids_G2018]$type   <- "mother-daughter"
V(g_G2018)[V(g_G2018)$name %in% daughter_ids_G2018]$type <- "mother-daughter"

plot(gfilter_G2018, edge.width = E(gfilter_G2018)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2018)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2018)$name,edge.color="plum3")
```

```{r}
md_lookup_G2018<- mother_daugther_G2018%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2018<- match(V(g_G2018)$name, md_lookup_G2018$ID)
V(g_G2018)$pairID <- md_lookup_G2018$pairID[IDs_G2018] 

pair_IDs_G2018<- factor(V(g_G2018)$pairID)

n_pairs_G2018<-length(levels(na.omit(pair_IDs_G2018)))
my_palette_G2018<-brewer.pal(n = max(3,n_pairs_G2018), name = "Paired") 

pair_colors_G2018<- my_palette_G2018[pair_IDs_G2018]
pair_colors_G2018[is.na(pair_colors_G2018)] <- "gray80"
```

####big plot

```{r}
plot(gfilter_G2018, edge.width = E(gfilter_G2018)$weight*6,edge.curved=F,vertex.color =pair_colors_G2018, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2018<- levels(na.omit(pair_IDs_G2018))
legend_colors_G2018 <- my_palette_G2018[seq_along(pair_levels_G2018)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2018), "random pair"),
  col    = c(legend_colors_G2018, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list
```{r}
edge_G2018 <- igraph::as_data_frame(g_G2018, what = "edges")

pid <- V(g_G2018)$pairID
names(pid) <- V(g_G2018)$name

from_pid <- pid[ edge_G2018$from ]
to_pid   <- pid[ edge_G2018$to ]

edge_G2018$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_G2018, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2018/full_edge_G2018.csv", row.names = FALSE)
```

# GB2 2019

```{r}
GB2_2019<-read.csv("gs_GB2_2019.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
```

####GBI
```{r}
colnames(GB2_2019)<-as.character(colnames(GB2_2019))
data_G2019<- GB2_2019 %>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2019$date<- as.Date(data_G2019$date, format = "%d.%m.%Y")
data_G2019$daygroup<-as.factor(paste(data_G2019$date, data_G2019$box, sep="_"))

ind_G2019<-data.table(ID=c(data_G2019$ID), group=data_G2019$daygroup, day=c(data_G2019$date)) #df with 3 col: ID, group and day 
GBI_G2019<- get_group_by_individual(ind_G2019, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2019<- as.matrix(GBI_G2019)
write.csv(GBI_G2019, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2019/full_GBI_G2019.csv", row.names = TRUE)
```

###network
```{r}
net_G2019<-get_network(GBI_G2019,data_format = "GBI",association_index = "SRI")
write.csv(net_G2019, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2019/full_net_G2019.csv", row.names = TRUE)
```

```{r}
g_G2019<- graph_from_adjacency_matrix(net_G2019,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2019,vertex.label = V(g_G2019)$name,edge.width = E(g_G2019)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```

plot + threshold

```{r}
threshold_G2019<- mean(net_G2019, na.rm = TRUE) +  sd(net_G2019, na.rm = TRUE)

gfilter_G2019<- delete_edges(g_G2019, E(g_G2019)[weight <threshold_G2019]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2019, edge.width = E(gfilter_G2019)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

###vertex attributes

```{r}
mother_ids_G2019<- mother_daugther_G2019$rfid_mom
daughter_ids_G2019<- mother_daugther_G2019$rfid_off

V(g_G2019)$type <- "other"  # default value
V(g_G2019)[V(g_G2019)$name %in% mother_ids_G2019]$type   <- "mother-daughter"
V(g_G2019)[V(g_G2019)$name %in% daughter_ids_G2019]$type <- "mother-daughter"

plot(gfilter_G2019, edge.width = E(gfilter_G2019)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2019)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2019)$name,edge.color="plum3")
```

```{r}
md_lookup_G2019<- mother_daugther_G2019%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2019<- match(V(g_G2019)$name, md_lookup_G2019$ID)  
V(g_G2019)$pairID <- md_lookup_G2019$pairID[IDs_G2019] 

pair_IDs_G2019<- factor(V(g_G2019)$pairID)

n_pairs_G2019<-length(levels(na.omit(pair_IDs_G2019)))
my_palette_G2019<-brewer.pal(n = max(3,n_pairs_G2019), name = "Paired") 

pair_colors_G2019<- my_palette_G2019[pair_IDs_G2019]
pair_colors_G2019[is.na(pair_colors_G2019)] <- "gray80"
```

big plot

```{r}
plot(gfilter_G2019, edge.width = E(gfilter_G2019)$weight*6,edge.curved=F,vertex.color =pair_colors_G2019, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2019<- levels(na.omit(pair_IDs_G2019))
legend_colors_G2019 <- my_palette_G2019[seq_along(pair_levels_G2019)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2019), "random pair"),
  col    = c(legend_colors_G2019, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

####edge list

```{r}
edge_G2019 <- igraph::as_data_frame(g_G2019, what = "edges")

pid <- V(g_G2019)$pairID
names(pid) <- V(g_G2019)$name

from_pid <- pid[ edge_G2019$from ]
to_pid   <- pid[ edge_G2019$to ]

edge_G2019$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_G2019, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2019/full_edge_G2019.csv", row.names = FALSE)
```

# GB2 2020

```{r import}
GB2_2020<-read.csv("gs_GB2_2020.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
```

Change some individual ID
```{r}
names(GB2_2020)[names(GB2_2020) == 'D9EF'] <- 'F429'
```

###GBI

```{r}
colnames(GB2_2020)<-as.character(colnames(GB2_2020))
data_G2020<- GB2_2020 %>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2020$date<- as.Date(data_G2020$date, format = "%d.%m.%Y")
data_G2020$daygroup<-as.factor(paste(data_G2020$date, data_G2020$box, sep="_"))

ind_G2020<-data.table(ID=c(data_G2020$ID), group=data_G2020$daygroup, day=c(data_G2020$date)) #df with 3 col: ID, group and day 
GBI_G2020<- get_group_by_individual(ind_G2020, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2020<- as.matrix(GBI_G2020)
write.csv(GBI_G2020, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2020/full_GBI_G2020.csv", row.names = TRUE)
```

###network

```{r}
net_G2020<-get_network(GBI_G2020,data_format = "GBI",association_index = "SRI")
write.csv(net_G2020, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2020/full_net_G2020.csv", row.names = TRUE)
```

```{r}
g_G2020<- graph_from_adjacency_matrix(net_G2020,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2020,vertex.label = V(g_G2020)$name,edge.width = E(g_G2020)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```

plot + threshold

```{r}
threshold_G2020<- mean(net_G2020, na.rm = TRUE) +  sd(net_G2020, na.rm = TRUE)

gfilter_G2020<- delete_edges(g_G2020, E(g_G2020)[weight <threshold_G2020]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2020, edge.width = E(gfilter_G2020)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

###vertex attributes

```{r}
mother_ids_G2020<- mother_daugther_G2020$rfid_mom
daughter_ids_G2020<- mother_daugther_G2020$rfid_off

V(g_G2020)$type <- "other"  # default value
V(g_G2020)[V(g_G2020)$name %in% mother_ids_G2020]$type   <- "mother-daughter"
V(g_G2020)[V(g_G2020)$name %in% daughter_ids_G2020]$type <- "mother-daughter"

plot(gfilter_G2020, edge.width = E(gfilter_G2020)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2020)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2020)$name,edge.color="plum3")
```

```{r}
md_lookup_G2020<- mother_daugther_G2020%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2020<- match(V(g_G2020)$name, md_lookup_G2020$ID)
V(g_G2020)$pairID <- md_lookup_G2020$pairID[IDs_G2020] 

pair_IDs_G2020<- factor(V(g_G2020)$pairID)

n_pairs_G2020<-length(levels(na.omit(pair_IDs_G2020)))
my_palette_G2020<-brewer.pal(n = max(3,n_pairs_G2020), name = "Paired") 

pair_colors_G2020<- my_palette_G2020[pair_IDs_G2020]
pair_colors_G2020[is.na(pair_colors_G2020)] <- "gray80"
```

big plot

```{r}
plot(gfilter_G2020, edge.width = E(gfilter_G2020)$weight*6,edge.curved=F,vertex.color =pair_colors_G2020, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2020<- levels(na.omit(pair_IDs_G2020))
legend_colors_G2020 <- my_palette_G2020[seq_along(pair_levels_G2020)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2020), "random pair"),
  col    = c(legend_colors_G2020, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_G2020 <- igraph::as_data_frame(g_G2020, what = "edges")

pid <- V(g_G2020)$pairID
names(pid) <- V(g_G2020)$name

from_pid <- pid[ edge_G2020$from ]
to_pid   <- pid[ edge_G2020$to ]

edge_G2020$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_G2020, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2020/full_edge_G2020.csv", row.names = FALSE)
```

# GB2 2021

```{r import}
GB2_2021<-read.csv("gs_GB2_2021.csv",header=TRUE, row.names=NULL,sep=",",check.names = F)
colnames(GB2_2021)[colnames(GB2_2021) == "Date"] <- "date"
```

####GBI
```{r}
colnames(GB2_2021)<-as.character(colnames(GB2_2021))
data_G2021<- GB2_2021%>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2021$date<- as.Date(data_G2021$date, format = "%d.%m.%Y")
data_G2021$daygroup<-as.factor(paste(data_G2021$date, data_G2021$box, sep="_"))

ind_G2021<-data.table(ID=c(data_G2021$ID), group=data_G2021$daygroup, day=c(data_G2021$date)) #df with 3 col: ID, group and day 
GBI_G2021<- get_group_by_individual(ind_G2021, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2021<- as.matrix(GBI_G2021)
write.csv(GBI_G2021, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2021/full_GBI_G2021.csv", row.names = TRUE)
```

###network

```{r}
net_G2021<-get_network(GBI_G2021,data_format = "GBI",association_index = "SRI")
write.csv(net_G2021, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2021/full_net_G2021.csv", row.names = TRUE)
```

```{r}
g_G2021<- graph_from_adjacency_matrix(net_G2021,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2021,vertex.label = V(g_G2021)$name,edge.width = E(g_G2021)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```

plot + threshold

```{r}
threshold_G2021<- mean(net_G2021, na.rm = TRUE) +  sd(net_G2021, na.rm = TRUE)

gfilter_G2021<- delete_edges(g_G2021, E(g_G2021)[weight <threshold_G2021]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2021, edge.width = E(gfilter_G2021)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

###vertex attributes

```{r}
mother_ids_G2021<- mother_daugther_G2021$rfid_mom
daughter_ids_G2021<- mother_daugther_G2021$rfid_off

V(g_G2021)$type <- "other"  # default value
V(g_G2021)[V(g_G2021)$name %in% mother_ids_G2021]$type   <- "mother-daughter"
V(g_G2021)[V(g_G2021)$name %in% daughter_ids_G2021]$type <- "mother-daughter"

plot(gfilter_G2021, edge.width = E(gfilter_G2021)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2021)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2021)$name,edge.color="plum3")
```

```{r}
md_lookup_G2021<- mother_daugther_G2021%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2021<- match(V(g_G2021)$name, md_lookup_G2021$ID)   
V(g_G2021)$pairID <- md_lookup_G2021$pairID[IDs_G2021] 

pair_IDs_G2021<- factor(V(g_G2021)$pairID)

n_pairs_G2021<-length(levels(na.omit(pair_IDs_G2021)))
my_palette_G2021<-brewer.pal(n = max(3,n_pairs_G2021), name = "Paired") 

pair_colors_G2021<- my_palette_G2021[pair_IDs_G2021]
pair_colors_G2021[is.na(pair_colors_G2021)] <- "gray80"
```

big plot

```{r}
plot(gfilter_G2021, edge.width = E(gfilter_G2021)$weight*6,edge.curved=F,vertex.color =pair_colors_G2021, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2021<- levels(na.omit(pair_IDs_G2021))
legend_colors_G2021 <- my_palette_G2021[seq_along(pair_levels_G2021)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2021), "random pair"),
  col    = c(legend_colors_G2021, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_G2021 <- igraph::as_data_frame(g_G2021, what = "edges")

pid <- V(g_G2021)$pairID
names(pid) <- V(g_G2021)$name

from_pid <- pid[ edge_G2021$from ]
to_pid   <- pid[ edge_G2021$to ]

edge_G2021$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_G2021, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2021/full_edge_G2021.csv", row.names = FALSE)
```

# GB2 2022

```{r import}
GB2_2022<-read.csv("gs_GB2_2022.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)
colnames(GB2_2022)[colnames(GB2_2022) == "Date"] <- "date"
GB2_2022 <- GB2_2022 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))
GB2_2022$date <- gsub("/", ".", GB2_2022$date)
```

Change some individual ID
```{r}
names(GB2_2022)[names(GB2_2022) == '1270'] <- 'EE51'
```

####GBI

```{r}
colnames(GB2_2022)<-as.character(colnames(GB2_2022))
data_G2022<- GB2_2022%>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

data_G2022$date<- as.Date(data_G2022$date, format = "%d.%m.%Y")
data_G2022$daygroup<-as.factor(paste(data_G2022$date, data_G2022$box, sep="_"))

ind_G2022<-data.table(ID=c(data_G2022$ID), group=data_G2022$daygroup, day=c(data_G2022$date)) #df with 3 col: ID, group and day 
GBI_G2022<- get_group_by_individual(ind_G2022, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2022<- as.matrix(GBI_G2022)
write.csv(GBI_G2022, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2022/full_GBI_G2022.csv", row.names = TRUE)
```

###network

```{r}
net_G2022<-get_network(GBI_G2022,data_format = "GBI",association_index = "SRI")
write.csv(net_G2022, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2022/full_net_G2022.csv", row.names = TRUE)
```

```{r}
g_G2022<- graph_from_adjacency_matrix(net_G2022,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2022,vertex.label = V(g_G2022)$name,edge.width = E(g_G2022)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3")
```

plot + threshold

```{r}
threshold_G2022<- mean(net_G2022, na.rm = TRUE) +  sd(net_G2022, na.rm = TRUE)

gfilter_G2022<- delete_edges(g_G2022, E(g_G2022)[weight <threshold_G2022]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2022, edge.width = E(gfilter_G2022)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

####vertex attributes

```{r}
mother_ids_G2022<- mother_daugther_G2022$rfid_mom
daughter_ids_G2022<- mother_daugther_G2022$rfid_off

V(g_G2022)$type <- "other"  # default value
V(g_G2022)[V(g_G2022)$name %in% mother_ids_G2022]$type   <- "mother-daughter"
V(g_G2022)[V(g_G2022)$name %in% daughter_ids_G2022]$type <- "mother-daughter"

plot(gfilter_G2022, edge.width = E(gfilter_G2022)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2022)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2022)$name,edge.color="plum3")
```

```{r}
md_lookup_G2022<- mother_daugther_G2022%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2022<- match(V(g_G2022)$name, md_lookup_G2022$ID)
V(g_G2022)$pairID <- md_lookup_G2022$pairID[IDs_G2022] 

pair_IDs_G2022<- factor(V(g_G2022)$pairID)

n_pairs_G2022<-length(levels(na.omit(pair_IDs_G2022)))
my_palette_G2022 <-  rainbow(19) 

pair_colors_G2022<- my_palette_G2022[pair_IDs_G2022]
pair_colors_G2022[is.na(pair_colors_G2022)] <- "gray80"
```

big plot

```{r}
plot(gfilter_G2022, edge.width = E(gfilter_G2022)$weight*6,edge.curved=F,vertex.color =pair_colors_G2022, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2022<- levels(na.omit(pair_IDs_G2022))
legend_colors_G2022 <- my_palette_G2022[seq_along(pair_levels_G2022)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2022), "random pair"),
  col    = c(legend_colors_G2022, "gray80"),
  pch    = 19,
  pt.cex = 2,
  cex = 0.6,
  bty    = "n"
)
```

####edge list

```{r}
edge_G2022<- igraph::as_data_frame(g_G2022, what = "edges")

pid <- V(g_G2022)$pairID
names(pid) <- V(g_G2022)$name

from_pid <- pid[ edge_G2022$from ]
to_pid   <- pid[ edge_G2022$to ]

edge_G2022$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
#write.csv(edge_G2022, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2022/full_edge_G2022.csv", row.names = FALSE)

```

# GB2 2023

```{r}
GB2_2023<-read.csv("gs_GB2_2023.csv",header=TRUE, row.names=NULL,sep=";",check.names = F)
colnames(GB2_2023)[colnames(GB2_2023) == "Date"] <- "date"
GB2_2023 <- GB2_2023 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))
GB2_2023$date <- gsub("/", ".", GB2_2023$date)
```

####GBI 

```{r}
colnames(GB2_2023)<-as.character(colnames(GB2_2023))
data_G2023<- GB2_2023%>%
  mutate(across(-date, as.character)) %>%
  pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_G2023$date<- as.Date(data_G2023$date, format = "%d.%m.%Y")
data_G2023$daygroup<-as.factor(paste(data_G2023$date, data_G2023$box, sep="_"))

ind_G2023<-data.table(ID=c(data_G2023$ID), group=data_G2023$daygroup, day=c(data_G2023$date)) #df with 3 col: ID, group and day 
GBI_G2023<- get_group_by_individual(ind_G2023, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_G2023<- as.matrix(GBI_G2023)
write.csv(GBI_G2023, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2023/full_GBI_G2023.csv", row.names = TRUE)
```

network

```{r}
net_G2023<-get_network(GBI_G2023,data_format = "GBI",association_index = "SRI")
write.csv(net_G2023, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2023/full_net_G2023.csv", row.names = TRUE)
```

```{r}
g_G2023<- graph_from_adjacency_matrix(net_G2023,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_G2023,vertex.label = V(g_G2023)$name,edge.width = E(g_G2023)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```

plot + threshold

```{r}
threshold_G2023<- mean(net_G2023, na.rm = TRUE) +  sd(net_G2023, na.rm = TRUE)

gfilter_G2023<- delete_edges(g_G2023, E(g_G2023)[weight <threshold_G2023]) #either the threshold or the mean? depending on the graph wanted
plot(gfilter_G2023, edge.width = E(gfilter_G2023)$weight*5,edge.curved=T,vertex.color = "lightblue", vertex.size = 15, vertex.label.cex = 0.7,edge.color = "plum3")
```

vertex attributes

```{r}
mother_ids_G2023<- mother_daugther_G2023$rfid_mom
daughter_ids_G2023<- mother_daugther_G2023$rfid_off

V(g_G2023)$type <- "other"  # default value
V(g_G2023)[V(g_G2023)$name %in% mother_ids_G2023]$type   <- "mother-daughter"
V(g_G2023)[V(g_G2023)$name %in% daughter_ids_G2023]$type <- "mother-daughter"

plot(gfilter_G2023, edge.width = E(gfilter_G2023)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_G2023)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_G2023)$name,edge.color="plum3")
```

```{r}
md_lookup_G2023<- mother_daugther_G2023%>%pivot_longer(cols = c("rfid_mom","rfid_off"), values_to = "ID") %>%select(ID, pairID)

IDs_G2023<- match(V(g_G2023)$name, md_lookup_G2023$ID) 
V(g_G2023)$pairID <- md_lookup_G2023$pairID[IDs_G2023] 

pair_IDs_G2023<- factor(V(g_G2023)$pairID)

n_pairs_G2023<-length(levels(na.omit(pair_IDs_G2023)))
my_palette_G2023 <-  rainbow(19) 

pair_colors_G2023<- my_palette_G2023[pair_IDs_G2023]
pair_colors_G2023[is.na(pair_colors_G2023)] <- "gray80"
```

big plot

```{r}
plot(gfilter_G2023, edge.width = E(gfilter_G2023)$weight*6,edge.curved=F,vertex.color =pair_colors_G2023, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_G2023<- levels(na.omit(pair_IDs_G2023))
legend_colors_G2023 <- my_palette_G2023[seq_along(pair_levels_G2023)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_G2023), "random pair"),
  col    = c(legend_colors_G2023, "gray80"),
  pch    = 19,
  pt.cex = 2,
  cex = 0.6,
  bty    = "n"
)
```

####edge list

```{r}
edge_G2023 <- igraph::as_data_frame(g_G2023, what = "edges")

pid <- V(g_G2023)$pairID
names(pid) <- V(g_G2023)$name

from_pid <- pid[ edge_G2023$from ]
to_pid   <- pid[ edge_G2023$to ]

edge_G2023$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_G2023, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2023/full_edge_G2023.csv", row.names = FALSE)
```

#ALL years
```{r}
edge_G2013$year<-2013;edge_G2014$year<-2014;edge_G2015$year<-2015;edge_G2016$year<-2016;edge_G2017$year<-2017;edge_G2018$year<-2018;edge_G2019$year<-2019;edge_G2020$year<-2020;edge_G2021$year<-2021;edge_G2022$year<-2022;edge_G2023$year<-2023

edges_GB2<- rbind(edge_G2013, edge_G2014, edge_G2015, edge_G2016, edge_G2017, edge_G2018,
                        edge_G2019, edge_G2020, edge_G2021, edge_G2022, edge_G2023)
write.csv(edges_GB2, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/full_edges_GB2.csv", row.names = FALSE)
```

###mother-daughter df with all the years
```{r}
library(dplyr)

years <- 2013:2023  
mother_daughter_GB2 <- data.frame()  

#loop through each year
for (year in years) {
  obj_name <- paste0("mother_daugther_G", year) 
  
  if (exists(obj_name)) { 
    mother_daugther_pairs <- get(obj_name)  
    
    #add pairID and year columns
    mother_daugther_pairs <- mother_daugther_pairs %>%
      mutate(pairID = paste(rfid_mom, rfid_off, sep = "-"), year = year)
    
    #store results in final table
    mother_daughter_GB2 <- bind_rows(mother_daughter_GB2, mother_daugther_pairs)
  } else {
    message(paste("Warning: Object", obj_name, "does not exist"))
  }
}

write.csv(mother_daughter_GB2, "/Users/flo_r/Desktop/master thesis/data/data in CSV/small_mother_daughter_GB2.csv", row.names = FALSE)
#write.csv(mother_daughter_GB2, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/small_mother_daughter_GB2.csv", row.names = FALSE)
```


# LAR analysis 
Data loaded per year, in "long format"

2013
```{r}
#group = date+box
data_G2013$daygroup<-as.factor(paste(data_G2013$date, data_G2013$box, sep="_"))
write.csv(data_G2013, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2013/LARdata_G2013.csv", row.names = FALSE)
```
2014
```{r}
data_G2014$daygroup<-as.factor(paste(data_G2014$date, data_G2014$box, sep="_"))
write.csv(data_G2014, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2014/LARdata_G2014.csv", row.names = FALSE)
```

2015
```{r}
data_G2015$daygroup<-as.factor(paste(data_G2015$date, data_G2015$box, sep="_"))
write.csv(data_G2015, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2015/LARdata_G2015.csv", row.names = FALSE)
```

2016
```{r}
data_G2016$daygroup<-as.factor(paste(data_G2016$date, data_G2016$box, sep="_"))
write.csv(data_G2016, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2016/LARdata_G2016.csv", row.names = FALSE)
```

2017
```{r}
data_G2017$daygroup<-as.factor(paste(data_G2017$date, data_G2017$box, sep="_"))
write.csv(data_G2017, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2017/LARdata_G2017.csv", row.names = FALSE)
```

2018
```{r}
data_G2018$daygroup<-as.factor(paste(data_G2018$date, data_G2018$box, sep="_"))
write.csv(data_G2018, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2018/LARdata_G2018.csv", row.names = FALSE)
```
2019
```{r}
data_G2019$daygroup<-as.factor(paste(data_G2019$date, data_G2019$box, sep="_"))
write.csv(data_G2019, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2019/LARdata_G2019.csv", row.names = FALSE)
```
2020
```{r}
data_G2020$daygroup<-as.factor(paste(data_G2020$date, data_G2020$box, sep="_"))
write.csv(data_G2020, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2020/LARdata_G2020.csv", row.names = FALSE)
```

2021
```{r}
data_G2021$daygroup<-as.factor(paste(data_G2021$date, data_G2021$box, sep="_"))
write.csv(data_G2021, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2021/LARdata_G2021.csv", row.names = FALSE)
```
2022
```{r}
data_G2022$daygroup<-as.factor(paste(data_G2022$date, data_G2022$box, sep="_"))
write.csv(data_G2022, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2022/LARdata_G2022.csv", row.names = FALSE)
```

2023
```{r}
data_G2023$daygroup<-as.factor(paste(data_G2023$date, data_G2023$box, sep="_"))
write.csv(data_G2023, "/Users/floremichel/Desktop/THESIS/analysis/variables_GB2/2023/LARdata_G2023.csv", row.names = FALSE)
```

#---------------------UA

```{r}
mother_daughter_UA<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/mother_daughter_UA.csv",header=TRUE,sep=",",check.names = F)
mother_daughter_UA<-as.data.frame(mother_daughter_UA)
```

#UA 2013

```{r, include= FALSE}
UA_2013<-read.csv("UA_2013_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2013<-UA_2013[c(1:31)]
UA_2013[UA_2013 == "juv"] <- NA

UA_2013 <- UA_2013 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2013 <- UA_2013 %>%
  mutate(across(-date, as.character))

```

### GBI matrix

```{r GBI matrix}
data_U2013<- UA_2013 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2013$date<- as.Date(data_U2013$date, format = "%d.%m.%Y")
data_U2013$daygroup<-as.factor(paste(data_U2013$date, data_U2013$box, sep="_"))

ind_U2013<-data.table(ID=c(data_U2013$ID), group=data_U2013$daygroup, day=c(data_U2013$date)) #df with 3 col: ID, group and day 
GBI_U2013<- get_group_by_individual(ind_U2013, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2013 <- as.matrix(GBI_U2013)

write.csv(GBI_U2013, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2013.csv", row.names = TRUE)

```
###network

```{r}
net_U2013<-get_network(GBI_U2013,data_format = "GBI",association_index = "SRI")

write.csv(net_U2013,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2013", row.names = TRUE)
```

```{r}
g_U2013<- graph_from_adjacency_matrix(net_U2013,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2013,vertex.label = V(g_U2013)$name,edge.width = E(g_U2013)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2013<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2013]
daughter_ids_U2013<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2013]

V(g_U2013)$type <- "other"  # default value
V(g_U2013)[V(g_U2013)$name %in% mother_ids_U2013]$type   <- "mother-daughter"
V(g_U2013)[V(g_U2013)$name %in% daughter_ids_U2013]$type <- "mother-daughter"

plot(g_U2013, edge.width = E(g_U2013)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2013)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2013)$name,edge.color="plum3")
```

```{r}
md_lookup_U2013 <- mother_daughter_UA %>%
  filter(year == 2013) %>%                  
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2013<- match(V(g_U2013)$name, md_lookup_U2013$ID)   
V(g_U2013)$pairID <- md_lookup_U2013$pairID[IDs_U2013] 

pair_IDs_U2013<- factor(V(g_U2013)$pairID)

n_pairs_U2013<-length(levels(na.omit(pair_IDs_U2013)))
my_palette_U2013<-brewer.pal(n = max(3,n_pairs_U2013), name = "Paired") 

pair_colors_U2013<- my_palette_U2013[pair_IDs_U2013]
pair_colors_U2013[is.na(pair_colors_U2013)] <- "gray80"
```

big plot

```{r}
plot(g_U2013, edge.width = E(g_U2013)$weight*6,edge.curved=F,vertex.color =pair_colors_U2013, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2013<- levels(na.omit(pair_IDs_U2013))
legend_colors_U2013 <- my_palette_U2013[seq_along(pair_levels_U2013)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2013), "random pair"),
  col    = c(legend_colors_U2013, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2013 <- igraph::as_data_frame(g_U2013, what = "edges")

pid <- V(g_U2013)$pairID
names(pid) <- V(g_U2013)$name

from_pid <- pid[ edge_U2013$from ]
to_pid   <- pid[ edge_U2013$to ]

edge_U2013$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2013, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2013.csv", row.names = FALSE)
```

# UA 2014

```{r, include= FALSE}
UA_2014<-read.csv("UA_2014_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2014<-UA_2014[c(1:33)]
UA_2014[UA_2014 == "juv"] <- NA

UA_2014 <- UA_2014 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2014 <- UA_2014 %>%
  mutate(across(-date, as.character))

```

### GBI matrix

```{r GBI matrix}
data_U2014<- UA_2014 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2014$date<- as.Date(data_U2014$date, format = "%d.%m.%Y")
data_U2014$daygroup<-as.factor(paste(data_U2014$date, data_U2014$box, sep="_"))

ind_U2014<-data.table(ID=c(data_U2014$ID), group=data_U2014$daygroup, day=c(data_U2014$date)) #df with 3 col: ID, group and day 
GBI_U2014<- get_group_by_individual(ind_U2014, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2014 <- as.matrix(GBI_U2014)

write.csv(GBI_U2014, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2014.csv", row.names = TRUE)

```
###network

```{r}
net_U2014<-get_network(GBI_U2014,data_format = "GBI",association_index = "SRI")

write.csv(net_U2014,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2014", row.names = TRUE)
```

```{r}
g_U2014<- graph_from_adjacency_matrix(net_U2014,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2014,vertex.label = V(g_U2014)$name,edge.width = E(g_U2014)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2014<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2014]
daughter_ids_U2014<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2014]

V(g_U2014)$type <- "other"  # default value
V(g_U2014)[V(g_U2014)$name %in% mother_ids_U2014]$type   <- "mother-daughter"
V(g_U2014)[V(g_U2014)$name %in% daughter_ids_U2014]$type <- "mother-daughter"

plot(g_U2014, edge.width = E(g_U2014)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2014)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2014)$name,edge.color="plum3")
```

```{r}
md_lookup_U2014 <- mother_daughter_UA %>%
  filter(year == 2014) %>%                      
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2014<- match(V(g_U2014)$name, md_lookup_U2014$ID)  
V(g_U2014)$pairID <- md_lookup_U2014$pairID[IDs_U2014] 

pair_IDs_U2014<- factor(V(g_U2014)$pairID)

n_pairs_U2014<-length(levels(na.omit(pair_IDs_U2014)))
my_palette_U2014<-brewer.pal(n = max(3,n_pairs_U2014), name = "Paired") 

pair_colors_U2014<- my_palette_U2014[pair_IDs_U2014]
pair_colors_U2014[is.na(pair_colors_U2014)] <- "gray80"
```

big plot

```{r}
plot(g_U2014, edge.width = E(g_U2014)$weight*6,edge.curved=F,vertex.color =pair_colors_U2014, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2014<- levels(na.omit(pair_IDs_U2014))
legend_colors_U2014 <- my_palette_U2014[seq_along(pair_levels_U2014)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2014), "random pair"),
  col    = c(legend_colors_U2014, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2014 <- igraph::as_data_frame(g_U2014, what = "edges")

pid <- V(g_U2014)$pairID
names(pid) <- V(g_U2014)$name

from_pid <- pid[ edge_U2014$from ]
to_pid   <- pid[ edge_U2014$to ]

edge_U2014$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2014, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2014.csv", row.names = FALSE)
```

# UA 2015

```{r, include= FALSE}
UA_2015<-read.csv("UA_2015_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2015<-UA_2015[c(1:37)]

UA_2015 <- UA_2015 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2015 <- UA_2015 %>%
  mutate(across(-date, as.character))
```

### GBI matrix

```{r GBI matrix}
data_U2015<- UA_2015 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2015$date<- as.Date(data_U2015$date, format = "%d.%m.%Y")
data_U2015$daygroup<-as.factor(paste(data_U2015$date, data_U2015$box, sep="_"))

ind_U2015<-data.table(ID=c(data_U2015$ID), group=data_U2015$daygroup, day=c(data_U2015$date)) #df with 3 col: ID, group and day 
GBI_U2015<- get_group_by_individual(ind_U2015, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2015 <- as.matrix(GBI_U2015)

write.csv(GBI_U2015, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2015.csv", row.names = TRUE)

```
###network

```{r}
net_U2015<-get_network(GBI_U2015,data_format = "GBI",association_index = "SRI")

write.csv(net_U2015,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2015", row.names = TRUE)
```

```{r}
g_U2015<- graph_from_adjacency_matrix(net_U2015,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2015,vertex.label = V(g_U2015)$name,edge.width = E(g_U2015)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2015<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2015]
daughter_ids_U2015<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2015]

V(g_U2015)$type <- "other"  # default value
V(g_U2015)[V(g_U2015)$name %in% mother_ids_U2015]$type   <- "mother-daughter"
V(g_U2015)[V(g_U2015)$name %in% daughter_ids_U2015]$type <- "mother-daughter"

plot(g_U2015, edge.width = E(g_U2015)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2015)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2015)$name,edge.color="plum3")
```

```{r}
md_lookup_U2015 <- mother_daughter_UA %>%
  filter(year == 2015) %>%                      
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2015<- match(V(g_U2015)$name, md_lookup_U2015$ID)   
V(g_U2015)$pairID <- md_lookup_U2015$pairID[IDs_U2015] 

pair_IDs_U2015<- factor(V(g_U2015)$pairID)

n_pairs_U2015<-length(levels(na.omit(pair_IDs_U2015)))
my_palette_U2015<-brewer.pal(n = max(3,n_pairs_U2015), name = "Paired") 

pair_colors_U2015<- my_palette_U2015[pair_IDs_U2015]
pair_colors_U2015[is.na(pair_colors_U2015)] <- "gray80"
```

big plot

```{r}
plot(g_U2015, edge.width = E(g_U2015)$weight*6,edge.curved=F,vertex.color =pair_colors_U2015, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2015<- levels(na.omit(pair_IDs_U2015))
legend_colors_U2015 <- my_palette_U2015[seq_along(pair_levels_U2015)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2015), "random pair"),
  col    = c(legend_colors_U2015, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2015 <- igraph::as_data_frame(g_U2015, what = "edges")

pid <- V(g_U2015)$pairID
names(pid) <- V(g_U2015)$name

from_pid <- pid[ edge_U2015$from ]
to_pid   <- pid[ edge_U2015$to ]

edge_U2015$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2015, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2015.csv", row.names = FALSE)
```

#UA 2016
```{r, include= FALSE}
UA_2016<-read.csv("UA_2016_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2016<-UA_2016[c(1:27)]
UA_2016[UA_2016 == "juv"] <- NA

UA_2016 <- UA_2016 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2016 <- UA_2016 %>%
  mutate(across(-date, as.character))
```

### GBI matrix

```{r GBI matrix}
data_U2016<- UA_2016 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2016$date<- as.Date(data_U2016$date, format = "%d.%m.%Y")
data_U2016$daygroup<-as.factor(paste(data_U2016$date, data_U2016$box, sep="_"))

ind_U2016<-data.table(ID=c(data_U2016$ID), group=data_U2016$daygroup, day=c(data_U2016$date)) #df with 3 col: ID, group and day 
GBI_U2016<- get_group_by_individual(ind_U2016, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2016 <- as.matrix(GBI_U2016)

write.csv(GBI_U2016, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2016.csv", row.names = TRUE)

```
###network

```{r}
net_U2016<-get_network(GBI_U2016,data_format = "GBI",association_index = "SRI")

write.csv(net_U2016,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2016", row.names = TRUE)
```

```{r}
g_U2016<- graph_from_adjacency_matrix(net_U2016,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2016,vertex.label = V(g_U2016)$name,edge.width = E(g_U2016)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2016<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2016]
daughter_ids_U2016<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2016]

V(g_U2016)$type <- "other"  # default value
V(g_U2016)[V(g_U2016)$name %in% mother_ids_U2016]$type   <- "mother-daughter"
V(g_U2016)[V(g_U2016)$name %in% daughter_ids_U2016]$type <- "mother-daughter"

plot(g_U2016, edge.width = E(g_U2016)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2016)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2016)$name,edge.color="plum3")
```

```{r}
md_lookup_U2016 <- mother_daughter_UA %>%
  filter(year == 2016) %>%                      
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2016<- match(V(g_U2016)$name, md_lookup_U2016$ID)  
V(g_U2016)$pairID <- md_lookup_U2016$pairID[IDs_U2016] 

pair_IDs_U2016<- factor(V(g_U2016)$pairID)

n_pairs_U2016<-length(levels(na.omit(pair_IDs_U2016)))
my_palette_U2016<-brewer.pal(n = max(3,n_pairs_U2016), name = "Paired") 

pair_colors_U2016<- my_palette_U2016[pair_IDs_U2016]
pair_colors_U2016[is.na(pair_colors_U2016)] <- "gray80"
```

big plot

```{r}
plot(g_U2016, edge.width = E(g_U2016)$weight*6,edge.curved=F,vertex.color =pair_colors_U2016, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2016<- levels(na.omit(pair_IDs_U2016))
legend_colors_U2016 <- my_palette_U2016[seq_along(pair_levels_U2016)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2016), "random pair"),
  col    = c(legend_colors_U2016, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2016 <- igraph::as_data_frame(g_U2016, what = "edges")

pid <- V(g_U2016)$pairID
names(pid) <- V(g_U2016)$name

from_pid <- pid[ edge_U2016$from ]
to_pid   <- pid[ edge_U2016$to ]

edge_U2016$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2016, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2016.csv", row.names = FALSE)
```

#UA 2017
```{r, include= FALSE}
UA_2017<-read.csv("UA_2017_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2017<-UA_2017[c(1:34)]
UA_2017[UA_2017 == "juv"] <- NA

UA_2017 <- UA_2017 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2017 <- UA_2017 %>%
  mutate(across(-date, as.character))
```

### GBI matrix

```{r GBI matrix}
data_U2017<- UA_2017 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2017$date<- as.Date(data_U2017$date, format = "%d.%m.%Y")
data_U2017$daygroup<-as.factor(paste(data_U2017$date, data_U2017$box, sep="_"))

ind_U2017<-data.table(ID=c(data_U2017$ID), group=data_U2017$daygroup, day=c(data_U2017$date)) #df with 3 col: ID, group and day 
GBI_U2017<- get_group_by_individual(ind_U2017, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2017 <- as.matrix(GBI_U2017)

write.csv(GBI_U2017, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2017.csv", row.names = TRUE)

```
###network

```{r}
net_U2017<-get_network(GBI_U2017,data_format = "GBI",association_index = "SRI")

write.csv(net_U2017,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2017", row.names = TRUE)
```

```{r}
g_U2017<- graph_from_adjacency_matrix(net_U2017,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2017,vertex.label = V(g_U2017)$name,edge.width = E(g_U2017)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2017<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2017]
daughter_ids_U2017<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2017]

V(g_U2017)$type <- "other"  # default value
V(g_U2017)[V(g_U2017)$name %in% mother_ids_U2017]$type   <- "mother-daughter"
V(g_U2017)[V(g_U2017)$name %in% daughter_ids_U2017]$type <- "mother-daughter"

plot(g_U2017, edge.width = E(g_U2017)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2017)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2017)$name,edge.color="plum3")
```

```{r}
md_lookup_U2017 <- mother_daughter_UA %>%
  filter(year == 2017) %>%                      
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2017<- match(V(g_U2017)$name, md_lookup_U2017$ID)   
V(g_U2017)$pairID <- md_lookup_U2017$pairID[IDs_U2017] 

pair_IDs_U2017<- factor(V(g_U2017)$pairID)

n_pairs_U2017<-length(levels(na.omit(pair_IDs_U2017)))
my_palette_U2017<-brewer.pal(n = max(3,n_pairs_U2017), name = "Paired") 

pair_colors_U2017<- my_palette_U2017[pair_IDs_U2017]
pair_colors_U2017[is.na(pair_colors_U2017)] <- "gray80"
```

big plot

```{r}
plot(g_U2017, edge.width = E(g_U2017)$weight*6,edge.curved=F,vertex.color =pair_colors_U2017, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2017<- levels(na.omit(pair_IDs_U2017))
legend_colors_U2017 <- my_palette_U2017[seq_along(pair_levels_U2017)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2017), "random pair"),
  col    = c(legend_colors_U2017, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2017 <- igraph::as_data_frame(g_U2017, what = "edges")

pid <- V(g_U2017)$pairID
names(pid) <- V(g_U2017)$name

from_pid <- pid[ edge_U2017$from ]
to_pid   <- pid[ edge_U2017$to ]

edge_U2017$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2017, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2017.csv", row.names = FALSE)
```

#UA 2018

```{r, include= FALSE}
UA_2018<-read.csv("UA_2018_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2018<-UA_2018[c(1:40)]
UA_2018[UA_2018 == "juv"] <- NA

UA_2018 <- UA_2018 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2018 <- UA_2018 %>%
  mutate(across(-date, as.character))

names(UA_2018)[names(UA_2018) == "D6CC"] <- "FA5B"
names(UA_2018)[names(UA_2018) == "0B43"] <- "101F"
```

### GBI matrix

```{r GBI matrix}
data_U2018<- UA_2018 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2018$date<- as.Date(data_U2018$date, format = "%d.%m.%Y")
data_U2018$daygroup<-as.factor(paste(data_U2018$date, data_U2018$box, sep="_"))

ind_U2018<-data.table(ID=c(data_U2018$ID), group=data_U2018$daygroup, day=c(data_U2018$date)) #df with 3 col: ID, group and day 
GBI_U2018<- get_group_by_individual(ind_U2018, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2018 <- as.matrix(GBI_U2018)

write.csv(GBI_U2018, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2018.csv", row.names = TRUE)

```
###network

```{r}
net_U2018<-get_network(GBI_U2018,data_format = "GBI",association_index = "SRI")

write.csv(net_U2018,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2018", row.names = TRUE)
```

```{r}
g_U2018<- graph_from_adjacency_matrix(net_U2018,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2018,vertex.label = V(g_U2018)$name,edge.width = E(g_U2018)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2018<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2018]
daughter_ids_U2018<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2018]

V(g_U2018)$type <- "other"  # default value
V(g_U2018)[V(g_U2018)$name %in% mother_ids_U2018]$type   <- "mother-daughter"
V(g_U2018)[V(g_U2018)$name %in% daughter_ids_U2018]$type <- "mother-daughter"

plot(g_U2018, edge.width = E(g_U2018)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2018)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2018)$name,edge.color="plum3")
```

```{r}
md_lookup_U2018 <- mother_daughter_UA %>%
  filter(year == 2018) %>%                      
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2018<- match(V(g_U2018)$name, md_lookup_U2018$ID)   
V(g_U2018)$pairID <- md_lookup_U2018$pairID[IDs_U2018] 

pair_IDs_U2018<- factor(V(g_U2018)$pairID)

n_pairs_U2018<-length(levels(na.omit(pair_IDs_U2018)))
my_palette_U2018<-brewer.pal(n = max(3,n_pairs_U2018), name = "Paired") 

pair_colors_U2018<- my_palette_U2018[pair_IDs_U2018]
pair_colors_U2018[is.na(pair_colors_U2018)] <- "gray80"
```

big plot

```{r}
plot(g_U2018, edge.width = E(g_U2018)$weight*6,edge.curved=F,vertex.color =pair_colors_U2018, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2018<- levels(na.omit(pair_IDs_U2018))
legend_colors_U2018 <- my_palette_U2018[seq_along(pair_levels_U2018)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2018), "random pair"),
  col    = c(legend_colors_U2018, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2018 <- igraph::as_data_frame(g_U2018, what = "edges")

pid <- V(g_U2018)$pairID
names(pid) <- V(g_U2018)$name

from_pid <- pid[ edge_U2018$from ]
to_pid   <- pid[ edge_U2018$to ]

edge_U2018$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2018, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2018.csv", row.names = FALSE)
```

#UA 2019


```{r, include= FALSE}
UA_2019<-read.csv("UA_2019_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2019<-UA_2019[c(1:27)]
UA_2019[UA_2019 == "juv"] <- NA

UA_2019 <- UA_2019 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2019 <- UA_2019 %>%
  mutate(across(-date, as.character))
```

### GBI matrix

```{r GBI matrix}
data_U2019<- UA_2019 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2019$date<- as.Date(data_U2019$date, format = "%d.%m.%Y")
data_U2019$daygroup<-as.factor(paste(data_U2019$date, data_U2019$box, sep="_"))

ind_U2019<-data.table(ID=c(data_U2019$ID), group=data_U2019$daygroup, day=c(data_U2019$date)) #df with 3 col: ID, group and day 
GBI_U2019<- get_group_by_individual(ind_U2019, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2019 <- as.matrix(GBI_U2019)

write.csv(GBI_U2019, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2019.csv", row.names = TRUE)

```
###network

```{r}
net_U2019<-get_network(GBI_U2019,data_format = "GBI",association_index = "SRI")

write.csv(net_U2019,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2019", row.names = TRUE)
```

```{r}
g_U2019<- graph_from_adjacency_matrix(net_U2019,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2019,vertex.label = V(g_U2019)$name,edge.width = E(g_U2019)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2019<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2019]
daughter_ids_U2019<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2019]

V(g_U2019)$type <- "other"  # default value
V(g_U2019)[V(g_U2019)$name %in% mother_ids_U2019]$type   <- "mother-daughter"
V(g_U2019)[V(g_U2019)$name %in% daughter_ids_U2019]$type <- "mother-daughter"

plot(g_U2019, edge.width = E(g_U2019)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2019)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2019)$name,edge.color="plum3")
```

```{r}
md_lookup_U2019 <- mother_daughter_UA %>%
  filter(year == 2019) %>%                       
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2019<- match(V(g_U2019)$name, md_lookup_U2019$ID)   
V(g_U2019)$pairID <- md_lookup_U2019$pairID[IDs_U2019] 

pair_IDs_U2019<- factor(V(g_U2019)$pairID)

n_pairs_U2019<-length(levels(na.omit(pair_IDs_U2019)))
my_palette_U2019<-brewer.pal(n = max(3,n_pairs_U2019), name = "Paired") 

pair_colors_U2019<- my_palette_U2019[pair_IDs_U2019]
pair_colors_U2019[is.na(pair_colors_U2019)] <- "gray80"
```

big plot

```{r}
plot(g_U2019, edge.width = E(g_U2019)$weight*6,edge.curved=F,vertex.color =pair_colors_U2019, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2019<- levels(na.omit(pair_IDs_U2019))
legend_colors_U2019 <- my_palette_U2019[seq_along(pair_levels_U2019)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2019), "random pair"),
  col    = c(legend_colors_U2019, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2019 <- igraph::as_data_frame(g_U2019, what = "edges")

pid <- V(g_U2019)$pairID
names(pid) <- V(g_U2019)$name

from_pid <- pid[ edge_U2019$from ]
to_pid   <- pid[ edge_U2019$to ]

edge_U2019$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2019, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2019.csv", row.names = FALSE)
```

#UA 2020


```{r, include= FALSE}
UA_2020<-read.csv("UA_2020_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2020<-UA_2020[c(1:31)]
UA_2020[UA_2020 == "juv"] <- NA

UA_2020 <- UA_2020 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2020 <- UA_2020 %>%
  mutate(across(-date, as.character))


```

### GBI matrix

```{r GBI matrix}
data_U2020<- UA_2020 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2020$date<- as.Date(data_U2020$date, format = "%d.%m.%Y")
data_U2020$daygroup<-as.factor(paste(data_U2020$date, data_U2020$box, sep="_"))

ind_U2020<-data.table(ID=c(data_U2020$ID), group=data_U2020$daygroup, day=c(data_U2020$date)) #df with 3 col: ID, group and day 
GBI_U2020<- get_group_by_individual(ind_U2020, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2020 <- as.matrix(GBI_U2020)

write.csv(GBI_U2020, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2020.csv", row.names = TRUE)

```
###network

```{r}
net_U2020<-get_network(GBI_U2020,data_format = "GBI",association_index = "SRI")

write.csv(net_U2020,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2020", row.names = TRUE)
```

```{r}
g_U2020<- graph_from_adjacency_matrix(net_U2020,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2020,vertex.label = V(g_U2020)$name,edge.width = E(g_U2020)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2020<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2020]
daughter_ids_U2020<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2020]

V(g_U2020)$type <- "other"  # default value
V(g_U2020)[V(g_U2020)$name %in% mother_ids_U2020]$type   <- "mother-daughter"
V(g_U2020)[V(g_U2020)$name %in% daughter_ids_U2020]$type <- "mother-daughter"

plot(g_U2020, edge.width = E(g_U2020)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2020)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2020)$name,edge.color="plum3")
```

```{r}
md_lookup_U2020 <- mother_daughter_UA %>%
  filter(year == 2020) %>%                    
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2020<- match(V(g_U2020)$name, md_lookup_U2020$ID)  
V(g_U2020)$pairID <- md_lookup_U2020$pairID[IDs_U2020] 

pair_IDs_U2020<- factor(V(g_U2020)$pairID)

n_pairs_U2020<-length(levels(na.omit(pair_IDs_U2020)))
my_palette_U2020<-brewer.pal(n = max(3,n_pairs_U2020), name = "Paired") 

pair_colors_U2020<- my_palette_U2020[pair_IDs_U2020]
pair_colors_U2020[is.na(pair_colors_U2020)] <- "gray80"
```

big plot

```{r}
plot(g_U2020, edge.width = E(g_U2020)$weight*6,edge.curved=F,vertex.color =pair_colors_U2020, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2020<- levels(na.omit(pair_IDs_U2020))
legend_colors_U2020 <- my_palette_U2020[seq_along(pair_levels_U2020)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2020), "random pair"),
  col    = c(legend_colors_U2020, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2020 <- igraph::as_data_frame(g_U2020, what = "edges")

pid <- V(g_U2020)$pairID
names(pid) <- V(g_U2020)$name

from_pid <- pid[ edge_U2020$from ]
to_pid   <- pid[ edge_U2020$to ]

edge_U2020$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2020, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2020.csv", row.names = FALSE)
```

#UA 2021

```{r, include= FALSE}
UA_2021<-read.csv("UA_2021_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2021<-UA_2021[c(1:38)]
UA_2021[UA_2021 == "juv"] <- NA

UA_2021 <- UA_2021 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2021 <- UA_2021 %>%
  mutate(across(-date, as.character))

names(UA_2021)[names(UA_2021) == "9B39"] <- "F50E"
names(UA_2021)[names(UA_2021) == "37F2"] <- "100F"
```

### GBI matrix

```{r GBI matrix}
data_U2021<- UA_2021 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2021$date<- as.Date(data_U2021$date, format = "%d.%m.%Y")
data_U2021$daygroup<-as.factor(paste(data_U2021$date, data_U2021$box, sep="_"))

ind_U2021<-data.table(ID=c(data_U2021$ID), group=data_U2021$daygroup, day=c(data_U2021$date)) #df with 3 col: ID, group and day 
GBI_U2021<- get_group_by_individual(ind_U2021, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2021 <- as.matrix(GBI_U2021)

write.csv(GBI_U2021, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2021.csv", row.names = TRUE)

```
###network

```{r}
net_U2021<-get_network(GBI_U2021,data_format = "GBI",association_index = "SRI")

write.csv(net_U2021,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2021", row.names = TRUE)
```

```{r}
g_U2021<- graph_from_adjacency_matrix(net_U2021,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2021,vertex.label = V(g_U2021)$name,edge.width = E(g_U2021)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2021<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2021]
daughter_ids_U2021<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2021]

V(g_U2021)$type <- "other"  # default value
V(g_U2021)[V(g_U2021)$name %in% mother_ids_U2021]$type   <- "mother-daughter"
V(g_U2021)[V(g_U2021)$name %in% daughter_ids_U2021]$type <- "mother-daughter"

plot(g_U2021, edge.width = E(g_U2021)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2021)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2021)$name,edge.color="plum3")
```

```{r}
md_lookup_U2021 <- mother_daughter_UA %>%
  filter(year == 2021) %>%                     
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2021<- match(V(g_U2021)$name, md_lookup_U2021$ID)   
V(g_U2021)$pairID <- md_lookup_U2021$pairID[IDs_U2021] 

pair_IDs_U2021<- factor(V(g_U2021)$pairID)

n_pairs_U2021<-length(levels(na.omit(pair_IDs_U2021)))
my_palette_U2021<-brewer.pal(n = max(3,n_pairs_U2021), name = "Paired") 

pair_colors_U2021<- my_palette_U2021[pair_IDs_U2021]
pair_colors_U2021[is.na(pair_colors_U2021)] <- "gray80"
```

big plot

```{r}
plot(g_U2021, edge.width = E(g_U2021)$weight*6,edge.curved=F,vertex.color =pair_colors_U2021, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2021<- levels(na.omit(pair_IDs_U2021))
legend_colors_U2021 <- my_palette_U2021[seq_along(pair_levels_U2021)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2021), "random pair"),
  col    = c(legend_colors_U2021, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2021 <- igraph::as_data_frame(g_U2021, what = "edges")

pid <- V(g_U2021)$pairID
names(pid) <- V(g_U2021)$name

from_pid <- pid[ edge_U2021$from ]
to_pid   <- pid[ edge_U2021$to ]

edge_U2021$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2021, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2021.csv", row.names = FALSE)
```

#UA 2022

```{r, include= FALSE}
UA_2022<-read.csv("UA_2022_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2022<-UA_2022[c(1:35)]
UA_2022[UA_2022 == "juv"] <- NA

UA_2022 <- UA_2022 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2022 <- UA_2022 %>%
  mutate(across(-date, as.character))
```

### GBI matrix

```{r GBI matrix}
data_U2022<- UA_2022 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2022$date<- as.Date(data_U2022$date, format = "%d.%m.%Y")
data_U2022$daygroup<-as.factor(paste(data_U2022$date, data_U2022$box, sep="_"))

ind_U2022<-data.table(ID=c(data_U2022$ID), group=data_U2022$daygroup, day=c(data_U2022$date)) #df with 3 col: ID, group and day 
GBI_U2022<- get_group_by_individual(ind_U2022, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2022 <- as.matrix(GBI_U2022)

write.csv(GBI_U2022, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2022.csv", row.names = TRUE)

```
###network

```{r}
net_U2022<-get_network(GBI_U2022,data_format = "GBI",association_index = "SRI")

write.csv(net_U2022,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2022", row.names = TRUE)
```

```{r}
g_U2022<- graph_from_adjacency_matrix(net_U2022,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2022,vertex.label = V(g_U2022)$name,edge.width = E(g_U2022)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2022<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2022]
daughter_ids_U2022<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2022]

V(g_U2022)$type <- "other"  # default value
V(g_U2022)[V(g_U2022)$name %in% mother_ids_U2022]$type   <- "mother-daughter"
V(g_U2022)[V(g_U2022)$name %in% daughter_ids_U2022]$type <- "mother-daughter"

plot(g_U2022, edge.width = E(g_U2022)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2022)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2022)$name,edge.color="plum3")
```

```{r}
md_lookup_U2022 <- mother_daughter_UA %>%
  filter(year == 2022) %>%                      
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2022<- match(V(g_U2022)$name, md_lookup_U2022$ID)  
V(g_U2022)$pairID <- md_lookup_U2022$pairID[IDs_U2022] 

pair_IDs_U2022<- factor(V(g_U2022)$pairID)

n_pairs_U2022<-length(levels(na.omit(pair_IDs_U2022)))
my_palette_U2022<-brewer.pal(n = max(3,n_pairs_U2022), name = "Paired") 

pair_colors_U2022<- my_palette_U2022[pair_IDs_U2022]
pair_colors_U2022[is.na(pair_colors_U2022)] <- "gray80"
```

big plot

```{r}
plot(g_U2022, edge.width = E(g_U2022)$weight*6,edge.curved=F,vertex.color =pair_colors_U2022, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2022<- levels(na.omit(pair_IDs_U2022))
legend_colors_U2022 <- my_palette_U2022[seq_along(pair_levels_U2022)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2022), "random pair"),
  col    = c(legend_colors_U2022, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2022 <- igraph::as_data_frame(g_U2022, what = "edges")

pid <- V(g_U2022)$pairID
names(pid) <- V(g_U2022)$name

from_pid <- pid[ edge_U2022$from ]
to_pid   <- pid[ edge_U2022$to ]

edge_U2022$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2022, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2022.csv", row.names = FALSE)
```

#UA 2023


```{r, include= FALSE}
UA_2023<-read.csv("UA_2023_full_dayroost.csv",header=TRUE,sep=";",check.names = F,row.names=NULL)

UA_2023<-UA_2023[c(1:30)]
UA_2023[UA_2023 == "juv"] <- NA

UA_2023 <- UA_2023 %>%
  mutate(across(everything(), ~ replace(., . == "", NA)))

UA_2023 <- UA_2023 %>%
  mutate(across(-date, as.character))

mother_daughter_UA$rfid_off[mother_daughter_UA$rfid_off == "AE33"] <- "F50E"
```

### GBI matrix

```{r GBI matrix}
data_U2023<- UA_2023 %>%pivot_longer(
    cols = -c(date),  
    names_to = "ID",                     
    values_to = "box"                       
  ) %>%
  drop_na(box)

#group = date+box
data_U2023$date<- as.Date(data_U2023$date, format = "%d.%m.%Y")
data_U2023$daygroup<-as.factor(paste(data_U2023$date, data_U2023$box, sep="_"))

ind_U2023<-data.table(ID=c(data_U2023$ID), group=data_U2023$daygroup, day=c(data_U2023$date)) #df with 3 col: ID, group and day 
GBI_U2023<- get_group_by_individual(ind_U2023, data_format = "individuals") #get group by individual matrix with 0 or 1 when present
GBI_U2023 <- as.matrix(GBI_U2023)

write.csv(GBI_U2023, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/fullGBI_U2023.csv", row.names = TRUE)

```
###network

```{r}
net_U2023<-get_network(GBI_U2023,data_format = "GBI",association_index = "SRI")

write.csv(net_U2023,"C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_net_U2023", row.names = TRUE)
```

```{r}
g_U2023<- graph_from_adjacency_matrix(net_U2023,mode = "undirected",weighted = TRUE,diag = FALSE) 
plot(g_U2023,vertex.label = V(g_U2023)$name,edge.width = E(g_U2023)$weight * 6, layout = layout_with_fr, vertex.color = "lightblue",edge.color = "plum3") 
```
###vertex attributes

```{r}
mother_ids_U2023<- mother_daughter_UA$rfid_mom[mother_daughter_UA$year == 2023]
daughter_ids_U2023<- mother_daughter_UA$rfid_off[mother_daughter_UA$year == 2023]

V(g_U2023)$type <- "other"  # default value
V(g_U2023)[V(g_U2023)$name %in% mother_ids_U2023]$type   <- "mother-daughter"
V(g_U2023)[V(g_U2023)$name %in% daughter_ids_U2023]$type <- "mother-daughter"

plot(g_U2023, edge.width = E(g_U2023)$weight*6,layout = layout_with_fr,vertex.color = ifelse(V(g_U2023)$type == "mother-daughter",   "darkcyan", "gray"),vertex.label = V(g_U2023)$name,edge.color="plum3")
```

```{r}
md_lookup_U2023 <- mother_daughter_UA %>%
  filter(year == 2023) %>%                    
  pivot_longer(cols = c("rfid_mom", "rfid_off"), values_to = "ID") %>%
  select(ID, pairID)

IDs_U2023<- match(V(g_U2023)$name, md_lookup_U2023$ID)   
V(g_U2023)$pairID <- md_lookup_U2023$pairID[IDs_U2023] 

pair_IDs_U2023<- factor(V(g_U2023)$pairID)

n_pairs_U2023<-length(levels(na.omit(pair_IDs_U2023)))
my_palette_U2023<-brewer.pal(n = max(3,n_pairs_U2023), name = "Paired") 


pair_colors_U2023<- my_palette_U2023[pair_IDs_U2023]
pair_colors_U2023[is.na(pair_colors_U2023)] <- "gray80"
```

big plot

```{r}
plot(g_U2023, edge.width = E(g_U2023)$weight*6,edge.curved=F,vertex.color =pair_colors_U2023, vertex.size = 15, vertex.label.cex = 0.7)
pair_levels_U2023<- levels(na.omit(pair_IDs_U2023))
legend_colors_U2023 <- my_palette_U2023[seq_along(pair_levels_U2023)]

legend(
  "topleft",
  legend = c(paste("Pair", pair_levels_U2023), "random pair"),
  col    = c(legend_colors_U2023, "gray80"),
  pch    = 19,
  pt.cex = 2,
  bty    = "n"
)
```

###edge list

```{r}
edge_U2023 <- igraph::as_data_frame(g_U2023, what = "edges")

pid <- V(g_U2023)$pairID
names(pid) <- V(g_U2023)$name

from_pid <- pid[ edge_U2023$from ]
to_pid   <- pid[ edge_U2023$to ]

edge_U2023$pair_type <- ifelse(
  !is.na(from_pid) & !is.na(to_pid) & (from_pid == to_pid),
  "mother-daughter",
  "other"
)
write.csv(edge_U2023, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2023.csv", row.names = FALSE)
```

#colony size
```{r}
gbi_objects <- ls(pattern = "^GBI_U\\d{4}$")

colony_size_UA <- data.frame(
  year = numeric(),
  colony_size = numeric()
)

for (obj in gbi_objects) {
  gbi_data <- get(obj)
  year <- as.numeric(gsub("GBI_U", "", obj))
  
  n_individuals <- ncol(gbi_data)
  
  colony_size_UA <- rbind(colony_size_UA, data.frame(
    year = year,
    colony_size = n_individuals
  ))
}

colony_size_UA <- colony_size_UA[order(colony_size_UA$year), ]

write.csv(colony_size_UA, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/colony_size_UA.csv", row.names = FALSE)
```
#import all full edges
```{r}
edge_U2013<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2013.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2014<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2014.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2015<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2015.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2016<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2016.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2017<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2017.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2018<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2018.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2019<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2019.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2020<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2020.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2021<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2021.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2022<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2022.csv",header=TRUE,sep=",",check.names = F,row.names=NULL); edge_U2023<-read.csv("C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_U2023.csv",header=TRUE,sep=",",check.names = F,row.names=NULL)
```
#ALL years
```{r}
edge_U2013$year<-2013;edge_U2014$year<-2014;edge_U2015$year<-2015;edge_U2016$year<-2016;edge_U2017$year<-2017;edge_U2018$year<-2018;edge_U2019$year<-2019;edge_U2020$year<-2020;edge_U2021$year<-2021;edge_U2022$year<-2022;edge_U2023$year<-2023

full_edges_UA<- rbind(edge_U2013, edge_U2014, edge_U2015, edge_U2016, edge_U2017, edge_U2018, edge_U2019, edge_U2020, edge_U2021, edge_U2022, edge_U2023)

write.csv(full_edges_UA, "C:/Users/flo_r/Desktop/master thesis/analysis/variables UA/full_UA/full_edges_UA.csv", row.names = FALSE)
```



